### 2010-04-19 plan.tex
initial commit

\documentclass{article}

\begin{document}

\section{Bikes}
Several bicycles will be measured to get a nice spread of property data from different bicycles. These are the possible choices:
\begin{itemize}
	\item Batavus Browser
	\item Batavus Browser with DAQ equipment
	\item Batavus Stratos Deluxe
	\item Batavus Crescendo with solid fork
	\item Yellow bike
	\item Bianchi Pista
	\item Jodi's mountain bike
	\item Miriam's road bike
\end{itemize}

\section{Geometry}
\begin{description}
	\item[$w$ wheelbase] fix the steer angle to zero by tying the wheel to the downtube. Using a ruler measure from the center of the front wheel to the center of the rear wheel.
	\item[$r_\mathrm{F}$ front wheel radius] Pump the tires to the recommended inflation pressure. Mark a fine line on the tire at the ground contact point (Use a square to find ground contact). Roll the bicycle through as many full revolutions that can be measured with a single measuring tape. Do this with and without a rider. Note the mass of the rider, revolutions and the distance. $r_\mathrm{F}=\frac{d}{2\pi n}$
	\item[$r_\mathrm{R}$ rear wheel radius] Same as the front wheel.
	\item[$\lambda$ steer axis angle] fix the bicycle such that the roll angle and steer angle are both zero. Affix a straightedge to the head tube such that extends to the ground (possibly design a universal attachment that takes the guessing out). Use the digital angle gage to measure the angle between the floor and the straightedge. $\lambda=\frac{\pi}{2}-\lambda_{ht}$
	\item[$c$ trail] After measuring the head tube angle, use a square to locate the ground contact point relative to the center of the wheel. Measure the distance from the ground contact to the intersection of the headtube straightedge and the ground. Also, remove the fork from the bicycle and mount it horizontally by the steer tube on a table. Measure the fork offset using a height gage.
\end{description}

\section{Mass}
Measure the mass of the entire bicycle by hanging it on a hook scale. Measure each individual part using a tabletop scale.

\section{Center of Mass}

\section{Moment of inertia}
\subsection{Torsional Pendulum}
The torsional pendulum will be calibrated with a long rod of known inertia each time the pendulum is set up to determine the coefficient of elasticity of the rod/fixture.
\subsection{Frame}
\end{document}

### 2010-04-19 report/parameterReport.tex
initial commit

\documentclass{article}
\usepackage{hyperref}
\usepackage{graphicx}

\begin{document}
\section{Bicycle}

\subsection{Geometry}
\subsubsection{Wheel radius}
The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were measured by measuring the linear distance traversed along the ground through eitehr 13 or 14 rotations of the wheel. Each wheel was measured separatley and the measurements were taken with a 72kg rider seated on the bicycle. A 30 meter tape measure (resolution: 2mm) was taped on a flat level smooth floor. The tire was marked with chalk and aligned with the tape measure. The accuracy of the distance measurement is approximately $\pm0.005$m. The tires were pump the recommended inflation pressure before the measurments. The wheel radius is calculated by
\begin{equation}
	r=\frac{\textrm{distance}}{2\cdot\pi\cdot\textrm{rotations}}
	\label{eq:wheelRadius}
\end{equation}
\begin{figure}[tb]
	\begin{center}
		\includegraphics[width=4in]{tireChalk.jpg}
	\end{center}
	\caption{Wheel and tire with chalk mark aligned to the tape measure.}
	\label{fig:tireChalk}
\end{figure}
\begin{table}
	\begin{tabular}{llllllll}
	Bicycle   & B      & B*     & C      & G      & P      & S      & Y \& Y*\\
	Front\\
	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 4      \\
	Distance  & 28.060 & 27.980 & 27.980 & 29.044 & 29.366 & 27.772 & 27.925 \\
	Rotations & 13     & 13     & 13     & 14     & 14     & 13     & 13     \\
	Radius    & 0.344  & 0.343  & 0.343  & 0.330  & 0.334  & 0.340  & 0.342  \\
	Rear\\
	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 3.5    \\
	Distance  & 27.850 & 27.835 & 27.768 & 29.788 & 29.212 & 29.774 & 27.884 \\
	Rotations & 13     & 13     & 13     & 14     & 14     & 14     & 13     \\
	Radius    & 0.341  & 0.341  & 0.340  & 0.339  & 0.332  & 0.338  & 0.341	
	\end{tabular}
	\caption{}
	\label{tab:wheelRadius}
\end{table}
\subsubsection{Head tube angle}
The head tube angle was measured directly using an electronic level with a $\pm0.2^{\circ}$ accuracy. The bicycle was fixed vertically, the front wheel was aligned with the rear frame and the bicycle was unloaded. The steer axis tilt $\lambda$ is the complement to the head tube angle.
\begin{equation}
	\lambda=\frac{\pi}{180}(90^{\circ}-\lambda_{ht})
\label{eq:headTubeAngle}
\end{equation}
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle   & B      & B*     & C      & G      & P      & S      & Y      & Y*\\
	Head tube angle & 67.1 & 67.1 & 69.0 & 71.1 & 74.2 & 73.1 & 72.7 & 70.6\\
	Steer axis tilt & 0.400 & 0.400 & 0.367 & 0.330 & 0.276 & 0.295 & 0.302 & 0.339
	\end{tabular}
	\caption{Head tube angle and steer axis tilt}
	\label{tab:lambda}
\end{table}

\subsubsection{Trail}
Trail is difficult to measure directly so we instead chose to measure the fork offset. The fork offset was measured by clamping the steer tube of the front fork into a v-block on a flat table. A ruler was used to measure the height of the center of the head tube and the height of the center of the axle axis. The fork blades where aligned such that the axle axis was parallel to the table surface.
\begin{equation}
	c=\frac{r_\mathrm{F}\sin{\lambda}-f_o}{\cos{\lambda}}
	\label{eq:trail}
\end{equation}
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle     & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
	Fork offset & 0.071 & 0.071 & 0.045 & 0.039 & 0.032 & 0.045 & 0.057 & -0.057\\
	Trail       & 0.069 & 0.068 & 0.083 & 0.072 & 0.062 & 0.056 & 0.047 & 0.180
	\end{tabular}
	\caption{Head tube angle and steer axis tilt}
	\label{tab:trail}
\end{table}
\subsubsection{Wheelbase}
The wheelbase was measured directly using a tape measure while the bicycle was in the same uprigth fixed position used when measuring the head tube angle.
\begin{table}
	\begin{tabular}{lllllllll}
	Bicycle    & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
	Wheel base & 1.121 & 1.121 & 1.101 & 1.070 & 0.989 & 1.037 & 1.089 & 0.985
	\end{tabular}
	\caption{Wheelbase values.}
	\label{tab:wheelbase}
\end{table}

\subsection{Mass}
The total mass of each bicycle was measured using a spring scale with a resolution of 100g. Then each of the four bicycle parts were measured using a Molen 20kg scale with a resulution of 20g.
\begin{table}
	\begin{tabular}{lllllllll}
		Bicycle     & B     & B*    & C     & G     & P    & S     & Y     & Y*\\
		total       & 18.5  & 23.5  & 20.9  & 10.5  & 9.9  & 17.6  & 10.2  & 10.2\\
		frame       & 9.86  & 14.71 & 9.18  & 4.48  & 4.49 & 7.22  & 3.31  & 3.31\\
		fork        &       & 3.22  & 4.57  & 2.52  & 2.27 & 3.04  & 2.45  & 2.45\\
		rear wheel  & 3.11  & 3.11  & 3.96  & 1.94  & 1.38 & 3.96  & 2.57  & 2.57\\
		front wheel & 2.02  & 2.02  & 3.55  & 1.50  & 1.58 & 3.33  & 1.90  & 1.90\\
		total sum   & 18.21 & 23.06 & 21.26 & 10.44 & 9.72 & 17.55 & 10.23 & 10.23
	\end{tabular}
	\caption{Mass of bicycles and parts}
	\label{tab:mass}
\end{table}
\subsection{Center of Mass}
The centers of mass of the wheels are assumed to be at their geometrical centers.
\subsubsection{Rear frame}
The rear frame was hung in three orientations from a torsional pendulum. The frame could rotate about a joint such that gravity aligned the center of mass with the pendulum axis. The new orientaion angle of the headtube $\alpha$ was measured using the digital level. A string was aligned with the pendulum axis and allowed to pass by the frame. The horizontal distance $d$ between the rear axle and the string (CoM line) was measured by aligning a ruler pendicular to the string. The distance $d$ was negative if the string fell to the right of the rear axle and positive if it fell to the left of the rear axle. These measurements allow for the calucation of the center of mass location in the global reference frame.
The frame rotation angle $\beta$ is defined as rotation of the frame in the nominal configuration to the hanging orientation, rotated about the $Y$ axis.
\begin{equation}
	\beta=\alpha-\lambda-\frac{\pi}{2}
\label{eq:frameRotAng}
\end{equation}
The CoM line is simply a line in the global reference frame with a slope $m$ and a z-intercept $b$. The slope can be shown to be
\begin{equation}
	m=\tan{\left(\beta-\frac{\pi}{2}\right)}
\label{eq:slope}
\end{equation}
The z-intercept can be shown to be
\begin{equation}
	b=-\frac{d}{\sin{\beta}}-r_\mathrm{R}
\label{eq:zInt}
\end{equation}
The CoM location can be calculated by finding the intersection of these three lines. The system is over determined so the least squares method is used to find a unique solution.
\begin{equation}
	\left[
	\begin{array}{cc}
		-m_1 & 1\\
		-m_2 & 1\\
		-m_3 & 1
	\end{array}
	\right]
	\left[
	\begin{array}{c}
		x\\
		z
	\end{array}
	\right]
	=
	\left[
	\begin{array}{c}
		b_1\\
		b_2\\
		b_3
	\end{array}
	\right]	
\label{eq:solveSlopeInt}
\end{equation}
\subsection{Moment of Inertia}
\section{Human}
\appendix
\subsection{Bicycle Descriptions}
\subsubsection{Yellow Bicycle}
The yellow is a bicycle used in the lab to demonstrate that a bicycle is stable at certain speeds. It is an aluminum road frame of unknown make with the majority of components removed. The wheels, drop handlebar, seat, seat post and bottom bracket are the only parts on the bike. The fork has been reversed to ''increase`` the stability of the bicycle. The bicycle was measured with both the fork in normal position and the fork reversed.
\begin{table}
	\begin{tabular}{ll}
		Style & Road\\
		Size & \\
		Wheels & 700c Aluminum Rims\\
		Tires & 28 x 1 3/8 x 1 5/8\\
		Sprocket & 6 speed free wheel\\
		Handlebars & Drop\\
	\end{tabular}
	\caption{}
	\label{yellowBikeSpecs}
\end{table}

\subsubsection{Batavus Browser}
The Batavus Browser is lower priced Dutch city bike. We measured the physical properities of the stock Browser model and also the same bike with the instrumentation used in our experiments.
\begin{table}
	\begin{tabular}{ll}
		Price          & € 449\\
		Wheel Size     & 28''\\
		Weight         & 18.7 kg\\
		Frame Size     & 54\\
		Frame material & High Tensile Steel\\
		Fork material  & Steel\\
		Rims           & Aluminum Airline\\
    Spokes         & Stainless steel\\
		Tires          & Trax Highway, 28'' x 1 5/8 x 1 3/8\\
		Handlebar      & Steel painted\\
		Handles        & Batavus Comfort\\
		Saddle         & Selle Royal 8274\\
		Dynamo         & Basta Trio\\
		Headlight      & Batavus Logic Light\\
		Rearlight      & Battery light\\
		Pedals         & Batavus plastic antislip\\
		Final          & Trelock RS42\\
		Theft chip     & Present\\
		Crank          & Aluminum\\
		Rack           & Steel\\
		Quick Binder   & Binder Trio\\
		Mudguards      & Steel\\
		Bell           & Kenlight\\
	\end{tabular}
	\caption{2008 Batavus Browser Specifications}
	\label{browserSpecs}
\end{table}

\subsubsection{2007 Bianchi Pista}
The Pista is a steel track bicycle. Specifications, taken from the Bianchi website~\cite{Bianchi2009}, are as follows:
\begin{table}
	\begin{tabular}{ll}
		Style            & Track Bike\\
  	Size             & 57cm\\
  	Color            & Gang Green\\
  	Frame            & Bianchi DB CrMo, rear entry track dropouts\\
  	Fork             & CrMo\\
  	Retail Price     & \$579.99\\
		Headset          & VP AheadSet, 1''threadless\\
  	Handlebar        & Bianchi steel track, 26.0mm\\
		Stem             & Bianchi alloy\\
    Brakes/Levers    & Shimano Tiagra front brake/lever\\
		Crankset         & Truvativ Touro 1.1, 48T\\
		Bottom Bracket   & Truvativ Power Spline cartridge\\
		Chain            & KMC\\
		Sprocket         & 16T fixed cog \& 16T Freewheel\\
		Pedals           & Campus pedals\\
		Wheels           & Bianchi Alex Solo wheelset, 28h rims\\
    Tires            & Continental UltraSport, 700x23C\\
    Saddle           & Bianchi Velo\\
		Seatpost         & Bianchi alloy, 27.2mm\\
		Size             & 57cm\\ 
		Seat Tube        & 570mm\\
		Top Tube-actual  & 560mm\\
		Top Tube-virtual & 560mm\\
		Chainstay        & 383mm\\
		Fork Rake        & 28mm\\
		Head Tube Angle  & 74.5$^{\circ}$\\
		Seat Tube Angle  & 75.5$^{\circ}$\\
		Wheelbase        & 967mm\\
		Standover Height & 32''  
	\end{tabular}
	\caption{}
	\label{pistaSpecs}
\end{table}
\subsubsection{Gary Fisher Mountain Bike}
\subsubsection{Batavus Crescendo Deluxe}

The Batavus Crescendo Deluxe is a contemporary of all-round cycling in cool black and white. A great bike for a good end to touring. And that works fine with 8 gears, suspension fork and suspension seatpost.
\begin{table}
	\begin{tabular}{ll}
		Price             & from € 869\\
		Wheel Size        & 28''\\
		Weight            & 21.5 kg\\
		Frame Size        & 53\\
		Performance/speed & Shimano Nexus 8 speed with roller brakes\\
		Color             & White / black\\
		Brakes            & Shimano roller brakes (BR-IM50)\\
		Frame type        & Ladies mono tube frame Gentlemen trend\\
		Frame material    & Aluminum 7005\\
		Front Fork        & spring, Batavus\\
		Rims              & Aluminum Rodi VR19\\
		Spokes            & Stainless steel (adapted spoke pattern)\\
		Tires             & CST End\\
		Handlebar         & Steel aluminum look\\
		Stem              & Manually adjustable, Batavus Ergo Matic III\\
		Handles           & adjustable Batavus Ergo Grip II\\
		Saddle            & Selle Royal 5160\\
		Seatpost          & spring, Post Modern Compact\\
		Dynamo            & Shimano Hub Dynamo\\
		Headlight         & Trelock Hi-Power automatic halogen\\
		Rear              & Smart battery automatically\\
		Pedals            & aluminum antislip\\
		Final							& AXA Defender\\
		Chip theft        & Present\\
		Crank             & Aluminum\\
		Rack              & Aluminum\\
		Quick Binder      & Binder Quattro\\
		Mudguards         & Batavus hyalite II\\
		Bell              & Bibia Touring\\
		Balhoofdset       & VP semi-integrated, oversized\\
		Shifters          & Batavus (8-speed)\\
		Saddlebag         & Batavus saddlebag with plakset\\
	\end{tabular}
	\caption{}
	\label{crescendoSpecs}
\end{table}
\subsection{Batavus Stratos Deluxe}
\begin{table}
	\begin{tabular}{ll}
		Switching           & Spin System (shifter)\\
		Monochromatic frame & Black\\
		Chain guard         & Semi-open\\
		Mudguards           & Aluminum\\
		Rack                & Yes\\
		Extra drive         & Regardless Motori degree\\
		Type                & Men's\\
		Frame Material      & Aluminum\\
		Rim Material        & Aluminum\\
		Rear braking system & Roller brake\\
		Front braking system & Roller Brake\\
		Number of gears      & 7 Gears\\
		Gear System          & Shimano Nexus\\
		Wheel Size           & 28 inch\\
		Tirewidth            & 37 mm\\
		Price                & 749,00\\
	\end{tabular}
	\caption{}
	\label{stratosSpecs}
\end{table}
\bibliographystyle{plain}
\bibliography{bicycle}
\end{document}

### 2010-04-19 acquireData.m

@@ -1,0 +1,36 @@
+% File: acquireData.m
+% Date: May 27, 2009
+% Author: Jason Moore
+% Description: Takes data from a single rate sensor and saves it.
+clear all
+close all
+clc
+bicycle = input('What is the bicycle? ', 's');
+part = input('What part are you measuring? ', 's');
+pendulum = input('What pendulum are you using? ', 's');
+angle = input('Which angle are you measuring at? ', 's');
+trial = input('What is the trial number? ', 's');
+notes = input('Any additional info? ','s');
+disp('Press any key to start recording')
+pause
+%s=daqhwinfo % find out what is plugged in
+%s.InstalledAdaptors % show the adaptors
+%out=daqhwinfo('nidaq')
+%out.ObjectConstructorName(:)
+ai=analoginput('nidaq','Dev1'); % set the analog input
+%daqhwinfo(ai)
+duration=30; % the sample time in seconds
+set(ai,'SampleRate',1000) % set the sample rate
+ActualRate = get(ai,'SampleRate');
+set(ai,'SamplesPerTrigger',duration*ActualRate) % 
+set(ai,'TriggerType','Manual')
+chan = addchannel(ai,0);
+start(ai)
+trigger(ai)
+wait(ai,duration + 1)
+data = getdata(ai);
+plot(data,'.-')
+delete(ai)
+clear ai chan
+filename = [bicycle part pendulum angle trial];
+save(filename) 

### 2010-04-19 arm.m

@@ -1,0 +1,2 @@
+function zero = arm(d,x19,x21,z19,z21,lla,lua)
+zero=(z21-z19)^2+(x19-x21)^2+lla^2-2*(z21-z19)*(lla^2-d.^2).^0.5-2*(x19-x21)*d-lua^2;

### 2010-04-19 batavus_browser_inertia.m

@@ -1,0 +1,226 @@
+% File: batavus_browser_inertia.m
+% Date: March 22, 2010
+% Author: Jason Moore
+% Description: Calculates the benchmark inertia matrices of the Batavus
+% Browser frame and fork/handlebar from the experiments.
+clear all
+close all
+clc
+% set the y component equal to zero
+IBFyy  = 0;
+IHyy   = 0;
+IBRAyy = 0;
+IBRJyy = 0;
+% Inertia tensors are defined as follows:
+% I = [ Ixx -Ixy -Ixz
+%      -Ixy  Iyy -Iyz
+%      -Ixz -Iyz  Izz]
+%------------Batavus Browser rear frame ----------------------------------%
+mBF = 9.86;  % [kg] mass of the rear frame
+xBF = 0.2799135; % [m] cg location in benchmark coordinate system
+zBF = -0.53478492; % [m] cg location in the benchmark coordinate system
+% a period of ten oscillations was measured with a pocket watch through 3
+% different planar axes each 3 different times and averaged
+TBF1 = 2.094431797861789679;   % [s] period of oscillation when hung from the top tube
+TBF2 = 2.376844356021118987;   % [s] period of oscillation when hung from the down tube
+TBF3 = 1.839869878050422791;   % [s] period of oscillation when hung from the fender
+k   =  5.6248209199363393;  % [Nm/rad] stiffness of the torsion rod, from page 43 of
+% moments of inertia [Nm*s^2] about the 1st, 2nd, and 3rd test axes
+IBFx1 = k*TBF1^2/4/pi^2;
+IBFx2 = k*TBF2^2/4/pi^2;
+IBFx3 = k*TBF3^2/4/pi^2;
+% angle [rad] between the test axes (x1,x2,x3) and the global benchmark
+% x-axis (positive rotation about the y-axis)
+alpha1 =  199.5/180*pi;
+alpha2 =  64.2/180*pi;
[…]

### 2010-04-19 batavus_stratos_inertia.m

@@ -1,0 +1,226 @@
+% File: batavus_stratos_inertia.m
+% Date: March 24, 2010
+% Author: Jason Moore
+% Description: Calculates the benchmark inertia matrices of the Batavus
+% Stratos frame and fork/handlebar from the experiments.
+clear all
+close all
+clc
+% set the y component equal to zero
+IBFyy  = 0;
+IHyy   = 0;
+IBRAyy = 0;
+IBRJyy = 0;
+% Inertia tensors are defined as follows:
+% I = [ Ixx -Ixy -Ixz
+%      -Ixy  Iyy -Iyz
+%      -Ixz -Iyz  Izz]
+%------------Batavus Stratos rear frame ----------------------------------%
+mBF = 7.2199999999999998;  % [kg] mass of the rear frame
+xBF = 0.32669395; % [m] cg location in benchmark coordinate system
+zBF = -0.48249332; % [m] cg location in the benchmark coordinate system
+% a period of ten oscillations was measured with a pocket watch through 3
+% different planar axes each 3 different times and averaged
+TBF1 = 1.81602013;   % [s] period of oscillation when hung from the top tube
+TBF2 = 1.58525347;   % [s] period of oscillation when hung from the down tube
+TBF3 = 1.66025349;   % [s] period of oscillation when hung from the fender
+k   =  5.6248209199363393;  % [Nm/rad] stiffness of the torsion rod, from page 43 of
+% moments of inertia [Nm*s^2] about the 1st, 2nd, and 3rd test axes
+IBFx1 = k*TBF1^2/4/pi^2;
+IBFx2 = k*TBF2^2/4/pi^2;
+IBFx3 = k*TBF3^2/4/pi^2;
+% angle [rad] between the test axes (x1,x2,x3) and the global benchmark
+% x-axis (positive rotation about the y-axis)
+alpha1 =  68/180*pi;
+alpha2 =  -24.7/180*pi;
+alpha3 =  12.4/180*pi;
[…]

### 2010-04-19 bike_inertia_jason_stratos.m

@@ -1,0 +1,661 @@
+% File: bike_inertia_jason_stratos.m
+% Date: March 24, 2010
+% Author: Jason Moore
+% Description: Determines the critical speeds for the weave and capsize
+% eigenmodes of a simplified bicycle model used in control experiments at 
+% TU Delft. In this case the Batavus Stratos Deluxe and Jason as the rider. 
+
+clear
+close all
+clc
+warning off MATLAB:fzero:UndeterminedSyntax
+%-------------------Frame Parameters (Stratos Deluxe)
+df      = 0.334*2;                         % front wheel diameter [m]
+dr      = 0.338*2;                         % rear wheel diameter [m]
+rf      = df/2;                          % front wheel radius [m]
+rr      = dr/2;                          % rear wheel radius [m]
+wb      = 1.037;                          % wheel base [m]
+hb      = 0.290;                         % bottom bracket height [m]
+lcs     = 0.445;                          % chain stay length [m]
+alphast = 75*pi/180;                   % seat tube angle [rad]
+lst     = 0.480;                          % seat tube length [m]
+tr      = 0.056;                         % trail [m]
+lambda  = 73.1*pi/180;                   % head tube angle [rad]
+fo      = rf*cos(lambda)-tr*sin(lambda); % fork offset [m]
+lf      = 0.455;                         % fork length [m]
+lsp     = 0.195;                          % seat post length [m]
+wr      = 0.13;                          % rear hub width [m]
+wf      = 0.10;                          % front hub width [m]
+whb     = 0.58;                          % handle bar width [m] (this isn't the actual value)
+lhb     = -0.090;                          % handle bar length [m]
+hs      = 0.190;                          % stem height [m]
+rho     = 7850;                          % frame density [kg/m^3]
+%-------------------Human Parameters (Jason Moore)
[…]

### 2010-04-19 fitData.py

@@ -1,0 +1,57 @@
+import scipy.io.matlab.mio as mio
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.optimize as op
+import os
+import pickle as p
+
+dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
+file = open('period.txt', 'w')
+filenames.sort()
+period = {}
+#for name in ['StratosFrameCompoundFirst2.mat']:
+for name in filenames:
+    pendDat = {}
+    mio.loadmat('data/pendDat/' + name, mdict=pendDat)
+    y = pendDat['data'].ravel()
+    x = np.linspace(0, 30, num=len(y))
+    # plot the original data
+    plt.plot(x, y, '.')
+    # decaying oscillating exponential function
+    fitfunc = lambda p, t: p[0] + np.exp(-p[3]*p[4]*t)*(p[1]*np.cos(p[4]*np.sqrt(1-p[3]**2)*t) + p[2]*np.sin(p[4]*np.sqrt(1-p[3]**2)*t))
+    # initial guesses
+    p0 = np.array([1.35, -.75, -.5, 0.01, 3.93])
+    # create the error function
+    errfunc = lambda p, t, y: fitfunc(p, t) - y
+    # minimize the error function
+    p1, success = op.leastsq(errfunc, p0[:], args=(x, y))
+    # plot the fitted curve
+    lscurve = fitfunc(p1, x)
+    plt.plot(x, lscurve, 'k')
+    plt.xlabel('Time [s]')
+    plt.ylabel('Amplitude')
+    #plt.show()
+    plt.savefig('data/pendDat/graphs/' + name[:-4] + '.png')
+    plt.close()
+    SSR = np.sum((lscurve - np.mean(y))**2)
+    SST = np.sum((y - np.mean(y))**2)
+    rsq = SSR/SST
+    if rsq <= 0.99:
+        rsq = str(rsq) + '*'
+    else:
+        pass
+    wo = p1[4]
+    f = wo/2./np.pi
+    T = 1./f
+    note = pendDat['notes']
+    line = name + ',' + str(T) + ',' + str(rsq) + ',' + str(note) + '\n'
+    file.write(line)
+    print line
+    if name[:-5] in period.keys():
+        period[name[:-5]].append(T)
+    else:
+        period[name[:-5]] = [T]
+file.close()
+file = open('period.p', 'w')
+p.dump(period, file)
+file.close()

### 2010-04-19 hunchAngle.py

@@ -1,0 +1,30 @@
+from numpy import *
+
+def uvec(x):
+    xhat = x/linalg.norm(x)
+    return xhat
+# Victor on the Stratos
+# name the four points from the inkscape superimposed lines
+a = array([190.716, 1052.483])
+b = array([457.633, 533.479])
+c = array([-312.107, -371.112])
+d = array([645.204, -352.409])
+# calculate the two vectors
+v1 = a - b
+v2 = c -d
+# calculate the angle between the two vectors
+theta = arccos(dot(uvec(v1), uvec(v2)))
+print "Victor on the Stratos =", rad2deg(theta)
+# Victor on the Browser
+# name the four points from the inkscape superimposed lines
+a = array([406.081, 989.949])
+b = array([512.147, 453.558])
+c = array([-334.360, -175.767])
+d = array([705.086, -125.259])
+# calculate the two vectors
+v1 = a - b
+v2 = c -d
+# calculate the angle between the two vectors
+theta = arccos(dot(uvec(v1), uvec(v2)))
+print "Victor on the Browser =", rad2deg(theta)
+

### 2010-04-19 leg.m

@@ -1,0 +1,2 @@
+function zero = leg(d,x5,x13,z5,z13,lth,lc)
+zero=lth^2-lc^2-(z13-z5)^2-(x5-x13)^2+2*(z13-z5)*(lc^2-d.^2).^0.5-2*(x5-x13).*d;

### 2010-04-19 plotData.m

@@ -1,0 +1,9 @@
+plot(t,data,'*')
+hold on
+[sse, FittedCurve] = model(estimates);
+plot(t, FittedCurve, 'r')
+xlabel('xdata')
+ylabel('f(estimates,xdata)')
+title(['Fitting to function ', func2str(model)]);
+legend('data', ['fit using ', func2str(model)])
+hold off

### 2010-04-19 plotRawData.py

@@ -1,0 +1,18 @@
+import os
+import scipy.io.matlab.mio as mio
+import matplotlib.pyplot as plt
+
+# make a list of the folder contents
+dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
+print filenames
+i = 0
+while i < len(filenames):
+    d = {}
+    mio.loadmat('data/pendDat/' + filenames[i], mdict=d)
+    print 'Plotting', filenames[i]
+    plt.plot(d['data'])
+    plt.title(filenames[i])
+    plt.show()
+    raw_input()
+    plt.close()
+    i += 1

### 2010-04-19 processData.m

@@ -1,0 +1,65 @@
+clear all
+close all
+clc
+load('data/data')
+% calculate the wheel radii
+rearWheelRadius = rearWheelDist./2./pi./rearWheelRot;
+frontWheelRadius = frontWheelDist./2./pi./frontWheelRot;
+% steer axis tilt in radians
+steerAxisTilt = pi./180.*(90-headTubeAngle);
+% calculate the front wheel trail
+trail = (frontWheelRadius.*sin(steerAxisTilt)-forkOffset)./cos(steerAxisTilt);
+% calculate the frame rotation angle
+beta = frameAngle.*pi./180-pi/2*ones(size(frameAngle))-[steerAxisTilt;steerAxisTilt;steerAxisTilt];
+% calculate the slope of the CoM line
+frameM = tan(beta-pi/2);
+% calculate the z-intercept of the CoM line
+frameB = -frameMassDist./sin(beta)-[rearWheelRadius;rearWheelRadius;rearWheelRadius];
+% calculate the center of mass position
+for i = 1:length(frameM)
+    CoM(1:2,i) = [-frameM(:,i) ones(3,1)]\frameB(:,i);
+end
+% calculate the fork rotation angle
+betaFork = forkAngle.*pi./180-pi/2*ones(size(forkAngle))-[steerAxisTilt;steerAxisTilt;steerAxisTilt];
+% calculate the slope of the fork CoM line
+forkM = tan(betaFork-pi/2);
+% calculate the z-intercept of the CoM line
+forkB = [wheelbase;wheelbase;wheelbase]./tan(betaFork)-forkMassDist./sin(betaFork)-[frontWheelRadius;frontWheelRadius;frontWheelRadius];
+% calculate the center of mass position
+for i = 1:length(forkM)
+    CoMFork(1:2,i) = [-forkM(:,i) ones(3,1)]\forkB(:,i);
+end
+% plot the CoM lines
+for i = 1:length(frameM)
+    figure(i)
+    hold on
+    axis([-1 2 -1 1])
+    rectangle('position',[-rearWheelRadius(i),-2*rearWheelRadius(i),2*rearWheelRadius(i),2*rearWheelRadius(i)],'curvature',[1 1])
+    first  = refline(frameM(1,i),frameB(1,i));
+    second = refline(frameM(2,i),frameB(2,i));
+    third  = refline(frameM(3,i),frameB(3,i));
+    xAxis = refline(0,0);
+    set(xAxis,'color','k')
[…]

### 2010-04-19 processData.py

@@ -1,0 +1,136 @@
+import scipy.io.matlab.mio as mio
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.optimize as op
+import os
+import re
+import pickle as p
+
+# load the main data file into a dictionary
+d = {}
+mio.loadmat('data/data.mat', mdict=d)
+nBk = len(d['bikes'])
+bikeNames = []
+for bike in d['bikes']:
+    bikeNames.append(bike[0][0].encode('ascii'))
+print "Number of bikes =", nBk
+# calculate the wheel radii
+rearWheelRadius = d['rearWheelDist'][0]/2/np.pi/d['rearWheelRot'][0]
+print "Rear wheel radii =\n", rearWheelRadius
+frontWheelRadius = d['frontWheelDist'][0]/2/np.pi/d['frontWheelRot'][0]
+print "Front wheel radii =\n", frontWheelRadius
+# steer axis tilt in radians
+steerAxisTilt = np.pi/180*(90-d['headTubeAngle'][0])
+print "Steer axis tilt [deg] =\n", steerAxisTilt*180./np.pi
+# calculate the front wheel trail
+trail = (frontWheelRadius*np.sin(steerAxisTilt) - d['forkOffset'][0])/np.cos(steerAxisTilt)
+print "Trail =\n", trail
+# calculate the frame rotation angle
+satMat = np.array([steerAxisTilt, steerAxisTilt, steerAxisTilt])
+betaFrame = d['frameAngle']*np.pi/180 - np.pi/2.*np.ones_like(d['frameAngle']) - satMat
+#betaFrame = np.pi + steerAxisTilt - d['frameAngle']*np.pi/180
+#for i, row in enumerate(betaFrame):
+#    for j, v in enumerate(row):
+#        if v < 0:
+#            betaFrame[i, j] = v + np.pi
+#        else:
+#            pass
+#for i, row in enumerate(betaFrame):
+#    for j, v in enumerate(row):
+#        if v > np.pi:
+#            betaFrame[i, j] = v - np.pi
+#        else:
+#            pass
+print "Frame rotation angle [deg] =\n", betaFrame/np.pi*180.
[…]

### 2010-04-19 torcom.py

@@ -1,0 +1,47 @@
+import pickle as p
+import numpy as np
+import re
+
+f = open('period.p', 'r')
+period = p.load(f)
+f.close()
+def space_out_camel_case(s):
+        """Adds spaces to a camel case string.  Failure to space out string
+        returns the original string.
+        >>> space_out_camel_case('DMLSServicesOtherBSTextLLC')
+        'DMLS Services Other BS Text LLC'
+        """
+
+        return re.sub('((?=[A-Z][a-z])|(?<=[a-z])(?=[A-Z]))', ' ', s)
+tor = np.zeros((12, 8))
+com = np.zeros((4, 8))
+bN = ['Browser', 'Browserins', 'Crescendo', 'Fisher', 'Pista', 'Stratos',
+        'Yellow', 'Yellowrev']
+fst = ['First', 'Second', 'Third']
+fffr = ['Frame', 'Fork', 'Fwheel', 'Rwheel']
+tc = ['Torsional', 'Compound']
+# average the periods
+for k, v in period.items():
+    km = re.sub('BrowserIns', 'Browserins', k)
+    km = re.sub('YellowRev', 'Yellowrev', km)
+    desc = space_out_camel_case(km).split()
+    print desc
+    if desc[0] == 'Rod':
+        rodPeriod = np.mean(v)
+    else:
+        # if torsional, put it in the torsional matrix
+        if desc[2] == tc[0]:
+            r = fffr.index(desc[1])*3 + fst.index(desc[3])
+            c = bN.index(desc[0])
+            tor[r, c] = np.mean(v)
+            print 'Added torsional to [', r, ',', c, ']'
+        # if compound, put in in the compound matrix
+        elif desc[2] == tc[1]:
+            com[fffr.index(desc[1]), bN.index(desc[0])] = np.mean(v)
+avgPer = {}
+avgPer['tor'] = tor
+avgPer['com'] = com
+avgPer['rodPer'] = rodPeriod
+f = open('avgPer.p', 'w')
+p.dump(avgPer, f)
+f.close()

### 2010-04-27 PhysicalParameters/abstract.tex
initial commit of the physical parameter abstract

\documentclass{bmd2010a}

\begin{document}

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  An Accurate Method of Measuring a Bicycle's Physical Parameters\\}
{\fontsize{16pt}{20pt}\selectfont%
  Motorcycle Dynamics Symposium\\}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{Jason K. Moore}\\}
  \textit{Mechanical and Aerospace Engineering\\
          University of California, Davis\\
          One Shields Avenue, Davis, CA 95616, USA\\
          e-mail: jkmoor@ucdavis.edu
}\vspace{10pt}\\
  {\fontsize{12}{14}{Mont Hubbard}\\}
  \textit{Affiliation}\vspace{10pt}\\
  {\fontsize{12}{14}{D. E. Author}\\}
  \textit{Affiliation}\vspace{10pt}\\
\end{flushleft}

\section*{Abstract}
Accurate measurments of a bicycle's physical parameters are required for
realistic dynamic simulations. For the most basic models the geometry, mass,
mass location and mass distributions must be measured. More complex models
require measurements of tire characteristics, stiffness, damping, etc. This
paper concerns itself with the measurement of bicycle parameters required for
simulations of the benchmark bicycle presented in \cite{Meijaard2007}. This
bicycle model is made of four rigid bodies, has ideal rolling, frictionless joints,
and is assumed to be symmetric about the frame center plane. A set of 25
physical parameters are used to describe the geometry, mass, mass location and
mass distribution of each of the rigid bodies. These parameters are presented
relative to a reference frame when the bicycle is in the nominal upright
configuration. The methods described herein are based primarily off of the work
done in \cite{Kooijman2006} but have been refined for improved accuracy and
methodology. \cite{Roland1971} measured an experimental bicycle frame in a
similar fashion and both \cite{Dohring1953} and \cite{Singh1971} have used
similar techniques with scooters. We measure the characteristics of six
different bicycles, two of which are setup in two different configurations, for
a total of eight different parameter sets that can be used with, but not
limited, to the benchmark bicycle model. The parameters set is also complete
for non-linear benchmark bicycle model.

The bicycles were chosen for both variety and convience. The six bicycles are as follows:
\begin{description}
    \item[Batavus Browser] A Dutch style city bicycle. The bicycle was measured
        in two configurations. One with the instrumentation described in
        \cite{Kooijman2009a} and one as from the manufacturer.
    \item[Batavus Stratos Deluxe] A Dutch style city bicycle
    \item[Batavus Crescendo Deluxe] A Dutch style city bicycle
    \item[Gary Fisher Mountain Bike] A hardtail mountain bicycle
    \item[Bianchi Pista] A modern steel frame track racing bicycle
    \item[Yellow Bicycle] A stripped down aluminum frame road bicycle. This
        bike was measured in two configurations, the second being the fork was
        rotated in the headtube 180 degrees for larger trail.
\end{description}
\subsection{Geometry}
The benchmark bicycle needs five basic geometric measurements: the wheelbase
$w$, steer axis tilt $\lambda$, trail $c$, and the wheel radii $r_\mathrm{F}$
and $r_\mathrm{R}$. The wheelbase and steer axis tilt were measured directly
with a scale and a digital level, respectively. The trail and wheel radii were
measured indirectly by fork offset and loaded wheel circumference.
\subsection{Mass}
The mass of each of the rigid bodies were measured with a precision scale.
\subsection{Mass location}
The center's of mass of each of the rigid bodies were measured by hanging the
rigid bodies through their centers of mass and assuming symmetry. The
orientation angle of the headtube angle and the point of suspension were
measured and the center's of mass of the frame and fork were calculated from
these values. The wheel centers of mass were assumed to be in their ideal
geometric center.
\subsection{Mass Distribution}
The moment's of inertia of the rigid bodies were measured with both a torsional
pendulum and a compound pendulum.
\bibliographystyle{acm}
\bibliography{bicycle}
\end{document}

### 2010-04-27 PhysicalParameters/abstract.tex
added two abstracts

@@
 \end{description}
-\subsection{Geometry}
+\subsection*{Geometry}
 The benchmark bicycle needs five basic geometric measurements: the wheelbase
@@
 measured indirectly by fork offset and loaded wheel circumference.
-\subsection{Mass}
+\subsection*{Mass}
 The mass of each of the rigid bodies were measured with a precision scale.
-\subsection{Mass location}
+\subsection*{Mass location}
 The center's of mass of each of the rigid bodies were measured by hanging the
@@
 geometric center.
-\subsection{Mass Distribution}
+\subsection*{Mass Distribution}
 The moment's of inertia of the rigid bodies were measured with both a torsional
-pendulum and a compound pendulum.
+pendulum and a compound pendulum. Once again the assumed symmetry of the
+bicycle was utlized to reduce the amount of measurements needed. The
+in-symmetric-plane moments of inertia were calculated from measurements of the
+rigid bodies period of oscillation when hung from a torsional pendulum at
+different orienatation angles. The out of plane moments of inertia were
+measured by swinging the rigid bodies as a compound pendulum.
+\subsection*{Results}
+Parameter sets for all six bicycles (and all eight
+configurations)~\ref{tab:par}  will be
+presented, along with basic comparisons of the open loop dynamics of the
+different types of bicycles. By examining the mass, damping and stiffness
+matrices of the system and the eigenvalues at different speeds anedoctal
+descriptions of handling are compared to the dynamics.
+\begin{table*}[tb]
+\centering
+\caption{COMBINED BICYCLE AND RIDER PARAMETER VALUES FOR THE BATAVUS BROWSER.}
+{\small
+\begin{tabular}{lcl}
+&&\\
+\hline
+Parameter & Symbol & Value \\
+\hline
+wheel base & $w$ & 1.121 m \\
+trail & $c$ & 0.069 m \\
+steer axis tilt & \emph{$\lambda$} & 0.400 rad \\
+gravity & \emph{g} & 9.81 N kg$^{-1}$ \\
+forward speed & \emph{v} & various m s$^{-1}$ \\
+
+\multicolumn{3}{l}{\textbf{R}\emph{ear wheel R}}\\
+radius & \emph{$r_\mathrm{R}$} & 0.341 m\\
+mass & \emph{$m_\mathrm{R}$} &  3.11 kg \\
+mass moments of inertia & \emph{$(I_{\mathrm{R}xx},\
+I_{\mathrm{R}yy})$} &
+(0.088,\ 0.152) $\mathrm{kg\ m}^2$ \\
+
+\multicolumn{3}{l}{\emph{rear} \textbf{B}\emph{ody and frame B}}\\
+position center of mass & \emph{$(x_\mathrm{B},\ z_\mathrm{B})$} &
+($0.276$,\ $-0.538$) m\\
+mass & \emph{$m_\mathrm{B}$} & 9.86 kg \\
+mass moments of inertia & $\left[ \begin{array}{ccc}
+I_{\mathrm{B}xx} &  0 & I_{\mathrm{B}xz}\\
+0 & I_{\mathrm{B}yy} & 0 \\
+I_{\mathrm{B}xz} & 0 & I_{\mathrm{B}zz}
+\end{array} \right] $ &
+$\left[ \begin{array}{ccc}
+0.527 &  0 & -0.114\\
+0 & 1.317 & 0 \\
+-0.114 & 0 & 0.759
+\end{array} \right] \ \mathrm{kg\ m}^{2}$\\
+
+\multicolumn{3}{l}{\emph{front} \textbf{H}\emph{andlebar and fork assembly H}}\\
+position center of mass & \emph{$(x_\mathrm{H},\ z_\mathrm{H})$} &
+($0.867$,\ $-0.748$) m\\
+mass & \emph{$m_\mathrm{H}$} & 3.22 kg \\
+mass moments of inertia & $\left[ \begin{array}{ccc}
+I_{\mathrm{H}xx} &  0 & I_{\mathrm{H}xz}\\
+0 & I_{\mathrm{H}yy} & 0 \\
+I_{\mathrm{H}xz} & 0 & I_{\mathrm{H}zz}
+\end{array} \right] $ &
+$\left[ \begin{array}{ccc}
+    0.253  &       0  & -0.072\\
+         0  & 0.246   &  0\\
+   -0.072  &       0  &  0.096
+\end{array} \right] \ \mathrm{kg\ m}^{2}$\\
+
+\multicolumn{3}{l}{\textbf{F}\emph{ront wheel F}}\\
+radius & \emph{$r_\mathrm{F}$} & 0.344 m\\
+mass & \emph{$m_\mathrm{F}$} & 2.02 kg \\
+mass moments of inertia & \emph{$(I_{\mathrm{F}xx},\
+I_{\mathrm{F}yy})$} &
+(0.090,\ 0.149) $\mathrm{kg\ m}^2$ \\
+\hline
+\end{tabular}
+}
+\label{tab:par}
+\end{table*}
+
 \bibliographystyle{acm}

### 2010-04-27 latex-abstracttemplate/abstract.tex
added two abstracts

\documentclass{bmd2010a}

\begin{document}

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Instructions for Preparing an Abstract for the Bicycle and\\}
{\fontsize{16pt}{20pt}\selectfont%
  Motorcycle Dynamics Symposium\\}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{A. B. Author}\\}
  \textit{Faculty of Mechanical Engineering\\
          University of Technology\\
          Address, Postcode City, Country\\
          e-mail: email1@address
}\vspace{10pt}\\
  {\fontsize{12}{14}{C. Author}\\}
  \textit{Affiliation}\vspace{10pt}\\
  {\fontsize{12}{14}{D. E. Author}\\}
  \textit{Affiliation}\vspace{10pt}\\
\end{flushleft}

\section*{Abstract}
Authors are requested to upload an abstract of two pages (including references
and figures) by the stated date in the call for papers to the symposium web
site http://bicycle.tudelft.nl/bmd2010. The uploaded file must use the
Portable Document Format (PDF) and the file should have a sensible name like: Surname1Surname2Surname3\_abstract.pdf.

The abstract must be written in English. It must contain the name, affiliation
and e-mail address of the authors. The header should be as in
this example. The text area is 24 by 15~cm, with left and right margins of
3~cm and a top margin (not including the header) of 3~cm. The paper used should have the size 210 x 297 mm which is the European A4 size. The title should be
in Times New Roman 16pt, and may extend over more than one line. The author's
name or authors' names should be in Times New Roman 12 pt, and the affiliation
should be in 11pt italics.

The abstract itself has the section heading \textbf{Abstract}, bold 11~pt
Times New Roman, and the text should be written in Times New Roman, 11~pt.

Equations should be numbered continuously according to the format
shown in Equation~(\ref{eq:equ1}):
\begin{equation} \label{eq:equ1}
e^{i\pi} + 1 = 0,
\end{equation}
where the unknown symbols are explained after the equation.

\begin{table}[h!]
\begin{center}
\caption{Example of a table with a short caption.} \label{tab:tab1}
\begin{tabular}{|c|ccc|}
\hline
    &  $x$  &  $y$  &  $z$ \\
\hline
$x'$  &  $\alpha_1$ & $\beta_1$ & $\gamma_1$ \\
$y'$  &  $\alpha_2$ & $\beta_2$ & $\gamma_2$ \\
$z'$  &  $\alpha_3$ & $\beta_3$ & $\gamma_3$ \\
\hline
\end{tabular}
\end{center}
\end{table}

Figures, graphs
and tables must be included using the same style as in Figure~\ref{fig:fig1}
and Table~\ref{tab:tab1}. Both figures and tables should be centred on the page. Use of at least 300 dpi resolution for pictures and 600 dpi for line art is required, 1 px wide lines in figures should be avoided as they may become invisible in print.

Bibliographical citations should be written in the
order in which they are cited, see the References section below, where
Reference~\cite{Pac02} exemplifies the case of a textbook, while
Reference~\cite{Ber07} is an article in conference proceedings and
Reference~\cite{Sha71} is an article in a journal. Use can be made of a
pre-existing bibtex style which uses a similar style.

The full paper will have a similar style, but with a different style file. Templates can be found at the symposium website in due time.

\begin{figure}[h!]
\begin{center}
  \includegraphics[width=55mm]{figure1}
  \caption{An example of a figure caption. Use 10~pt Times New Roman.
           For long captions, use a text width of 13~cm.
           Use the same style for the tables.} \label{fig:fig1}
\end{center}
\end{figure}

We very much look forward to welcoming you in Delft! Best wishes and the warmest regards from the Organizing Committee of Bicycle and Motorcycle Dynamics 2010.

% \bibliographystyle{acm}
% \bibliography{biblio}
\begin{thebibliography}{6}
% \begin{thebibliography}{66} % use this for more than 9 references
\bibitem{Pac02} H.~B.~Pacejka,
\textit{Tyre and Vehicle Dynamics},
Butterworth and Heinemann, Oxford, 2002.

\bibitem{Ber07} E.~Bertolazzi, F.~Biral, M.~Da~Lio and V.~Cossalter,
``The influence of rider's upper body motions on motorcycle minimum time
  maneuvering'',
in C.~L.~Bottasso, P.~Masarati and L.~Trainelli (eds),
\textit{Proceedings, Multibody Dynamics 2007, ECCOMAS Thematic Conference},
Milano, Italy, 25--28 June 2007,
Politecnico di Milano, Milano, 2007, 15~pp.

\bibitem{Sha71} R.~S.~Sharp,
``The stability and control of motorcycles'',
\textit{Proceedings of the IMechE, Part C, Journal of Mechanical Engineering
  Science} \textbf{13} (1971), pp.~316--329.

\end{thebibliography}

\end{document}

### 2010-04-30 PhysicalParameters/abstract.tex
initial commit of the blender animation files

@@
 {\fontsize{16pt}{20pt}\selectfont%
-  An Accurate Method of Measuring a Bicycle's Physical Parameters\\}
-{\fontsize{16pt}{20pt}\selectfont%
-  Motorcycle Dynamics Symposium\\}
+  An Accurate Method of Measuring and Comparing a Bicycle's Physical Parameters\\}
 \end{flushleft}
@@
 \begin{flushleft}
-  {\fontsize{12}{14}{Jason K. Moore}\\}
+  {\fontsize{12}{14}{Jason K. Moore and Mont Hubbard}\\}
   \textit{Mechanical and Aerospace Engineering\\
@@
           One Shields Avenue, Davis, CA 95616, USA\\
-          e-mail: jkmoor@ucdavis.edu
+          e-mail: jkmoor@ucdavis.edu and mhubbard@ucdavis.edu
   }\vspace{10pt}\\
-  {\fontsize{12}{14}{Mont Hubbard}\\}
-  \textit{Affiliation}\vspace{10pt}\\
-  {\fontsize{12}{14}{D. E. Author}\\}
-  \textit{Affiliation}\vspace{10pt}\\
+  {\fontsize{12}{14}{A. L. Schwab and J. D. G. Kooijman}\\}
+  \textit{Laboratory for Engineering Mechanics\\
+          Delft University of Technology, Mekelweg 2, 2628CD Delft, The
+          Netherlands\\
+          e-mail: a.l.schwab@tudelft.nl and jodikooijman@gmail.com
+  }\vspace{10pt}\\
 \end{flushleft}
@@
 \section*{Abstract}
-Accurate measurments of a bicycle's physical parameters are required for
+Accurate measurements of a bicycle's physical parameters are required for
 realistic dynamic simulations. For the most basic models the geometry, mass,
-mass location and mass distributions must be measured. More complex models
+mass center location and mass distributions must be measured. More complex models
 require measurements of tire characteristics, stiffness, damping, etc. This
-paper concerns itself with the measurement of bicycle parameters required for
+paper concerns the measurement of bicycle parameters required for
 simulations of the benchmark bicycle presented in \cite{Meijaard2007}. This
-bicycle model is made of four rigid bodies, has ideal rolling, frictionless joints,
-and is assumed to be symmetric about the frame center plane. A set of 25
+model is composed of four rigid bodies, has ideal rolling and frictionless joints,
+and is assumed to be laterally symmetric. A set of 25
 physical parameters are used to describe the geometry, mass, mass location and

### 2010-04-30 BenchmarkBikeTools.py
changed the code to print text files readable in windows

@@ -4,7 +4,7 @@ def bmp2cm(filename):
     f = open(filename, 'r')
     p = {}
     for i, line in enumerate(f):
-        list = line[:-1].split(',')
+        list = line[:-3].split(',')
         pi = np.pi
         p[list[0]] = eval(list[1])
     mT = p['mR'] + p['mB'] + p['mH'] + p['mF']
@@ -50,9 +50,9 @@ def bmp2cm(filename):
     C1dp = -(mu*ST + SF*np.cos(p['lambda']))
     C1dd = IAlz/p['w']*np.cos(p['lambda']) + mu*(SA + ITzz/p['w']*np.cos(p['lambda']))
     C1 = np.array([[C1pp, C1pd], [C1dp, C1dd]])
-    return M, K0, K2, C1, p
+    return M, C1, K0, K2, p
 
-def aMatrix(M, K0, K2, C1, p):
+def aMatrix(M, C1, K0, K2, p):
     '''Calculates the A matrix from the canonical matrices for the benchmark
     bicycle'''
     from numpy import eye, zeros, vstack, hstack, dot

### 2010-04-30 processData.py

@@ -189,17 +189,20 @@ par['IBzz'] = np.array(par['IBzz'])
 par['v'] = np.ones_like(par['rR'])
 # write the parameter files
 for i, name in enumerate(bikeNames):
+    dir = 'bikeParameters/'
     fname = ''.join(name.split()) + 'Par.txt'
-    file = open(fname, 'w')
+    file = open(dir + fname, 'w')
     for k, v in par.items():
-        line = k + ',' + str(v[i]) + '\n'
+        line = k + ',' + str(v[i]) + '\r\n'
         file.write(line)
     file.close()
-    M, K0, K2, C1, p = bmp2cm(fname)
-    file = open(fname[:-7] + 'Can.txt', 'w')
-    for mat in ['M', 'K0', 'K2', 'C1']:
-        file.write(mat + '\n')
-        file.write(str(eval(mat)) + '\n')
+    M, C1, K0, K2, p = bmp2cm(dir + fname)
+    A = aMatrix(M, C1, K0, K2, p)
+    dir = 'bikeCanonical/'
+    file = open(dir + fname[:-7] + 'Can.txt', 'w')
+    for mat in ['M','C1', 'K0', 'K2', 'A']:
+        file.write(mat + '\r\n')
+        file.write(str(eval(mat)) + '\r\n')
     file.close()
 # Jason's parameters (sitting on the browser)
 IBJ = np.array([[7.9985, 0 , -1.9272], [0, 8.0689, 0], [ -1.9272, 0, 2.3624]])
@@ -220,7 +223,6 @@ for i in range(nBk):
     IB[1] = np.array([0., par['IByy'][i], 0.])
     IB[2] = np.array([par['IBxz'][i], 0., par['IBzz'][i]])
     I = parallel_axis(IBJ, mBJ, dJ[:, i]) + parallel_axis(IB, par['mB'][i], dB[:, i])
-    print I
     par['IBxx'][i] = I[0, 0]
     par['IBxz'][i] = I[0, 2]
     par['IByy'][i] = I[1, 1]
@@ -230,15 +232,18 @@ for i in range(nBk):
     par['zB'][i] = zB[i]
 # write the parameter files
 for i, name in enumerate(bikeNames):
+    dir = 'bikeRiderParameters/'
     fname = ''.join(name.split()) + 'RiderPar.txt'
-    file = open(fname, 'w')
+    file = open(dir + fname, 'w')
     for k, v in par.items():
[…]

### 2010-05-04 PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
renamed the file

--- 2010-04-30 PhysicalParameters/abstract.tex
@@
           One Shields Avenue, Davis, CA 95616, USA\\
-          e-mail: jkmoor@ucdavis.edu and mhubbard@ucdavis.edu
+          e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu
   }\vspace{10pt}\\
@@
           Netherlands\\
-          e-mail: a.l.schwab@tudelft.nl and jodikooijman@gmail.com
+          e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com
   }\vspace{10pt}\\
@@
 and is assumed to be laterally symmetric. A set of 25
-physical parameters are used to describe the geometry, mass, mass location and
+physical parameters is used to describe the geometry, mass, mass location and
 mass distribution of each of the rigid bodies. These parameters are presented
 relative to a reference frame when the bicycle is in the nominal upright
-configuration. The methods described herein are based primarily off of the work
+configuration. The methods described herein are based primarily on the work
 done in \cite{Kooijman2006} but have been refined for improved accuracy and
-methodology. \cite{Roland1971} measured an experimental bicycle frame in a
+methodology. Previously, \cite{Roland1971} measured an experimental bicycle frame in a
 similar fashion and both \cite{Dohring1953} and \cite{Singh1971} have used
-similar techniques with scooters. We measure the characteristics of six
-different bicycles, two of which are setup in two different configurations, for
-a total of eight different parameter sets that can be used with, but not
-limited, to the benchmark bicycle model. The parameters set is also complete
+similar techniques with scooters. We measured the characteristics of six
+different bicycles, two of which were set up in two different configurations.
+This is a total of eight different parameter sets that can be used with, but not
+limited to, the benchmark bicycle model. The parameters set is also complete
 for non-linear benchmark bicycle model.
 
-The bicycles were chosen for both variety and convience. The six bicycles are as follows:
+The six bicycles, chosen for both variety and convenience, are as follows:
 \begin{description}
-    \item[Batavus Browser] A Dutch style city bicycle. The bicycle was measured
+    \item[Batavus Browser] The Dutch style city bicycle was measured
         in two configurations. One with the instrumentation described in
-        \cite{Kooijman2009a} and one as from the manufacturer.
-    \item[Batavus Stratos Deluxe] A Dutch style city bicycle
-    \item[Batavus Crescendo Deluxe] A Dutch style city bicycle
-    \item[Gary Fisher Mountain Bike] A hardtail mountain bicycle
-    \item[Bianchi Pista] A modern steel frame track racing bicycle
+        \cite{Kooijman2009a} and one as received from the manufacturer.
+    \item[Batavus Stratos Deluxe] A Dutch style sporty city bicycle.
+    \item[Batavus Crescendo Deluxe] A Dutch style city bicycle with a
+        suspension fork.
+    \item[Gary Fisher Mountain Bike] A hardtail mountain bicycle.
+    \item[Bianchi Pista] A modern steel frame track racing bicycle.
     \item[Yellow Bicycle] A stripped down aluminum frame road bicycle. This
-        bike was measured in two configurations, the second being the fork was
+        bike was measured in two configurations, the second with the fork
         rotated in the headtube 180 degrees for larger trail.
@@
 \subsection*{Geometry}
-The benchmark bicycle needs five basic geometric measurements: the wheelbase
-$w$, steer axis tilt $\lambda$, trail $c$, and the wheel radii $r_\mathrm{F}$
+The benchmark bicycle requires five basic geometric measurements: wheelbase
+$w$, steer axis tilt $\lambda$, trail $c$, and wheel radii $r_\mathrm{F}$
 and $r_\mathrm{R}$. The wheelbase and steer axis tilt were measured directly
-with a scale and a digital level, respectively. The trail and wheel radii were
-measured indirectly by fork offset and loaded wheel circumference.
-\subsection*{Mass}
+with a meter scale and a digital level, respectively. The trail and wheel radii were
+measured indirectly using fork offset and loaded wheel circumference.
+\subsection*{Mass, Mass Center Location and Mass Distribution}
 The mass of each of the rigid bodies were measured with a precision scale.
-\subsection*{Mass location}
-The center's of mass of each of the rigid bodies were measured by hanging the
-rigid bodies through their centers of mass and assuming symmetry. The
-orientation angle of the headtube angle and the point of suspension were
-measured and the center's of mass of the frame and fork were calculated from
-these values. The wheel centers of mass were assumed to be in their ideal
+The centers of mass of each of the rigid bodies were measured by hanging each
+rigid body at the mass center and assuming lateral symmetry. The mass centers
+were calculated from the orientation angle of the headtube and the location of
+the point of suspension relative to the wheel center.
+The wheel centers of mass were assumed to be in their ideal
 geometric center.
-\subsection*{Mass Distribution}
-The moment's of inertia of the rigid bodies were measured with both a torsional
-pendulum and a compound pendulum. Once again the assumed symmetry of the
-bicycle was utlized to reduce the amount of measurements needed. The
+
+The moments of inertia of the rigid bodies were measured in both torsional
+pendulum and compound pendulum configurations. Again the assumed symmetry of the
+bicycle was utilized to reduce the number of measurements needed. The
 in-symmetric-plane moments of inertia were calculated from measurements of the
-rigid bodies period of oscillation when hung from a torsional pendulum at
-different orienatation angles. The out of plane moments of inertia were
+rigid bodies' periods of oscillation when hung as a torsional pendulum at
+different orienatation angles. The out-of-plane moments of inertia were
 measured by swinging the rigid bodies as a compound pendulum.
 \subsection*{Results}
-Parameter sets for all six bicycles (and all eight
-configurations)~\ref{tab:par}  will be
-presented, along with basic comparisons of the open loop dynamics of the
-different types of bicycles. By examining the mass, damping and stiffness
-matrices of the system and the eigenvalues at different speeds anedoctal
-descriptions of handling are compared to the dynamics.
+Parameter sets for all six bicycles and all eight
+configurations, example in Table~\ref{tab:par},  will be
+presented, together with basic comparisons of the open loop dynamics of the
+different types of bicycles. The physical parameters, the mass, damping and stiffness
+matrices of the system, the eigenvalues and open loop transfer functions at
+different speeds will be compared with typical anedoctal
+descriptions of handling.
 \begin{table*}[tb]
 \centering
-\caption{COMBINED BICYCLE AND RIDER PARAMETER VALUES FOR THE BATAVUS BROWSER.}
+\caption{PARAMETER VALUES FOR THE BATAVUS BROWSER.}
 {\small

### 2010-05-04 BikeTaskComparison/MooreHessHubbard.tex
initial commit of the abstract

\documentclass{bmd2010a}

\begin{document}

\begin{flushleft}
{\fontsize{16pt}{20pt}\selectfont%
  Manually Controlled Lane Change Manuever with 8 Bicycles\\}
\end{flushleft}

%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
\begin{flushleft}
  {\fontsize{12}{14}{Jason K. Moore, Ronald Hess, Mont Hubbard and Dale L.
  Peterson}\\}
  \textit{Mechanical and Aerospace Engineering\\
          University of California, Davis\\
          One Shields Avenue, Davis, CA 95616, USA\\
          e-mail: jkmoor@ucdavis.edu, rahess@ucdavis.edu, mhubbard@ucdavis.edu,
          dlpeterson@ucdavis.edu
  }\vspace{10pt}\\
\end{flushleft}

\section*{Abstract}

One of the most sought after answers to questions in bicycle dynamics is ``How
will the bicycle handle differently if its phyiscal characteristics are
changed?'' or ``Why do different bicycles handle differently''. The answer to
these questions are not trivial. Many factors play into the picture such as,
but not limited to, the bicycle's dynamics, the rider's dynamics, the rider's
control system, the manuever, the performance criteria and the rider's
subjective opinion of the handling. Of these only the bicycle's dynamics and
the rider's control system are somewhat understoof. Bicycle models have been
shown to predict basic motion through some experientation and much research has
been done in manual control for aircraft pilots.

This work follows in the foot
steps of \cite{Weir1972} and \cite{Lunteren1970} by using the well defined
cross-over model to predict the control action of a rigid person with only the
steering action as an input. We use the physical parameters measured from eight
different bicycles and one rider along with a multi-loop manual control model to simulate a
lane change manuever at three speeds: one below, in and above the linear stable
speed range of the bicycle. Two types of approaches are used. The first is to
tune the control gains for one of the bicycles for good performance. The same
controller is then used for the same manuever for each of the other
bicycle/rider combinations. The overshoot, (other things) are compared and the
bicycles are ranked according to how good or bad the perform under these
criteria. Secondly, a separate controller is tuned for each bicycle/rider for
good performance. We use the same methodology to reach good performance for
each controller design. We then compare peak inputs, gains, and controller
energy? of each bicycle and rank them based on these metrics. I the
relationship between the performance and the phyiscal parameter differences are
analyzed and compared with some anecdotal bicycle lore on how these parameters
changes affect handling.

An ability to predict the handling qualities of bicycles with different
physical characteristics remains an important research issue in the study of
manual control of single-track vehicles. Numerous factors affect bicycle
handling qualities as perceived by the rider, e.g., the dynamics of the bicycle
itself, the dynamics of the rider, the control characteristics of the rider,
and the manner in which the rider quantifies his/her opinion of the bicycle's
handling qualities. The work to be described follows in the footsteps
of~\cite{Lunteren1970} and \cite{Weir1972}, and utilizes a human operator model
discussed in~\cite{Hess2006} as applied to
piloted aircraft control. Figure~\ref{fig:block} is a block diagram representation of the
rider model for roll control of a bicycle, with appropriate sensory modalities
noted. As presented in Fig.~\ref{fig:block}, only three gains and a simple neuromuscular
system model parameterize the rider model. By way of exposition in this brief
abstract, six bicycle models were chosen, five from existing bicycles as
parameterized by the first author, and the sixth the ``benchmark'' bicycle from
~\cite{Meijaard2007}. All models are linear and valid for a forward velocity of 5 m/sec. A
simple 2 m lane change and lane return maneuver was selected for simplicity.
Figure~\ref{fig:path} shows the maneuver paths of the six rider/bicycles, where the bicycle
coordinate shown is associated with the rear wheel contact point. The complete
rider model includes two additional outer-loop closures not shown in
Fig.~\ref{fig:plots} (heading and lateral deviation). In addition, a simple model of rider preview
is included. Figure~\ref{fig:inputs} shows the rider steering inputs for the
task. Figure~\ref{fig:handling} shows the ``Handling Qualities Sensitivity
Function'' (HQSFs) for each vehicle.
The magnitude of each of these functions has been shown to correlate well with
handling qualities ratings for aircraft flight control~\cite{Hess2006}. As an example here,
the two functions with the lowest magnitudes in the frequency range of
importance for manual control ($0<\omega$  10 rad/sec) are associated with the two
bicycles models that are either stable or marginally unstable (one root just
into the right-half plane). Finally, Fig.~\ref{fig:bode} shows Bode diagrams for the
rider/bicycle open-loop transfer function for each application, showing that
the rider model follows the dictates of the well-known ``crossover model'' of the
human operator~\cite{McRuer1974}, appropriate for the pure-gain like
$\frac{\phi}{T_\delta}$ dynamics of the
bicycles in question. The final paper will include more bicycle models and
maneuvers and a thorough discussion of applying the HQSF to the prediction of
bicycle handling qualities.
\bibliographystyle{acm}
\bibliography{bicycle}
\end{document}

### 2010-05-04 PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
initial commit of the abstract

@@
 \begin{flushleft}
-  {\fontsize{12}{14}{Jason K. Moore and Mont Hubbard}\\}
+  {\fontsize{12}{14}{Jason K. Moore, Mont Hubbard and Dale L. Peterson}\\}
   \textit{Mechanical and Aerospace Engineering\\
@@
           One Shields Avenue, Davis, CA 95616, USA\\
-          e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu
+          e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu,
+          dlpeterson@ucdavis.edu
   }\vspace{10pt}\\
@@
 rigid bodies' periods of oscillation when hung as a torsional pendulum at
-different orienatation angles. The out-of-plane moments of inertia were
+different orientation angles. The out-of-plane moments of inertia were
 measured by swinging the rigid bodies as a compound pendulum.
@@
 matrices of the system, the eigenvalues and open loop transfer functions at
-different speeds will be compared with typical anedoctal
+different speeds will be compared with typical anecdotal
 descriptions of handling.

### 2010-05-05 BikeTaskComparison/MooreHessHubbardPeterson_abstract.tex
fixed a couple of errors

--- 2010-05-04 BikeTaskComparison/MooreHessHubbard.tex
@@
 \documentclass{bmd2010a}
+
+\usepackage{subfig}
 
@@
 {\fontsize{16pt}{20pt}\selectfont%
-  Manually Controlled Lane Change Manuever with 8 Bicycles\\}
+  Manually Controlled Lane Change Maneuver with 8 Bicycles\\}
 \end{flushleft}
@@
 Fig.~\ref{fig:plots} (heading and lateral deviation). In addition, a simple model of rider preview
-is included. Figure~\ref{fig:inputs} shows the rider steering inputs for the
+is included. Figure~\ref{fig:steer} shows the rider steering inputs for the
 task. Figure~\ref{fig:handling} shows the ``Handling Qualities Sensitivity
@@
 bicycle handling qualities.
+\begin{figure}[tb]
+    \centering
+    \includegraphics[width=\columnwidth]{block.jpg}
+    \caption{Block Diagram of Rider/Bicycle System}
+    \label{fig:block}
+\end{figure}
+\begin{figure}[tb]
+    \centering
+    \subfloat[Lane Change Performance]{\label{fig:path}\includegraphics[width=0.5\columnwidth]{path.png}}
+    \subfloat[Handling Qualities Sensitivity Functions]{\label{fig:handling}\includegraphics[width=0.5\columnwidth]{handling.png}}
+
+    \subfloat[Rider Steering Inputs]{\label{fig:steer}\includegraphics[width=0.5\columnwidth]{steer.png}}
+    \subfloat[Bode Diagrams of Rider/Bicycle Open-Loop Transfer
+    Functions]{\label{fig:bode}\includegraphics[width=0.5\columnwidth]{bode.png}}
+    \caption{Linear simulation results and compartive metrics.}
+    \label{fig:plots}
+\end{figure}
+
 \bibliographystyle{acm}

### 2010-05-05 PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
fixed a couple of errors

@@
 
-\section*{Abstract}
+\section*{Introduction}
 Accurate measurements of a bicycle's physical parameters are required for
@@
 limited to, the benchmark bicycle model. The parameters set is also complete
-for non-linear benchmark bicycle model.
+for the non-linear benchmark bicycle model.
 
@@
     \item[Batavus Browser] The Dutch style city bicycle was measured
-        in two configurations. One with the instrumentation described in
+        in two configurations, one with the instrumentation described in
         \cite{Kooijman2009a} and one as received from the manufacturer.

### 2010-05-05 BikeTaskComparison/MooreHessHubbardPeterson_abstract.tex
manual control abstract final draft

--- 2010-05-04 BikeTaskComparison/MooreHessHubbard.tex
@@
 {\fontsize{16pt}{20pt}\selectfont%
-  Manually Controlled Lane Change Manuever with 8 Bicycles\\}
+  Modeling Manually Controlled Bicycle Maneuvers\\}
 \end{flushleft}
@@
 \begin{flushleft}
-  {\fontsize{12}{14}{Jason K. Moore, Ronald Hess, Mont Hubbard and Dale L.
+  {\fontsize{12}{14}{Ronald Hess, Jason K. Moore, Mont Hubbard and Dale L.
   Peterson}\\}
@@
           One Shields Avenue, Davis, CA 95616, USA\\
-          e-mail: jkmoor@ucdavis.edu, rahess@ucdavis.edu, mhubbard@ucdavis.edu,
+          e-mail: rahess@ucdavis.edu, jkmoor@ucdavis.edu, mhubbard@ucdavis.edu,
           dlpeterson@ucdavis.edu
@@
 \section*{Abstract}
-
-One of the most sought after answers to questions in bicycle dynamics is ``How
-will the bicycle handle differently if its phyiscal characteristics are
-changed?'' or ``Why do different bicycles handle differently''. The answer to
-these questions are not trivial. Many factors play into the picture such as,
-but not limited to, the bicycle's dynamics, the rider's dynamics, the rider's
-control system, the manuever, the performance criteria and the rider's
-subjective opinion of the handling. Of these only the bicycle's dynamics and
-the rider's control system are somewhat understoof. Bicycle models have been
-shown to predict basic motion through some experientation and much research has
-been done in manual control for aircraft pilots.
-
-This work follows in the foot
-steps of \cite{Weir1972} and \cite{Lunteren1970} by using the well defined
-cross-over model to predict the control action of a rigid person with only the
-steering action as an input. We use the physical parameters measured from eight
-different bicycles and one rider along with a multi-loop manual control model to simulate a
-lane change manuever at three speeds: one below, in and above the linear stable
-speed range of the bicycle. Two types of approaches are used. The first is to
-tune the control gains for one of the bicycles for good performance. The same
-controller is then used for the same manuever for each of the other
-bicycle/rider combinations. The overshoot, (other things) are compared and the
-bicycles are ranked according to how good or bad the perform under these
-criteria. Secondly, a separate controller is tuned for each bicycle/rider for
-good performance. We use the same methodology to reach good performance for
-each controller design. We then compare peak inputs, gains, and controller
-energy? of each bicycle and rank them based on these metrics. I the
-relationship between the performance and the phyiscal parameter differences are
-analyzed and compared with some anecdotal bicycle lore on how these parameters
-changes affect handling.
-
 An ability to predict the handling qualities of bicycles with different
@@
 discussed in~\cite{Hess2006} as applied to
-piloted aircraft control. Figure~\ref{fig:block} is a block diagram representation of the
+piloted aircraft control.
+\begin{figure}[hb]
+    \centering
+    \includegraphics[width=0.7\columnwidth]{block.jpg}
+    \caption{Block Diagram of Rider/Bicycle System}
+    \label{fig:block}
+\end{figure}
+
+Figure~\ref{fig:block} is a block diagram representation of the
 rider model for roll control of a bicycle, with appropriate sensory modalities
@@
 abstract, six bicycle models were chosen, five from existing bicycles as
-parameterized by the first author, and the sixth the ``benchmark'' bicycle from
+parameterized by the second author in~\cite{Moore2010}, and the sixth the ``benchmark'' bicycle from
 ~\cite{Meijaard2007}. All models are linear and valid for a forward velocity of 5 m/sec. A
-simple 2 m lane change and lane return maneuver was selected for simplicity.
+simple two meter lane change and lane return maneuver was selected for simplicity.
 Figure~\ref{fig:path} shows the maneuver paths of the six rider/bicycles, where the bicycle
@@
 rider model includes two additional outer-loop closures not shown in
-Fig.~\ref{fig:plots} (heading and lateral deviation). In addition, a simple model of rider preview
-is included. Figure~\ref{fig:inputs} shows the rider steering inputs for the
+Fig.~\ref{fig:block} (heading and lateral deviation). In addition, a simple model of rider preview
+is included. Figure~\ref{fig:steer} shows the rider steering inputs for the
 task. Figure~\ref{fig:handling} shows the ``Handling Qualities Sensitivity
-Function'' (HQSFs) for each vehicle.
+Functions'' (HQSFs) for each vehicle.
 The magnitude of each of these functions has been shown to correlate well with
@@
 the two functions with the lowest magnitudes in the frequency range of
-importance for manual control ($0<\omega$  10 rad/sec) are associated with the two
-bicycles models that are either stable or marginally unstable (one root just
+importance for manual control ($0<\omega<10$ rad/sec) are associated with the two
+bicycle models that are either stable or marginally unstable (one root just
 into the right-half plane). Finally, Fig.~\ref{fig:bode} shows Bode diagrams for the
@@
 bicycle handling qualities.
+\begin{figure}[tb]
+    \centering
+    \subfloat[Lane Change Performance]{\label{fig:path}\includegraphics[width=0.4\columnwidth]{path.png}}
+    \subfloat[Handling Qualities Sensitivity Functions]{\label{fig:handling}\includegraphics[width=0.4\columnwidth]{handling.png}}
+
+    \subfloat[Rider Steering Inputs]{\label{fig:steer}\includegraphics[width=0.4\columnwidth]{steer.png}}
+    \subfloat[Bode Diagrams of Rider/Bicycle Open-Loop Transfer
+    Functions]{\label{fig:bode}\includegraphics[width=0.4\columnwidth]{bode.png}}
+    \caption{Linear simulation results and comparative metrics for six bicycles.}
+    \label{fig:plots}
+\end{figure}
 \bibliographystyle{acm}

### 2010-05-06 BikeTaskComparison/MooreHessHubbardPeterson_abstract.tex
updated the abstract

@@
 manual control of single-track vehicles. Numerous factors affect bicycle
-handling qualities as perceived by the rider, e.g., the dynamics of the bicycle
+handling qualities as perceived by the rider, e.g. the dynamics of the bicycle
 itself, the dynamics of the rider, the control characteristics of the rider,
@@
 of~\cite{Lunteren1970} and \cite{Weir1972}, and utilizes a human operator model
-discussed in~\cite{Hess2006} as applied to
-piloted aircraft control.
+discussed in~\cite{Hess2006} as applied to piloted aircraft control.
 \begin{figure}[hb]
     \centering
-    \includegraphics[width=0.7\columnwidth]{block.jpg}
+    \includegraphics[width=0.65\columnwidth]{block.jpg}
     \caption{Block Diagram of Rider/Bicycle System}
@@
 noted. As presented in Fig.~\ref{fig:block}, only three gains and a simple neuromuscular
-system model parameterize the rider model. By way of exposition in this brief
+system model parameterize the rider. By way of exposition in this brief
 abstract, six bicycle models were chosen, five from existing bicycles as
-parameterized by the second author in~\cite{Moore2010}, and the sixth the ``benchmark'' bicycle from
+parameterized by the second author in~\cite{Moore2010}, and the sixth being the ``benchmark'' bicycle from
 ~\cite{Meijaard2007}. All models are linear and valid for a forward velocity of 5 m/sec. A
-simple two meter lane change and lane return maneuver was selected for simplicity.
+simple two meter lane change and lane return maneuver was selected as a simple
+task.
 Figure~\ref{fig:path} shows the maneuver paths of the six rider/bicycles, where the bicycle
@@
 the rider model follows the dictates of the well-known ``crossover model'' of the
-human operator~\cite{McRuer1974}, appropriate for the pure-gain like
+human operator~\cite{McRuer1974}, appropriate for the
 $\frac{\phi}{T_\delta}$ dynamics of the
-bicycles in question. The final paper will include more bicycle models and
+bicycles in question. The final paper will include nine bicycle models and
 maneuvers and a thorough discussion of applying the HQSF to the prediction of
@@
     \centering
-    \subfloat[Lane Change Performance]{\label{fig:path}\includegraphics[width=0.4\columnwidth]{path.png}}
-    \subfloat[Handling Qualities Sensitivity Functions]{\label{fig:handling}\includegraphics[width=0.4\columnwidth]{handling.png}}
+    \subfloat[Lane Change Performance]{\label{fig:path}\includegraphics[width=0.425\columnwidth]{path.png}}
+    \subfloat[Handling Qualities Sensitivity Functions]{\label{fig:handling}\includegraphics[width=0.425\columnwidth]{handling.png}}
 
-    \subfloat[Rider Steering Inputs]{\label{fig:steer}\includegraphics[width=0.4\columnwidth]{steer.png}}
+    \subfloat[Rider Steering Inputs]{\label{fig:steer}\includegraphics[width=0.425\columnwidth]{steer.png}}
     \subfloat[Bode Diagrams of Rider/Bicycle Open-Loop Transfer
-    Functions]{\label{fig:bode}\includegraphics[width=0.4\columnwidth]{bode.png}}
+    Functions]{\label{fig:bode}\includegraphics[width=0.425\columnwidth]{bode.png}}
     \caption{Linear simulation results and comparative metrics for six bicycles.}

### 2010-05-06 BenchmarkBikeTools.py
doc string updated and change the pi import

@@ -1,11 +1,22 @@
 def bmp2cm(filename):
-    ''' Converts the benchmark bicycle parameters to the canonical matrices'''
-    import numpy as np
+    '''Return the benchmark canonical matrices from the bicycle parameters. Formulated from Meijaard et al. 2007.
+
+    Input:
+        filename is a text file with the 27 parameters listed in csv format. One
+        parameter per line in the form: lamba, 10/pi
+    Output:
+        M is the mass matrix
+        C1 is the damping like matrix that is proportional to v
+        K0 is the stiffness matrix proportional to gravity
+        K2 is the stiffness matrix proportional to v**2
+        p is a dictionary of the parameters. the keys are the variable names
+
+        '''
+    from numpy import pi
     f = open(filename, 'r')
     p = {}
     for i, line in enumerate(f):
         list = line[:-3].split(',')
-        pi = np.pi
         p[list[0]] = eval(list[1])
     mT = p['mR'] + p['mB'] + p['mH'] + p['mF']
     xT = (p['xB']*p['mB'] + p['xH']*p['mH'] + p['w']*p['mF'])/mT
@@ -35,7 +46,7 @@ def bmp2cm(filename):
     Mdp = Mpd
     Mdd = IAll + 2*mu*IAlz + mu**2*ITzz
     M = np.array([[Mpp, Mpd], [Mdp, Mdd]])
-    K0pp = mT*zT # this value only reports to 13 digit precision?
+    K0pp = mT*zT # this value only reports to 13 digit precision it seems?
     K0pd = -SA
     K0dp = K0pd
     K0dd = -SA*np.sin(p['lambda'])

### 2010-05-06 BenchmarkBikeTools.py
added the B matrix calc

@@ -64,8 +64,18 @@ def bmp2cm(filename):
     return M, C1, K0, K2, p
 
 def aMatrix(M, C1, K0, K2, p):
-    '''Calculates the A matrix from the canonical matrices for the benchmark
-    bicycle'''
+    '''Calculate the A and B matrices for the benchmark bicycle
+
+    Input:
+        M is the mass matrix
+        C1 is the damping like matrix that is proportional to v
+        K0 is the stiffness matrix proportional to gravity
+        K2 is the stiffness matrix proportional to v**2
+        p is a dictionary of the parameters. the keys are the variable names
+    Returns:
+        A system dynamic matrix
+        B control matrix
+    '''
     from numpy import eye, zeros, vstack, hstack, dot
     from numpy.linalg import inv
     a11 = -p['v']*C1
@@ -73,7 +83,8 @@ def aMatrix(M, C1, K0, K2, p):
     a21 = eye(2)
     a22 = zeros((2, 2))
     A = vstack((dot(inv(M), hstack((a11, a12))), hstack((a21, a22))))
-    return A
+    B = vstack(inv(M), zeros((2, 2)))
+    return A, B
 
 def tor_inertia(k, T):
     '''Calculates the moment of interia for an ideal torsional pendulm'''

### 2010-05-06 BenchmarkBikeTools.py
fixed import

@@ -12,9 +12,10 @@ def bmp2cm(filename):
         p is a dictionary of the parameters. the keys are the variable names
 
         '''
-    from numpy import pi
+    from numpy import pi, cos, sin, array
     f = open(filename, 'r')
     p = {}
+    # parse the text file
     for i, line in enumerate(f):
         list = line[:-3].split(',')
         p[list[0]] = eval(list[1])
@@ -32,11 +33,11 @@ def bmp2cm(filename):
     IAxx = p['IHxx'] + p['IFxx'] + p['mH']*(p['zH'] - zA)**2 + p['mF']*(p['rF'] + zA)**2
     IAxz = p['IHxz'] - p['mH']*(p['xH'] - xA)*(p['zH'] - zA) + p['mF']*(p['w'] - xA)*(p['rF'] + zA)
     IAzz = p['IHzz'] + p['IFzz'] + p['mH']*(p['xH'] - xA)**2 + p['mF']*(p['w'] - xA)**2
-    uA = (xA - p['w'] - p['c'])*np.cos(p['lambda']) - zA*np.sin(p['lambda'])
-    IAll = mA*uA**2 + IAxx*np.sin(p['lambda'])**2 + 2*IAxz*np.sin(p['lambda'])*np.cos(p['lambda']) + IAzz*np.cos(p['lambda'])**2
-    IAlx = -mA*uA*zA + IAxx*np.sin(p['lambda']) + IAxz*np.cos(p['lambda'])
-    IAlz = mA*uA*xA + IAxz*np.sin(p['lambda']) + IAzz*np.cos(p['lambda'])
-    mu = p['c']/p['w']*np.cos(p['lambda'])
+    uA = (xA - p['w'] - p['c'])*cos(p['lambda']) - zA*sin(p['lambda'])
+    IAll = mA*uA**2 + IAxx*sin(p['lambda'])**2 + 2*IAxz*sin(p['lambda'])*cos(p['lambda']) + IAzz*cos(p['lambda'])**2
+    IAlx = -mA*uA*zA + IAxx*sin(p['lambda']) + IAxz*cos(p['lambda'])
+    IAlz = mA*uA*xA + IAxz*sin(p['lambda']) + IAzz*cos(p['lambda'])
+    mu = p['c']/p['w']*cos(p['lambda'])
     SR = p['IRyy']/p['rR']
[…]

### 2010-05-27 MakeParTables.py
added the make tables python code

@@ -1,0 +1,22 @@
+import pickle
+import re
+import uncertainties as un
+
+# load in the base data file
+f = open('data/data.p', 'r')
+data = pickle.load(f)
+f.close()
+
+nBk = len(data['bikes'])
+
+# load in the parameter data file
+f = open('data/par.p', 'r')
+par = pickle.load(f)
+f.close()
+
+for i, name in enumerate(data['bikes']):
+    fname = ''.join(name.split()) + 'RiderCan.p'
+    file = open(dir + fname)
+    can = pickle.load(file)
+    file.close()
+

### 2010-05-27 MakeParTables.py
successfully replacing quantities in the tex docs

@@ -1,5 +1,6 @@
 import pickle
 import re
+import os
 import uncertainties as un
 
 # load in the base data file
@@ -14,9 +15,39 @@ f = open('data/par.p', 'r')
 par = pickle.load(f)
 f.close()
 
+direct = 'ParTables'
 for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split()) + 'RiderCan.p'
-    file = open(dir + fname)
-    can = pickle.load(file)
-    file.close()
-
+    fname = ''.join(name.split()) 
+    # open the new file
+    f = open(direct + '/ParameterTable.tex', 'r')
+    fn = open(direct + '/' + fname + 'RiderPar.tex', 'w') 
+    for line in f:
+        print line
+        # find all of the matches in the line
+        test = re.findall('\|(\w*.)\|', line)
+        print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
+        print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
+        # if there are matches
+        if test:
+            print 'Found this!\n', test
+            # go through each match and make a substitution
+            for match in test:
+                print "replace this: ", match
+                if match[-1] == '?':
+                    try:
+                        line = re.sub('\|(\w*\?)\|', '%-6.3f' % par[match[:-1]][i].std_dev(), line,
+                                count=1)
+                        print line
+                    except:
+                        pass
+                else:
+                    try:
+                        line = re.sub('\|(\w*(?!\?))\|',
+                                str(par[match][i].nominal_value), line, count=1)
+                        print line
+                    except:
+                        pass
+        print line
+        fn.write(line)
+    f.close()
+    fn.close()

### 2010-05-27 MakeParTables.py
makes the pdfs now

@@ -42,8 +42,7 @@ for i, name in enumerate(data['bikes']):
                         pass
                 else:
                     try:
-                        line = re.sub('\|(\w*(?!\?))\|',
-                                str(par[match][i].nominal_value), line, count=1)
+                        line = re.sub('\|(\w*(?!\?))\|', '%.3f' % par[match][i].nominal_value, line, count=1)
                         print line
                     except:
                         pass
@@ -51,3 +50,6 @@ for i, name in enumerate(data['bikes']):
         fn.write(line)
     f.close()
     fn.close()
+    os.system('pdflatex -output-directory=ParTables ' + direct + '/' + fname + 'RiderPar.tex')
+    os.system('rm ParTables/*.aux')
+    os.system('rm ParTables/*.log')

### 2010-05-27 MakeParTables.py
adds the bike name to the pdf

@@ -22,31 +22,33 @@ for i, name in enumerate(data['bikes']):
     f = open(direct + '/ParameterTable.tex', 'r')
     fn = open(direct + '/' + fname + 'RiderPar.tex', 'w') 
     for line in f:
-        print line
+        #print line
         # find all of the matches in the line
         test = re.findall('\|(\w*.)\|', line)
-        print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
-        print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
+        #print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
+        #print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
         # if there are matches
         if test:
             print 'Found this!\n', test
             # go through each match and make a substitution
             for match in test:
-                print "replace this: ", match
+                #print "replace this: ", match
                 if match[-1] == '?':
                     try:
                         line = re.sub('\|(\w*\?)\|', '%-6.3f' % par[match[:-1]][i].std_dev(), line,
                                 count=1)
-                        print line
+                        #print line
                     except:
                         pass
+                elif match == 'bikename':
+                    line = re.sub('\|bikename\|', name, line)
                 else:
                     try:
                         line = re.sub('\|(\w*(?!\?))\|', '%.3f' % par[match][i].nominal_value, line, count=1)
-                        print line
+                        #print line
                     except:
                         pass
-        print line
+        #print line
         fn.write(line)
     f.close()
     fn.close()

### 2010-05-27 MakeParTables.py
makes the parameters for the bikes only

@@ -17,10 +17,10 @@ f.close()
 
 direct = 'ParTables'
 for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split()) 
+    fname = ''.join(name.split())
     # open the new file
     f = open(direct + '/ParameterTable.tex', 'r')
-    fn = open(direct + '/' + fname + 'RiderPar.tex', 'w') 
+    fn = open(direct + '/' + fname + 'Par.tex', 'w')
     for line in f:
         #print line
         # find all of the matches in the line
@@ -29,7 +29,7 @@ for i, name in enumerate(data['bikes']):
         #print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
         # if there are matches
         if test:
-            print 'Found this!\n', test
+            #print 'Found this!\n', test
             # go through each match and make a substitution
             for match in test:
                 #print "replace this: ", match
@@ -41,17 +41,17 @@ for i, name in enumerate(data['bikes']):
                     except:
                         pass
                 elif match == 'bikename':
-                    line = re.sub('\|bikename\|', name, line)
+                    line = re.sub('\|bikename\|', name.upper(), line)
                 else:
                     try:
                         line = re.sub('\|(\w*(?!\?))\|', '%.3f' % par[match][i].nominal_value, line, count=1)
                         #print line
                     except:
-                        pass
+                        line = re.sub('\|(\w*(?!\?))\|', '%.3f' % par[match][i], line, count=1)
         #print line
         fn.write(line)
     f.close()
     fn.close()
-    os.system('pdflatex -output-directory=ParTables ' + direct + '/' + fname + 'RiderPar.tex')
+    os.system('pdflatex -output-directory=ParTables ' + direct + '/' + fname + 'Par.tex')
     os.system('rm ParTables/*.aux')
     os.system('rm ParTables/*.log')

### 2010-05-28 report/parameterReport.tex
typed in the report and adding trail functions

@@
 \documentclass{article}
-\usepackage{hyperref}
+\usepackage[colorlinks=true, linkcolor=blue]{hyperref}
 \usepackage{graphicx}
 
+\author{Jason K. Moore}
+\title{Accurate Measurement of Eight Bicycles' Physical Parameters}
+
 \begin{document}
-\section{Bicycle}
+\maketitle
+\tableofcontents
 
-\subsection{Geometry}
-\subsubsection{Wheel radius}
-The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were measured by measuring the linear distance traversed along the ground through eitehr 13 or 14 rotations of the wheel.
-Each wheel was measured separatley and the measurements were taken with a 72kg rider seated on the bicycle.
-A 30 meter tape measure (resolution: 2mm) was taped on a flat level smooth floor.
-The tire was marked with chalk and aligned with the tape measure.
-The accuracy of the distance measurement is approximately $\pm0.005$m.
-The tires were pump the recommended inflation pressure before the measurments.
-The wheel radius is calculated by
+\section{Introduction}
+D\"{o}hring measured the physical parameters of a scooter~\cite{Dohring1953}.
+Singh and Goel measured the physical parameters of a scooter~\cite{Singh1971}.
+Roland and Massing measured the physical parameters of a bicycle in much the
+same was as presented, including calculations of uncertainty from the indirect
+measurement techniques~\cite{Roland1971}. This work is based off of the work
+done by Kooijman~\cite{Kooijman2006}. I worked with Jodi Kooijman for a year and
+was able to use much of the same apparatus and refine the measurement
+technique.
+\section{Accuracy}
+Error propagation theory was utilized to calculate what the accuracy the
+benchmark ideal parameters are based on the measurements taken.
+
+\section{Geometry}
+
+\subsection{Wheel radius}
+The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were
+measured by measuring the linear distance traversed along the ground through
+either 13 or 14 rotations of the wheel. Each wheel was measured separately and
+the measurements were taken with a 72kg rider seated on the bicycle. A 30 meter
+tape measure (resolution: 2mm) was taped on a flat level smooth floor. The tire
+was marked with chalk and aligned with the tape measure. The accuracy of the
+distance measurement is approximately $\pm0.005$m. The tires were pumped to the
+recommended inflation pressure before the measurements. The wheel radius is
+calculated by
 \begin{equation}
@@
 	\begin{center}
-		\includegraphics[width=4in]{tireChalk.jpg}
+		\includegraphics[width=4in]{images/tireChalk.jpg}
 	\end{center}
@@
 \end{table}
-\subsubsection{Head tube angle}
-The head tube angle was measured directly using an electronic level with a $\pm0.2^{\circ}$ accuracy.
-The bicycle was fixed vertically, the front wheel was aligned with the rear frame and the bicycle was unloaded.
-The steer axis tilt $\lambda$ is the complement to the head tube angle.
+\subsection{Head tube angle}
+The head tube angle was measured directly using an electronic level with a
+$\pm0.2^{\circ}$ accuracy. The bicycle was fixed vertically, the front wheel
+was aligned with the rear frame and the bicycle was unloaded. The steer axis
+tilt $\lambda$ is the complement to the head tube angle.
 \begin{equation}
@@
 
-\subsubsection{Trail}
-Trail is difficult to measure directly so we instead chose to measure the fork offset.
-The fork offset was measured by clamping the steer tube of the front fork into a v-block on a flat table.
-A ruler was used to measure the height of the center of the head tube and the height of the center of the axle axis.
-The fork blades where aligned such that the axle axis was parallel to the table surface.
+\subsection{Trail}
+Trail is difficult to measure directly so we instead chose to measure the fork
+offset. The fork offset was measured by clamping the steer tube of the front
+fork into a v-block on a flat table. A ruler was used to measure the height of
+the center of the head tube and the height of the center of the axle axis. The
+fork blades where aligned such that the axle axis was parallel to the table
+surface.
 \begin{equation}
@@
 \end{table}
-\subsubsection{Wheelbase}
-The wheelbase was measured directly using a tape measure while the bicycle was in the same uprigth fixed position used when measuring the head tube angle.
+\subsection{Wheelbase}
+The wheelbase was measured directly using a tape measure while the bicycle was
+in the same upright fixed position used when measuring the head tube angle.
 \begin{table}
@@
 
-\subsection{Mass}
-The total mass of each bicycle was measured using a spring scale with a resolution of 100g. Then each of the four bicycle parts were measured using a Molen 20kg scale with a resulution of 20g.
+\section{Mass}
+The total mass of each bicycle was measured using a spring scale with a
+resolution of 100g. Then each of the four bicycle parts were measured using a
+Molen 20kg scale with a resolution of 20g.
 \begin{table}
@@
 \end{table}
-\subsection{Center of Mass}
+\section{Center of Mass}
+\subsection{Wheels}
 The centers of mass of the wheels are assumed to be at their geometrical centers.
-\subsubsection{Rear frame}
-The rear frame was hung in three orientations from a torsional pendulum.
-The frame could rotate about a joint such that gravity aligned the center of mass with the pendulum axis.
-The new orientaion angle of the headtube $\alpha$ was measured using the digital level.
-A string was aligned with the pendulum axis and allowed to pass by the frame.
-The horizontal distance $d$ between the rear axle and the string (CoM line) was measured by aligning a ruler pendicular to the string.
-The distance $d$ was negative if the string fell to the right of the rear axle and positive if it fell to the left of the rear axle.
-These measurements allow for the calucation of the center of mass location in the global reference frame.
-The frame rotation angle $\beta$ is defined as rotation of the frame in the nominal configuration to the hanging orientation, rotated about the $Y$ axis.
+\subsection{Rear frame}
+The rear frame was hung in three orientations as a torsional pendulum (the
+pendulum was used to measure moments of inertia). I
+assumed that the frame was laterally symmetric. The
+frame could rotate about a joint such that gravity aligned the center of mass
+with the pendulum axis. The orientation angle of the headtube $\alpha$
+Figure~\ref{fig:angles} relative
+to the earth was
+measured using the digital level Figure~\ref{fig:level}. A string was aligned with the pendulum axis
+and allowed to pass by the frame. The horizontal distance $a_\mathrm{B}$ between the rear
+axle and the string (CoM line) was measured by aligning a ruler perpendicular to
+the string. The distance $a$ was negative if the string fell to the right of
+the rear axle and positive if it fell to the left of the rear axle. These
+measurements allow for the calculation of the center of mass location in the
+global reference frame.
+\begin{figure}[tb]
+	\begin{center}
+		\includegraphics[width=5in]{images/level.jpg}
+	\end{center}
+	\caption{The digital level was mounted to a straight edge that was aligned
+    with the headtube of the bicycle frame. This was done without allowing the
+    straight edge to touch the frame. The frame wasn't completely stationary so
+    this was difficult. The light frame oscillations could be damped out by
+    submerging a low hanging area of the frame into a bucket of water to
+    decrease the oscillation.}
+	\label{fig:level}
+\end{figure}
+The frame rotation angle $\beta$ is defined as rotation of the frame in the
+nominal configuration to the hanging orientation, rotated about the $Y$ axis.
+\begin{figure}[tb]
+	\begin{center}
+		\includegraphics[width=4in]{figures/angles.pdf}
+	\end{center}
+	\caption{Pictorial description of the angles and dimensions that related
+    the nominal bicycle reference frame $XYZ_B$ with the pendulum reference frame
+    $XYZ_P$.}
+	\label{fig:angles}
+\end{figure}
 \begin{equation}
-	\beta=\alpha-\lambda-\frac{\pi}{2}
+	\beta=\lambda-\alpha
     \label{eq:frameRotAng}
 \end{equation}
-The CoM line is simply a line in the global reference frame with a slope $m$ and a z-intercept $b$. The slope can be shown to be
+The pendulum axis $X_P$ is simply a line in the nominal bicycle reference frame with a slope $m$ and a z-intercept $b$. The slope can be shown to be
 \begin{equation}
-	m=\tan{\left(\beta-\frac{\pi}{2}\right)}
+	m_i=-\tan{\beta_i}
 \label{eq:slope}
@@
 \begin{equation}
-	b=-\frac{d}{\sin{\beta}}-r_\mathrm{R}
+	b_i=\frac{d_i}{\cos{\beta_i}}-r_\mathrm{R}
     \label{eq:zInt}
 \end{equation}
-The CoM location can be calculated by finding the intersection of these three lines. The system is over determined so the least squares method is used to find a unique solution.
+Only two lines are required to calculate the center of mass of the laterally
+symmetric frame, but more orientations increase the center of mass measurement
+accuracy. The three lines are defined as:
+\begin{equation}
+   y = m_ix+b_i
+   \label{eq:line}
+\end{equation}
+The CoM location can be calculated by finding the intersection of these three
+lines. Two approaches were used used to calculate the center of mass. Intuition
+leads me to think that the center of mass is located at the centroid of the
+triangle made by the three intersecting lines. The centroid can be found by
+calculating the intersection point of each pair of lines and then averaging the
+three intersection points.
+\begin{equation}
+	\left[
+	\begin{array}{cc}
+		-m_1 & 1\\
+		-m_2 & 1
+	\end{array}
+	\right]
+	\left[
+	\begin{array}{c}
+		x_a\\
+		z_a
+	\end{array}
+	\right]
+	=
+	\left[
+	\begin{array}{c}
+		b_1\\
+		b_2
+	\end{array}
+	\right]
+\label{eq:linearSystem}
+\end{equation}
+\begin{equation}
+    x_\mathrm{B} = \frac{x_a + x_b + x_c}{3}
+\end{equation}
+\begin{equation}
+    z_\mathrm{B} = \frac{z_a + z_b + z_c}{3}
+\end{equation}
+Alternatively the three lines can be treated as an over determined linear
+system and the least squares method is used to find a unique solution. This
+solution is not at the centroid of the triangle made by the intersecting lines.
+I am not sure which solution is theoretically the correct one for the location
+of the center of mass.
 \begin{equation}
@@
 	\begin{array}{c}
-		x\\
-		z
+        x_\mathrm{B}\\
+        z_\mathrm{B}
 	\end{array}
@@
 	\right]	
-\label{eq:solveSlopeInt}
+\label{eq:leastSquares}
 \end{equation}
-\subsection{Moment of Inertia}
-\section{Human}
+\subsection{Fork}
+The center of mass of the fork is calculated in the same fashion, only the
+z-intercept is different.
+\begin{equation}
+    b = r_\mathrm{F} + \frac{a}{\cos{\beta}} + w\tan{\beta}
+\end{equation}
+\section{Moment of Inertia}
+The moments of inertia of the wheels, frame and fork were measured by taking
+advantage of the assumed symmetry of the parts and by hanging the parts as both
+compound and torsional pendulums and measuring the period of oscillation when
+perturb at small angles. The rate of oscillation was measured was measured
+using a \href{http://www.siliconsensing.com/CRS03}{Silicon Sensing CRS03 100
+deg/s rate gyro}. The rate gyro was sampled at
+1000hz with a
+\href{http://sine.ni.com/nips/cds/view/p/lang/en/nid/14604}{National
+Instruments USB-6008 12 bit data acquisition unit} and
+Matlab. The measurement duration were either 15 or 30 secs and each moment of
+inertia measurement was performed three times. No extra care was taken to
+calibrate the rate gyro, maintain a constant power source (i.e. the battery
+drains slowly), or account for drift. The raw voltage signal was used to
+determine only the period of oscillation which is need for the moment of inertia
+calculations.
+\subsection{Torsional Pendulum}
+A torsional pendulum was used to measure all of moments of inertia about axes
+in the lateral symmetric plane of each of the wheels, fork and frame. The
+pendulum is made up of a rigid mount, an upper clamp, a torsion rod, and
+various lower clamps.
+A 5mm diameter, 1m long mild steel rod was used as the torsion
+spring. A lightweight, low relative moment of inertia clamp was constructed
+that could clamp the rim and the tire. The wheel was hung freely such that the
+center of mass aligned with the torsional pendulum axis and then it was
+secured. The wheel was then perturbed and it oscillated about the pendulum axis.
+The rate gyro was mounted on the clamp in line with the pendulum axis. The
+radial moment of inertia can can calculated as such:
+
+\subsection{Wheels}
+The wheels are assumed to be symmetric laterally and about any radial axis. Thus only two
+moment's of inertia are required. The moment of inertia about the axle was
+measured by hanging the wheel as a compound pendulum. The wheel was hung on a
+horizontal rod and perturbed to oscillate about the axis of the rod. This rate
+gyro was attached to the spokes near the hub and orientated mostly along the
+axle axis. The wheels tended to precess at the contact point about the vertical
+axis. This adds a very low frequency component of rate along the vertical axis,
+but should not affect the compound pendulum axis. A better fixture for the
+wheel could prevent the precession. The pendulum arm length is the distance
+from the rod/rim contact point to the mass center of the wheel. The inner
+diameter of the rim was measured and divided by two to get $l$. The moment of
+inertia about the axle is calculated by:
+\begin{equation}
+    I_{\mathrm{R}yy} = \left(\frac{\bar{T}}{2\pi}\right)^2m_\mathrm{R}gl_\mathrm{R} -
+    m_\mathrm{R}l^2
+\end{equation}
+
+The radial moment of inertia was measured by hanging the wheel as a torsional
+pendulum. A 5mm diameter, 1m long mild steel rod was used as the torsion
+spring. A lightweight, low relative moment of inertia clamp was constructed
+that could clamp the rim and the tire. The wheel was hung freely such that the
+center of mass aligned with the torsional pendulum axis and then it was
+secured. The wheel was then perturbed and it oscillated about the pendulum axis.
+The rate gyro was mounted on the clamp in line with the pendulum axis. The
+radial moment of inertia can can calculated as such:
+\begin{equation}
+    I_{xx} = \frac{k\bar{T}^2}{4\pi^2}
+\end{equation}
+\subsection{Frame}
+\subsection{Fork}
 \appendix
-\subsection{Bicycle Descriptions}
-\subsubsection{Yellow Bicycle}
-The yellow is a bicycle used in the lab to demonstrate that a bicycle is stable at certain speeds.
-It is an aluminum road frame of unknown make with the majority of components removed.
-The wheels, drop handlebar, seat, seat post and bottom bracket are the only parts on the bike.
-The fork has been reversed to ''increase`` the stability of the bicycle.
-The bicycle was measured with both the fork in normal position and the fork reversed.
+\section{Bicycle Descriptions}
+\subsection{Yellow Bicycle}
+The yellow is a bicycle used in the lab to demonstrate that a bicycle is stable
+at certain speeds. It is an aluminum road frame of unknown make with the
+majority of components removed. The wheels, drop handlebar, seat, seat post and
+bottom bracket are the only parts on the bike. The fork has been reversed to
+''increase`` the stability of the bicycle. The bicycle was measured with both
+the fork in normal position and the fork reversed.
 \begin{table}
@@
 
-\subsubsection{Batavus Browser}
-The Batavus Browser is lower priced Dutch city bike. We measured the physical properities of the stock Browser model and also the same bike with the instrumentation used in our experiments.
+\subsection{Batavus Browser}
+The Batavus Browser is lower priced Dutch city bike. We measured the physical
+properities of the stock Browser model and also the same bike with the
+instrumentation used in our experiments.
 \begin{table}
@@
 
-\subsubsection{2007 Bianchi Pista}
-The Pista is a steel track bicycle. Specifications, taken from the Bianchi website~\cite{Bianchi2009}, are as follows:
+\subsection{2007 Bianchi Pista}
+The Pista is a steel track bicycle. Specifications, taken from the Bianchi
+website~\cite{Bianchi2009}, are as follows:
 \begin{table}
@@
 \end{table}
-\subsubsection{Gary Fisher Mountain Bike}
-\subsubsection{Batavus Crescendo Deluxe}
+\subsection{Gary Fisher Mountain Bike}
+\subsection{Batavus Crescendo Deluxe}
 

### 2010-05-28 benchmark_bike_tools.py

@@ -508,3 +508,27 @@ def parallel_axis(Ic, m, d):
     dMat[1] = array([-a*b, c**2 + a**2, -b*c])
     dMat[2] = array([-a*c, -b*c, a**2 + b**2])
     return Ic + m*dMat
+
+def trail(rF, lam, fo):
+    '''Caluculate the trail and mechanical trail
+
+    Parameters:
+    -----------
+    rF: float
+        The front wheel radius
+    lam: float
+        The steer axis tilt (pi/2 - headtube angle)
+    fo: float
+        The fork offset
+
+    Returns:
+    --------
+    c: float
+        Trail
+    c: float
+        Mechanical Trail
+
+    '''
+    from math import sin, cos
+
+    c = (rF*sin(lam) - fo)/cos(lam)

### 2010-05-29 benchmark_bike_tools.py
trail works

@@ -532,3 +532,5 @@ def trail(rF, lam, fo):
     from math import sin, cos
 
     c = (rF*sin(lam) - fo)/cos(lam)
+    cm = c*cos(lam)
+    return c, cm

### 2010-06-01 report/parameterReport.tex
typed more on the report

@@
 \documentclass{article}
+\usepackage[lmargin=1in,rmargin=1in]{geometry}
 \usepackage[colorlinks=true, linkcolor=blue]{hyperref}
@@
 Error propagation theory was utilized to calculate what the accuracy the
-benchmark ideal parameters are based on the measurements taken.
+benchmark ideal parameters are based on the measurements taken. If $x$ is a
+random variable and is defined as $x=f(u,v,\ldots)$ then the sample variance of
+$x$ is defined as
+\begin{equation}
+    s_x^2 = \frac{1}{N-1}\sum^N_{i=1}
+    \left[(u_i - \bar{u})^2\left(\frac{\partial x}{\partial u}\right)^2 +
+    (v_i - \bar{v})^2\left(\frac{\partial x}{\partial v}\right)^2 +
+    2(u_i - \bar{u})(v_i - \bar{v})\left(\frac{\partial x}{\partial u}\right)\left(\frac{\partial x}{\partial v}\right)
+    + \ldots\right]
+    \label{eqn:sampleVariance}
+\end{equation}
+Using the definitions for variance and covariance
+Equation~\ref{eqn:sampleVariance} can be simplified to
+\begin{equation}
+    s_x^2 = s_u^2\left(\frac{\partial x}{\partial u}\right)^2 +
+            s_v^2\left(\frac{\partial x}{\partial v}\right)^2 +
+            2s_{uv}\left(\frac{\partial x}{\partial u}\right)\left(\frac{\partial x}{\partial v}\right)
+            + \ldots
+    \label{eqn:variance}
+\end{equation}
+If $u$ and $v$ are uncorrelated then $s_{uv}=0$. We assume that the all the
+random variables hereafter are uncorrelated. Equation~\ref{eqn:variance} can be
+used to calculated the variance of all types of functions. The most basic are
+functions such as addtion
+\begin{eqnarray}
+    \label{eqn:addition}
+    x &=&  au + bv\\
+    s_x &=& a^2s_u^2 + b^2s_v^2
+\end{eqnarray}
 
@@
 	Bicycle   & B      & B*     & C      & G      & P      & S      & Y \& Y*\\
+    \hline
 	Front\\
@@
 	\begin{tabular}{lllllllll}
+    \hline
 	Bicycle   & B      & B*     & C      & G      & P      & S      & Y      & Y*\\
@@
 	Bicycle     & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
+    \hline
 	Fork offset & 0.071 & 0.071 & 0.045 & 0.039 & 0.032 & 0.045 & 0.057 & -0.057\\
@@
 	Bicycle    & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
+    \hline
 	Wheel base & 1.121 & 1.121 & 1.101 & 1.070 & 0.989 & 1.037 & 1.089 & 0.985
@@
 \begin{table}
-	\begin{tabular}{lllllllll}
-		Bicycle     & B     & B*    & C     & G     & P    & S     & Y     & Y*\\
-		total       & 18.5  & 23.5  & 20.9  & 10.5  & 9.9  & 17.6  & 10.2  & 10.2\\
-		frame       & 9.86  & 14.71 & 9.18  & 4.48  & 4.49 & 7.22  & 3.31  & 3.31\\
-		fork        &       & 3.22  & 4.57  & 2.52  & 2.27 & 3.04  & 2.45  & 2.45\\
-		rear wheel  & 3.11  & 3.11  & 3.96  & 1.94  & 1.38 & 3.96  & 2.57  & 2.57\\
-		front wheel & 2.02  & 2.02  & 3.55  & 1.50  & 1.58 & 3.33  & 1.90  & 1.90\\
-		total sum   & 18.21 & 23.06 & 21.26 & 10.44 & 9.72 & 17.55 & 10.23 & 10.23
+    \centering
+	\begin{tabular}{llllllllll}
+		Bicycle &     $\sigma_m$ & B     & B*    & C     & G     & P    & S     & Y     & Y*\\
+        \hline
+		total   &    0.1       & 18.5  & 23.5  & 20.9  & 10.5  & 9.9  & 17.6  & 10.2  & 10.2\\
+        \hline
+		frame   &    0.02      & 9.86  & 14.71 & 9.18  & 4.48  & 4.49 & 7.22  & 3.31  & 3.31\\
+		fork    &    0.02      & 3.22  & 3.22  & 4.57  & 2.52  & 2.27 & 3.04  & 2.45  & 2.45\\
+		rear wheel & 0.02      & 3.11  & 3.11  & 3.96  & 1.94  & 1.38 & 3.96  & 2.57  & 2.57\\
+		front wheel& 0.02      & 2.02  & 2.02  & 3.55  & 1.50  & 1.58 & 3.33  & 1.90  & 1.90\\
+		total sum  & 0.04      & 18.21 & 23.06 & 21.26 & 10.44 & 9.72 & 17.55 & 10.23 & 10.23
 	\end{tabular}
@@
 advantage of the assumed symmetry of the parts and by hanging the parts as both
-compound and torsional pendulums and measuring the period of oscillation when
+compound and torsional pendulums and measuring their period of oscillation when
 perturb at small angles. The rate of oscillation was measured was measured
@@
 calculations.
+\begin{figure}[tbp]
+    \begin{center}
+        \includegraphics[width=5in]{../data/pendDat/graphs/BrowserFrameCompoundFirst1.png}
+    \end{center}
+    \caption{Example of the raw voltage data taken during a 30 second
+    measurement of the oscillation of one of the components.}
+    \label{fig:voltage}
+\end{figure}
 \subsection{Torsional Pendulum}

### 2010-06-01 fit_data.py

@@ -22,7 +22,7 @@ for name in filenames:
     time = float(pendDat['duration'][0])
     x = np.linspace(0, time, num=len(y))
     # plot the original data
-    plt.plot(x, y, '.')
+    plt.plot(x, y, '.', markersize=14)
     # decaying oscillating exponential function
     fitfunc = lambda p, t: p[0] + np.exp(-p[3]*p[4]*t)*(p[1]*np.sin(p[4]*np.sqrt(1-p[3]**2)*t) + p[2]*np.cos(p[4]*np.sqrt(1-p[3]**2)*t))
     # initial guesses
@@ -33,9 +33,9 @@ for name in filenames:
     p1, success = op.leastsq(errfunc, p0[:], args=(x, y))
     # plot the fitted curve
     lscurve = fitfunc(p1, x)
-    plt.plot(x, lscurve, 'k')
+    plt.plot(x, lscurve, 'k-')
     plt.xlabel('Time [s]')
-    plt.ylabel('Amplitude')
+    plt.ylabel('Amplitude [V]')
     #plt.show()
     plt.savefig('data/pendDat/graphs/' + name[:-4] + '.png')
     plt.close()

### 2010-06-01 benchmark_bike_tools.py
osciallation plotting function

@@ -1,3 +1,43 @@
+def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
+    '''Plot fitted data over the measured
+
+    Parameters:
+    -----------
+    t : ndarray (n,)
+        Measurement time in seconds
+    ym : ndarray (n,)
+        The measured voltage
+    yf : ndarray (n,)
+    p : ndarray (5,)
+        The fit parameters for the decaying osicallation fucntion
+    rsq : float
+        The r squared value of y (the fit)
+    T : float
+        The period
+
+    Returns:
+    --------
+    fig : the figure
+
+    '''
+    from matplotlib.pyplot import figure, plot, xlabel, ylabel, text
+    if fig:
+        fig = fig
+    else:
+        fig = figure()
+    plot(t, ym, '.', markersize=14)
+    plot(t, yf, 'k-')
+    xlabel('Time [s]')
+    ylabel('Amplitude [V]')
+    equation = 'f(t) =
+    {0}+e^{{-({3})({4})t}}\left[{1}\sin{{\sqrt{{1-{4}^2}}{3}t}}+{2}\cos{{\sqrt{{1-{4}^2}}{3}t}}\right]'.format(p[0],
+    p[1], p[2], p[3], p[4])
+    print equation
+    text(0.5, 0.5, equation)
+    text(0.5, 0.45, '$T =$' + str(T))
+    text(0.5, 0.4, '$r^2 =$' + str(rsq))
+    return fig
+
 def space_out_camel_case(s):
         """Adds spaces to a camel case string.  Failure to space out string
         returns the original string.
@@ -17,7 +57,7 @@ def bode(ABCD=None, numden=None, w=None, fig=None, n=None, label=None,
 
     Returns magnitude and phase vectors, and figure object.
     """
-    import numpy as np
+    i mport numpy as np
     import scipy as sp
     import matplotlib.pyplot as plt
     if fig == None:

### 2010-06-02 benchmark_bike_tools.py
fit plotting is now better

@@ -20,22 +20,21 @@ def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
     fig : the figure
 
     '''
-    from matplotlib.pyplot import figure, plot, xlabel, ylabel, text
+    from matplotlib.pyplot import figure, plot, xlabel, ylabel, title, legend, axes
     if fig:
         fig = fig
     else:
-        fig = figure()
-    plot(t, ym, '.', markersize=14)
+        fig = figure(2)
+    ax1 = axes([0.1, 0.1, 0.8, 0.7])
+    ax1.plot(t, ym, '.')
     plot(t, yf, 'k-')
     xlabel('Time [s]')
     ylabel('Amplitude [V]')
-    equation = 'f(t) =
-    {0}+e^{{-({3})({4})t}}\left[{1}\sin{{\sqrt{{1-{4}^2}}{3}t}}+{2}\cos{{\sqrt{{1-{4}^2}}{3}t}}\right]'.format(p[0],
-    p[1], p[2], p[3], p[4])
-    print equation
-    text(0.5, 0.5, equation)
-    text(0.5, 0.45, '$T =$' + str(T))
-    text(0.5, 0.4, '$r^2 =$' + str(rsq))
+    equation = r'$f(t)={0:1.2f}+e^{{-({3:1.3f})({4:1.1f})t}}\left[{1:1.2f}\sin{{\sqrt{{1-{3:1.3f}^2}}{4:1.1f}t}}+{2:1.2f}\cos{{\sqrt{{1-{3:1.3f}^2}}{4:1.1f}t}}\right]$'.format(p[0], p[1], p[2], p[3], p[4])
+    rsquare = '$r^2={0:1.3f}$'.format(rsq)
+    period = '$T={0} s$'.format(T)
+    title(equation + '\n' + rsquare + ', ' + period)
+    legend(['Measured', 'Fit'])
     return fig
 
 def space_out_camel_case(s):
@@ -57,7 +56,7 @@ def bode(ABCD=None, numden=None, w=None, fig=None, n=None, label=None,
 
     Returns magnitude and phase vectors, and figure object.
     """
-    i mport numpy as np
+    import numpy as np
     import scipy as sp
     import matplotlib.pyplot as plt
     if fig == None:

### 2010-06-02 fit_data.py

@@ -6,23 +6,24 @@ import os
 import pickle as p
 from scipy.optimize import approx_fprime
 from uncertainties import ufloat
-from benchmark_bike_tools import fit_goodness, jac_fitfunc
+from benchmark_bike_tools import fit_goodness, jac_fitfunc, plot_osfit
 
 dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
 file = open('data/period.txt', 'w')
 filenames.sort()
 period = {}
-#for name in ['YellowRevForkTorsionalFirst1.mat']:
+for name in ['YellowRevForkTorsionalFirst1.mat']:
 #for name in ['YellowFwheelCompoundFirst1.mat']:
 #for name in ['StratosFrameCompoundFirst2.mat']:
-for name in filenames:
+#for name in filenames:
     pendDat = {}
     mio.loadmat('data/pendDat/' + name, mdict=pendDat)
     y = pendDat['data'].ravel()
     time = float(pendDat['duration'][0])
     x = np.linspace(0, time, num=len(y))
+    plt.figure(1)
     # plot the original data
-    plt.plot(x, y, '.', markersize=14)
+    plt.plot(x, y, '.')
     # decaying oscillating exponential function
     fitfunc = lambda p, t: p[0] + np.exp(-p[3]*p[4]*t)*(p[1]*np.sin(p[4]*np.sqrt(1-p[3]**2)*t) + p[2]*np.cos(p[4]*np.sqrt(1-p[3]**2)*t))
     # initial guesses
@@ -60,6 +61,10 @@ for name in filenames:
     wd = (1. - zeta**2.)**(1./2.)*wo
     f = wd/2./np.pi
     T = 1./f
+    fig = plt.figure(2)
+    plot_osfit(x, y, lscurve, p1, rsq, T, fig=fig)
+    plt.savefig('data/pendDat/graphs/atest.png')
+    plt.close()
     # include the notes for the experiment
     note = pendDat['notes']
     line = name + ',' + str(T) + ',' + str(rsq) + ',' + str(sigma) + ',' + str(note) + '\n'

### 2010-06-02 report/parameterReport.tex
osciallation plotting function
typed some more

@@
 three intersection points.
+\begin{figure}[htpb]
+    \begin{center}
+        \includegraphics{figures/triangle.pdf}
+    \end{center}
+    \caption{Exaggerated intersection of the three pendulum axes and the
+    location of the center of mass.}
+    \label{fig:triangle}
+\end{figure}
 \begin{equation}
@@
 \end{figure}
+The following function Eqn~\ref{eqn:decayOs} was fit to the data using a nonlinear least squares fit
+routine for each experiment.
+\begin{equation}
+    f(t) = A + e^{-\zeta\omega t}\left[B\sin{\sqrt{1-\zeta^2}\omega t} +
+    C\cos{\sqrt{1-\zeta^2}\omega t}\right]
+    \label{eqn:decayOs}
+\end{equation}
+Most of the data fit the damped oscillation function well with very light (and
+ignorable) damping. There were several instances of beating-like phenomena for
+some of the parts at particular orienations. Roland and
+Massing~\cite{Roland1971} also encountered this problem and used a bearing to
+prevent the torsional pendulum from swinging. Figure~\ref{fig:beating} shows an
+example of the beating like phenomena.
+\begin{figure}[htpb]
+    \begin{center}
+        \includegraphics[width=5in]{../data/pendDat/graphs/CrescendoForkTorsionalFirst2.png}
+    \end{center}
+    \caption{An example of the beating-like phenomena observed on 5\% of the
+    experiments.}
+    \label{fig:beating}
+\end{figure}
+The physical phenonena that was observed for data sets such as these were that
+the bicycle frame or fork was perturbed torsionally. After set into motion the
+torionsal motion died out and a longitudinal swinging motion increased. The
+motions alternated back and forth with neither ever reaching zero. The
+frequencies of these motions were very close to one another and it is not
+apparent how disect the two. We explored fitting to a function such as
+\begin{equation}
+    f(t) = A\sin{(\omega_1 t)} + B\sin{(\omega_2 t + \phi)} + C
+    \label{eqn:sumSines}
+\end{equation}
+But the fit predicts that $\omega_1$ and $\omega_2$ have very similar
+frequencies. There was no easy way to choose which of the two $\omega$'s was
+the one associated with the torionsal osciallation. Some work was done to model
+the torsional pedulum as a laterally flexible beam to determine this, but we
+thought the improvement in the accuracy of the period calculation would not
+improve enough for the effort required. Future experiements should simply
+prevent the swinging motion of the pendulum without damping the torsional
+motion. Roland and Massing~\cite{Roland1971} noted that they had this same problem and solved it
+by hold the torsional rod in a fixed bearing.
+
+The period for a damped osciallation is
+\begin{equation}
+    T = \frac{2\pi}{\sqrt{1-\zeta^2}\omega_n}
+    \label{eqn:periodDamped}
+\end{equation}
+and for light damping
+\begin{equation}
+    T \approx \frac{2\pi}{\omega_n}
+    \label{eqn:period}
+\end{equation}
+The uncertaintity in the period, $T$, can be determined from the fit. Firstly,
+the variance of the fit is
+\begin{equation}
+    \sigma_y^2 =
+    \frac{1}{N-5}\sum_{i=1}^N(y_{mi}-\bar{y}_m)^2-(y_{pi}-\bar{y}_m)^2
+    \label{eqn:fitVariance}
+\end{equation}
+The covariance matrix of the fit function can be formed
+\begin{equation}
+    \mathbf{U} = \sigma_y^2\mathbf{H}^{-1}
+    \label{eqn:covariance}
+\end{equation}
+where $\mathbf{H}$ is the Hessian. $\mathbf{U}$ is a $5\times5$ matrix with the
+variances of each of the five fit parameters along the diagonal. This assumes
+that the parameters are uncorrelated [note for Jason: you may have to take into
+account correlation for this..maybe checkout the R function for nls]. The
+variance of $T$ can be computed using the variance of $\zeta$ and $\omega$. It
+is important to note that the uncertainties in the period are very low
+($<10e-4$), even for the fits with low $r^2$ values due to the beating issue.
 \subsection{Torsional Pendulum}

### 2010-06-02 benchmark_bike_tools.py

@@ -20,13 +20,13 @@ def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
     fig : the figure
 
     '''
-    from matplotlib.pyplot import figure, plot, xlabel, ylabel, title, legend, axes
+    from matplotlib.pyplot import figure, plot, xlabel, ylabel, title, legend, axes, xlim
     if fig:
         fig = fig
     else:
         fig = figure(2)
     ax1 = axes([0.1, 0.1, 0.8, 0.7])
-    ax1.plot(t, ym, '.')
+    ax1.plot(t, ym, '.', markersize=2)
     plot(t, yf, 'k-')
     xlabel('Time [s]')
     ylabel('Amplitude [V]')
@@ -35,6 +35,7 @@ def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
     period = '$T={0} s$'.format(T)
     title(equation + '\n' + rsquare + ', ' + period)
     legend(['Measured', 'Fit'])
+    #xlim((0, 1))
     return fig
 
 def space_out_camel_case(s):

### 2010-06-02 fit_data.py

@@ -12,18 +12,15 @@ dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
 file = open('data/period.txt', 'w')
 filenames.sort()
 period = {}
-for name in ['YellowRevForkTorsionalFirst1.mat']:
+#for name in ['YellowRevForkTorsionalFirst1.mat']:
 #for name in ['YellowFwheelCompoundFirst1.mat']:
 #for name in ['StratosFrameCompoundFirst2.mat']:
-#for name in filenames:
+for name in filenames:
     pendDat = {}
     mio.loadmat('data/pendDat/' + name, mdict=pendDat)
     y = pendDat['data'].ravel()
     time = float(pendDat['duration'][0])
     x = np.linspace(0, time, num=len(y))
-    plt.figure(1)
-    # plot the original data
-    plt.plot(x, y, '.')
     # decaying oscillating exponential function
     fitfunc = lambda p, t: p[0] + np.exp(-p[3]*p[4]*t)*(p[1]*np.sin(p[4]*np.sqrt(1-p[3]**2)*t) + p[2]*np.cos(p[4]*np.sqrt(1-p[3]**2)*t))
     # initial guesses
@@ -34,12 +31,6 @@ for name in ['YellowRevForkTorsionalFirst1.mat']:
     p1, success = op.leastsq(errfunc, p0[:], args=(x, y))
     # plot the fitted curve
     lscurve = fitfunc(p1, x)
-    plt.plot(x, lscurve, 'k-')
-    plt.xlabel('Time [s]')
-    plt.ylabel('Amplitude [V]')
-    #plt.show()
-    plt.savefig('data/pendDat/graphs/' + name[:-4] + '.png')
-    plt.close()
     rsq, SSE, SST, SSR = fit_goodness(y, lscurve)
     sigma = np.sqrt(SSE/(len(y)-len(p0))) # DoF for non-lin fit may be different
     # calculate the jacobian
@@ -50,21 +41,21 @@ for name in ['YellowRevForkTorsionalFirst1.mat']:
     U = sigma**2.*np.linalg.inv(H)
     # the standard deviations
     sigp = np.sqrt(U.diagonal())
-    # add a star in the R value is low
-    if rsq <= 0.99:
-        rsq = str(rsq) + '*'
-    else:
-        pass
     # frequency and period
     wo = ufloat((p1[4], sigp[4]))
     zeta = ufloat((p1[3], sigp[3]))
[…]

### 2010-06-02 benchmark_bike_tools.py
uround function ready for testing

@@ -1,3 +1,47 @@
+def uround(value):
+    '''Round values according to their uncertainity
+
+    Parameters:
+    -----------
+    value: float with uncertainty
+
+    Returns:
+    --------
+    s: string that is properly rounded
+
+    2.4563752289999+/-0.0003797273827
+
+    becomes
+
+    2.4564+/-0.0004
+
+    This doesn't work for large uncertainties.
+    '''
+    from uncertainties import ufloat
+    try:
+        # grab the nominal value and the uncertainty
+        nom = value.nominal_value
+        uncert = value.std_dev()
+        # convert the uncertainty to a string
+        s = str(uncert)
+        # find the first non-zero character
+        for j, number in enumerate(s):
+            if number == '0' or number == '.':
+                pass
+            else:
+                digit = j
+                break
+        newUncert = round(uncert, digit-1)
+        newNom = round(nom, len(str(newUncert)) - 2)
+        newValue = ufloat((newNom, newUncert))
+        if len(str(newNom)) < len(str(newUncert)):
+            s = str(newNom) + '0+/-' +str(newUncert)
+        else:
+            s = str(newValue)
+    except:
+        s = str(value)
+    return s
+
 def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
     '''Plot fitted data over the measured
 

### 2010-06-02 benchmark_bike_tools.py
uround function works pretty good, doesn't handle all cases though

@@ -34,11 +34,16 @@ def uround(value):
         newUncert = round(uncert, digit-1)
         newNom = round(nom, len(str(newUncert)) - 2)
         newValue = ufloat((newNom, newUncert))
-        if len(str(newNom)) < len(str(newUncert)):
-            s = str(newNom) + '0+/-' +str(newUncert)
+        print newValue
+        diff = len(str(newUncert)) - len(str(newNom))
+        if diff > 0:
+            print 'diff is greater than zero'
+            s = str(newNom) + int(diff)*'0' + '+/-' +str(newUncert)
         else:
+            print 'not greater than zero'
             s = str(newValue)
     except:
+        print 'the exception'
         s = str(value)
     return s
 

### 2010-06-02 benchmark_bike_tools.py
changed make par tables to use uround

@@ -15,7 +15,7 @@ def uround(value):
 
     2.4564+/-0.0004
 
-    This doesn't work for large uncertainties.
+    This doesn't work for weird cases like large uncertainties.
     '''
     from uncertainties import ufloat
     try:
@@ -34,16 +34,12 @@ def uround(value):
         newUncert = round(uncert, digit-1)
         newNom = round(nom, len(str(newUncert)) - 2)
         newValue = ufloat((newNom, newUncert))
-        print newValue
         diff = len(str(newUncert)) - len(str(newNom))
         if diff > 0:
-            print 'diff is greater than zero'
             s = str(newNom) + int(diff)*'0' + '+/-' +str(newUncert)
         else:
-            print 'not greater than zero'
             s = str(newValue)
     except:
-        print 'the exception'
         s = str(value)
     return s
 

### 2010-06-02 make_par_tables.py

@@ -3,6 +3,8 @@ import re
 import os
 import uncertainties as un
 
+from benchmark_bike_tools import uround
+
 # load in the base data file
 f = open('data/data.p', 'r')
 data = pickle.load(f)
@@ -31,19 +33,21 @@ for i, name in enumerate(data['bikes']):
             # go through each match and make a substitution
             for match in test:
                 #print "replace this: ", match
+                # if the match has a question mark it is an uncertainty value
                 if match[-1] == '?':
                     try:
                         line = re.sub('\|(\w*\?)\|',
-                                str(par[match[:-1]][i].std_dev()), line, count=1)
+                                uround(par[match[:-1]][i]).split('+/-')[1], line, count=1)
                         #print line
                     except:
                         pass
+                # else if the match is the bicycle name
                 elif match == 'bikename':
                     line = re.sub('\|bikename\|', name.upper(), line)
                 else:
                     try:
                         line = re.sub('\|(\w*(?!\?))\|',
-                                str(par[match][i].nominal_value), line, count=1)
+                                uround(par[match][i]).split('+/-')[0], line, count=1)
                         #print line
                     except:
                         line = re.sub('\|(\w*(?!\?))\|', str(par[match][i]), line, count=1)

### 2010-06-02 calc_parameters.py
took rounding code out

@@ -238,30 +238,6 @@ par['IBxz'] = np.array(par['IBxz'])
 par['IBzz'] = np.array(par['IBzz'])
 par['v'] = np.ones_like(d['forkMass'])
 
-# round the numbers according to the calculated uncertainty
-par_test = {}
-for k, v in par.items():
-    par_test[k] = np.zeros_like(v)
-    for i, value in enumerate(v):
-        try:
-            uncert = value.std_dev()
-            nom = value.nominal_value
-            s = str(uncert)
-            for j, number in enumerate(s):
-                if number == '0' or number == '.':
-                    pass
-                else:
-                    digit = j
-                    break
-            newUncert = round(uncert, digit-1)
-            newNom = round(nom, len(str(newUncert)) - 2)
-            newValue = u.ufloat((newNom, newUncert))
-            par_test[k][i] = newValue
-        except:
-            pass
-del par
-par = par_test
-
 # make a dictionary with only the nominal values
 par_n = {}
 for k, v in par.items():

### 2010-06-02 report/parameterReport.tex
typing more on the report

@@
 \documentclass{article}
+
 \usepackage[lmargin=1in,rmargin=1in]{geometry}
@@
 \begin{document}
+
 \maketitle
@@
 \section{Introduction}
+This work is intended to document the indirect measurement of eight real
+bicycles' physical parameters. The physical parameters measured are those that
+are needed for the benchmark bicycle model~\cite{Meijaard2007}. The work is
+based off of the techiques used to measure the instrumented bicycle in
+\cite{Kooijman2006} and \cite{Kooijman2008}. I used the same techniques as
+Kooijman to measure a second instrumented bicycle~\cite{Moore2009a}. This
+method had low accuracy and the $I_{yy}$ moments of inertia were
+not measured for the fork and frame. Furthermore, very little data exists on
+the physical parameters of different types of bicycles and this work aims to
+provide a small sample of bicycles.
+
 D\"{o}hring measured the physical parameters of a scooter~\cite{Dohring1953}.

### 2010-06-03 report/parameterReport.tex
more typing

@@
 
+The torsional pendulum was calibrated using a known moment of inertia.
+
 \subsection{Wheels}
@@
 \subsection{Frame}
+Three measurements were made to estimate the globally referenced moments and
+products of inertia ($I_{xx}$, $I_{xz}$ and $I_{zz}$) of the rear frame. The
+frame was typically hung from the three main tubes: seat tube, down tube
+and top tube. The rear fender prevented easy connection to the seat tube on
+some of the bikes and the clamp was attached to the fender instead which was
+less rigid. For best accuracy with only three orientation angles, the frame
+should be hung at three angles that are $120^\circ$ apart. The three tubes on
+the frame generally provide that the orientation angles were spread evenly.
+Taking data at more orientation angles could improve the accuracy further and
+is gernally possible with standard diamond frame bicylces.
+ The same torsional pendulum used in
+\cite{Kooijman2006} was used to measure the averaged period $\overline{T}_i$ of
+oscillation of the rear frame and handlebar/fork assembly at three different
+orientation angles $\beta_i$, where $i=1$, $2$, $3$, as shown in
+Fig.~\ref{fig:angles}. The parts were perturbed lightly, less than 1 degree,
+and allowed to oscillate about the pendulum axis through at least ten periods.
+The time of oscillation was recorded via a stopwatch ($\pm1$~s). This was done
+three times for each frame and the recorded times were averaged. The
+coefficient of elasticity $k$ for the torsional pendulum had previously been
+measured in \cite{Kooijman2006} and found to be $k=5.01\pm0.01$
+$\frac{\textrm{Nm}}{\textrm{rad}}$. Three moments of inertia $J_{i}$ about the
+pendulum axes were calculated with
+\begin{equation}
+	J_i=\frac{k\overline{T}_i^2}{4\pi^2}
+\label{eq:torPend}
+\end{equation}
+
+The moments and products of inertia of the rear frame and handlebar/fork assembly with reference to the benchmark coordinate system were calculated by formulating the relationship between inertial frames
+\begin{equation}
+	\mathbf{J}_i=\mathbf{R}_i^T\mathbf{IR}_i
+\label{eq:rotIn}
+\end{equation}
+where $\mathbf{J}_i$ is the inertia tensor about the pendulum axes, $\mathbf{I}$, is the inertia tensor in the global reference frame and $\mathbf{R}$ is the rotation matrix relating the two frames.
+The global inertia tensor is defined as
+\begin{equation}
+	\mathbf{I}=
+	\left[
+	\begin{array}{rr}
+		I_{xx}  & -I_{xz}\\
+		-I_{xz} & I_{zz}
+	\end{array}
+	\right]\textrm{.}
+	\label{eq:MoI}
+\end{equation}
+The inertia tensor can be reduced to a $2\times2$ matrix because the $I_{yy}$ component is not needed in the linear formulation of the benchmark bicycle\footnote{The pitch of the rear frame and handlebar/fork assembly are quadratic functions of the lean and steer\cite{PetersenHubbard2007}, so the pitch becomes zero in the linear model.} and the bicycle is assumed to be symmetric about the $XZ$-plane.
+The simple rotation matrix about the $Y$-axis can similarly be reduced to a $2\times2$ matrix where $s_{\beta i}$ and $c_{\beta i}$ are defined as $\sin{\beta_i}$ and $\cos{\beta_i}$, respectively.
+\begin{equation}
+	\mathbf{R}=
+	\left[
+	\begin{array}{rr}
+		c_{\beta i} & s_{\beta i}\\
+		-s_{\beta i} & c_{\beta i}
+	\end{array}
+	\right]
+	\label{eq:rotMat}
+\end{equation}
+The first entry of $\mathbf{J}_i$ in Eq.~\ref{eq:rotIn} is the moment of inertia about the pendulum axis and is written explicitly as 
+\begin{equation}
+	J_{i}=c^{2}_{\beta i}I_{xx}+2s_{\beta i}c_{\beta i}I_{xz}+s^{2}_{\beta i}I_{zz}\textrm{.}
+\label{eq:inRelComp}
+\end{equation}
+Calculating all three $J_{i}$ allows one to form
+\begin{equation}
+	\left[
+	\begin{array}{c}
+		J_{1}\\
+		J_{2}\\
+		J_{3}
+	\end{array}
+	\right]
+	=
+	\left[
+	\begin{array}{ccc}
+		c_{\beta 1}^2 & -2s_{\beta 1}c_{\beta 1} & s_{\beta 1}^2\\
+		c_{\beta 2}^2 & -2s_{\beta 2}c_{\beta 2} & s_{\beta 2}^2\\
+		c_{\beta 3}^2 & -2s_{\beta 3}c_{\beta 3} & s_{\beta 3}^2
+	\end{array}
+	\right]
+	\left[
+	\begin{array}{c}
+		I_{xx}\\
+		I_{xz}\\
+		I_{zz}
+	\end{array}
+	\right]
+\label{eq:inRel}
+\end{equation}
+and the unknown global inertia tensor can be solved for.
+
+Finding the inertia tensors of the wheels is less complex because the wheels are symmetric about three orthogonal planes so there are no products of inertia.
+The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring the averaged period of oscillation about an axis in the $XZ$-plane using the torsional pendulum setup and Eq.~\ref{eq:torPend}.
+The $I_{yy}$ moment of inertia was calculated with a compound pendulum as described in \cite{Kooijman2006} and shown in Fig.~\ref{fig:comPend} using
+\begin{equation}
+	I_{yy}=\left(\frac{\overline{T}}{2\pi}\right)^2mgl-ml^2
+\label{eq:comPend}
+\end{equation}
+where $l=0.303\pm0.002$ m is the pendulum length, $m$ is the mass of the wheel, $\overline{T}$ is the averaged period and $g$ is the local acceleration due to gravity.
+Table \ref{tab:wheeliInertiaComp} gives the calculated values.
+
 \subsection{Fork}

### 2010-06-04 report/parameterReport.tex
starting to work on calculating the uncertainties in the canonical matrices and the eigenvalues

@@
 
-\usepackage[lmargin=1in,rmargin=1in]{geometry}
+\usepackage[lmargin=1in,rmargin=1in, tmargin=1in, bmargin=1in]{geometry}
 \usepackage[colorlinks=true, linkcolor=blue]{hyperref}
 \usepackage{graphicx}
+\usepackage{eurosans}
 
@@
 same was as presented, including calculations of uncertainty from the indirect
-measurement techniques~\cite{Roland1971}. This work is based off of the work
+measurement techniques~\cite{Roland1971}. Patterson used a swing to measure the
+inertia of a bicycle and rider~\cite{Patterson2004}. This work is based off of the work
 done by Kooijman~\cite{Kooijman2006}. I worked with Jodi Kooijman for a year and
@@
 technique.
+\section{Bicycle Descriptions}
+\subsection{Yellow Bicycle}
+The yellow is a bicycle used in the lab to demonstrate that a bicycle is stable
+at certain speeds. It is an aluminum road frame of unknown make with the
+majority of components removed. The wheels, drop handlebar, seat, seat post and
+bottom bracket are the only parts on the bike. The fork has been reversed to
+''increase`` the stability of the bicycle. The bicycle was measured with both
+the fork in normal position and the fork reversed.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/yellow.jpg}
+    \end{center}
+    \caption{Side view of the yellow bicycle with the fork in the reversed
+    posistion.}
+    \label{fig:YellowBike}
+\end{figure}
+\begin{table}[htbp]
+	\begin{tabular}{ll}
+		Style & Road\\
+		Size & \\
+		Wheels & 700c Aluminum Rims\\
+		Tires & 28 x 1 3/8 x 1 5/8\\
+		Sprocket & 6 speed free wheel\\
+		Handlebars & Drop\\
+	\end{tabular}
+	\caption{Specifications of the Yellow Bicycle}
+	\label{yellowBikeSpecs}
+\end{table}
+\subsection{Batavus Browser}
+The Batavus Browser is lower priced Dutch city bike. We measured the physical
+proprieties of the stock Browser model and also the same bike with the
+instrumentation used in our experiments.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/browser.jpg}
+    \end{center}
+    \caption{Side view of the Batavus Browser as from the factory.}
+    \label{fig:browser}
+\end{figure}
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/browserIns.jpg}
+    \end{center}
+    \caption{The Batavus Browser with instrumentation as used
+    in~\cite{Kooijman2009}}
+    \label{fig:browserIns}
+\end{figure}
+\begin{table}[htbp]
+	\begin{tabular}{ll}
+        Price          & euro symbol 449\\
+        Wheel Size     & 28''\\
+        Weight         & 18.7 kg\\
+        Frame Size     & 54\\
+        Frame material & High Tensile Steel\\
+        Fork material  & Steel\\
+        Rims           & Aluminum Airline\\
+        Spokes         & Stainless steel\\
+        Tires          & Trax Highway, 28'' x 1 5/8 x 1 3/8\\
+        Handlebar      & Steel painted\\
+        Handles        & Batavus Comfort\\
+        Saddle         & Selle Royal 8274\\
+        Dynamo         & Basta Trio\\
+        Headlight      & Batavus Logic Light\\
+        Rearlight      & Battery light\\
+        Pedals         & Batavus plastic antislip\\
+        Final          & Trelock RS42\\
+        Theft chip     & Present\\
+        Crank          & Aluminum\\
+        Rack           & Steel\\
+        Quick Binder   & Binder Trio\\
+        Mudguards      & Steel\\
+        Bell           & Kenlight\\
+	\end{tabular}
+	\caption{2008 Batavus Browser Specifications.}
+	\label{browserSpecs}
+\end{table}
+
+\subsection{2007 Bianchi Pista}
+The Pista is a modern steel track bicycle. Specifications, taken from the Bianchi
+website~\cite{Bianchi2009}, are as follows:
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/pista.jpg}
+    \end{center}
+    \caption{The Bianchi Pista hanging from a spring scale.}
+    \label{fig:pista}
+\end{figure}
+\begin{table}
+	\begin{tabular}{ll}
+        Style            & Track Bike\\
+        Size             & 57cm\\
+        Color            & Gang Green\\
+        Frame            & Bianchi DB CrMo, rear entry track dropouts\\
+        Fork             & CrMo\\
+        Retail Price     & \$579.99\\
+        Headset          & VP AheadSet, 1''threadless\\
+        Handlebar        & Bianchi steel track, 26.0mm\\
+        Stem             & Bianchi alloy\\
+        Brakes/Levers    & Shimano Tiagra front brake/lever\\
+        Crankset         & Truvativ Touro 1.1, 48T\\
+        Bottom Bracket   & Truvativ Power Spline cartridge\\
+        Chain            & KMC\\
+        Sprocket         & 16T fixed cog \& 16T Freewheel\\
+        Pedals           & Campus pedals\\
+        Wheels           & Bianchi Alex Solo wheelset, 28h rims\\
+        Tires            & Continental UltraSport, 700x23C\\
+        Saddle           & Bianchi Velo\\
+        Seatpost         & Bianchi alloy, 27.2mm\\
+        Size             & 57cm\\ 
+        Seat Tube        & 570mm\\
+        Top Tube-actual  & 560mm\\
+        Top Tube-virtual & 560mm\\
+        Chainstay        & 383mm\\
+        Fork Rake        & 28mm\\
+        Head Tube Angle  & 74.5$^{\circ}$\\
+        Seat Tube Angle  & 75.5$^{\circ}$\\
+        Wheelbase        & 967mm\\
+        Standover Height & 32''  
+	\end{tabular}
+	\caption{}
+	\label{pistaSpecs}
+\end{table}
+\subsection{Gary Fisher Mountain Bike}
+The Gary Fisher Ziggurat is a modern front suspended mountain bike.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/fisher.jpg}
+    \end{center}
+    \caption{The Gary Fisher mountain bike hanging from a spring scale.}
+    \label{fig:fisher}
+\end{figure}
+\begin{table}
+    \centering
+    \begin{tabular}{ll}
+        Geometry & Genisis\\
+        Frame & Aluminum, Damping, AC/DC, Carbon, Composite\\
+        Front Wheel & DT Swiss XR4.1 Ceramic 559 17\\
+        Tires & Michelin 26'' X 2.00 (5.2-559)\\
+        Rear Wheel & DT Swiss 240s\\
+        Rear hub & 9 speed\\
+        Rear derailleur & SRAM X-0\\
+        Crank & Bonterager 3-speed\\
+        Seat post & Ritchey\\
+        Fork & SID Rock Shox
+    \end{tabular}
+    \caption{Specifications of the Gary Fisher Mountain Bike}
+    \label{tab:gary}
+\end{table}
+\subsection{Batavus Crescendo Deluxe}
+The Batavus Crescendo Deluxe is a contemporary of all-round cycling in cool
+black and white. A great bike for a good end to touring. And that works fine
+with 8 gears, suspension fork and suspension seatpost.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/crescendo.jpg}
+    \end{center}
+    \caption{A side view of the Batavus Crescendo Deluxe.}
+    \label{fig:crescendo}
+\end{figure}
+\begin{table}
+	\begin{tabular}{ll}
+		Price             & from € 869\\
+		Wheel Size        & 28''\\
+		Weight            & 21.5 kg\\
+		Frame Size        & 53\\
+		Performance/speed & Shimano Nexus 8 speed with roller brakes\\
+		Color             & White / black\\
+		Brakes            & Shimano roller brakes (BR-IM50)\\
+		Frame type        & Ladies mono tube frame Gentlemen trend\\
+		Frame material    & Aluminum 7005\\
+		Front Fork        & spring, Batavus\\
+		Rims              & Aluminum Rodi VR19\\
+		Spokes            & Stainless steel (adapted spoke pattern)\\
+		Tires             & CST End\\
+		Handlebar         & Steel aluminum look\\
+		Stem              & Manually adjustable, Batavus Ergo Matic III\\
+		Handles           & adjustable Batavus Ergo Grip II\\
+		Saddle            & Selle Royal 5160\\
+		Seatpost          & spring, Post Modern Compact\\
+		Dynamo            & Shimano Hub Dynamo\\
+		Headlight         & Trelock Hi-Power automatic halogen\\
+		Rear              & Smart battery automatically\\
+		Pedals            & aluminum antislip\\
+		Final							& AXA Defender\\
+		Chip theft        & Present\\
+		Crank             & Aluminum\\
+		Rack              & Aluminum\\
+		Quick Binder      & Binder Quattro\\
+		Mudguards         & Batavus hyalite II\\
+		Bell              & Bibia Touring\\
+		Balhoofdset       & VP semi-integrated, oversized\\
+		Shifters          & Batavus (8-speed)\\
+		Saddlebag         & Batavus saddlebag with plakset\\
+	\end{tabular}
+	\caption{}
+	\label{crescendoSpecs}
+\end{table}
+\subsection{Batavus Stratos Deluxe}
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/stratos.jpg}
+    \end{center}
+    \caption{A side view of the Batavus Stratos Deluxe}
+    \label{fig:stratos}
+\end{figure}
+\begin{table}
+	\begin{tabular}{ll}
+		Switching           & Spin System (shifter)\\
+		Monochromatic frame & Black\\
+		Chain guard         & Semi-open\\
+		Mudguards           & Aluminum\\
+		Rack                & Yes\\
+		Extra drive         & Regardless Motori degree\\
+		Type                & Men's\\
+		Frame Material      & Aluminum\\
+		Rim Material        & Aluminum\\
+		Rear braking system & Roller brake\\
+		Front braking system & Roller Brake\\
+		Number of gears      & 7 Gears\\
+		Gear System          & Shimano Nexus\\
+		Wheel Size           & 28 inch\\
+		Tirewidth            & 37 mm\\
+		Price                & 749,00\\
+	\end{tabular}
+	\caption{}
+	\label{stratosSpecs}
+\end{table}
 \section{Accuracy}
@@
 Molen 20kg scale with a resolution of 20g.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/massScale.jpg}
+    \end{center}
+    \caption{The scale used to measure the mass of each bicycle component.}
+    \label{fig:massScale}
+\end{figure}
 \begin{table}
@@
 global reference frame.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/pendDist.jpg}
+    \end{center}
+    \caption{Measuring the distance from the pendulum axis to the rear wheel
+    axle using level ruler.}
+    \label{fig:PendDist}
+\end{figure}
 \begin{figure}[tb]
@@
 \begin{equation}
-	b_i=\frac{d_i}{\cos{\beta_i}}-r_\mathrm{R}
+    b_i=-\left(\frac{a_\mathrm{B}}{\cos{\beta_i}}+r_\mathrm{R}\right)
     \label{eq:zInt}
@@
 \subsection{Fork}
-The center of mass of the fork is calculated in the same fashion, only the
-z-intercept is different.
+The fork and handlebars are a bit trickier to hang in three different
+orientations. Typically two angles can be obtained by clamping to the steer
+tube at the top and the bottom. The third angle can be obtained by clamping to
+the stem. The center of mass of the fork is calculated in the same fashion. The
+slope of the line in the benchmark reference frame is the same as for the
+frame but the z-intercept is different:
 \begin{equation}
-    b = r_\mathrm{F} + \frac{a}{\cos{\beta}} + w\tan{\beta}
+    b = w\tan{\beta} - r_\mathrm{F} - \frac{a}{\cos{\beta}} 
 \end{equation}
@@
 ignorable) damping. There were several instances of beating-like phenomena for
-some of the parts at particular orienations. Roland and
+some of the parts at particular orientations. Roland and
 Massing~\cite{Roland1971} also encountered this problem and used a bearing to
@@
 \end{figure}
-The physical phenonena that was observed for data sets such as these were that
+The physical phenomena that was observed for data sets such as these were that
 the bicycle frame or fork was perturbed torsionally. After set into motion the
-torionsal motion died out and a longitudinal swinging motion increased. The
+torsional motion died out and a longitudinal swinging motion increased. The
 motions alternated back and forth with neither ever reaching zero. The
 frequencies of these motions were very close to one another and it is not
-apparent how disect the two. We explored fitting to a function such as
+apparent how dissect the two. We explored fitting to a function such as
 \begin{equation}
@@
 frequencies. There was no easy way to choose which of the two $\omega$'s was
-the one associated with the torionsal osciallation. Some work was done to model
-the torsional pedulum as a laterally flexible beam to determine this, but we
+the one associated with the torsional oscillation. Some work was done to model
+the torsional pendulum as a laterally flexible beam to determine this, but we
 thought the improvement in the accuracy of the period calculation would not
-improve enough for the effort required. Future experiements should simply
+improve enough for the effort required. Future experiments should simply
 prevent the swinging motion of the pendulum without damping the torsional
@@
 
-The period for a damped osciallation is
+The period for a damped oscillation is
 \begin{equation}
@@
 \end{equation}
-The uncertaintity in the period, $T$, can be determined from the fit. Firstly,
+The uncertainty in the period, $T$, can be determined from the fit. Firstly,
 the variance of the fit is
@@
 various lower clamps.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/fixture.jpg}
+    \end{center}
+    \caption{The rigid pendulumfixture mounted to a concrete column.}
+    \label{fig:fixture}
+\end{figure}
 A 5mm diameter, 1m long mild steel rod was used as the torsion
@@
 The torsional pendulum was calibrated using a known moment of inertia. 
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/rod.jpg}
+    \end{center}
+    \caption{The steel calibration rod.}
+    \label{fig:rod}
+\end{figure}
+A torsional pendulum almost identical to the one used in
+\cite{Kooijman2006} was used to measure the averaged period $\overline{T}_i$ of
+oscillation of the rear frame at three different
+orientation angles $\beta_i$, where $i=1$, $2$, $3$, as shown in
+Fig.~\ref{fig:triangle}. The parts were perturbed lightly, less than 1 degree,
+and allowed to oscillate about the pendulum axis through at least ten periods.
+The time of oscillation was recorded via a stopwatch ($\pm1$~s). This was done
+three times for each frame and the recorded times were averaged. The
+coefficient of elasticity $k$ for the torsional pendulum had previously been
+measured in \cite{Kooijman2006} and found to be $k=5.01\pm0.01$
+$\frac{\textrm{Nm}}{\textrm{rad}}$.
 
@@
 \end{equation}
-
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/CrescendoFwheelTorsionalFirst.jpg}
+    \end{center}
+    \caption{The front wheel of the Crescendo hung as a torsional pendulum.}
+    \label{fig:FwheelTor}
+\end{figure}
 The radial moment of inertia was measured by hanging the wheel as a torsional
@@
 \end{equation}
+Finding the inertia tensors of the wheels is less complex because the wheels
+are symmetric about three orthogonal planes so there are no products of
+inertia. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring
+the averaged period of oscillation about an axis in the $XZ$-plane using the
+torsional pendulum setup and Eq.~\ref{eq:torPend}. The $I_{yy}$ moment of
+inertia was calculated with a compound pendulum as described in
+\cite{Kooijman2006} and shown in Fig. using
+\begin{equation}
+	I_{yy}=\left(\frac{\overline{T}}{2\pi}\right)^2mgl-ml^2
+\label{eq:comPend}
+\end{equation}
+where $l=0.303\pm0.002$ m is the pendulum length, $m$ is the mass of the wheel,
+$\overline{T}$ is the averaged period and $g$ is the local acceleration due to
+gravity. Table  gives the calculated values.
+\begin{figure}[htpb]
+    \begin{center}
+        \includegraphics[width=5in]{images/wheelIyy.jpg}
+    \end{center}
+    \caption{A wheel hung as a compoun pendulum.}
+    \label{fig:wheelIyy}
+\end{figure}
 \subsection{Frame}
@@
 and top tube. The rear fender prevented easy connection to the seat tube on
-some of the bikes and the clamp was attached to the fender instead which was
-less rigid. For best accuracy with only three orientation angles, the frame
+some of the bikes and the clamp was attached to the fender. The fender was
+generally less rigid than the frame tube. For best accuracy with only three orientation angles, the frame
 should be hung at three angles that are $120^\circ$ apart. The three tubes on
-the frame generally provide that the orientation angles were spread evenly.
-Taking data at more orientation angles could improve the accuracy further and
+the frame generally provide that the orientation angles were spread evenly at
+about $120^\circ$. Taking data at more orientation angles could improve the accuracy further and
 is gernally possible with standard diamond frame bicylces.
- The same torsional pendulum used in
-\cite{Kooijman2006} was used to measure the averaged period $\overline{T}_i$ of
-oscillation of the rear frame and handlebar/fork assembly at three different
-orientation angles $\beta_i$, where $i=1$, $2$, $3$, as shown in
-Fig.~\ref{fig:angles}. The parts were perturbed lightly, less than 1 degree,
-and allowed to oscillate about the pendulum axis through at least ten periods.
-The time of oscillation was recorded via a stopwatch ($\pm1$~s). This was done
-three times for each frame and the recorded times were averaged. The
-coefficient of elasticity $k$ for the torsional pendulum had previously been
-measured in \cite{Kooijman2006} and found to be $k=5.01\pm0.01$
-$\frac{\textrm{Nm}}{\textrm{rad}}$. Three moments of inertia $J_{i}$ about the
-pendulum axes were calculated with
+
+Three moments of inertia $J_{i}$ about the pendulum axes were calculated with
 \begin{equation}
@@
 
-The moments and products of inertia of the rear frame and handlebar/fork assembly with reference to the benchmark coordinate system were calculated by formulating the relationship between inertial frames
+The moments and products of inertia of the rear frame and handlebar/fork
+assembly with reference to the benchmark coordinate system were calculated by
+formulating the relationship between inertial frames
 \begin{equation}
@@
 \end{equation}
-where $\mathbf{J}_i$ is the inertia tensor about the pendulum axes, $\mathbf{I}$, is the inertia tensor in the global reference frame and $\mathbf{R}$ is the rotation matrix relating the two frames.
-The global inertia tensor is defined as
+where $\mathbf{J}_i$ is the inertia tensor about the pendulum axes,
+$\mathbf{I}$, is the inertia tensor in the global reference frame and
+$\mathbf{R}$ is the rotation matrix relating the two frames. The global inertia
+tensor is defined as
 \begin{equation}
@@
 \end{equation}
-The inertia tensor can be reduced to a $2\times2$ matrix because the $I_{yy}$ component is not needed in the linear formulation of the benchmark bicycle\footnote{The pitch of the rear frame and handlebar/fork assembly are quadratic functions of the lean and steer\cite{PetersenHubbard2007}, so the pitch becomes zero in the linear model.} and the bicycle is assumed to be symmetric about the $XZ$-plane.
-The simple rotation matrix about the $Y$-axis can similarly be reduced to a $2\times2$ matrix where $s_{\beta i}$ and $c_{\beta i}$ are defined as $\sin{\beta_i}$ and $\cos{\beta_i}$, respectively.
+The inertia tensor can be reduced to a $2\times2$ matrix because the frame is
+assumed to be laterally symmetric and the $y$ axis of the pendulum reference is
+the same as the $y$ axis of the benchmark reference frame. The simple rotation matrix about the $Y$-axis
+can similarly be reduced to a $2\times2$ matrix where $s_{\beta i}$ and
+$c_{\beta i}$ are defined as $\sin{\beta_i}$ and $\cos{\beta_i}$,
+respectively.
 \begin{equation}
@@
 and the unknown global inertia tensor can be solved for.
-
-Finding the inertia tensors of the wheels is less complex because the wheels are symmetric about three orthogonal planes so there are no products of inertia.
-The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring the averaged period of oscillation about an axis in the $XZ$-plane using the torsional pendulum setup and Eq.~\ref{eq:torPend}.
-The $I_{yy}$ moment of inertia was calculated with a compound pendulum as described in \cite{Kooijman2006} and shown in Fig.~\ref{fig:comPend} using
-\begin{equation}
-	I_{yy}=\left(\frac{\overline{T}}{2\pi}\right)^2mgl-ml^2
-\label{eq:comPend}
-\end{equation}
-where $l=0.303\pm0.002$ m is the pendulum length, $m$ is the mass of the wheel, $\overline{T}$ is the averaged period and $g$ is the local acceleration due to gravity.
-Table \ref{tab:wheeliInertiaComp} gives the calculated values.
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/YellowFrameCompoundFirst.jpg}
+    \end{center}
+    \caption{The rear frame hung as a compound pendulum.}
+    \label{fig:FrameCompound}
+\end{figure}
 
 \subsection{Fork}
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=5in]{images/StratosForkTorsionalThird.jpg}
+    \end{center}
+    \caption{The Stratos fork and handlebar assembly hung as a torsional
+    pendulum.}
+    \label{fig:StratosFork}
+\end{figure}
+\section{Linear Analysis}
+\subsection{Canonical Matrices}
+\subsection{Eigenvalues}
+\subsection{Frequency Response}
 \appendix
-\section{Bicycle Descriptions}
-\subsection{Yellow Bicycle}
-The yellow is a bicycle used in the lab to demonstrate that a bicycle is stable
-at certain speeds. It is an aluminum road frame of unknown make with the
-majority of components removed. The wheels, drop handlebar, seat, seat post and
-bottom bracket are the only parts on the bike. The fork has been reversed to
-''increase`` the stability of the bicycle. The bicycle was measured with both
-the fork in normal position and the fork reversed.
-\begin{table}
-	\begin{tabular}{ll}
-		Style & Road\\
-		Size & \\
-		Wheels & 700c Aluminum Rims\\
-		Tires & 28 x 1 3/8 x 1 5/8\\
-		Sprocket & 6 speed free wheel\\
-		Handlebars & Drop\\
-	\end{tabular}
-	\caption{}
-	\label{yellowBikeSpecs}
-\end{table}
-
-\subsection{Batavus Browser}
-The Batavus Browser is lower priced Dutch city bike. We measured the physical
-properities of the stock Browser model and also the same bike with the
-instrumentation used in our experiments.
-\begin{table}
-	\begin{tabular}{ll}
-		Price          & € 449\\
-		Wheel Size     & 28''\\
-		Weight         & 18.7 kg\\
-		Frame Size     & 54\\
-		Frame material & High Tensile Steel\\
-		Fork material  & Steel\\
-		Rims           & Aluminum Airline\\
-    Spokes         & Stainless steel\\
-		Tires          & Trax Highway, 28'' x 1 5/8 x 1 3/8\\
-		Handlebar      & Steel painted\\
-		Handles        & Batavus Comfort\\
-		Saddle         & Selle Royal 8274\\
-		Dynamo         & Basta Trio\\
-		Headlight      & Batavus Logic Light\\
-		Rearlight      & Battery light\\
-		Pedals         & Batavus plastic antislip\\
-		Final          & Trelock RS42\\
-		Theft chip     & Present\\
-		Crank          & Aluminum\\
-		Rack           & Steel\\
-		Quick Binder   & Binder Trio\\
-		Mudguards      & Steel\\
-		Bell           & Kenlight\\
-	\end{tabular}
-	\caption{2008 Batavus Browser Specifications}
-	\label{browserSpecs}
-\end{table}
-
-\subsection{2007 Bianchi Pista}
-The Pista is a steel track bicycle. Specifications, taken from the Bianchi
-website~\cite{Bianchi2009}, are as follows:
-\begin{table}
-	\begin{tabular}{ll}
-		Style            & Track Bike\\
-  	Size             & 57cm\\
-  	Color            & Gang Green\\
-  	Frame            & Bianchi DB CrMo, rear entry track dropouts\\
-  	Fork             & CrMo\\
-  	Retail Price     & \$579.99\\
-		Headset          & VP AheadSet, 1''threadless\\
-  	Handlebar        & Bianchi steel track, 26.0mm\\
-		Stem             & Bianchi alloy\\
-    Brakes/Levers    & Shimano Tiagra front brake/lever\\
-		Crankset         & Truvativ Touro 1.1, 48T\\
-		Bottom Bracket   & Truvativ Power Spline cartridge\\
-		Chain            & KMC\\
-		Sprocket         & 16T fixed cog \& 16T Freewheel\\
-		Pedals           & Campus pedals\\
-		Wheels           & Bianchi Alex Solo wheelset, 28h rims\\
-    Tires            & Continental UltraSport, 700x23C\\
-    Saddle           & Bianchi Velo\\
-		Seatpost         & Bianchi alloy, 27.2mm\\
-		Size             & 57cm\\ 
-		Seat Tube        & 570mm\\
-		Top Tube-actual  & 560mm\\
-		Top Tube-virtual & 560mm\\
-		Chainstay        & 383mm\\
-		Fork Rake        & 28mm\\
-		Head Tube Angle  & 74.5$^{\circ}$\\
-		Seat Tube Angle  & 75.5$^{\circ}$\\
-		Wheelbase        & 967mm\\
-		Standover Height & 32''  
-	\end{tabular}
-	\caption{}
-	\label{pistaSpecs}
-\end{table}
-\subsection{Gary Fisher Mountain Bike}
-\subsection{Batavus Crescendo Deluxe}
-
-The Batavus Crescendo Deluxe is a contemporary of all-round cycling in cool black and white.
-A great bike for a good end to touring.
-And that works fine with 8 gears, suspension fork and suspension seatpost.
-\begin{table}
-	\begin{tabular}{ll}
-		Price             & from € 869\\
-		Wheel Size        & 28''\\
-		Weight            & 21.5 kg\\
-		Frame Size        & 53\\
-		Performance/speed & Shimano Nexus 8 speed with roller brakes\\
-		Color             & White / black\\
-		Brakes            & Shimano roller brakes (BR-IM50)\\
-		Frame type        & Ladies mono tube frame Gentlemen trend\\
-		Frame material    & Aluminum 7005\\
-		Front Fork        & spring, Batavus\\
-		Rims              & Aluminum Rodi VR19\\
-		Spokes            & Stainless steel (adapted spoke pattern)\\
-		Tires             & CST End\\
-		Handlebar         & Steel aluminum look\\
-		Stem              & Manually adjustable, Batavus Ergo Matic III\\
-		Handles           & adjustable Batavus Ergo Grip II\\
-		Saddle            & Selle Royal 5160\\
-		Seatpost          & spring, Post Modern Compact\\
-		Dynamo            & Shimano Hub Dynamo\\
-		Headlight         & Trelock Hi-Power automatic halogen\\
-		Rear              & Smart battery automatically\\
-		Pedals            & aluminum antislip\\
-		Final							& AXA Defender\\
-		Chip theft        & Present\\
-		Crank             & Aluminum\\
-		Rack              & Aluminum\\
-		Quick Binder      & Binder Quattro\\
-		Mudguards         & Batavus hyalite II\\
-		Bell              & Bibia Touring\\
-		Balhoofdset       & VP semi-integrated, oversized\\
-		Shifters          & Batavus (8-speed)\\
-		Saddlebag         & Batavus saddlebag with plakset\\
-	\end{tabular}
-	\caption{}
-	\label{crescendoSpecs}
-\end{table}
-\subsection{Batavus Stratos Deluxe}
-\begin{table}
-	\begin{tabular}{ll}
-		Switching           & Spin System (shifter)\\
-		Monochromatic frame & Black\\
-		Chain guard         & Semi-open\\
-		Mudguards           & Aluminum\\
-		Rack                & Yes\\
-		Extra drive         & Regardless Motori degree\\
-		Type                & Men's\\
-		Frame Material      & Aluminum\\
-		Rim Material        & Aluminum\\
-		Rear braking system & Roller brake\\
-		Front braking system & Roller Brake\\
-		Number of gears      & 7 Gears\\
-		Gear System          & Shimano Nexus\\
-		Wheel Size           & 28 inch\\
-		Tirewidth            & 37 mm\\
-		Price                & 749,00\\
-	\end{tabular}
-	\caption{}
-	\label{stratosSpecs}
-\end{table}
 \bibliographystyle{plain}

### 2010-06-04 benchmark_bike_tools.py

@@ -371,12 +371,19 @@ def bmp2cm(filename):
 
         '''
     from numpy import pi, cos, sin, array
+    from uncertainties import ufloat
     f = open(filename, 'r')
     p = {}
     # parse the text file
     for i, line in enumerate(f):
         list1 = line[:-1].split(',')
-        p[list1[0]] = eval(list1[1])
+        # if there is an uncertainty value try to make a ufloat
+        try:
+            p[list1[0]] = ufloat(eval(list1[1]), eval(list1[2]))
+        # else keep it as a float
+        except:
+            p[list1[0]] = eval(list1[1])
+    print p
     mT = p['mR'] + p['mB'] + p['mH'] + p['mF']
     xT = (p['xB']*p['mB'] + p['xH']*p['mH'] + p['w']*p['mF'])/mT
     zT = (-p['rR']*p['mR'] + p['zB']*p['mB'] + p['zH']*p['mH'] - p['rF']*p['mF'])/mT

### 2010-06-04 benchmark_bike_tools.py
bmp2cm now works with uncertainties

@@ -370,8 +370,9 @@ def bmp2cm(filename):
         p is a dictionary of the parameters. the keys are the variable names
 
         '''
-    from numpy import pi, cos, sin, array
+    from numpy import pi, array
     from uncertainties import ufloat
+    from uncertainties.umath import sin, cos
     f = open(filename, 'r')
     p = {}
     # parse the text file
@@ -379,11 +380,10 @@ def bmp2cm(filename):
         list1 = line[:-1].split(',')
         # if there is an uncertainty value try to make a ufloat
         try:
-            p[list1[0]] = ufloat(eval(list1[1]), eval(list1[2]))
+            p[list1[0]] = ufloat((eval(list1[1]), eval(list1[2])))
         # else keep it as a float
         except:
             p[list1[0]] = eval(list1[1])
-    print p
     mT = p['mR'] + p['mB'] + p['mH'] + p['mF']
     xT = (p['xB']*p['mB'] + p['xH']*p['mH'] + p['w']*p['mF'])/mT
     zT = (-p['rR']*p['mR'] + p['zB']*p['mB'] + p['zH']*p['mH'] - p['rF']*p['mF'])/mT

### 2010-06-04 benchmark_bike_tools.py
the canonical matrices for the bikes now have uncertainties

@@ -443,6 +443,8 @@ def abMatrix(M, C1, K0, K2, v, g):
     Returns:
         A : system dynamic matrix
         B : control matrix
+
+    The states are [roll rate, steer rate, roll angle, steer angle]
     '''
     from numpy import eye, zeros, vstack, hstack, dot
     from numpy.linalg import inv

### 2010-06-04 calc_canon.py

@@ -1,4 +1,5 @@
 import pickle as p
+from uncertainties.unumpy import nominal_values as nom
 
 from benchmark_bike_tools import *
 
@@ -12,7 +13,11 @@ for i, name in enumerate(data['bikes']):
     dir = 'data/bikeParameters/'
     fname = ''.join(name.split())
     M, C1, K0, K2, param = bmp2cm(dir + fname + 'Par.txt')
-    A, B = abMatrix(M, C1, K0, K2, param['v'], param['g'])
+    try:
+        A, B = abMatrix(nom(M), nom(C1), nom(K0), nom(K2), nom(param['v']),
+                nom(param['g']))
+    except:
+        A, B = abMatrix(M, C1, K0, K2, param['v'], param['g'])
     dir = 'data/bikeCanonical/'
     file = open(dir + fname + 'Can.txt', 'w')
     for mat in ['M','C1', 'K0', 'K2', 'A', 'B']:

### 2010-06-04 benchmark_bike_tools.py
doc editing

@@ -358,16 +358,21 @@ def bike_eig(M, C1, K0, K2, v, g):
 def bmp2cm(filename):
     '''Return the benchmark canonical matrices from the bicycle parameters. Formulated from Meijaard et al. 2007.
 
-    Input:
-        filename is a text file with the 27 parameters listed in csv format. One
+    Parameters:
+    -----------
+    filename: file
+        is a text file with the 27 parameters listed in csv format. One
         parameter per line in the form: 'lamba,10/pi' or 'v,5'. Use the
         benchmark paper's units!
-    Output:
+    Returns:
+    --------
         M is the mass matrix
         C1 is the damping like matrix that is proportional to v
         K0 is the stiffness matrix proportional to gravity
         K2 is the stiffness matrix proportional to v**2
         p is a dictionary of the parameters. the keys are the variable names
+
+    This function handles parameters with uncertanties.
 
         '''
     from numpy import pi, array

### 2010-06-04 benchmark_bike_tools.py
just wrote this weird tex replacing function for variables in a dictionary, hope it works

@@ -1,3 +1,87 @@
+def replace_values(directory, template, newfile, replacers):
+    '''
+    Replaces variables with values in a text file.
+
+    Parameters:
+    -----------
+    directory: string that locates the template file
+    template: string text file
+    newfile: string  text file
+    replacers: dictionary
+
+    Returns:
+    --------
+
+    This looks for |varible_name| and replaces it with the value stored in
+    variable_name. It also looks for |variable_name?| and replaces it with the
+    uncertainty of the value of variable_name.
+
+    '''
+    import re
+    import os
+
+    from benchmark_bike_tools import uround
+
+    # open the template file
+    f = open(template, 'r')
+    # open the new file
+    fn = open(newfile, 'w')
+    for line in f:
+        #print line
+        # find all of the matches in the line
+        test = re.findall('\|(\w*.)\|', line)
+        #print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
+        #print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
+        # if there are matches
+        if test:
+            #print 'Found this!\n', test
+            # go through each match and make a substitution
+            for match in test:
+                #print "replace this: ", match
+                # if there is a '[' then the match is an array entry
+                if '[' in match:
+                    var = match.split('[')[0] # this returns the variable
+                    loc = '[' + match.split('[')[1] # may include question mark
+                    # if the match has a question mark it is an uncertainty value
+                    if loc[-1] == '?':
+                        try:
+                            line = re.sub('\|(\w*\?)\|',
+                                    uround(eval('replacers[' + var + ']' + loc[:-1])).split('+/-')[1], line, count=1)
+                            #print line
+                        except: # there is no uncertainty
+                            pass
+                    else:
+                        try:
+                            line = re.sub('\|(\w*(?!\?))\|',
+                                    uround(eval('replacers[' + var + ']' + loc)).split('+/-')[0], line, count=1)
+                            #print line
+                        except: # there is no uncertainty
+                            line = re.sub('\|(\w*(?!\?))\|', str(eval('replacers[' + var + ']' + loc)), line, count=1)
+                # else the match is just a scalar
+                else:
+                    # if the match has a question mark it is an uncertainty value
+                    if match[-1] == '?':
+                        try:
+                            line = re.sub('\|(\w*\?)\|',
+                                    uround(replacers[match[:-1]]).split('+/-')[1], line, count=1)
[…]

### 2010-06-04 benchmark_bike_tools.py
still working on this silly function, seems like there should be something that does this, maybe I should look around

@@ -23,38 +23,43 @@ def replace_values(directory, template, newfile, replacers):
     from benchmark_bike_tools import uround
 
     # open the template file
-    f = open(template, 'r')
+    f = open(directory + template, 'r')
     # open the new file
-    fn = open(newfile, 'w')
+    fn = open(directory + newfile, 'w')
     for line in f:
-        #print line
+        print 'This is the template line:\n', line
         # find all of the matches in the line
         test = re.findall('\|(\w*.)\|', line)
-        #print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
-        #print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
+        print 'these are all the matches', test
+        print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
+        print 'search for un: ',  re.findall('\|(\w*\?)\|', line)
         # if there are matches
         if test:
-            #print 'Found this!\n', test
+            print 'Found this!\n', test
             # go through each match and make a substitution
             for match in test:
-                #print "replace this: ", match
-                # if there is a '[' then the match is an array entry
-                if '[' in match:
-                    var = match.split('[')[0] # this returns the variable
-                    loc = '[' + match.split('[')[1] # may include question mark
+                print "replace this: ", match
+                # if there is a '_' then the match is an array entry
+                if '_' in match:
+                    ques = None
+                    try:
+                        var, row, col, ques = match.split('_') # this returns the variable
+                    except:
+                        var, row, col = match.split('_') # this returns the variable
+
                     # if the match has a question mark it is an uncertainty value
-                    if loc[-1] == '?':
+                    if ques:
                         try:
                             line = re.sub('\|(\w*\?)\|',
                                     uround(eval('replacers[' + var + ']' + loc[:-1])).split('+/-')[1], line, count=1)
-                            #print line
+                            print line
                         except: # there is no uncertainty
                             pass
                     else:
                         try:
                             line = re.sub('\|(\w*(?!\?))\|',
                                     uround(eval('replacers[' + var + ']' + loc)).split('+/-')[0], line, count=1)
-                            #print line
+                            print line
                         except: # there is no uncertainty
                             line = re.sub('\|(\w*(?!\?))\|', str(eval('replacers[' + var + ']' + loc)), line, count=1)
                 # else the match is just a scalar
@@ -64,23 +69,23 @@ def replace_values(directory, template, newfile, replacers):
[…]

### 2010-06-04 calc_canon.py

@@ -10,16 +10,16 @@ f.close()
 
 # write the parameter files
 for i, name in enumerate(data['bikes']):
-    dir = 'data/bikeParameters/'
+    directory = 'data/bikeParameters/'
     fname = ''.join(name.split())
-    M, C1, K0, K2, param = bmp2cm(dir + fname + 'Par.txt')
+    M, C1, K0, K2, param = bmp2cm(directory + fname + 'Par.txt')
     try:
         A, B = abMatrix(nom(M), nom(C1), nom(K0), nom(K2), nom(param['v']),
                 nom(param['g']))
     except:
         A, B = abMatrix(M, C1, K0, K2, param['v'], param['g'])
-    dir = 'data/bikeCanonical/'
-    file = open(dir + fname + 'Can.txt', 'w')
+    directory = 'data/bikeCanonical/'
+    file = open(directory + fname + 'Can.txt', 'w')
     for mat in ['M','C1', 'K0', 'K2', 'A', 'B']:
         if mat == 'A' or mat == 'B':
             file.write(mat + ' (v = ' + str(param['v']) + ')\n')
@@ -27,7 +27,9 @@ for i, name in enumerate(data['bikes']):
             file.write(mat + '\n')
         file.write(str(eval(mat)) + '\n')
     file.close()
-    file = open(dir + fname + 'Can.p', 'w')
-    p.dump({'M':M, 'C1':C1, 'K0':K0, 'K2':K2, 'A':A, 'B':B, 'v':param['v']},
-            file)
+    file = open(directory + fname + 'Can.p', 'w')
+    canon = {'M':M, 'C1':C1, 'K0':K0, 'K2':K2, 'A':A, 'B':B, 'v':param['v']}
+    p.dump(canon, file)
     file.close()
+    replace_values(directory, 'CanonTemplate.tex', fname + 'Can.tex', canon)
+    stop

### 2010-06-04 benchmark_bike_tools.py
replacement seems to work for the canonical matrices

@@ -22,6 +22,8 @@ def replace_values(directory, template, newfile, replacers):
 
     from benchmark_bike_tools import uround
 
+    print replacers
+
     # open the template file
     f = open(directory + template, 'r')
     # open the new file
@@ -43,25 +45,24 @@ def replace_values(directory, template, newfile, replacers):
                 if '_' in match:
                     ques = None
                     try:
-                        var, row, col, ques = match.split('_') # this returns the variable
+                        var, row, col, ques = match.split('_')
                     except:
-                        var, row, col = match.split('_') # this returns the variable
-
+                        var, row, col = match.split('_')
+                    valStr = "replacers['" + var + "']" + '[' + row + ', ' + col + ']'
+                    print valStr
+                    value = eval(valStr)
+                    print 'value =', value
                     # if the match has a question mark it is an uncertainty value
                     if ques:
                         try:
-                            line = re.sub('\|(\w*\?)\|',
-                                    uround(eval('replacers[' + var + ']' + loc[:-1])).split('+/-')[1], line, count=1)
-                            print line
+                            line = re.sub('\|(\w*\?)\|', uround(value).split('+/-')[1], line, count=1)
                         except: # there is no uncertainty
                             pass
                     else:
                         try:
-                            line = re.sub('\|(\w*(?!\?))\|',
-                                    uround(eval('replacers[' + var + ']' + loc)).split('+/-')[0], line, count=1)
-                            print line
+                            line = re.sub('\|(\w*(?!\?))\|', uround(value).split('+/-')[0], line, count=1)
                         except: # there is no uncertainty
-                            line = re.sub('\|(\w*(?!\?))\|', str(eval('replacers[' + var + ']' + loc)), line, count=1)
+                            line = re.sub('\|(\w*(?!\?))\|', str(value), line, count=1)
                 # else the match is just a scalar
                 else:
                     # if the match has a question mark it is an uncertainty value

### 2010-06-04 benchmark_bike_tools.py
replace_values seems to be working

@@ -37,7 +37,6 @@ def replace_values(directory, template, newfile, replacers):
         print 'search for un: ',  re.findall('\|(\w*\?)\|', line)
         # if there are matches
         if test:
-            print 'Found this!\n', test
             # go through each match and make a substitution
             for match in test:
                 print "replace this: ", match
@@ -49,20 +48,29 @@ def replace_values(directory, template, newfile, replacers):
                     except:
                         var, row, col = match.split('_')
                     valStr = "replacers['" + var + "']" + '[' + row + ', ' + col + ']'
-                    print valStr
+                    #print valStr
                     value = eval(valStr)
                     print 'value =', value
                     # if the match has a question mark it is an uncertainty value
                     if ques:
                         try:
-                            line = re.sub('\|(\w*\?)\|', uround(value).split('+/-')[1], line, count=1)
+                            replacement = uround(value).split('+/-')[1]
+                            print "with this:", replacement
+                            line = re.sub('\|(\w*\?)\|', replacement, line, count=1)
                         except: # there is no uncertainty
-                            pass
+                            replacement = '0.0'
+                            print "with this:", replacement
+                            line = re.sub('\|(\w*\?)\|', replacement, line, count=1)
                     else:
                         try:
-                            line = re.sub('\|(\w*(?!\?))\|', uround(value).split('+/-')[0], line, count=1)
+                            replacement = uround(value).split('+/-')[0]
+                            print "with this:", replacement
+                            line = re.sub('\|(\w*(?!\?))\|', replacement, line, count=1)
                         except: # there is no uncertainty
-                            line = re.sub('\|(\w*(?!\?))\|', str(value), line, count=1)
+                            replacement = str(value)
+                            print "with this:", replacement
+                            line = re.sub('\|(\w*(?!\?))\|', replacement, line, count=1)
+                    del var, row, col
                 # else the match is just a scalar
                 else:
                     # if the match has a question mark it is an uncertainty value
@@ -84,9 +92,9 @@ def replace_values(directory, template, newfile, replacers):
         fn.write(line)
     f.close()
     fn.close()
-    # os.system('pdflatex -output-directory=' + directory + ' ' + newfile)
-    # os.system('rm ' + directory + '*.aux')
-    # os.system('rm ' + directory + '*.log')
+    os.system('pdflatex -output-directory=' + directory + ' ' + newfile)
+    os.system('rm ' + directory + '*.aux')
+    os.system('rm ' + directory + '*.log')
 
 def uround(value):
     '''Round values according to their uncertainity

### 2010-06-04 calc_canon.py
the pdfs look good

@@ -31,5 +31,5 @@ for i, name in enumerate(data['bikes']):
     canon = {'M':M, 'C1':C1, 'K0':K0, 'K2':K2, 'A':A, 'B':B, 'v':param['v']}
     p.dump(canon, file)
     file.close()
+    # build some pretty tex files
     replace_values(directory, 'CanonTemplate.tex', fname + 'Can.tex', canon)
-    stop

### 2010-06-05 report/parameterReport.tex
canonical template is ready
updated for the subfiles package
changed the parameter tables to be smaller

@@
 \usepackage{eurosans}
+\usepackage{subfiles}
 
@@
 \end{equation}
-where $\mathbf{H}$ is the Hessian. $\mathbf{U}$ is a $5\times5$ matrix with the
+where $\mathbf{H}$ is the Hessian~\cite{Hubbard1989b}. $\mathbf{U}$ is a $5\times5$ matrix with the
 variances of each of the five fit parameters along the diagonal. This assumes
@@
 \end{figure}
+\section{Parameters}
+\subfile{../parTables/BatavusBrowserPar.tex}
+\subfile{../parTables/BatavusBrowserInstrumentedPar.tex}
+\subfile{../parTables/BatavusCrescendoDeluxePar.tex}
+\subfile{../parTables/GaryFisherPar.tex}
+\subfile{../parTables/BianchiPistaPar.tex}
+\subfile{../parTables/BatavusStratosDeluxePar.tex}
+\subfile{../parTables/YellowBikePar.tex}
+\subfile{../parTables/YellowBikeReversedForkPar.tex}
 \section{Linear Analysis}
 \subsection{Canonical Matrices}
+The final results are presented in the form used by the benchmark model. These
+can be used to populate the canonical form
+\begin{equation}
+\mathbf{M\ddot{q}}+v\mathbf{C}_1\mathbf{\dot{q}}+\left[g\mathbf{K}_0+v^2\mathbf{K}_2\right]\mathbf{q}=0
+\label{eq:canonical}
+\end{equation}
+of the linear benchmark equations of motion presented in
+\cite{Meijaard2007}. The coefficient matrices for the
+example rider and bicycle follow in Eqs.~\ref{eq:M}-\ref{eq:K2}.
+\subsubsection{Batavus Browser}
+\subfile{../data/bikeCanonical/BatavusBrowserCan.tex}
+
 \subsection{Eigenvalues}

### 2010-06-08 report/parameterReport.tex
change file to f and dir to direct

@@
     \begin{center}
-        \includegraphics[width=5in]{images/yellow.jpg}
+        \includegraphics[width=4in]{images/yellow.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/browser.jpg}
+        \includegraphics[width=4in]{images/browser.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/browserIns.jpg}
+        \includegraphics[width=4in]{images/browserIns.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/pista.jpg}
+        \includegraphics[width=4in]{images/pista.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/fisher.jpg}
+        \includegraphics[width=4in]{images/fisher.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/crescendo.jpg}
+        \includegraphics[width=4in]{images/crescendo.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/stratos.jpg}
+        \includegraphics[width=4in]{images/stratos.jpg}
     \end{center}
@@
 \end{equation}
+\begin{equation}
+    \sigma_r^{2}=\left(\frac{\sigma_d}{2\pi n}\right)^{2}
+    \label{eq:WheelRadVar}
+\end{equation}
 \begin{figure}[tb]
@@
 \begin{equation}
-	\lambda=\frac{\pi}{180}(90^{\circ}-\lambda_{ht})
+    \lambda=\frac{\pi}{180^{\circ}}(90^{\circ}-\lambda_{ht})
 \label{eq:headTubeAngle}
+\end{equation}
+\begin{equation}
+    \sigma_\lambda^{2}=-\sigma_{\lambda_{ht}}^{2}\left(\frac{\pi}{180^\circ}\right)^2
+    \label{eq:HeadTubeAngleVar}
 \end{equation}
@@
 	\label{eq:trail}
+\end{equation}
+\begin{equation}
+    \sigma_{c}^{2}=\sigma_{r_{\mathrm{F}}}^{2}\tan^2{\lambda} -
+    \sigma_{f_o}^{2}\sec^2{\lambda} +
+    \sigma_{\lambda}^{2}\left(r_\mathrm{F}\sec^2{\lambda} -
+    f_o\sec{\lambda}\tan{\lambda}\right)^2
+    \label{eq:TrailVar}
 \end{equation}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/massScale.jpg}
+        \includegraphics[width=4in]{images/massScale.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/pendDist.jpg}
+        \includegraphics[width=4in]{images/pendDist.jpg}
     \end{center}
@@
 	\begin{center}
-		\includegraphics[width=5in]{images/level.jpg}
+		\includegraphics[width=4in]{images/level.jpg}
 	\end{center}
@@
 \end{equation}
+\begin{equation}
+    \sigma_{\beta}^{2} = \sigma_{\lambda}^{2} + \sigma_{\alpha}^{2}
+    \label{eq:FrameRotAngVar}
+\end{equation}
 The pendulum axis $X_P$ is simply a line in the nominal bicycle reference frame with a slope $m$ and a z-intercept $b$. The slope can be shown to be
@@
 \end{equation}
+\begin{equation}
+    \sigma_{m}^{2} = \sigma_{\beta}^{2}\sec^{4}{\beta}
+    \label{eq:SlopeVar}
+\end{equation}
 The z-intercept can be shown to be
@@
     \label{eq:zInt}
+\end{equation}
+\begin{equation}
+    \sigma_{b}^{2} = \sigma_{a}^{2}\sec^{2}{\beta} +
+    \sigma_{r_\mathrm{R}}^{2} +
+    \sigma_{\beta}^{2}a^{2}\sec^{2}{\beta}\tan^{2}{\beta}
+    \label{eq:zIntvar}
 \end{equation}
@@
     b = w\tan{\beta} - r_\mathrm{F} - \frac{a}{\cos{\beta}} 
+    \label{eq:zIntFork}
+\end{equation}
+\begin{equation}
+    \sigma_{b}^{2} = \sigma_{w}^{2}\tan^{2}\beta +
+    \sigma_{\beta}^{2}\left(w\sec^{2}\beta -
+    a\sec\beta\tan\beta\right)^{2} + \sigma_{r_\mathrm{F}}^{2} +
+    \sigma_{a}^{2}\sec^{2}\beta
+    \label{eq:zIntForkVar}
 \end{equation}
@@
     \begin{center}
-        \includegraphics[width=5in]{../data/pendDat/graphs/BrowserFrameCompoundFirst1.png}
+        \includegraphics[width=4in]{../data/pendDat/graphs/BrowserFrameCompoundFirst1.png}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{../data/pendDat/graphs/CrescendoForkTorsionalFirst2.png}
+        \includegraphics[width=4in]{../data/pendDat/graphs/CrescendoForkTorsionalFirst2.png}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/fixture.jpg}
+        \includegraphics[width=4in]{images/fixture.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/rod.jpg}
+        \includegraphics[width=4in]{images/rod.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/CrescendoFwheelTorsionalFirst.jpg}
+        \includegraphics[width=4in]{images/CrescendoFwheelTorsionalFirst.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/wheelIyy.jpg}
+        \includegraphics[width=4in]{images/wheelIyy.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/YellowFrameCompoundFirst.jpg}
+        \includegraphics[width=4in]{images/YellowFrameCompoundFirst.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=5in]{images/StratosForkTorsionalThird.jpg}
+        \includegraphics[width=4in]{images/StratosForkTorsionalThird.jpg}
     \end{center}
@@
 \section{Parameters}
+\subsection{Batavus Browser}
 \subfile{../parTables/BatavusBrowserPar.tex}
+\subsection{Batavus Browser Instrumented}
 \subfile{../parTables/BatavusBrowserInstrumentedPar.tex}
+\subsection{Batavus Crescendo Deluxe}
 \subfile{../parTables/BatavusCrescendoDeluxePar.tex}
+\subsection{Gary Fisher}
 \subfile{../parTables/GaryFisherPar.tex}
+\subsection{Bianchi Pista}
 \subfile{../parTables/BianchiPistaPar.tex}
+\subsection{Batavus Stratos Deluxe}
 \subfile{../parTables/BatavusStratosDeluxePar.tex}
+\subsection{Yellow Bike}
 \subfile{../parTables/YellowBikePar.tex}
+\subsection{Yellow Bike Reversed Fork}
 \subfile{../parTables/YellowBikeReversedForkPar.tex}
@@
 \subfile{../data/bikeCanonical/BatavusBrowserCan.tex}
+\subsubsection{Batavus Browser Instrumented}
+\subfile{../data/bikeCanonical/BatavusBrowserInstrumentedCan.tex}
+\subsubsection{Batavus Crescendo Deluxe}
+\subfile{../data/bikeCanonical/BatavusCrescendoDeluxeCan.tex}
+\subsubsection{Gary Fisher}
+\subfile{../data/bikeCanonical/GaryFisherCan.tex}
+\subsubsection{Bianchi Pista}
+\subfile{../data/bikeCanonical/BianchiPistaCan.tex}
+\subsubsection{Batavus Stratos Deluxe}
+\subfile{../data/bikeCanonical/BatavusStratosDeluxeCan.tex}
+\subsubsection{Yellow Bike}
+\subfile{../data/bikeCanonical/YellowBikeCan.tex}
+\subsubsection{Yellow Bike Reversed Fork}
+\subfile{../data/bikeCanonical/YellowBikeReversedForkCan.tex}
 

### 2010-06-08 bike_eig.py

@@ -16,13 +16,13 @@ vd = np.zeros(nBk)
 vw = np.zeros(nBk)
 vc = np.zeros(nBk)
 eigFig = plt.figure(num=1)
-dir = 'data/bikeRiderCanonical/'
+direct = 'data/bikeRiderCanonical/'
 vel = np.linspace(0, 20, num=1000)
 for i, name in enumerate(data['bikes']):
     fname = ''.join(name.split()) + 'RiderCan.p'
-    file = open(dir + fname)
-    can = pickle.load(file)
-    file.close()
+    f = open(direct + fname)
+    can = pickle.load(f)
+    f.close()
     evals, evecs = bike_eig(can['M'], can['C1'], can['K0'], can['K2'], vel, 9.81)
     wea, cap, cas = sort_modes(evals, evecs)
     plt.figure(1)

### 2010-06-08 plot_eigvec.py
adding a new eigenvector plotting script

@@ -1,0 +1,29 @@
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+import numpy as np
+import pickle
+
+from benchmark_bike_tools import bike_eig, sort_modes
+
+f = open('data/bikeRiderCanonical/BianchiPistaRiderCan.p')
+can = pickle.load(f)
+f.close()
+
+vel = np.linspace(0, 20, num=500)
+evals, evecs = bike_eig(can['M'], can['C1'], can['K0'], can['K2'], vel, 9.81)
+wea, cap, cas = sort_modes(evals, evecs)
+
+colors = ['blue', 'red', 'green', 'orange']
+
+fig = plt.figure()
+ax = Axes3D(fig)
+for i, row in enumerate(wea['evecs'][:, :, 1]):
+    for k, component in enumerate(row[:2]):
+        point1 = np.array([0., 0., vel[i]])
+        point2 = np.array([np.abs(np.real(component)),
+            np.abs(np.imag(component)), vel[i]])
+        ax.plot(np.hstack((point1[0], point2[0])), np.hstack((point1[1],
+            point2[1])), zs=vel[i], color=colors[k])
+            #ax.plot(np.abs(np.real(wea['evecs'][:, :, 1])),
+                #np.abs(np.imag(wea['evecs'][:, :, 1])), zs=vel)
+plt.show()

### 2010-09-08 bike_eig.py
turned the bike_eig back to the bike only

@@ -11,14 +11,18 @@ f = open('data/data.p', 'r')
 data = pickle.load(f)
 f.close()
 
-#data['bikes'].append('Jodi Bike')
+data['bikes'].append('Jodi Bike')
 
 nBk = len(data['bikes'])
 
+# one color for each bike
 colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange', 'red']
+
 vd = np.zeros(nBk)
 vw = np.zeros(nBk)
 vc = np.zeros(nBk)
+
+# figure properties
 figwidth = 4. # in inches
 goldenMean = (np.sqrt(5)-1.0)/2.0
 figsize = [figwidth, figwidth*goldenMean]
@@ -34,10 +38,11 @@ figsize = [figwidth, figwidth*goldenMean]
 eigFig = plt.figure(num=nBk + 1)#, figsize=figsize)
 #plt.clf()
 #plt.axes([0.125,0.2,0.95-0.125,0.95-0.2])
-direct = 'data/bikeLegsCanonical/'
+
+direct = 'data/bikeCanonical/'
 vel = np.linspace(0, 20, num=1000)
 for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split()) + 'LegsCan.p'
+    fname = ''.join(name.split()) + 'Can.p'
     f = open(direct + fname)
     can = pickle.load(f)
     f.close()

### 2010-09-08 bike_eig.py
bike_eig has option to choose rider, no rider and legs

@@ -6,12 +6,25 @@ from uncertainties.unumpy import std_devs
 
 from benchmark_bike_tools import *
 
+# choose the bike alone, bike with rigid legs or bike with rider
+# bike alone
+#direct = 'data/bikeCanonical/'
+#fend = 'Can.p'
+# bike with rider's legs
+#direct = 'data/bikeLegsCanonical/'
+#fend = 'LegsCan.p'
+# bike with rider
+direct = 'data/bikeRiderCanonical/'
+fend = 'RiderCan.p'
+
+
 # load in the base data file
 f = open('data/data.p', 'r')
 data = pickle.load(f)
 f.close()
 
-data['bikes'].append('Jodi Bike')
+# add Jodi's thesis bike for comparison (only for bikeCanonical)
+#data['bikes'].append('Jodi Bike')
 
 nBk = len(data['bikes'])
 
@@ -39,10 +52,9 @@ eigFig = plt.figure(num=nBk + 1)#, figsize=figsize)
 #plt.clf()
 #plt.axes([0.125,0.2,0.95-0.125,0.95-0.2])
 
-direct = 'data/bikeCanonical/'
 vel = np.linspace(0, 20, num=1000)
 for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split()) + 'Can.p'
+    fname = ''.join(name.split()) + fend
     f = open(direct + fname)
     can = pickle.load(f)
     f.close()

### 2010-09-08 PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
added the bike dynamics and control software map, among other things

@@
 Accurate measurements of a bicycle's physical parameters are required for
-realistic dynamic simulations. For the most basic models the geometry, mass,
-mass center location and mass distributions must be measured. More complex models
+realistic dynamic simulations and anaysis. For the most basic models the geometry, mass,
+mass location and mass distributions must be measured. More complex models
 require measurements of tire characteristics, stiffness, damping, etc. This
-paper concerns the measurement of bicycle parameters required for
-simulations of the benchmark bicycle presented in \cite{Meijaard2007}. This
+paper concerns the measurement of the minimal bicycle parameters required for
+the benchmark bicycle presented in~\cite{Meijaard2007}. This
 model is composed of four rigid bodies, has ideal rolling and frictionless joints,
 and is assumed to be laterally symmetric. A set of 25
-physical parameters is used to describe the geometry, mass, mass location and
+parameters is used to describe the geometry, mass, mass location and
 mass distribution of each of the rigid bodies. These parameters are presented
@@
 configuration. The methods described herein are based primarily on the work
-done in \cite{Kooijman2006} but have been refined for improved accuracy and
+done in~\cite{Kooijman2006} but have been refined for improved accuracy and
 methodology. Previously, \cite{Roland1971} measured an experimental bicycle frame in a
-similar fashion and both \cite{Dohring1953} and \cite{Singh1971} have used
+similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} have used
 similar techniques with scooters. We measured the characteristics of six
@@
 The six bicycles, chosen for both variety and convenience, are as follows:
-\begin{description}
-    \item[Batavus Browser] The Dutch style city bicycle was measured
-        in two configurations, one with the instrumentation described in
-        \cite{Kooijman2009a} and one as received from the manufacturer.
-    \item[Batavus Stratos Deluxe] A Dutch style sporty city bicycle.
-    \item[Batavus Crescendo Deluxe] A Dutch style city bicycle with a
-        suspension fork.
-    \item[Gary Fisher Mountain Bike] A hardtail mountain bicycle.
-    \item[Bianchi Pista] A modern steel frame track racing bicycle.
-    \item[Yellow Bicycle] A stripped down aluminum frame road bicycle. This
-        bike was measured in two configurations, the second with the fork
-        rotated in the headtube 180 degrees for larger trail.
-\end{description}
+\emph{Batavus Browser}, a Dutch style city bicycle measured with and without
+instrumentation as described in~\cite{Kooijman2009a}, \emph{Batavus Stratos
+Deluxe}, a Dutch style sporty city bicycle, \emph{Batavus Crescendo Deluxe} a
+Dutch style city bicycle with a suspened fork, \emph{Gary Fisher Mountain
+Bike}, a hardtail mountain bicycle, \emph{Bianchi Pista}, a modern steel frame
+track racing bicycle, \emph{Yellow Bicycle}, a stripped down aluminum frame
+road bicycle measured in two configurations, the second with the fork rotated
+in the headtube 180 degrees for larger trail.
 \subsection*{Geometry}
@@
 Parameter sets for all six bicycles and all eight
-configurations, example in Table~\ref{tab:par},  will be
+configurations  will be
 presented, together with basic comparisons of the open loop dynamics of the
@@
 descriptions of handling.
-\begin{table*}[tb]
-\centering
-\caption{PARAMETER VALUES FOR THE BATAVUS BROWSER.}
-{\small
-\begin{tabular}{lcl}
-&&\\
-\hline
-Parameter & Symbol & Value \\
-\hline
-wheel base & $w$ & 1.121 m \\
-trail & $c$ & 0.069 m \\
-steer axis tilt & \emph{$\lambda$} & 0.400 rad \\
-gravity & \emph{g} & 9.81 N kg$^{-1}$ \\
-forward speed & \emph{v} & various m s$^{-1}$ \\
-
-\multicolumn{3}{l}{\textbf{R}\emph{ear wheel R}}\\
-radius & \emph{$r_\mathrm{R}$} & 0.341 m\\
-mass & \emph{$m_\mathrm{R}$} &  3.11 kg \\
-mass moments of inertia & \emph{$(I_{\mathrm{R}xx},\
-I_{\mathrm{R}yy})$} &
-(0.088,\ 0.152) $\mathrm{kg\ m}^2$ \\
-
-\multicolumn{3}{l}{\emph{rear} \textbf{B}\emph{ody and frame B}}\\
-position center of mass & \emph{$(x_\mathrm{B},\ z_\mathrm{B})$} &
-($0.276$,\ $-0.538$) m\\
-mass & \emph{$m_\mathrm{B}$} & 9.86 kg \\
-mass moments of inertia & $\left[ \begin{array}{ccc}
-I_{\mathrm{B}xx} &  0 & I_{\mathrm{B}xz}\\
-0 & I_{\mathrm{B}yy} & 0 \\
-I_{\mathrm{B}xz} & 0 & I_{\mathrm{B}zz}
-\end{array} \right] $ &
-$\left[ \begin{array}{ccc}
-0.527 &  0 & -0.114\\
-0 & 1.317 & 0 \\
--0.114 & 0 & 0.759
-\end{array} \right] \ \mathrm{kg\ m}^{2}$\\
-
-\multicolumn{3}{l}{\emph{front} \textbf{H}\emph{andlebar and fork assembly H}}\\
-position center of mass & \emph{$(x_\mathrm{H},\ z_\mathrm{H})$} &
-($0.867$,\ $-0.748$) m\\
-mass & \emph{$m_\mathrm{H}$} & 3.22 kg \\
-mass moments of inertia & $\left[ \begin{array}{ccc}
-I_{\mathrm{H}xx} &  0 & I_{\mathrm{H}xz}\\
-0 & I_{\mathrm{H}yy} & 0 \\
-I_{\mathrm{H}xz} & 0 & I_{\mathrm{H}zz}
-\end{array} \right] $ &
-$\left[ \begin{array}{ccc}
-    0.253  &       0  & -0.072\\
-         0  & 0.246   &  0\\
-   -0.072  &       0  &  0.096
-\end{array} \right] \ \mathrm{kg\ m}^{2}$\\
-
-\multicolumn{3}{l}{\textbf{F}\emph{ront wheel F}}\\
-radius & \emph{$r_\mathrm{F}$} & 0.344 m\\
-mass & \emph{$m_\mathrm{F}$} & 2.02 kg \\
-mass moments of inertia & \emph{$(I_{\mathrm{F}xx},\
-I_{\mathrm{F}yy})$} &
-(0.090,\ 0.149) $\mathrm{kg\ m}^2$ \\
-\hline
-\end{tabular}
-}
-\label{tab:par}
-\end{table*}
 

### 2010-09-08 bike_eig.py
brought over the BMD2010 stuff

@@ -8,14 +8,14 @@ from benchmark_bike_tools import *
 
 # choose the bike alone, bike with rigid legs or bike with rider
 # bike alone
-#direct = 'data/bikeCanonical/'
-#fend = 'Can.p'
+direct = 'data/bikeCanonical/'
+fend = 'Can.p'
 # bike with rider's legs
 #direct = 'data/bikeLegsCanonical/'
 #fend = 'LegsCan.p'
 # bike with rider
-direct = 'data/bikeRiderCanonical/'
-fend = 'RiderCan.p'
+#direct = 'data/bikeRiderCanonical/'
+#fend = 'RiderCan.p'
 
 
 # load in the base data file
@@ -26,7 +26,7 @@ f.close()
 # add Jodi's thesis bike for comparison (only for bikeCanonical)
 #data['bikes'].append('Jodi Bike')
 
-nBk = len(data['bikes'])
+nBk = len(data['bikes']) # number of bikes
 
 # one color for each bike
 colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange', 'red']
@@ -52,6 +52,7 @@ eigFig = plt.figure(num=nBk + 1)#, figsize=figsize)
 #plt.clf()
 #plt.axes([0.125,0.2,0.95-0.125,0.95-0.2])
 
+# caluculate eigenvalues/vectors over this range
 vel = np.linspace(0, 20, num=1000)
 for i, name in enumerate(data['bikes']):
     fname = ''.join(name.split()) + fend
@@ -62,47 +63,16 @@ for i, name in enumerate(data['bikes']):
         can[k] = nom(v)
     evals, evecs = bike_eig(can['M'], can['C1'], can['K0'], can['K2'], vel, 9.81)
     wea, cap, cas = sort_modes(evals, evecs)
-    #nwea = []
-    #swea = []
-    #for num in wea['evals'].flatten():
-        #nwea.append(num.nominal_value)
-        #swea.append(num.std_dev())
-    #nwea = np.array(nwea).reshape(wea['evals'].shape[0],
-            #wea['evals'].shape[1])
-    #swea = np.array(swea).reshape(wea['evals'].shape[0],
-            #wea['evals'].shape[1])
-    #ncap = np.zeros_like(cap['evals'])
-    #scap = np.zeros_like(cap['evals'])
-    #ncas = np.zeros_like(cas['evals'])
-    #scas = np.zeros_like(cas['evals'])
[…]

### 2010-09-08 BMD2010/PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
added a figure

--- 2010-09-08 PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
@@
 Accurate measurements of a bicycle's physical parameters are required for
-realistic dynamic simulations and anaysis. For the most basic models the geometry, mass,
-mass location and mass distributions must be measured. More complex models
-require measurements of tire characteristics, stiffness, damping, etc. This
+realistic dynamic simulations and analysis. For the most basic models the
+geometry, mass, mass location and mass distributions must be measured. More complex models
+require measurements of tire characteristics, human characteristics, friction, stiffness, damping, etc. This
 paper concerns the measurement of the minimal bicycle parameters required for
@@
 relative to a reference frame when the bicycle is in the nominal upright
-configuration. The methods described herein are based primarily on the work
+configuration. The experimental methods described herein are based primarily on the work
 done in~\cite{Kooijman2006} but have been refined for improved accuracy and
-methodology. Previously, \cite{Roland1971} measured an experimental bicycle frame in a
-similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} have used
+methodology. Koojiman's was preceeded by \cite{Roland1971} who measured an experimental bicycle frame in a
+similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
 similar techniques with scooters. We measured the characteristics of six
@@
 This is a total of eight different parameter sets that can be used with, but not
-limited to, the benchmark bicycle model. The parameters set is also complete
-for the non-linear benchmark bicycle model.
+limited to, the benchmark bicycle model. The accuracy of all the measurements
+are presented up through the eigenvalue prediction of the linear model and are
+based error propogation theory with correlations taken into account.
 
@@
 \subsection*{Results}
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=4in]{../../report/figures/bike_eig.png}
+    \end{center}
+    \caption{Eigenvalues versus speed for all eight bicycle configurations.}
+    \label{fig:bike_eig}
+\end{figure}
+
 Parameter sets for all six bicycles and all eight
-configurations  will be
-presented, together with basic comparisons of the open loop dynamics of the
+configurations  will be presented, together with basic comparisons of the open loop dynamics of the
 different types of bicycles. The physical parameters, the mass, damping and stiffness

### 2010-09-08 bike_eig.py

@@ -39,15 +39,16 @@ vc = np.zeros(nBk)
 figwidth = 4. # in inches
 goldenMean = (np.sqrt(5)-1.0)/2.0
 figsize = [figwidth, figwidth*goldenMean]
-#params = {#'backend': 'ps',
+params = {#'backend': 'ps',
     #'axes.labelsize': 10,
     #'text.fontsize': 10,
-    #'legend.fontsize': 10,
+    'legend.fontsize': 10,
     #'xtick.labelsize': 8,
     #'ytick.labelsize': 8,
     #'text.usetex': True,
-    #'figure.figsize': figsize}
-#plt.rcParams.update(params)
+    #'figure.figsize': figsize
+    }
+plt.rcParams.update(params)
 eigFig = plt.figure(num=nBk + 1)#, figsize=figsize)
 #plt.clf()
 #plt.axes([0.125,0.2,0.95-0.125,0.95-0.2])

### 2010-09-08 BMD2010/PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
almost got the abstract shaped up

@@
 {\fontsize{16pt}{20pt}\selectfont%
-  An Accurate Method of Measuring and Comparing a Bicycle's Physical Parameters\\}
+  An Accurate Method of Measuring a Bicycle's Physical Parameters\\}
 \end{flushleft}
@@
 done in~\cite{Kooijman2006} but have been refined for improved accuracy and
-methodology. Koojiman's was preceeded by \cite{Roland1971} who measured an experimental bicycle frame in a
+methodology. Koojiman's was preceded by \cite{Roland1971} who measured an experimental bicycle frame in a
 similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
@@
 Deluxe}, a Dutch style sporty city bicycle, \emph{Batavus Crescendo Deluxe} a
-Dutch style city bicycle with a suspened fork, \emph{Gary Fisher Mountain
+Dutch style city bicycle with a suspended fork, \emph{Gary Fisher Mountain
 Bike}, a hardtail mountain bicycle, \emph{Bianchi Pista}, a modern steel frame
@@
 in the headtube 180 degrees for larger trail.
-\subsection*{Geometry}
+
+\section*{Experimental Methodology}
 The benchmark bicycle requires five basic geometric measurements: wheelbase
 $w$, steer axis tilt $\lambda$, trail $c$, and wheel radii $r_\mathrm{F}$
-and $r_\mathrm{R}$. The wheelbase and steer axis tilt were measured directly
-with a meter scale and a digital level, respectively. The trail and wheel radii were
-measured indirectly using fork offset and loaded wheel circumference.
-\subsection*{Mass, Mass Center Location and Mass Distribution}
-The mass of each of the rigid bodies were measured with a precision scale.
-The centers of mass of each of the rigid bodies were measured by hanging each
-rigid body at the mass center and assuming lateral symmetry. The mass centers
-were calculated from the orientation angle of the headtube and the location of
-the point of suspension relative to the wheel center.
-The wheel centers of mass were assumed to be in their ideal
-geometric center.
+and $r_\mathrm{R}$. For each geometric measurement care was taken to measure
+the parameter directly as possible with high accuracy measurement equipment
+such as height gages, scales, etc. The mass of each of the rigid bodies were
+measured with a precision scale. The centers of mass of each of the rigid
+bodies were measured by hanging each rigid body at the mass center and
+assuming lateral symmetry. The mass centers were calculated from the
+orientation angle of the headtube and the location of the point of suspension
+relative to the wheel center. The wheel centers of mass were assumed to be in
+their ideal geometric center.
 
@@
 different orientation angles. The out-of-plane moments of inertia were
-measured by swinging the rigid bodies as a compound pendulum.
-\subsection*{Results}
+measured by swinging the rigid bodies as a compound pendulum. The periods were
+estimated from precision rate gyro measurements.
+\section*{Results}
 \begin{figure}[htbp]
     \begin{center}
-        \includegraphics[width=4in]{../../report/figures/bike_eig.png}
+        \includegraphics[width=3.5in]{../../report/figures/bike_eig.png}
     \end{center}
@@
 
-Parameter sets for all six bicycles and all eight
-configurations  will be presented, together with basic comparisons of the open loop dynamics of the
-different types of bicycles. The physical parameters, the mass, damping and stiffness
-matrices of the system, the eigenvalues and open loop transfer functions at
-different speeds will be compared with typical anecdotal
-descriptions of handling.
+Parameter sets with measurement uncertainties for all six bicycles and all
+eight configurations  will be presented, together with basic comparisons of the
+open loop dynamics of the different types of bicycles (eigenvalues,
+eigenvectors, bode plots, etc). Notable results include high accuracy
+measurements, dual oscillatory eigenmodes, bicycles without riders may often
+have no stable speed range, and it is possible to calculate the uncertainties
+in the eigenvalues and eigenvectors.
 

### 2010-09-08 BMD2010/PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
got it on two pages

@@
 parameters is used to describe the geometry, mass, mass location and
-mass distribution of each of the rigid bodies. These parameters are presented
-relative to a reference frame when the bicycle is in the nominal upright
-configuration. The experimental methods described herein are based primarily on the work
+mass distribution of each of the rigid bodies. The experimental methods
+described herein are based primarily on the work
 done in~\cite{Kooijman2006} but have been refined for improved accuracy and
-methodology. Koojiman's was preceded by \cite{Roland1971} who measured an experimental bicycle frame in a
+methodology. Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
 similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
@@
 measured with a precision scale. The centers of mass of each of the rigid
-bodies were measured by hanging each rigid body at the mass center and
-assuming lateral symmetry. The mass centers were calculated from the
+bodies were measured by hanging each rigid body at the mass center under the
+assumption of lateral symmetry. The mass centers were calculated from the
 orientation angle of the headtube and the location of the point of suspension
-relative to the wheel center. The wheel centers of mass were assumed to be in
-their ideal geometric center.
+relative to the wheel center. The wheel centers of mass were assumed to be
+aligned with the geometric center.
 
@@
 rigid bodies' periods of oscillation when hung as a torsional pendulum at
-different orientation angles. The out-of-plane moments of inertia were
+different orientation angles. The out-of-symmetric-plane moments of inertia were
 measured by swinging the rigid bodies as a compound pendulum. The periods were
@@
     \begin{center}
-        \includegraphics[width=3.5in]{../../report/figures/bike_eig.png}
+        \includegraphics[width=3.25in]{../../report/figures/bike_eig.pdf}
     \end{center}
@@
 open loop dynamics of the different types of bicycles (eigenvalues,
-eigenvectors, bode plots, etc). Notable results include high accuracy
+eigenvectors, bode plots, etc), Fig.~\ref{fig:bike_eig}. Notable results include high accuracy
 measurements, dual oscillatory eigenmodes, bicycles without riders may often

### 2010-09-08 bike_eig.py

@@ -36,7 +36,7 @@ vw = np.zeros(nBk)
 vc = np.zeros(nBk)
 
 # figure properties
-figwidth = 4. # in inches
+figwidth = 3.25 # in inches
 goldenMean = (np.sqrt(5)-1.0)/2.0
 figsize = [figwidth, figwidth*goldenMean]
 params = {#'backend': 'ps',
@@ -96,7 +96,7 @@ plt.title('Eigenvalues vs Speed')
 plt.xlabel('Speed [m/s]')
 plt.ylabel('Real and Imaginary Parts of the Eigenvalue [1/s]')
 try:
-    plt.savefig('report/figures/bike_eig.png')
+    plt.savefig('report/figures/bike_eig.pdf')
 except:
     pass
 # make a plot comparing the critical speeds of each bike

### 2010-09-10 BMD2010/PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
worked on hubbard's comment

@@
 {\fontsize{16pt}{20pt}\selectfont%
-  An Accurate Method of Measuring a Bicycle's Physical Parameters\\}
+  Accurate Measurement of Bicycle Physical Parameters\\}
 \end{flushleft}
@@
 are presented up through the eigenvalue prediction of the linear model and are
-based error propogation theory with correlations taken into account.
+based error propagation theory with correlations taken into account.
 
@@
 \emph{Batavus Browser}, a Dutch style city bicycle measured with and without
-instrumentation as described in~\cite{Kooijman2009a}, \emph{Batavus Stratos
-Deluxe}, a Dutch style sporty city bicycle, \emph{Batavus Crescendo Deluxe} a
+instrumentation as described in~\cite{Kooijman2009a}; \emph{Batavus Stratos
+Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
 Dutch style city bicycle with a suspended fork, \emph{Gary Fisher Mountain
-Bike}, a hardtail mountain bicycle, \emph{Bianchi Pista}, a modern steel frame
-track racing bicycle, \emph{Yellow Bicycle}, a stripped down aluminum frame
+Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
+track racing bicycle; \emph{Yellow Bicycle}, a stripped down aluminum frame
 road bicycle measured in two configurations, the second with the fork rotated
@@
 $w$, steer axis tilt $\lambda$, trail $c$, and wheel radii $r_\mathrm{F}$
-and $r_\mathrm{R}$. For each geometric measurement care was taken to measure
-the parameter directly as possible with high accuracy measurement equipment
-such as height gages, scales, etc. The mass of each of the rigid bodies were
-measured with a precision scale. The centers of mass of each of the rigid
-bodies were measured by hanging each rigid body at the mass center under the
-assumption of lateral symmetry. The mass centers were calculated from the
-orientation angle of the headtube and the location of the point of suspension
-relative to the wheel center. The wheel centers of mass were assumed to be
-aligned with the geometric center.
+and $r_\mathrm{R}$. For each geometric measurement, care was taken to measure
+the parameter directly as possible with high accuracy measurement equipment.
+The mass of each of the rigid bodies was
+measured with a precision scale (to so many grams). The center of mass of each rigid
+body was measured by hanging each it from several points. The mass
+center is the intersection of the lines of support under the
+assumption of lateral symmetry. Wheel centers of mass were assumed to be
+at their geometric center.
 
@@
 pendulum and compound pendulum configurations. Again the assumed symmetry of the
-bicycle was utilized to reduce the number of measurements needed. The
-in-symmetric-plane moments of inertia were calculated from measurements of the
-rigid bodies' periods of oscillation when hung as a torsional pendulum at
-different orientation angles. The out-of-symmetric-plane moments of inertia were
-measured by swinging the rigid bodies as a compound pendulum. The periods were
-estimated from precision rate gyro measurements.
+bicycle was utilized to reduce the number of measurements needed.
+In-symmetric-plane moments of inertia were calculated from measurements of the
+periods of oscillation when the rigid bodies were hung as a torsional pendulum about the vertical at
+different orientation angles. Out-of-symmetric-plane moments of inertia were
+measured by swinging the rigid bodies as a compound pendulum. Periods were
+estimated from zero crossings of precision rate gyro measurements.
 \section*{Results}
@@
 open loop dynamics of the different types of bicycles (eigenvalues,
-eigenvectors, bode plots, etc), Fig.~\ref{fig:bike_eig}. Notable results include high accuracy
+eigenvectors, Bode plots, etc.), Fig.~\ref{fig:bike_eig}. Notable results
+include: high accuracy
 measurements, dual oscillatory eigenmodes, bicycles without riders may often
-have no stable speed range, and it is possible to calculate the uncertainties
+have no stable speed range, and it is possible to calculate uncertainties
 in the eigenvalues and eigenvectors.

### 2010-10-08 make_par_tables.py
tables are working but still need to check rounding' '

@@ -56,10 +56,8 @@ for i, name in enumerate(data['bikes']):
     f.close()
     fn.close()
     os.system('pdflatex -output-directory=parTables ' + direct + fname + 'Par.tex')
-    os.system('rm parTables/*.aux')
-    os.system('rm parTables/*.log')
 
-# make the master table
+# make the master parameter table
 template = open('parTables/MasterParTableTemplate.tex', 'r')
 final = open('parTables/MasterParTable.tex', 'w')
 
@@ -83,5 +81,42 @@ for line in template:
 template.close()
 final.close()
 os.system('pdflatex -output-directory=parTables ' + direct + 'MasterParTable.tex')
+
+
+# make the master canonical matrix table
+template = open('parTables/MasterCanTableTemplate.tex', 'r')
+final = open('parTables/MasterCanTable.tex', 'w')
+
+abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
+for line in template:
+    if line[0] == '%':
+        varname, indice, fline = line[1:].split('%')
+        # remove the \n
+        fline = fline[:-1]
+        for i, bike in enumerate(data['bikes']):
+            fcan = open('data/bikeCanonical/' + ''.join(bike.split()) + 'Can.p')
+            can = pickle.load(fcan)
+            fcan.close()
+            if varname == 'bike':
+                fline = fline + ' & \multicolumn{2}{c}{' + abbrev[i] + '}'
+            else:
+                try:
+                    val, sig = uround(can[varname][indice[0], indice[1]]).split('+/-')
+                except:
+                    val = str(can[varname][indice[0], indice[1]])
+                    sig = '0.0'
+                fline = fline + ' & ' + val + ' & ' + sig
+        fline = fline + r'\\' + '\n'
+        final.write(fline)
+    else:
+        final.write(line)
+
+template.close()
+final.close()
+os.system('pdflatex -output-directory=parTables ' + direct + 'MasterCanTable.tex')
+
 os.system('rm parTables/*.aux')
 os.system('rm parTables/*.log')
+os.system('rm parTables/*.out')
+os.system('rm parTables/*.blg')
+os.system('rm parTables/*.bbl')

### 2010-10-08 PhysicalParameters.py
about to delete the originals

@@ -1,5 +1,5 @@
 import pickle
-from re
+import re
 from os import system, walk
 from copy import copy
 from numpy.linalg import inv, eig, lstsq
@@ -20,6 +20,7 @@ from uncertainties.umath import sin as usin
 from uncertainties.umath import cos as ucos
 from uncertainties.umath import tan as utan
 from scipy.optimize import leastsq
+from scipy.io import savemat
 from matplotlib.pyplot import figure, plot, xlabel, ylabel, title, legend, gca
 from matplotlib.pyplot import axes, xlim, setp, close, savefig, subplot, Circle
 from matplotlib.pyplot import axis, ylim, xticks, show
@@ -421,6 +422,90 @@ def calc_canon():
         f.close()
         # build some pretty tex files
         replace_values(directory, 'CanonTemplate.tex', fname + 'Can.tex', canon)
+
+def calc_jason_on_bikes():
+
+    # load in the base data file
+    f = open('data/data.p', 'r')
+    data = pickle.load(f)
+    f.close()
+
+    # load in the parameter file
+    f = open('data/par.p', 'r')
+    par = pickle.load(f)
+    f.close()
+
+    nBk = len(data['bikes'])
+
+    # remove the uncertainties
+    par_n = {}
+    for k, v in par.items():
+        if type(v[0]) == type(par['rF'][0]) or type(v[0]) == type(par['mF'][0]):
+            par_n[k] = nominal_values(v)
+        else:
+            par_n[k] = par[k]
+
+    # Jason's parameters (sitting on the Batavus Browser)
+    IBJ = array([[7.9985, 0 , -1.9272], [0, 8.0689, 0], [ -1.9272, 0, 2.3624]])
+    mBJ = 72.
+    xBJ = 0.2909
+    zBJ = -1.1091
+
+    # compute the total mass
+    mB = par_n['mB'] + mBJ
+    # compute the new CoM
+    xB = (mBJ*xBJ + par_n['mB']*par_n['xB'])/mB
+    zB = (mBJ*zBJ + par_n['mB']*par_n['zB'])/mB
+    # compute the new moment of inertia
+    dJ = vstack((xB, zeros(nBk), zB)) - vstack((xBJ, 0., zBJ))
+    dB = vstack((xB, zeros(nBk), zB)) - vstack((par_n['xB'], zeros(nBk), par_n['zB']))
+    IB = zeros((3, 3))
[…]

### 2010-10-08 PhysicalParameters.py
added in hunch angle and deleted some

@@ -2,12 +2,13 @@ import pickle
 import re
 from os import system, walk
 from copy import copy
-from numpy.linalg import inv, eig, lstsq
+from numpy.linalg import inv, eig, lstsq, norm
 from numpy import sin, cos, vstack, array, zeros, dot, diag
 from numpy import exp, sqrt, ones_like, sum, mean, pi, unwrap
 from numpy import argmin, abs, real, imag, zeros_like, max
 from numpy import hsplit, finfo, eye, hstack, log10, arctan2
 from numpy import poly1d, linspace, shape, ones, arange
+from numpy import arccos, rad2deg
 from uncertainties import ufloat
 from uncertainties.unumpy.ulinalg import inv as uinv
 from uncertainties.unumpy import nominal_values, std_devs, uarray
@@ -24,6 +25,36 @@ from scipy.io import savemat
 from matplotlib.pyplot import figure, plot, xlabel, ylabel, title, legend, gca
 from matplotlib.pyplot import axes, xlim, setp, close, savefig, subplot, Circle
 from matplotlib.pyplot import axis, ylim, xticks, show
+
+def hunch_angle():
+
+    def uvec(x):
+        xhat = x/norm(x)
+        return xhat
+    # Victor on the Stratos
+    # name the four points from the inkscape superimposed lines
+    a = array([190.716, 1052.483])
+    b = array([457.633, 533.479])
+    c = array([-312.107, -371.112])
+    d = array([645.204, -352.409])
+    # calculate the two vectors
+    v1 = a - b
+    v2 = c -d
+    # calculate the angle between the two vectors
+    theta = arccos(dot(uvec(v1), uvec(v2)))
+    print "Victor on the Stratos =", rad2deg(theta)
+    # Victor on the Browser
+    # name the four points from the inkscape superimposed lines
+    a = array([406.081, 989.949])
+    b = array([512.147, 453.558])
+    c = array([-334.360, -175.767])
+    d = array([705.086, -125.259])
+    # calculate the two vectors
+    v1 = a - b
+    v2 = c -d
+    # calculate the angle between the two vectors
+    theta = arccos(dot(uvec(v1), uvec(v2)))
+    print "Victor on the Browser =", rad2deg(theta)
 
 def fit_data():
 

### 2010-10-08 calc_JasonOnBikes.py

@@ -1,86 +1,0 @@
-import numpy as np
-import pickle as p
-from uncertainties import unumpy
-
-from benchmark_bike_tools import *
-
-# load in the base data file
-f = open('data/data.p', 'r')
-data = p.load(f)
-f.close()
-
-# load in the parameter file
-f = open('data/par.p', 'r')
-par = p.load(f)
-f.close()
-
-nBk = len(data['bikes'])
-
-# remove the uncertainties
-par_n = {}
-for k, v in par.items():
-    if type(v[0]) == type(par['rF'][0]) or type(v[0]) == type(par['mF'][0]):
-        par_n[k] = unumpy.nominal_values(v)
-    else:
-        par_n[k] = par[k]
-
-# Jason's parameters (sitting on the Batavus Browser)
-IBJ = np.array([[7.9985, 0 , -1.9272], [0, 8.0689, 0], [ -1.9272, 0, 2.3624]])
-mBJ = 72.
-xBJ = 0.2909
-zBJ = -1.1091
-
-# compute the total mass
-mB = par_n['mB'] + mBJ
-# compute the new CoM
-xB = (mBJ*xBJ + par_n['mB']*par_n['xB'])/mB
-zB = (mBJ*zBJ + par_n['mB']*par_n['zB'])/mB
-# compute the new moment of inertia
-dJ = np.vstack((xB, np.zeros(nBk), zB)) - np.vstack((xBJ, 0., zBJ))
-dB = np.vstack((xB, np.zeros(nBk), zB)) - np.vstack((par_n['xB'], np.zeros(nBk), par_n['zB']))
-IB = np.zeros((3, 3))
-for i in range(nBk):
-    IB[0] = np.array([par_n['IBxx'][i], 0., par_n['IBxz'][i]])
-    IB[1] = np.array([0., par_n['IByy'][i], 0.])
-    IB[2] = np.array([par_n['IBxz'][i], 0., par_n['IBzz'][i]])
-    I = parallel_axis(IBJ, mBJ, dJ[:, i]) + parallel_axis(IB, par_n['mB'][i], dB[:, i])
-    par_n['IBxx'][i] = I[0, 0]
-    par_n['IBxz'][i] = I[0, 2]
-    par_n['IByy'][i] = I[1, 1]
-    par_n['IBzz'][i] = I[2, 2]
-    par_n['mB'][i] = mB[i]
-    par_n['xB'][i] = xB[i]
[…]

### 2010-10-08 calc_parameters.py

@@ -1,256 +1,0 @@
-import os
-import re
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-import uncertainties as u
-import uncertainties.umath as umath
-import uncertainties.unumpy as unumpy
-from math import pi
-
-from benchmark_bike_tools import *
-
-# load the data file
-f = open('data/udata.p', 'r')
-ddU = pickle.load(f)
-f.close()
-
-# the number of different bikes
-nBk = len(ddU['bikes'])
-
-# make a list of the bikes' names
-bikeNames = ddU['bikes']
-
-# calculate all the benchmark parameters
-par = {}
-
-# calculate the wheel radii
-par['rR'] = ddU['rearWheelDist']/2./pi/ddU['rearWheelRot']
-par['rF'] = ddU['frontWheelDist']/2./pi/ddU['frontWheelRot']
-
-# steer axis tilt in radians
-par['lambda'] = pi/180.*(90. - ddU['headTubeAngle'])
-
-# calculate the front wheel trail
-forkOffset = ddU['forkOffset']
-par['c'] = (par['rF']*unumpy.sin(par['lambda'])
-              - forkOffset)/unumpy.cos(par['lambda'])
-
-# wheelbase
-par['w'] = ddU['wheelbase']
-
-# calculate the dees
-par['d1'] = unumpy.cos(par['lambda'])*(par['c']+par['w']-par['rR']*unumpy.tan(par['lambda']))
-par['d3'] = -unumpy.cos(par['lambda'])*(par['c']-par['rF']*unumpy.tan(par['lambda']))
-
-# calculate the frame rotation angle
-# alpha is the angle between the negative z pendulum (horizontal) and the
-# positive (up) steer axis, rotation about positive y
-alphaFrame = ddU['frameAngle']
-# beta is the angle between the x bike frame and the x pendulum frame, rotation
-# about positive y
-betaFrame = par['lambda'] - alphaFrame*pi/180
-
-# calculate the slope of the CoM line
-frameM = -unumpy.tan(betaFrame)
-
-# calculate the z-intercept of the CoM line
-# frameMassDist is positive according to the pendulum ref frame
-frameMassDist = ddU['frameMassDist']
-cb = unumpy.cos(betaFrame)
-frameB = -frameMassDist/cb - par['rR']
[…]

### 2010-10-08 fit_data.py

@@ -1,78 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import scipy.optimize as op
-import os
-import pickle as p
-from uncertainties import ufloat
-
-from benchmark_bike_tools import fit_goodness, jac_fitfunc, plot_osfit
-
-dirs, subdirs, filenames = list(os.walk('data/pendDat/p'))[0]
-file = open('data/period.txt', 'w')
-filenames.sort()
-period = {}
-#for name in ['YellowRevForkTorsionalFirst1.p']:
-#for name in ['YellowFwheelCompoundFirst1.p']:
-#for name in ['StratosFrameCompoundFirst2.p']:
-for name in filenames:
-    df = open('data/pendDat/p/' + name)
-    pendDat = p.load(df)
-    df.close()
-    y = pendDat['data'].ravel()
-    time = pendDat['duration']
-    x = np.linspace(0, time, num=len(y))
-    # decaying oscillating exponential function
-    fitfunc = lambda p, t: p[0] + np.exp(-p[3]*p[4]*t)*(p[1]*np.sin(p[4]*np.sqrt(1-p[3]**2)*t) + p[2]*np.cos(p[4]*np.sqrt(1-p[3]**2)*t))
-    # initial guesses
-    p0 = np.array([1.35, -.5, -.75, 0.01, 3.93])
-    # create the error function
-    errfunc = lambda p, t, y: fitfunc(p, t) - y
-    # minimize the error function
-    p1, success = op.leastsq(errfunc, p0[:], args=(x, y))
-    # plot the fitted curve
-    lscurve = fitfunc(p1, x)
-    rsq, SSE, SST, SSR = fit_goodness(y, lscurve)
-    sigma = np.sqrt(SSE/(len(y)-len(p0)))
-    # calculate the jacobian
-    L = jac_fitfunc(p1, x)
-    # the Hessian
-    H = np.dot(L.T, L)
-    # the covariance matrix
-    U = sigma**2.*np.linalg.inv(H)
-    # the standard deviations
-    sigp = np.sqrt(U.diagonal())
-    # frequency and period
-    wo = ufloat((p1[4], sigp[4]))
-    zeta = ufloat((p1[3], sigp[3]))
-    wd = (1. - zeta**2.)**(1./2.)*wo
-    f = wd/2./np.pi
-    T = 1./f
-    fig = plt.figure(1)
[…]

### 2010-10-08 tor_com.py

@@ -1,46 +1,0 @@
-import pickle as p
-import numpy as np
-import re
-from benchmark_bike_tools import space_out_camel_case
-
-f = open('data/period.p', 'r')
-period = p.load(f)
-f.close()
-
-tor = np.zeros((12, 8), dtype='object')
-com = np.zeros((4, 8), dtype='object')
-# list of bike names (only first letter is capatilized)
-bN = ['Browser', 'Browserins', 'Crescendo', 'Fisher', 'Pista', 'Stratos',
-        'Yellow', 'Yellowrev']
-# list of the orientation angles
-fst = ['First', 'Second', 'Third']
-# list of the bicycle parts
-fffr = ['Frame', 'Fork', 'Fwheel', 'Rwheel']
-# list of type of pendulums
-tc = ['Torsional', 'Compound']
-# average the periods
-for k, v in period.items():
-    # substitute names so the camel case function works
-    km = re.sub('BrowserIns', 'Browserins', k)
-    km = re.sub('YellowRev', 'Yellowrev', km)
-    desc = space_out_camel_case(km).split()
-    print desc
-    if desc[0] == 'Rod':
-        rodPeriod = np.mean(v)
-    else:
-        # if torsional, put it in the torsional matrix
-        if desc[2] == tc[0]:
-            r = fffr.index(desc[1])*3 + fst.index(desc[3])
-            c = bN.index(desc[0])
-            tor[r, c] = np.mean(v)
-            print 'Added torsional to [', r, ',', c, ']'
-        # if compound, put in in the compound matrix
-        elif desc[2] == tc[1]:
-            com[fffr.index(desc[1]), bN.index(desc[0])] = np.mean(v)
-avgPer = {}
-avgPer['tor'] = tor
-avgPer['com'] = com
-avgPer['rodPer'] = rodPeriod
-f = open('data/avgPer.p', 'w')
-p.dump(avgPer, f)
-f.close()

### 2010-10-08 PhysicalParameters.py
cleaned up the imports

@@ -1,64 +1,47 @@
 import pickle
 import re
-from os import system, walk
 from copy import copy
-from numpy.linalg import inv, eig, lstsq, norm
-from numpy import sin, cos, vstack, array, zeros, dot, diag
-from numpy import exp, sqrt, ones_like, sum, mean, pi, unwrap
-from numpy import argmin, abs, real, imag, zeros_like, max
-from numpy import hsplit, finfo, eye, hstack, log10, arctan2
-from numpy import poly1d, linspace, shape, ones, arange
-from numpy import arccos, rad2deg
-from uncertainties import ufloat
-from uncertainties.unumpy.ulinalg import inv as uinv
-from uncertainties.unumpy import nominal_values, std_devs, uarray
-from uncertainties.unumpy import matrix as umatrix
+from math import pi
+import numpy as np
+from uncertainties import ufloat, unumpy, umath
 from uncertainties.unumpy.core import wrap_array_func
-from uncertainties.unumpy import sin as unsin
-from uncertainties.unumpy import cos as uncos
-from uncertainties.unumpy import tan as untan
-from uncertainties.umath import sin as usin
-from uncertainties.umath import cos as ucos
-from uncertainties.umath import tan as utan
 from scipy.optimize import leastsq
 from scipy.io import savemat
-from matplotlib.pyplot import figure, plot, xlabel, ylabel, title, legend, gca
-from matplotlib.pyplot import axes, xlim, setp, close, savefig, subplot, Circle
-from matplotlib.pyplot import axis, ylim, xticks, show
+import matplotlib.pyplot as plt
 
 def hunch_angle():
 
     def uvec(x):
-        xhat = x/norm(x)
+        xhat = x/np.linalg.norm(x)
         return xhat
     # Victor on the Stratos
     # name the four points from the inkscape superimposed lines
-    a = array([190.716, 1052.483])
-    b = array([457.633, 533.479])
-    c = array([-312.107, -371.112])
-    d = array([645.204, -352.409])
+    a = np.array([190.716, 1052.483])
+    b = np.array([457.633, 533.479])
+    c = np.array([-312.107, -371.112])
+    d = np.array([645.204, -352.409])
     # calculate the two vectors
     v1 = a - b
     v2 = c -d
[…]

### 2010-10-08 calc_canon.py

@@ -1,32 +1,0 @@
-import pickle as p
-
-from benchmark_bike_tools import *
-
-# load in the base data file
-f = open('data/data.p', 'r')
-data = p.load(f)
-f.close()
-
-data['bikes'].append('Jodi Bike')
-
-# write the parameter files
-for i, name in enumerate(data['bikes']):
-    directory = 'data/bikeParameters/'
-    fname = ''.join(name.split())
-    M, C1, K0, K2, param = bmp2cm(directory + fname + 'Par.txt')
-    A, B = abMatrix(M, C1, K0, K2, param['v'], param['g'])
-    directory = 'data/bikeCanonical/'
-    file = open(directory + fname + 'Can.txt', 'w')
-    for mat in ['M','C1', 'K0', 'K2', 'A', 'B']:
-        if mat == 'A' or mat == 'B':
-            file.write(mat + ' (v = ' + str(param['v']) + ')\n')
-        else:
-            file.write(mat + '\n')
-        file.write(str(eval(mat)) + '\n')
-    file.close()
-    file = open(directory + fname + 'Can.p', 'w')
-    canon = {'M':M, 'C1':C1, 'K0':K0, 'K2':K2, 'A':A, 'B':B, 'v':param['v']}
-    p.dump(canon, file)
-    file.close()
-    # build some pretty tex files
-    replace_values(directory, 'CanonTemplate.tex', fname + 'Can.tex', canon)

### 2010-10-08 test_BicycleParameters.py

@@ -1,6 +1,0 @@
-import BicycleParameters as BP
-
-# see what happens when you don't have a directory with data
-booger = BP.bicycle('booger')
-
-Browser = BP.bicycle('Browser')

### 2010-10-08 PhysicalParameters.py
bike eig plots works

@@ -9,6 +9,127 @@ from scipy.optimize import leastsq
 from scipy.io import savemat
 import matplotlib.pyplot as plt
 
+def bike_eig_plots():
+
+    # choose the bike alone, bike with rigid legs or bike with rider
+    # bike alone
+    direct = 'data/bikeCanonical/'
+    fend = 'Can.p'
+    # bike with rider's legs
+    #direct = 'data/bikeLegsCanonical/'
+    #fend = 'LegsCan.p'
+    # bike with rider
+    #direct = 'data/bikeRiderCanonical/'
+    #fend = 'RiderCan.p'
+
+
+    # load in the base data file
+    f = open('data/data.p', 'r')
+    data = pickle.load(f)
+    f.close()
+
+    # add Jodi's thesis bike for comparison (only for bikeCanonical)
+    #data['bikes'].append('Jodi Bike')
+
+    nBk = len(data['bikes']) # number of bikes
+
+    # one color for each bike
+    colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange', 'red']
+
+    vd = np.zeros(nBk)
+    vw = np.zeros(nBk)
+    vc = np.zeros(nBk)
+
+    # figure properties
+    figwidth = 3.25 # in inches
+    goldenMean = (np.sqrt(5)-1.0)/2.0
+    figsize = [figwidth, figwidth*goldenMean]
+    params = {#'backend': 'ps',
+        #'axes.labelsize': 10,
+        #'text.fontsize': 10,
+        'legend.fontsize': 10,
+        #'xtick.labelsize': 8,
+        #'ytick.labelsize': 8,
+        #'text.usetex': True,
+        #'figure.figsize': figsize
+        }
+    plt.rcParams.update(params)
+    eigFig = plt.figure(num=nBk + 1)#, figsize=figsize)
+    #plt.clf()
+    #plt.axes([0.125,0.2,0.95-0.125,0.95-0.2])
+
+    # caluculate eigenvalues/vectors over this range
+    vel = np.linspace(0, 10, num=1000)
+
+    # color scale for root loci plot
+
+
+    for i, name in enumerate(data['bikes']):
+        fname = ''.join(name.split()) + fend
+        f = open(direct + fname)
+        can = pickle.load(f)
+        f.close()
+        for k, v in can.items():
+            can[k] = unumpy.nominal_values(v)
[…]

### 2010-10-08 PhysicalParameters.py
merged bike_bode

@@ -9,6 +9,47 @@ from scipy.optimize import leastsq
 from scipy.io import savemat
 import matplotlib.pyplot as plt
 
+def bike_bode_plots():
+
+    # load in the base data file
+    f = open('data/data.p', 'r')
+    data = pickle.load(f)
+    f.close()
+
+    nBk = len(data['bikes'])
+
+    colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange']
+
+    Tdel2phi = plt.figure(num=1)
+    Tdel2del = plt.figure(num=2)
+    Tphi2phi = plt.figure(num=3)
+    Tphi2del = plt.figure(num=4)
+
+    dir = 'data/bikeRiderCanonical/'
+    for i, name in enumerate(data['bikes']):
+        fname = ''.join(name.split()) + 'RiderCan.p'
+        file = open(dir + fname)
+        can = pickle.load(file)
+        file.close()
+        # make some bode plots
+        A, B = abMatrix(can['M'], can['C1'], can['K0'], can['K2'], 1., 9.81)
+        # y is [phidot, deldot, phi, del]
+        C = np.eye(A.shape[0])
+        freq = np.logspace(0, 2, 5000)
+        plt.figure(1)
+        bode(ABCD=(A, B[:, 1], C[2], 0.), w=freq, fig=Tdel2phi)
+        plt.figure(2)
+        bode(ABCD=(A, B[:, 1], C[3], 0.), w=freq, fig=Tdel2del)
+        plt.figure(3)
+        bode(ABCD=(A, B[:, 0], C[2], 0.), w=freq, fig=Tphi2phi)
+        plt.figure(4)
+        bode(ABCD=(A, B[:, 0], C[3], 0.), w=freq, fig=Tphi2del)
+    #for i, line in enumerate(Tdel2phi.ax1.lines):
+    #    plt.setp(line, color=colors[i])
+    #    plt.setp(Tdel2phi.ax2.lines[i], color=colors[i])
+    # plot the bike names on the eigenvalue plot
+    plt.show()
+
 def bike_eig_plots():
 
     # choose the bike alone, bike with rigid legs or bike with rider

### 2010-10-08 bike_eig.py

@@ -1,126 +1,0 @@
-import pickle
-import numpy as np
-import matplotlib.pyplot as plt
-from uncertainties.unumpy import nominal_values as nom
-from uncertainties.unumpy import std_devs
-
-from benchmark_bike_tools import *
-
-# choose the bike alone, bike with rigid legs or bike with rider
-# bike alone
-direct = 'data/bikeCanonical/'
-fend = 'Can.p'
-# bike with rider's legs
-#direct = 'data/bikeLegsCanonical/'
-#fend = 'LegsCan.p'
-# bike with rider
-#direct = 'data/bikeRiderCanonical/'
-#fend = 'RiderCan.p'
-
-
-# load in the base data file
-f = open('data/data.p', 'r')
-data = pickle.load(f)
-f.close()
-
-# add Jodi's thesis bike for comparison (only for bikeCanonical)
-#data['bikes'].append('Jodi Bike')
-
-nBk = len(data['bikes']) # number of bikes
-
-# one color for each bike
-colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange', 'red']
-
-vd = np.zeros(nBk)
-vw = np.zeros(nBk)
-vc = np.zeros(nBk)
-
-# figure properties
-figwidth = 3.25 # in inches
-goldenMean = (np.sqrt(5)-1.0)/2.0
-figsize = [figwidth, figwidth*goldenMean]
-params = {#'backend': 'ps',
-    #'axes.labelsize': 10,
-    #'text.fontsize': 10,
-    'legend.fontsize': 10,
-    #'xtick.labelsize': 8,
-    #'ytick.labelsize': 8,
-    #'text.usetex': True,
-    #'figure.figsize': figsize
-    }
-plt.rcParams.update(params)
-eigFig = plt.figure(num=nBk + 1)#, figsize=figsize)
-#plt.clf()
-#plt.axes([0.125,0.2,0.95-0.125,0.95-0.2])
-
-# caluculate eigenvalues/vectors over this range
-vel = np.linspace(0, 10, num=1000)
-
-# color scale for root loci plot
-
-
-for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split()) + fend
-    f = open(direct + fname)
-    can = pickle.load(f)
-    f.close()
-    for k, v in can.items():
-        can[k] = nom(v)
[…]

### 2010-10-08 PhysicalParameters.py
merged make_par_tables

@@ -1,5 +1,6 @@
 import pickle
 import re
+import os
 from copy import copy
 from math import pi
 import numpy as np
@@ -8,6 +9,124 @@ from uncertainties.unumpy.core import wrap_array_func
 from scipy.optimize import leastsq
 from scipy.io import savemat
 import matplotlib.pyplot as plt
+
+def make_tables():
+
+    # load in the base data file
+    f = open('data/data.p', 'r')
+    data = pickle.load(f)
+    f.close()
+
+    # load in the parameter data file
+    f = open('data/par.p', 'r')
+    par = pickle.load(f)
+    f.close()
+
+    direct = 'parTables/'
+    for i, name in enumerate(data['bikes']):
+        fname = ''.join(name.split())
+        # open the new file
+        f = open(direct + 'ParameterTable.tex', 'r')
+        fn = open(direct + fname + 'Par.tex', 'w')
+        for line in f:
+            #print line
+            # find all of the matches in the line
+            test = re.findall('\|(\w*.)\|', line)
+            #print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
+            #print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
+            # if there are matches
+            if test:
+                #print 'Found this!\n', test
+                # go through each match and make a substitution
+                for match in test:
+                    #print "replace this: ", match
+                    # if the match has a question mark it is an uncertainty value
+                    if match[-1] == '?':
+                        try:
+                            line = re.sub('\|(\w*\?)\|',
+                                    uround(par[match[:-1]][i]).split('+/-')[1], line, count=1)
+                            #print line
+                        except:
+                            pass
+                    # else if the match is the bicycle name
+                    elif match == 'bikename':
+                        line = re.sub('\|bikename\|', name.upper(), line)
+                    else:
+                        try:
+                            line = re.sub('\|(\w*(?!\?))\|',
+                                    uround(par[match][i]).split('+/-')[0], line, count=1)
+                            #print line
+                        except:
+                            line = re.sub('\|(\w*(?!\?))\|', str(par[match][i]), line, count=1)
+            #print line
+            fn.write(line)
+        f.close()
+        fn.close()
[…]

### 2010-10-08 bike_bode.py

@@ -1,44 +1,0 @@
-import numpy as np
-import pickle
-import matplotlib.pyplot as plt
-
-from benchmark_bike_tools import *
-
-# load in the base data file
-f = open('data/data.p', 'r')
-data = pickle.load(f)
-f.close()
-
-nBk = len(data['bikes'])
-
-colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange']
-
-Tdel2phi = plt.figure(num=1)
-Tdel2del = plt.figure(num=2)
-Tphi2phi = plt.figure(num=3)
-Tphi2del = plt.figure(num=4)
-
-dir = 'data/bikeRiderCanonical/'
-for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split()) + 'RiderCan.p'
-    file = open(dir + fname)
-    can = pickle.load(file)
-    file.close()
-    # make some bode plots
-    A, B = abMatrix(can['M'], can['C1'], can['K0'], can['K2'], 1., 9.81)
-    # y is [phidot, deldot, phi, del]
-    C = np.eye(A.shape[0])
-    freq = np.logspace(0, 2, 5000)
-    plt.figure(1)
-    bode(ABCD=(A, B[:, 1], C[2], 0.), w=freq, fig=Tdel2phi)
-    plt.figure(2)
-    bode(ABCD=(A, B[:, 1], C[3], 0.), w=freq, fig=Tdel2del)
-    plt.figure(3)
-    bode(ABCD=(A, B[:, 0], C[2], 0.), w=freq, fig=Tphi2phi)
-    plt.figure(4)
-    bode(ABCD=(A, B[:, 0], C[3], 0.), w=freq, fig=Tphi2del)
-#for i, line in enumerate(Tdel2phi.ax1.lines):
-#    plt.setp(line, color=colors[i])
-#    plt.setp(Tdel2phi.ax2.lines[i], color=colors[i])
-# plot the bike names on the eigenvalue plot
-plt.show()

### 2010-10-08 make_par_tables.py

@@ -1,122 +1,0 @@
-import pickle
-import re
-import os
-import uncertainties as un
-
-from benchmark_bike_tools import uround
-
-# load in the base data file
-f = open('data/data.p', 'r')
-data = pickle.load(f)
-f.close()
-
-# load in the parameter data file
-f = open('data/par.p', 'r')
-par = pickle.load(f)
-f.close()
-
-direct = 'parTables/'
-for i, name in enumerate(data['bikes']):
-    fname = ''.join(name.split())
-    # open the new file
-    f = open(direct + 'ParameterTable.tex', 'r')
-    fn = open(direct + fname + 'Par.tex', 'w')
-    for line in f:
-        #print line
-        # find all of the matches in the line
-        test = re.findall('\|(\w*.)\|', line)
-        #print 'search for u: ',  re.findall('\|(\w*\?)\|', line)
-        #print 'search for nom: ', re.findall('\|(\w*)(?!\?)\|', line)
-        # if there are matches
-        if test:
-            #print 'Found this!\n', test
-            # go through each match and make a substitution
-            for match in test:
-                #print "replace this: ", match
-                # if the match has a question mark it is an uncertainty value
-                if match[-1] == '?':
-                    try:
-                        line = re.sub('\|(\w*\?)\|',
-                                uround(par[match[:-1]][i]).split('+/-')[1], line, count=1)
-                        #print line
-                    except:
-                        pass
-                # else if the match is the bicycle name
-                elif match == 'bikename':
-                    line = re.sub('\|bikename\|', name.upper(), line)
-                else:
-                    try:
-                        line = re.sub('\|(\w*(?!\?))\|',
-                                uround(par[match][i]).split('+/-')[0], line, count=1)
-                        #print line
-                    except:
-                        line = re.sub('\|(\w*(?!\?))\|', str(par[match][i]), line, count=1)
-        #print line
-        fn.write(line)
-    f.close()
-    fn.close()
-    os.system('pdflatex -output-directory=parTables ' + direct + fname + 'Par.tex')
-
-# make the master parameter table
-template = open('parTables/MasterParTableTemplate.tex', 'r')
-final = open('parTables/MasterParTable.tex', 'w')
-
-abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
-for line in template:
[…]

### 2010-10-08 benchmark_bike_tools.py
removed the orginal module

@@ -1,881 +1,0 @@
-from numpy.linalg import inv
-from numpy import dot
-
-def ab_to_mck(A, B):
-    '''
-    Returns the spring-mass-damper matrices for a general system.
-
-    Parameters:
-    -----------
-    A : ndarray, shape(n,n)
-    B : ndarray, shape(n/2,n)
-
-    Returns:
-    --------
-    M : ndarray, shape(n/2,n/2)
-    C : ndarray, shape(n/2,n/2)
-    K : ndarray, shape(n/2/n/2)
-
-    This function converts the linear set of equations x' = Ax + Bu to the
-    canonical form Mq'' + Cq' + Kq = f. The states, x, must be ordered
-    sequentially with the configuration variables proceeding the rates (i.e.
-    x = [x1, ..., xn, x1', ..., xn'] and q = [x1, ..., xn]).
-
-    '''
-    numRow, numCol = A.shape
-    M = inv(B[numRow/2:, :])
-    C = -dot(M, A[numRow/2:, numCol/2:])
-    K = -dot(M, A[numRow/2:, :numCol/2])
-
-    return M, C, K
-
-def ueig2(uA):
-    from uncertainties.unumpy.core import wrap_array_func
-    import numpy
-
-    ueig0 = wrap_array_func(lambda m: numpy.linalg.eig(m)[0])
-    ueig1 = wrap_array_func(lambda m: numpy.linalg.eig(m)[1])
-
-    def eig(m):
-        """
-        Version of numpy.linalg.eig that works on matrices that contain
-        numbers with uncertainties.
-        """
-        return (ueig0(m), ueig1(m))
-
-    return eig(uA)
-
-def ueig(uA):
-    '''
-    Returns eigenvalues and eigenvectors with uncertainties.
-
-    Parameters:
-    -----------
-    uA : ndarry, shape(n x n)
-
-    Returns:
-    --------
-    uw : ndarry, shape = (n,)
-    uv : ndarray, shape = (n, n)
-
-    uses both numpy.linalg.eig and the uncertainties package to calculate eigen
-    values and eigenvectors with uncertainties
-
-    '''
-    from copy import copy
-    from uncertainties import ufloat
-    from uncertainties.unumpy import nominal_values, std_devs, uarray
-    from numpy.linalg import eig
-    from numpy import zeros, diag, zeros_like, hsplit, dot, vstack, finfo, sqrt
-
-    # separate the nominal values from the uncertainties
[…]

### 2010-10-08 PhysicalParameters.py
eee

@@ -22,7 +22,7 @@ def make_tables():
     par = pickle.load(f)
     f.close()
 
-    direct = 'parTables/'
+    direct = 'tables/'
     for i, name in enumerate(data['bikes']):
         fname = ''.join(name.split())
         # open the new file
@@ -62,11 +62,11 @@ def make_tables():
             fn.write(line)
         f.close()
         fn.close()
-        os.system('pdflatex -output-directory=parTables ' + direct + fname + 'Par.tex')
+        os.system('pdflatex -output-directory=tables ' + direct + fname + 'Par.tex')
 
     # make the master parameter table
-    template = open('parTables/MasterParTableTemplate.tex', 'r')
-    final = open('parTables/MasterParTable.tex', 'w')
+    template = open('tables/MasterParTableTemplate.tex', 'r')
+    final = open('tables/MasterParTable.tex', 'w')
 
     abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
     for line in template:
@@ -87,12 +87,12 @@ def make_tables():
 
     template.close()
     final.close()
-    os.system('pdflatex -output-directory=parTables ' + direct + 'MasterParTable.tex')
+    os.system('pdflatex -output-directory=tables ' + direct + 'MasterParTable.tex')
 
 
     # make the master canonical matrix table
-    template = open('parTables/MasterCanTableTemplate.tex', 'r')
-    final = open('parTables/MasterCanTable.tex', 'w')
+    template = open('tables/MasterCanTableTemplate.tex', 'r')
+    final = open('tables/MasterCanTable.tex', 'w')
 
     abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
     for line in template:
@@ -120,13 +120,13 @@ def make_tables():
 
     template.close()
     final.close()
-    os.system('pdflatex -output-directory=parTables ' + direct + 'MasterCanTable.tex')
+    os.system('pdflatex -output-directory=tables ' + direct + 'MasterCanTable.tex')
 
-    os.system('rm parTables/*.aux')
-    os.system('rm parTables/*.log')
-    os.system('rm parTables/*.out')
-    os.system('rm parTables/*.blg')
-    os.system('rm parTables/*.bbl')
+    os.system('rm tables/*.aux')
+    os.system('rm tables/*.log')
+    os.system('rm tables/*.out')
+    os.system('rm tables/*.blg')
+    os.system('rm tables/*.bbl')
 
 def bike_bode_plots():
 

### 2010-10-08 physical_parameters.py

@@ -1,19 +1,0 @@
-import tarfile
-import os
-
-# uncompress the data file
-tar = tarfile.open('data.tar.gz')
-tar.extractall()
-tar.close()
-
-# create data directory and move the pendat folder inside
-os.system('mkdir data')
-os.system('mv pendDat data/pendDat')
-os.system('mv data.mat data/data.mat')
-os.system('mv MeasUncert.txt data/MeasUncert.txt')
-
-# run the scripts to process the data
-os.system('python fit_data.py')
-os.system('python tor_com.py')
-os.system('python calc_parameters.py')
-os.system('python calc_canon.py')

### 2010-10-08 PhysicalParameters.py
merged some more

@@ -8,7 +8,50 @@ from uncertainties import ufloat, unumpy, umath
 from uncertainties.unumpy.core import wrap_array_func
 from scipy.optimize import leastsq
 from scipy.io import savemat
+import scipy.io.matlab.mio as mio
 import matplotlib.pyplot as plt
+from mpl_toolkits.mplot3d import Axes3D
+
+def plot_raw_data():
+    # make a list of the folder contents
+    dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
+    print filenames
+    i = 0
+    while i < len(filenames):
+        d = {}
+        mio.loadmat('data/pendDat/' + filenames[i], mdict=d)
+        print 'Plotting', filenames[i]
+        plt.plot(d['data'])
+        plt.title(filenames[i])
+        plt.show()
+        raw_input()
+        plt.close()
+        i += 1
+
+def plot_evecs():
+
+    f = open('data/bikeRiderCanonical/BianchiPistaRiderCan.p')
+    can = pickle.load(f)
+    f.close()
+
+    vel = np.linspace(0, 20, num=500)
+    evals, evecs = bike_eig(can['M'], can['C1'], can['K0'], can['K2'], vel, 9.81)
+    wea, cap, cas = sort_modes(evals, evecs)
+
+    colors = ['blue', 'red', 'green', 'orange']
+
+    fig = plt.figure()
+    ax = Axes3D(fig)
+    for i, row in enumerate(wea['evecs'][:, :, 1]):
+        for k, component in enumerate(row[:2]):
+            point1 = np.array([0., 0., vel[i]])
+            point2 = np.array([np.abs(np.real(component)),
+                np.abs(np.imag(component)), vel[i]])
+            ax.plot(np.hstack((point1[0], point2[0])), np.hstack((point1[1],
+                point2[1])), zs=vel[i], color=colors[k])
+                #ax.plot(np.abs(np.real(wea['evecs'][:, :, 1])),
+                    #np.abs(np.imag(wea['evecs'][:, :, 1])), zs=vel)
+    plt.show()
 
 def make_tables():
 

### 2010-10-08 plot_eigvec.py

@@ -1,29 +1,0 @@
-from mpl_toolkits.mplot3d import Axes3D
-import matplotlib.pyplot as plt
-import numpy as np
-import pickle
-
-from benchmark_bike_tools import bike_eig, sort_modes
-
-f = open('data/bikeRiderCanonical/BianchiPistaRiderCan.p')
-can = pickle.load(f)
-f.close()
-
-vel = np.linspace(0, 20, num=500)
-evals, evecs = bike_eig(can['M'], can['C1'], can['K0'], can['K2'], vel, 9.81)
-wea, cap, cas = sort_modes(evals, evecs)
-
-colors = ['blue', 'red', 'green', 'orange']
-
-fig = plt.figure()
-ax = Axes3D(fig)
-for i, row in enumerate(wea['evecs'][:, :, 1]):
-    for k, component in enumerate(row[:2]):
-        point1 = np.array([0., 0., vel[i]])
-        point2 = np.array([np.abs(np.real(component)),
-            np.abs(np.imag(component)), vel[i]])
-        ax.plot(np.hstack((point1[0], point2[0])), np.hstack((point1[1],
-            point2[1])), zs=vel[i], color=colors[k])
-            #ax.plot(np.abs(np.real(wea['evecs'][:, :, 1])),
-                #np.abs(np.imag(wea['evecs'][:, :, 1])), zs=vel)
-plt.show()

### 2010-10-08 PhysicalParameters.py
mat2dic

@@ -12,21 +12,105 @@ import scipy.io.matlab.mio as mio
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 
-def plot_raw_data():
-    # make a list of the folder contents
+def mat2dic():
+
+    # load the main data file into a dictionary
+    d = {}
+    mio.loadmat('data/data.mat', mdict=d)
+
+    # make a list of the bikes' names
+    bikeNames = []
+    for bike in d['bikes']:
+        # get rid of the weird matlab unicoding
+        bikeNames.append(bike[0][0].encode('ascii'))
+
+    # clean up the matlab imports
+    d['bikes'] = bikeNames
+    del(d['__globals__'], d['__header__'], d['__version__'])
+    for k, v in d.items():
+        if np.shape(v)[0] == 1:
+            d[k] = v[0]
+
+    # make a dictionary for the measurement standard deviations
+    dU = {}
+    f = open('data/MeasUncert.txt', 'r')
+    for line in f:
+        l = line.split(',')
+        dU[l[0]] = eval(l[1])
+    f.close()
+
+    # add the uncertainties to the data
+    ddU = {}
+    for k, v in dU.items():
+        for pair in zip(d[k].flatten(), np.ones_like(d[k].flatten())*v):
+            if k in ddU.keys():
+                ddU[k].append(ufloat((float(pair[0]), pair[1])))
+            else:
+                ddU[k] = []
+                ddU[k].append(ufloat((float(pair[0]), pair[1])))
+        ddU[k] = np.array(ddU[k])
+        if ddU[k].shape[0] > 8:
+            ddU[k] = ddU[k].reshape((ddU[k].shape[0]/8, -1))
+
+    ddU['bikes'] = d['bikes']
+
+    # bring in the calibration rod data
+    f = open('data/CalibrationRod.txt', 'r')
+    for line in f:
+        var, val, unc = line.split(',')
+        d[var] = float(val)
+        ddU[var] = ufloat((float(val), float(unc)))
+    f.close()
+
+    # pickle the data dictionary
+    f = open('data/data.p', 'w')
+    pickle.dump(d, f)
+    f.close()
+
+    # pickle the data with the uncertainties
+    f = open('data/udata.p', 'w')
+    pickle.dump(ddU, f)
+    f.close()
+
     dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
-    print filenames
-    i = 0
[…]

### 2010-10-08 mat2dic.py

@@ -1,88 +1,0 @@
-import pickle
-import scipy.io.matlab.mio as mio
-import uncertainties as u
-import numpy as np
-import os
-
-# load the main data file into a dictionary
-d = {}
-mio.loadmat('data/data.mat', mdict=d)
-
-# make a list of the bikes' names
-bikeNames = []
-for bike in d['bikes']:
-    # get rid of the weird matlab unicoding
-    bikeNames.append(bike[0][0].encode('ascii'))
-
-# clean up the matlab imports
-d['bikes'] = bikeNames
-del(d['__globals__'], d['__header__'], d['__version__'])
-for k, v in d.items():
-    if np.shape(v)[0] == 1:
-        d[k] = v[0]
-
-# make a dictionary for the measurement standard deviations
-dU = {}
-f = open('data/MeasUncert.txt', 'r')
-for line in f:
-    l = line.split(',')
-    dU[l[0]] = eval(l[1])
-f.close()
-
-# add the uncertainties to the data
-ddU = {}
-for k, v in dU.items():
-    for pair in zip(d[k].flatten(), np.ones_like(d[k].flatten())*v):
-        if k in ddU.keys():
-            ddU[k].append(u.ufloat((float(pair[0]), pair[1])))
-        else:
-            ddU[k] = []
-            ddU[k].append(u.ufloat((float(pair[0]), pair[1])))
-    ddU[k] = np.array(ddU[k])
-    if ddU[k].shape[0] > 8:
-        ddU[k] = ddU[k].reshape((ddU[k].shape[0]/8, -1))
-
-ddU['bikes'] = d['bikes']
-
-# bring in the calibration rod data
-f = open('data/CalibrationRod.txt', 'r')
-for line in f:
-    var, val, unc = line.split(',')
-    d[var] = float(val)
-    ddU[var] = u.ufloat((float(val), float(unc)))
-f.close()
-
-# pickle the data dictionary
-f = open('data/data.p', 'w')
-pickle.dump(d, f)
-f.close()
-
-# pickle the data with the uncertainties
-f = open('data/udata.p', 'w')
-pickle.dump(ddU, f)
-f.close()
-
-dirs, subdirs, filenames = list(os.walk('data/pendDat'))[0]
-filenames.sort()
-for name in filenames:
-    pendDat = {}
-    mio.loadmat('data/pendDat/' + name, mdict=pendDat)
-    # clean up the matlab imports
[…]

### 2010-10-08 papers/BMD2010/BikeTaskComparison/MooreHessHubbardPeterson_abstract.tex
cleaning house

--- 2010-05-06 BikeTaskComparison/MooreHessHubbardPeterson_abstract.tex
@@
 Fig.~\ref{fig:block} (heading and lateral deviation). In addition, a simple model of rider preview
-is included. Figure~\ref{fig:steer} shows the rider steering inputs for the
+is included~\cite{Hess2003a}. Figure~\ref{fig:steer} shows the rider steering inputs for the
 task. Figure~\ref{fig:handling} shows the ``Handling Qualities Sensitivity
@@
 the rider model follows the dictates of the well-known ``crossover model'' of the
-human operator~\cite{McRuer1974}, appropriate for the
-$\frac{\phi}{T_\delta}$ dynamics of the
+human operator~\cite{McRuer1974}, appropriate for the dynamics of the
 bicycles in question. The final paper will include nine bicycle models and
@@
     \centering
-    \subfloat[Lane Change Performance]{\label{fig:path}\includegraphics[width=0.425\columnwidth]{path.png}}
-    \subfloat[Handling Qualities Sensitivity Functions]{\label{fig:handling}\includegraphics[width=0.425\columnwidth]{handling.png}}
+    \subfloat[Lane Change Performance]{\label{fig:path}\includegraphics[width=0.38\columnwidth]{path.png}}
+    \subfloat[Handling Qualities Sensitivity Functions]{\label{fig:handling}\includegraphics[width=0.38\columnwidth]{handling.png}}
 
-    \subfloat[Rider Steering Inputs]{\label{fig:steer}\includegraphics[width=0.425\columnwidth]{steer.png}}
+    \subfloat[Rider Steering Inputs]{\label{fig:steer}\includegraphics[width=0.38\columnwidth]{steer.png}}
     \subfloat[Bode Diagrams of Rider/Bicycle Open-Loop Transfer
-    Functions]{\label{fig:bode}\includegraphics[width=0.425\columnwidth]{bode.png}}
+    Functions]{\label{fig:bode}\includegraphics[width=0.38\columnwidth]{bode.png}}
     \caption{Linear simulation results and comparative metrics for six bicycles.}

### 2010-10-08 PhysicalParameters.py
merged it all

@@ -12,6 +12,100 @@ import scipy.io.matlab.mio as mio
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 
+def calc_legs_on_bikes():
+
+    # load in the base data file
+    f = open('data/data.p', 'r')
+    data = pickle.load(f)
+    f.close()
+
+    # load in the parameter file
+    f = open('data/par.p', 'r')
+    par = pickle.load(f)
+    f.close()
+
+    nBk = len(data['bikes'])
+
+    # remove the uncertainties
+    par_n = {}
+    for k, v in par.items():
+        if type(v[0]) == type(par['rF'][0]) or type(v[0]) == type(par['mF'][0]):
+            par_n[k] = unumpy.nominal_values(v)
+        else:
+            par_n[k] = par[k]
+
+    # Jason's leg parameters (sitting on the Batavus Browser)
+    IBJ = np.array([[ 1.371635407777139, 0. ,               -0.355678341037248],
+                    [ 0.,                1.248341633510700,  0.               ],
+                    [-0.355678341037248, 0.,                 0.721984553624527]])
+    mBJ = 23.183999999999997
+    xBJ = 0.41651301147631
+    zBJ = -0.742728766690162
+
+    # compute the total mass
+    mB = par_n['mB'] + mBJ
+    # compute the new CoM
+    xB = (mBJ*xBJ + par_n['mB']*par_n['xB'])/mB
+    zB = (mBJ*zBJ + par_n['mB']*par_n['zB'])/mB
+    # compute the new moment of inertia
+    dJ = np.vstack((xB, np.zeros(nBk), zB)) - np.vstack((xBJ, 0., zBJ))
+    dB = np.vstack((xB, np.zeros(nBk), zB)) - np.vstack((par_n['xB'], np.zeros(nBk), par_n['zB']))
+    IB = np.zeros((3, 3))
+    for i in range(nBk):
+        IB[0] = np.array([par_n['IBxx'][i], 0., par_n['IBxz'][i]])
+        IB[1] = np.array([0., par_n['IByy'][i], 0.])
+        IB[2] = np.array([par_n['IBxz'][i], 0., par_n['IBzz'][i]])
+        I = parallel_axis(IBJ, mBJ, dJ[:, i]) + parallel_axis(IB, par_n['mB'][i], dB[:, i])
+        par_n['IBxx'][i] = I[0, 0]
[…]

### 2010-10-08 calc_JasonsLegsOnBikes.py

@@ -1,99 +1,0 @@
-import numpy as np
-import pickle as p
-from uncertainties import unumpy
-import os
-from scipy.io.matlab.mio import savemat
-
-from benchmark_bike_tools import *
-
-# load in the base data file
-f = open('data/data.p', 'r')
-data = p.load(f)
-f.close()
-
-# load in the parameter file
-f = open('data/par.p', 'r')
-par = p.load(f)
-f.close()
-
-nBk = len(data['bikes'])
-
-# remove the uncertainties
-par_n = {}
-for k, v in par.items():
-    if type(v[0]) == type(par['rF'][0]) or type(v[0]) == type(par['mF'][0]):
-        par_n[k] = unumpy.nominal_values(v)
-    else:
-        par_n[k] = par[k]
-
-# Jason's leg parameters (sitting on the Batavus Browser)
-IBJ = np.array([[ 1.371635407777139, 0. ,               -0.355678341037248],
-                [ 0.,                1.248341633510700,  0.               ],
-                [-0.355678341037248, 0.,                 0.721984553624527]])
-mBJ = 23.183999999999997
-xBJ = 0.41651301147631
-zBJ = -0.742728766690162
-
-# compute the total mass
-mB = par_n['mB'] + mBJ
-# compute the new CoM
-xB = (mBJ*xBJ + par_n['mB']*par_n['xB'])/mB
-zB = (mBJ*zBJ + par_n['mB']*par_n['zB'])/mB
-# compute the new moment of inertia
-dJ = np.vstack((xB, np.zeros(nBk), zB)) - np.vstack((xBJ, 0., zBJ))
-dB = np.vstack((xB, np.zeros(nBk), zB)) - np.vstack((par_n['xB'], np.zeros(nBk), par_n['zB']))
-IB = np.zeros((3, 3))
-for i in range(nBk):
-    IB[0] = np.array([par_n['IBxx'][i], 0., par_n['IBxz'][i]])
-    IB[1] = np.array([0., par_n['IByy'][i], 0.])
-    IB[2] = np.array([par_n['IBxz'][i], 0., par_n['IBzz'][i]])
-    I = parallel_axis(IBJ, mBJ, dJ[:, i]) + parallel_axis(IB, par_n['mB'][i], dB[:, i])
-    par_n['IBxx'][i] = I[0, 0]
-    par_n['IBxz'][i] = I[0, 2]
[…]

### 2010-10-09 PhysicalParameters.py
working on the tables output

@@ -231,7 +231,15 @@ def plot_evecs():
                     #np.abs(np.imag(wea['evecs'][:, :, 1])), zs=vel)
     plt.show()
 
-def make_tables():
+def make_tables(typ='Bike'):
+    '''
+
+    typ : string
+        'Bike'
+        'BikeRider'
+        'BikeLegs'
+
+    '''
 
     # load in the base data file
     f = open('data/data.p', 'r')
@@ -243,11 +251,14 @@ def make_tables():
     par = pickle.load(f)
     f.close()
 
-    direct = 'tables/'
+    direct = 'tables/' + typ + '/Parameters/'
+    if not os.path.isdir(direct):
+        os.system('mkdir ' + direct)
+
     for i, name in enumerate(data['bikes']):
         fname = ''.join(name.split())
         # open the new file
-        f = open(direct + 'ParameterTable.tex', 'r')
+        f = open('tables/' + 'ParameterTable.tex', 'r')
         fn = open(direct + fname + 'Par.tex', 'w')
         for line in f:
             #print line
@@ -283,11 +294,11 @@ def make_tables():
             fn.write(line)
         f.close()
         fn.close()
-        os.system('pdflatex -output-directory=tables ' + direct + fname + 'Par.tex')
+        os.system('pdflatex -output-directory=' + direct[:-1] + ' ' + direct + fname + 'Par.tex')
 
     # make the master parameter table
     template = open('tables/MasterParTableTemplate.tex', 'r')
-    final = open('tables/MasterParTable.tex', 'w')
+    final = open(direct + 'MasterParTable.tex', 'w')
 
     abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
     for line in template:
@@ -308,12 +319,15 @@ def make_tables():
 
     template.close()
     final.close()
-    os.system('pdflatex -output-directory=tables ' + direct + 'MasterParTable.tex')
-
+    os.system('pdflatex -output-directory=' + direct[:-1] + ' ' + direct + 'MasterParTable.tex')
 
     # make the master canonical matrix table
+    direct = 'tables/' + typ + '/Canonical/'
+    if not os.path.isdir(direct):
+        os.system('mkdir ' + direct)
+
     template = open('tables/MasterCanTableTemplate.tex', 'r')
-    final = open('tables/MasterCanTable.tex', 'w')
[…]

### 2010-10-11 papers/BMD2010/PhysicalParametersAbstract/MooreHubbardSchwabKooijman_abstract.tex
bode plots are working and the paper is started

--- 2010-09-10 BMD2010/PhysicalParameters/MooreHubbardSchwabKooijman_abstract.tex
@@
 are presented up through the eigenvalue prediction of the linear model and are
-based error propagation theory with correlations taken into account.
+based on error propagation theory with correlations taken into account.
 
@@
 Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
-Dutch style city bicycle with a suspended fork, \emph{Gary Fisher Mountain
+Dutch style city bicycle with a suspended fork; \emph{Gary Fisher Mountain
 Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
-track racing bicycle; \emph{Yellow Bicycle}, a stripped down aluminum frame
+track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
 road bicycle measured in two configurations, the second with the fork rotated
@@
 The mass of each of the rigid bodies was
-measured with a precision scale (to so many grams). The center of mass of each rigid
-body was measured by hanging each it from several points. The mass
+measured with a precision scale. The center of mass of each rigid
+body was measured by hanging each body from several points. The mass
 center is the intersection of the lines of support under the
@@
 eigenvectors, Bode plots, etc.), Fig.~\ref{fig:bike_eig}. Notable results
-include: high accuracy
-measurements, dual oscillatory eigenmodes, bicycles without riders may often
-have no stable speed range, and it is possible to calculate uncertainties
-in the eigenvalues and eigenvectors.
+include high accuracy measurements and uncertainty propagation in the eigenvalues and eigenvectors.
 

### 2010-10-11 PhysicalParameters.py

@@ -280,46 +280,98 @@ def make_tables(typ='Bike'):
     os.system('rm tables/' + typ + '/Parameters/*.blg')
     os.system('rm tables/' + typ + '/Parameters/*.bbl')
 
-def bike_bode_plots():
+def bike_bode_plots(typ='Bike', speeds=None):
+    '''
+    Makes several bode plots for the bicycles.
+
+    Parameters:
+    -----------
+    typ : string
+        'Bike', 'BikeRider', 'BikeLegs'
+
+    '''
+    if speeds == None:
+        speeds = [0., 2.5, 4., 5., 5.8, 7.5, 12.0]
 
     # load in the base data file
     f = open('data/data.p', 'r')
     data = pickle.load(f)
     f.close()
 
+    # figure properties
+    figwidth = 5. # in inches
+    goldenMean = (np.sqrt(5)-1.0)/2.0
+    figsize = [figwidth, figwidth*goldenMean]
+    params = {#'backend': 'ps',
+        'axes.labelsize': 8,
+        'text.fontsize': 8,
+        'legend.fontsize': 6,
+        'xtick.labelsize': 8,
+        'ytick.labelsize': 8,
+        #'text.usetex': True,
+        'figure.figsize': figsize
+        }
+    plt.rcParams.update(params)
+
     nBk = len(data['bikes'])
 
-    colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange']
+    colors = ['k', 'r', 'b', 'g', 'y', 'm', 'c', 'orange', 'red']
+    #colors = ['#000000',
+              #'#FF0000',
+              #'#00FF00',
+              #'#0000FF',
+              #'#FFFF00',
+              #'#00FFFF',
+              #'#FF00FF',
+              #'#C0C0C0']
 
-    Tdel2phi = plt.figure(num=1)
-    Tdel2del = plt.figure(num=2)
-    Tphi2phi = plt.figure(num=3)
-    Tphi2del = plt.figure(num=4)
+    plotlist = ['Tdel2phi', 'Tdel2del', 'Tphi2phi', 'Tphi2del']
 
-    dir = 'data/bikeRiderCanonical/'
-    for i, name in enumerate(data['bikes']):
-        fname = ''.join(name.split()) + 'RiderCan.p'
-        file = open(dir + fname)
-        can = pickle.load(file)
-        file.close()
-        # make some bode plots
[…]

### 2010-10-11 papers/BMD2010/PhysicalParametersPaper/paper.tex
paper started

--- 2010-06-08 report/parameterReport.tex
@@
-\documentclass{article}
+\documentclass{bmd2010p}
 
-\usepackage[lmargin=1in,rmargin=1in, tmargin=1in, bmargin=1in]{geometry}
-\usepackage[colorlinks=true, linkcolor=blue]{hyperref}
+%\usepackage[colorlinks=true, linkcolor=blue]{hyperref}
 \usepackage{graphicx}
-\usepackage{eurosans}
 \usepackage{subfiles}
-
-\author{Jason K. Moore}
-\title{Accurate Measurement of Eight Bicycles' Physical Parameters}
 
@@
 
-\maketitle
-\tableofcontents
+\begin{center}
+\fontsize{14}{20}{\bf Accurate Measurement of Bicycle Parameter[3pt]}
+\end{center}
 
-\section{Introduction}
+%%%%%%%%%%%%%%%% authors %%%%%%%%%%%%%%%
+\begin{center}
+\normalsize{\bf{Jason K. Moore$^{*}$, Mont Hubbard, A. L. Schwab$^\#$,
+            J. D. G. Kooijman$^\dag$}}
+\end{center}
+
+\begin{center}
+\begin{tabular}{c}
+$^*$ Mechanical and Aerospace Engineering\\
+University of California, Davis\\
+One Shields Avenue, Davis, CA, 95616, USA\\
+e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu\\
+\end{tabular}
+\begin{tabular}{c}
+$^\#$ Laboratory for Engineering Mechanics\\
+Delft University of Technology\\
+Mekelweg 2, 2628CD Delft, The Netherlands\\
+e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com\\
+\end{tabular} \\ \vspace{2.5ex}
+
+\end{center}
+
+\section*{ABSTRACT}
+Accurate measurements of a bicycle's physical parameters are required for
+realistic dynamic simulations and analysis. For the most basic models the
+geometry, mass, mass location and mass distributions must be measured. More complex models
+require measurements of tire characteristics, human characteristics, friction, stiffness, damping, etc. This
+paper concerns the measurement of the minimal bicycle parameters required for
+the benchmark bicycle presented in~\cite{Meijaard2007}. This
+model is composed of four rigid bodies, has ideal rolling and frictionless joints,
+and is assumed to be laterally symmetric. A set of 25
+parameters is used to describe the geometry, mass, mass location and
+mass distribution of each of the rigid bodies. The experimental methods
+described herein are based primarily on the work
+done in~\cite{Kooijman2006} but have been refined for improved accuracy and
+methodology. Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
+similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
+similar techniques with scooters. We measured the characteristics of six
+different bicycles, two of which were set up in two different configurations.
+This is a total of eight different parameter sets that can be used with, but not
+limited to, the benchmark bicycle model. The accuracy of all the measurements
+are presented up through the eigenvalue prediction of the linear model and are
+based on error propagation theory with correlations taken into account.
+
+The six bicycles, chosen for both variety and convenience, are as follows:
+\emph{Batavus Browser}, a Dutch style city bicycle measured with and without
+instrumentation as described in~\cite{Kooijman2009a}; \emph{Batavus Stratos
+Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
+Dutch style city bicycle with a suspended fork; \emph{Gary Fisher Mountain
+Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
+track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
+road bicycle measured in two configurations, the second with the fork rotated
+in the headtube 180 degrees for larger trail.
+\begin{keywords}
+bicycle,
+parameters,
+eigenvalues,
+Bode.
+\end{keywords}
+
+\section{INTRODUCTION}
+
 This work is intended to document the indirect measurement of eight real
@@
 technique.
+
 \section{Bicycle Descriptions}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/yellow.jpg}
+        \includegraphics[width=4in]{../../../images/yellow.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}[htbp]
-	\begin{tabular}{ll}
-		Style & Road\\
-		Size & \\
-		Wheels & 700c Aluminum Rims\\
-		Tires & 28 x 1 3/8 x 1 5/8\\
-		Sprocket & 6 speed free wheel\\
-		Handlebars & Drop\\
-	\end{tabular}
-	\caption{Specifications of the Yellow Bicycle}
-	\label{yellowBikeSpecs}
-\end{table}
 \subsection{Batavus Browser}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/browser.jpg}
+        \includegraphics[width=4in]{../../../images/browser.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/browserIns.jpg}
+        \includegraphics[width=4in]{../../../images/browserIns.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}[htbp]
-	\begin{tabular}{ll}
-        Price          & euro symbol 449\\
-        Wheel Size     & 28''\\
-        Weight         & 18.7 kg\\
-        Frame Size     & 54\\
-        Frame material & High Tensile Steel\\
-        Fork material  & Steel\\
-        Rims           & Aluminum Airline\\
-        Spokes         & Stainless steel\\
-        Tires          & Trax Highway, 28'' x 1 5/8 x 1 3/8\\
-        Handlebar      & Steel painted\\
-        Handles        & Batavus Comfort\\
-        Saddle         & Selle Royal 8274\\
-        Dynamo         & Basta Trio\\
-        Headlight      & Batavus Logic Light\\
-        Rearlight      & Battery light\\
-        Pedals         & Batavus plastic antislip\\
-        Final          & Trelock RS42\\
-        Theft chip     & Present\\
-        Crank          & Aluminum\\
-        Rack           & Steel\\
-        Quick Binder   & Binder Trio\\
-        Mudguards      & Steel\\
-        Bell           & Kenlight\\
-	\end{tabular}
-	\caption{2008 Batavus Browser Specifications.}
-	\label{browserSpecs}
-\end{table}
 
@@
     \begin{center}
-        \includegraphics[width=4in]{images/pista.jpg}
+        \includegraphics[width=4in]{../../../images/pista.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}
-	\begin{tabular}{ll}
-        Style            & Track Bike\\
-        Size             & 57cm\\
-        Color            & Gang Green\\
-        Frame            & Bianchi DB CrMo, rear entry track dropouts\\
-        Fork             & CrMo\\
-        Retail Price     & \$579.99\\
-        Headset          & VP AheadSet, 1''threadless\\
-        Handlebar        & Bianchi steel track, 26.0mm\\
-        Stem             & Bianchi alloy\\
-        Brakes/Levers    & Shimano Tiagra front brake/lever\\
-        Crankset         & Truvativ Touro 1.1, 48T\\
-        Bottom Bracket   & Truvativ Power Spline cartridge\\
-        Chain            & KMC\\
-        Sprocket         & 16T fixed cog \& 16T Freewheel\\
-        Pedals           & Campus pedals\\
-        Wheels           & Bianchi Alex Solo wheelset, 28h rims\\
-        Tires            & Continental UltraSport, 700x23C\\
-        Saddle           & Bianchi Velo\\
-        Seatpost         & Bianchi alloy, 27.2mm\\
-        Size             & 57cm\\ 
-        Seat Tube        & 570mm\\
-        Top Tube-actual  & 560mm\\
-        Top Tube-virtual & 560mm\\
-        Chainstay        & 383mm\\
-        Fork Rake        & 28mm\\
-        Head Tube Angle  & 74.5$^{\circ}$\\
-        Seat Tube Angle  & 75.5$^{\circ}$\\
-        Wheelbase        & 967mm\\
-        Standover Height & 32''  
-	\end{tabular}
-	\caption{}
-	\label{pistaSpecs}
-\end{table}
 \subsection{Gary Fisher Mountain Bike}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/fisher.jpg}
+        \includegraphics[width=4in]{../../../images/fisher.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}
-    \centering
-    \begin{tabular}{ll}
-        Geometry & Genisis\\
-        Frame & Aluminum, Damping, AC/DC, Carbon, Composite\\
-        Front Wheel & DT Swiss XR4.1 Ceramic 559 17\\
-        Tires & Michelin 26'' X 2.00 (5.2-559)\\
-        Rear Wheel & DT Swiss 240s\\
-        Rear hub & 9 speed\\
-        Rear derailleur & SRAM X-0\\
-        Crank & Bonterager 3-speed\\
-        Seat post & Ritchey\\
-        Fork & SID Rock Shox
-    \end{tabular}
-    \caption{Specifications of the Gary Fisher Mountain Bike}
-    \label{tab:gary}
-\end{table}
 \subsection{Batavus Crescendo Deluxe}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/crescendo.jpg}
+        \includegraphics[width=4in]{../../../images/crescendo.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}
-	\begin{tabular}{ll}
-		Price             & from € 869\\
-		Wheel Size        & 28''\\
-		Weight            & 21.5 kg\\
-		Frame Size        & 53\\
-		Performance/speed & Shimano Nexus 8 speed with roller brakes\\
-		Color             & White / black\\
-		Brakes            & Shimano roller brakes (BR-IM50)\\
-		Frame type        & Ladies mono tube frame Gentlemen trend\\
-		Frame material    & Aluminum 7005\\
-		Front Fork        & spring, Batavus\\
-		Rims              & Aluminum Rodi VR19\\
-		Spokes            & Stainless steel (adapted spoke pattern)\\
-		Tires             & CST End\\
-		Handlebar         & Steel aluminum look\\
-		Stem              & Manually adjustable, Batavus Ergo Matic III\\
-		Handles           & adjustable Batavus Ergo Grip II\\
-		Saddle            & Selle Royal 5160\\
-		Seatpost          & spring, Post Modern Compact\\
-		Dynamo            & Shimano Hub Dynamo\\
-		Headlight         & Trelock Hi-Power automatic halogen\\
-		Rear              & Smart battery automatically\\
-		Pedals            & aluminum antislip\\
-		Final							& AXA Defender\\
-		Chip theft        & Present\\
-		Crank             & Aluminum\\
-		Rack              & Aluminum\\
-		Quick Binder      & Binder Quattro\\
-		Mudguards         & Batavus hyalite II\\
-		Bell              & Bibia Touring\\
-		Balhoofdset       & VP semi-integrated, oversized\\
-		Shifters          & Batavus (8-speed)\\
-		Saddlebag         & Batavus saddlebag with plakset\\
-	\end{tabular}
-	\caption{}
-	\label{crescendoSpecs}
-\end{table}
 \subsection{Batavus Stratos Deluxe}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/stratos.jpg}
+        \includegraphics[width=4in]{../../../images/stratos.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}
-	\begin{tabular}{ll}
-		Switching           & Spin System (shifter)\\
-		Monochromatic frame & Black\\
-		Chain guard         & Semi-open\\
-		Mudguards           & Aluminum\\
-		Rack                & Yes\\
-		Extra drive         & Regardless Motori degree\\
-		Type                & Men's\\
-		Frame Material      & Aluminum\\
-		Rim Material        & Aluminum\\
-		Rear braking system & Roller brake\\
-		Front braking system & Roller Brake\\
-		Number of gears      & 7 Gears\\
-		Gear System          & Shimano Nexus\\
-		Wheel Size           & 28 inch\\
-		Tirewidth            & 37 mm\\
-		Price                & 749,00\\
-	\end{tabular}
-	\caption{}
-	\label{stratosSpecs}
-\end{table}
 \section{Accuracy}
@@
 	\begin{center}
-		\includegraphics[width=4in]{images/tireChalk.jpg}
+		\includegraphics[width=4in]{../../../images/tireChalk.jpg}
 	\end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/massScale.jpg}
+        \includegraphics[width=4in]{../../../images/massScale.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/pendDist.jpg}
+        \includegraphics[width=4in]{../../../images/pendDist.jpg}
     \end{center}
@@
 	\begin{center}
-		\includegraphics[width=4in]{images/level.jpg}
+		\includegraphics[width=4in]{../../../images/level.jpg}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=4in]{figures/angles.pdf}
+		\includegraphics[width=4in]{../../../figures/angles.pdf}
 	\end{center}
@@
     \begin{center}
-        \includegraphics{figures/triangle.pdf}
+        \includegraphics{../../../figures/triangle.pdf}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{../data/pendDat/graphs/BrowserFrameCompoundFirst1.png}
+        \includegraphics[width=4in]{../../../plots/PendFit/BrowserFrameCompoundFirst1.png}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{../data/pendDat/graphs/CrescendoForkTorsionalFirst2.png}
+        \includegraphics[width=4in]{../../../plots/PendFit/CrescendoForkTorsionalFirst2.png}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/fixture.jpg}
+        \includegraphics[width=4in]{../../../images/fixture.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/rod.jpg}
+        \includegraphics[width=4in]{../../../images/rod.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/CrescendoFwheelTorsionalFirst.jpg}
+        \includegraphics[width=4in]{../../../images/CrescendoFwheelTorsionalFirst.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/wheelIyy.jpg}
+        \includegraphics[width=4in]{../../../images/wheelIyy.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/YellowFrameCompoundFirst.jpg}
+        \includegraphics[width=4in]{../../../images/YellowFrameCompoundFirst.jpg}
     \end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{images/StratosForkTorsionalThird.jpg}
+        \includegraphics[width=4in]{../../../images/StratosForkTorsionalThird.jpg}
     \end{center}
@@
 \section{Parameters}
-\subsection{Batavus Browser}
-\subfile{../parTables/BatavusBrowserPar.tex}
-\subsection{Batavus Browser Instrumented}
-\subfile{../parTables/BatavusBrowserInstrumentedPar.tex}
-\subsection{Batavus Crescendo Deluxe}
-\subfile{../parTables/BatavusCrescendoDeluxePar.tex}
-\subsection{Gary Fisher}
-\subfile{../parTables/GaryFisherPar.tex}
-\subsection{Bianchi Pista}
-\subfile{../parTables/BianchiPistaPar.tex}
-\subsection{Batavus Stratos Deluxe}
-\subfile{../parTables/BatavusStratosDeluxePar.tex}
-\subsection{Yellow Bike}
-\subfile{../parTables/YellowBikePar.tex}
-\subsection{Yellow Bike Reversed Fork}
-\subfile{../parTables/YellowBikeReversedForkPar.tex}
 \section{Linear Analysis}
@@
 example rider and bicycle follow in Eqs.~\ref{eq:M}-\ref{eq:K2}.
-\subsubsection{Batavus Browser}
-\subfile{../data/bikeCanonical/BatavusBrowserCan.tex}
-\subsubsection{Batavus Browser Instrumented}
-\subfile{../data/bikeCanonical/BatavusBrowserInstrumentedCan.tex}
-\subsubsection{Batavus Crescendo Deluxe}
-\subfile{../data/bikeCanonical/BatavusCrescendoDeluxeCan.tex}
-\subsubsection{Gary Fisher}
-\subfile{../data/bikeCanonical/GaryFisherCan.tex}
-\subsubsection{Bianchi Pista}
-\subfile{../data/bikeCanonical/BianchiPistaCan.tex}
-\subsubsection{Batavus Stratos Deluxe}
-\subfile{../data/bikeCanonical/BatavusStratosDeluxeCan.tex}
-\subsubsection{Yellow Bike}
-\subfile{../data/bikeCanonical/YellowBikeCan.tex}
-\subsubsection{Yellow Bike Reversed Fork}
-\subfile{../data/bikeCanonical/YellowBikeReversedForkCan.tex}
+\subsection{Eigenvalues}
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=4in]{../../../plots/Bike/CrescendoRootLoci.png}
+    \end{center}
+    \caption{The root loci with speed as the parameter for the Crescendo.}
+    \label{fig:StratosFork}
+\end{figure}
 
-\subsection{Eigenvalues}
 \subsection{Frequency Response}
-\appendix
-\bibliographystyle{plain}
+\bibliographystyle{acm}
 \bibliography{bicycle}

### 2010-10-12 papers/BMD2010/PhysicalParametersPaper/paper.tex
bicycle subfigure done

@@
 \usepackage{subfiles}
+\usepackage{subfig}
 
@@
 \section{Bicycle Descriptions}
-\subsection{Yellow Bicycle}
-The yellow is a bicycle used in the lab to demonstrate that a bicycle is stable
-at certain speeds. It is an aluminum road frame of unknown make with the
-majority of components removed. The wheels, drop handlebar, seat, seat post and
-bottom bracket are the only parts on the bike. The fork has been reversed to
-''increase`` the stability of the bicycle. The bicycle was measured with both
-the fork in normal position and the fork reversed.
+\begin{description}
+    \item[Batavus Browser]{The Batavus Browser is lower priced Dutch city bike.
+        We measured the physical proprieties of the stock Browser model and
+        also the same bike with the instrumentation used in our experiments.}
+    \item[Batavus Crescendo Deluxe]{The Batavus Crescendo Deluxe is a
+        contemporary of all-round cycling in cool black and white. A great bike
+        for a good end to touring. And that works fine with 8 gears, suspension
+        fork and suspension seatpost.}
+    \item[Gary Fisher Mountain Bike]{The Gary Fisher Ziggurat is a modern front
+        suspended mountain bike.}
+    \item[2007 Bianchi Pista]{The Pista is a modern steel track bicycle.}
+    \item[Batavus Stratos Deluxe]{Is a Dutch sport bicycle.}
+    \item[Yellow Bicycle]{The yellow is a bicycle used in the lab to
+        demonstrate that a bicycle is stable at certain speeds. It is an
+        aluminum road frame of unknown make with the majority of components
+        removed. The wheels, drop handlebar, seat, seat post and bottom bracket
+        are the only parts on the bike. The fork has been reversed to
+        ''increase``~\cite{Kooijman2006} the stability of the bicycle. The
+        bicycle was measured with both the fork in normal position and the fork
+        reversed.}
+\end{description}
 \begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/yellow.jpg}
-    \end{center}
-    \caption{Side view of the yellow bicycle with the fork in the reversed
-    posistion.}
-    \label{fig:YellowBike}
-\end{figure}
-\subsection{Batavus Browser}
-The Batavus Browser is lower priced Dutch city bike. We measured the physical
-proprieties of the stock Browser model and also the same bike with the
-instrumentation used in our experiments.
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/browser.jpg}
-    \end{center}
-    \caption{Side view of the Batavus Browser as from the factory.}
-    \label{fig:browser}
-\end{figure}
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/browserIns.jpg}
-    \end{center}
-    \caption{The Batavus Browser with instrumentation as used
-    in~\cite{Kooijman2009}}
-    \label{fig:browserIns}
-\end{figure}
-
-\subsection{2007 Bianchi Pista}
-The Pista is a modern steel track bicycle. Specifications, taken from the Bianchi
-website~\cite{Bianchi2009}, are as follows:
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/pista.jpg}
-    \end{center}
-    \caption{The Bianchi Pista hanging from a spring scale.}
-    \label{fig:pista}
-\end{figure}
-\subsection{Gary Fisher Mountain Bike}
-The Gary Fisher Ziggurat is a modern front suspended mountain bike.
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/fisher.jpg}
-    \end{center}
-    \caption{The Gary Fisher mountain bike hanging from a spring scale.}
-    \label{fig:fisher}
-\end{figure}
-\subsection{Batavus Crescendo Deluxe}
-The Batavus Crescendo Deluxe is a contemporary of all-round cycling in cool
-black and white. A great bike for a good end to touring. And that works fine
-with 8 gears, suspension fork and suspension seatpost.
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/crescendo.jpg}
-    \end{center}
-    \caption{A side view of the Batavus Crescendo Deluxe.}
-    \label{fig:crescendo}
-\end{figure}
-\subsection{Batavus Stratos Deluxe}
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/stratos.jpg}
-    \end{center}
-    \caption{A side view of the Batavus Stratos Deluxe}
-    \label{fig:stratos}
+    \centering
+    \subfloat[Batavus Browser]{
+        \label{fig:browser}\includegraphics[width=1.75in]{../../../images/browserIns_sub.jpg}}
+    \subfloat[Batavus Crescendo Deluxe]{
+        \label{fig:crescendo}\includegraphics[width=1.75in]{../../../images/crescendo_sub.jpg}}
+    \subfloat[Gary Fisher]{
+        \label{fig:fisher}\includegraphics[width=1.75in]{../../../images/fisher_sub.jpg}}
+        \\
+    \subfloat[Bianchi Pista]{
+        \label{fig:pista}\includegraphics[width=1.75in]{../../../images/pista_sub.jpg}}
+    \subfloat[Batavus Stratos Deluxe]{
+        \label{fig:stratos}\includegraphics[width=1.75in]{../../../images/stratos_sub.jpg}}
+    \subfloat[Yellow Bicycle]{
+        \label{fig:yellow}\includegraphics[width=1.75in]{../../../images/yellow_sub.jpg}}
+    \caption{The six bicycles measured in the experiments. The Batavus
+    Browser~\subref{fig:browser} is shown with the instrumentation and the
+    Yellow Bicycle~\subref{fig:yellow} is shown with its fork reversed.}
+    \label{fig:bicycles}
 \end{figure}

### 2010-10-12 PhysicalParameters.py

@@ -1345,8 +1345,9 @@ def bode(ABCD=None, numden=None, w=None, fig=None, n=None, label=None,
                   x=-0.01)
 
     fig.axprops = {}
-    fig.ax1 = fig.add_axes([0.125, 0.5, 0.95-0.125, 0.475-0.2], **fig.axprops)
-    fig.ax2 = fig.add_axes([0.125, 0.2, 0.95-0.125, 0.475-0.2], **fig.axprops)
+    # axes [left, bottom, width, height]
+    fig.ax1 = fig.add_axes([.125, .525, .825, .275], **fig.axprops)
+    fig.ax2 = fig.add_axes([.125, .2, .825, .275], **fig.axprops)
 
     if (ABCD):
         A, B, C, D = ABCD

### 2010-10-12 papers/BMD2010/PhysicalParametersPaper/paper.tex
the bicycle descriptions are good

@@
 \begin{center}
-\fontsize{14}{20}{\bf Accurate Measurement of Bicycle Parameter[3pt]}
+\fontsize{14}{20}{\bf Accurate Measurement of Bicycle Parameters}
 \end{center}
@@
 \begin{center}
-\normalsize{\bf{Jason K. Moore$^{*}$, Mont Hubbard, A. L. Schwab$^\#$,
+\normalsize{\bf{Jason K. Moore$^{*}$, Mont Hubbard$^{*}$, A. L. Schwab$^\dag$,
             J. D. G. Kooijman$^\dag$}}
@@
 \begin{tabular}{c}
-$^\#$ Laboratory for Engineering Mechanics\\
+$^\dag$ Laboratory for Engineering Mechanics\\
 Delft University of Technology\\
@@
 \section{Bicycle Descriptions}
+We choose to measure the physical parameters of six bicycles
+Figs.~\ref{fig:bicycles}. The three Batavus bicycles were donated by the
+manufacturer. We asked for a bicycle that they considered stable and one that
+they did not. They claimed the Browser was a ``stable'' bicycle and that the
+Stratos was ``nervous''. The Fisher and the Pista were chosen to provide some
+variety, a mountatin and road bike. The yellow bike is used to demonstrate
+bicycle stability.
 \begin{description}
-    \item[Batavus Browser]{The Batavus Browser is lower priced Dutch city bike.
-        We measured the physical proprieties of the stock Browser model and
-        also the same bike with the instrumentation used in our experiments.}
-    \item[Batavus Crescendo Deluxe]{The Batavus Crescendo Deluxe is a
-        contemporary of all-round cycling in cool black and white. A great bike
-        for a good end to touring. And that works fine with 8 gears, suspension
-        fork and suspension seatpost.}
-    \item[Gary Fisher Mountain Bike]{The Gary Fisher Ziggurat is a modern front
-        suspended mountain bike.}
-    \item[2007 Bianchi Pista]{The Pista is a modern steel track bicycle.}
-    \item[Batavus Stratos Deluxe]{Is a Dutch sport bicycle.}
-    \item[Yellow Bicycle]{The yellow is a bicycle used in the lab to
+    \item[Batavus Browser]{The Batavus Browser Fig.~\ref{fig:browser} is an averaged priced Dutch city
+        bike. It has a steel frame, a three speed internal rear hub, handle
+        bars for an upright posture and includes various accerories for utility
+        puposes. We measured the physical proprieties of the stock Browser model and
+        also equipped with the instrumentation used in our other
+        experiments~\cite{Kooijman2009}.}
+    \item[Batavus Crescendo Deluxe]{The Batavus Crescendo Deluxe
+        Fig.~\ref{fig:crescendo} is also a
+        Dutch city bike that is setup for touring. It has an aluminum frame, an
+        eight speed internal hub, upright handlebars, accesories for utility
+        and a suspension fork and suspension seatpost.}
+    \item[Gary Fisher Ziggurat]{The Gary Fisher Ziggurat Fig.~\ref{fig:fisher} is a modern
+        lightweight front suspended mountain bike. It has a ?? frame, large low
+        pressure mountain bike tires, and is setup for racing with few extra accerories.}
+    \item[2007 Bianchi Pista]{The Bianchi Pista Fig.~\ref{fig:pista} is a modern lightweight steel track
+        bicycle. It has a single gear ratio and minimal extras to keep the
+        weight low. It has drop handlebars and high pressure racing tires.}
+    \item[Batavus Stratos Deluxe]{The Batavus Stratos Deluxe
+        Fig.~\ref{fig:stratos} is a sporty Dutch city bicycle. The frame is
+        made of aluminum and it has a seven speed internal hub. It has mountain
+        style handle bars for a less upright seating posture, but also includes
+        accesories for utility such as a rear rack, fenders, light and
+        chainguard.}
+    \item[Yellow Bicycle]{The yellow bicycle Fig.~\ref{fig:yellow} used in the lab to
         demonstrate that a bicycle is stable at certain speeds. It is an
@@
         removed. The wheels, drop handlebar, seat, seat post and bottom bracket
-        are the only parts on the bike. The fork has been reversed to
-        ''increase``~\cite{Kooijman2006} the stability of the bicycle. The
-        bicycle was measured with both the fork in normal position and the fork
-        reversed.}
+        are the only parts on the bike. The bicycle was measured with both the
+        fork in normal position and the fork reversed. The fork was reversed to
+        ''decrease the minimal stable speed``~\cite{Kooijman2006} of the bicycle.}
 \end{description}

### 2010-10-13 papers/BMD2010/PhysicalParametersPaper/paper.tex
down to the headtube angle

@@
 \usepackage{subfig}
+\usepackage{amsmath}
 
@@
 technique.
-
-\section{Bicycle Descriptions}
+\section{BENCHMARK BICYCLE MODEL}
+Recently, the Whipple bicycle model has been benchmarked~\cite{Meijaard2007}
+and this model is widely used for bicycle dynamics studies. The unforced two
+degree-of-freedom, $\mathbf{q}$ = [steer and roll], model takes the form:
+\begin{equation}
+    \mathbf{M\ddot{q}}
+    +v\mathbf{C}_1\mathbf{\dot{q}}
+    +\left[g\mathbf{K}_0
+    +v^2\mathbf{K}_2\right]\mathbf{q}
+    =0
+    \label{eq:canonical}
+\end{equation}
+Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
+matrices are combinations of 25 of the bicycle's physical parameters that
+include the geometry, mass, mass location and mass distribution of the four
+rigid bodies. The 25 parameters presented in \cite{Meijaard2007} are not
+necessarily a minimun set for the Whipple model, as shown in \cite{Sharp2008},
+but are useful as they represent more easily measurable quantities.
+Furthermore, many more parameters are not needed due to the assumptions of the
+Whipple model such as no-slip tires, lateral symmetry, knife edge wheels, etc.
+\section{BICYCLE DESCRIPTIONS}
 We choose to measure the physical parameters of six bicycles
@@
 \begin{description}
-    \item[Batavus Browser]{The Batavus Browser Fig.~\ref{fig:browser} is an averaged priced Dutch city
+    \item[Batavus Browser (B, B*)]{The Batavus Browser Fig.~\ref{fig:browser} is an averaged priced Dutch city
         bike. It has a steel frame, a three speed internal rear hub, handle
@@
         experiments~\cite{Kooijman2009}.}
-    \item[Batavus Crescendo Deluxe]{The Batavus Crescendo Deluxe
+    \item[Batavus Crescendo Deluxe (C)]{The Batavus Crescendo Deluxe
         Fig.~\ref{fig:crescendo} is also a
@@
         and a suspension fork and suspension seatpost.}
-    \item[Gary Fisher Ziggurat]{The Gary Fisher Ziggurat Fig.~\ref{fig:fisher} is a modern
+    \item[Gary Fisher Ziggurat (G)]{The Gary Fisher Ziggurat Fig.~\ref{fig:fisher} is a modern
         lightweight front suspended mountain bike. It has a ?? frame, large low
         pressure mountain bike tires, and is setup for racing with few extra accerories.}
-    \item[2007 Bianchi Pista]{The Bianchi Pista Fig.~\ref{fig:pista} is a modern lightweight steel track
+    \item[Bianchi Pista (P)]{The Bianchi Pista Fig.~\ref{fig:pista} is a modern lightweight steel track
         bicycle. It has a single gear ratio and minimal extras to keep the
         weight low. It has drop handlebars and high pressure racing tires.}
-    \item[Batavus Stratos Deluxe]{The Batavus Stratos Deluxe
+    \item[Batavus Stratos Deluxe (S)]{The Batavus Stratos Deluxe
         Fig.~\ref{fig:stratos} is a sporty Dutch city bicycle. The frame is
@@
         chainguard.}
-    \item[Yellow Bicycle]{The yellow bicycle Fig.~\ref{fig:yellow} used in the lab to
+    \item[Yellow Bicycle (Y, Y*)]{The yellow bicycle Fig.~\ref{fig:yellow} used in the lab to
         demonstrate that a bicycle is stable at certain speeds. It is an
@@
     \subfloat[Batavus Browser]{
-        \label{fig:browser}\includegraphics[width=1.75in]{../../../images/browserIns_sub.jpg}}
+        \label{fig:browser}
+        \includegraphics[width=1.75in]{../../../images/browserIns_sub.jpg}
+        }
     \subfloat[Batavus Crescendo Deluxe]{
-        \label{fig:crescendo}\includegraphics[width=1.75in]{../../../images/crescendo_sub.jpg}}
+        \label{fig:crescendo}
+        \includegraphics[width=1.75in]{../../../images/crescendo_sub.jpg}
+        }
     \subfloat[Gary Fisher]{
-        \label{fig:fisher}\includegraphics[width=1.75in]{../../../images/fisher_sub.jpg}}
-        \\
+        \label{fig:fisher}
+        \includegraphics[width=1.75in]{../../../images/fisher_sub.jpg}
+        }\\
     \subfloat[Bianchi Pista]{
-        \label{fig:pista}\includegraphics[width=1.75in]{../../../images/pista_sub.jpg}}
+        \label{fig:pista}
+        \includegraphics[width=1.75in]{../../../images/pista_sub.jpg}
+        }
     \subfloat[Batavus Stratos Deluxe]{
-        \label{fig:stratos}\includegraphics[width=1.75in]{../../../images/stratos_sub.jpg}}
+        \label{fig:stratos}
+        \includegraphics[width=1.75in]{../../../images/stratos_sub.jpg}
+        }
     \subfloat[Yellow Bicycle]{
-        \label{fig:yellow}\includegraphics[width=1.75in]{../../../images/yellow_sub.jpg}}
+        \label{fig:yellow}
+        \includegraphics[width=1.75in]{../../../images/yellow_sub.jpg}
+        }
     \caption{The six bicycles measured in the experiments. The Batavus
@@
 \end{figure}
-\section{Accuracy}
-Error propagation theory was utilized to calculate what the accuracy the
-benchmark ideal parameters are based on the measurements taken. If $x$ is a
-random variable and is defined as $x=f(u,v,\ldots)$ then the sample variance of
-$x$ is defined as
+\section{ACCURACY}
+We took great care too improve upon the accuracy and report the accuracy of the
+measurements of the bicycle parameters. Following the vain of
+~\cite{Roland1971} we used error propagation theory to calculate accuracy of
+the 25 benchmark parameters. We start by estimating the standard deviation of
+the actual measurements taken. If $x$ is a parameter and is a function of
+the measurments, $u,v,\ldots$, then $x$ is a random variable defined as
+$x=f(u,v,\ldots)$. The sample variance of $x$ is defined as
 \begin{equation}
-    s_x^2 = \frac{1}{N-1}\sum^N_{i=1}
+    s_x^2 =
+    \frac{1}{N-1}\sum^N_{i=1}
     \left[(u_i - \bar{u})^2\left(\frac{\partial x}{\partial u}\right)^2 +
@@
 \end{equation}
-If $u$ and $v$ are uncorrelated then $s_{uv}=0$. We assume that the all the
-random variables hereafter are uncorrelated. Equation~\ref{eqn:variance} can be
-used to calculated the variance of all types of functions. The most basic are
-functions such as addtion
+If $u$ and $v$ are uncorrelated then $s_{uv}=0$. Most of the calculations
+hereafter have uncorrelate variables but a few do not and the covariance has to
+be taken into account. Equation~\ref{eqn:variance} can be
+used to calculated the variance of all types of functions. Simple addition of
+two random variables may be the most basic example:
 \begin{eqnarray}
@@
 \end{eqnarray}
-
-\section{Geometry}
-
-\subsection{Wheel radius}
+\section{GEOMETRY}
+\subsection{WHEEL RADII}
 The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were
-measured by measuring the linear distance traversed along the ground through
+estimated by measuring the linear distance traversed along the ground through
 either 13 or 14 rotations of the wheel. Each wheel was measured separately and
 the measurements were taken with a 72kg rider seated on the bicycle. A 30 meter
-tape measure (resolution: 2mm) was taped on a flat level smooth floor. The tire
-was marked with chalk and aligned with the tape measure. The accuracy of the
-distance measurement is approximately $\pm0.005$m. The tires were pumped to the
+tape measure (resolution: 2mm) was pulled taught and taped on a flat level smooth floor. The tire
+was marked with chalk and aligned with the tape measure
+Fig.~\ref{fig:tireChalk}. The accuracy of the
+distance measurement is approximately $\pm0.01$m. The tires were pumped to the
 recommended inflation pressure before the measurements. The wheel radius is
@@
 \begin{equation}
-	r=\frac{\textrm{distance}}{2\cdot\pi\cdot\textrm{rotations}}
+	r\pm\sigma_r=
+    \frac{d}{2\pi n}
+    \pm\left(\frac{\sigma_d}{2\pi n}\right)
 	\label{eq:wheelRadius}
-\end{equation}
-\begin{equation}
-    \sigma_r^{2}=\left(\frac{\sigma_d}{2\pi n}\right)^{2}
-    \label{eq:WheelRadVar}
 \end{equation}
@@
 \begin{table}
+    \centering
 	\begin{tabular}{llllllll}
@@
 	Front\\
-	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 4      \\
-	Distance  & 28.060 & 27.980 & 27.980 & 29.044 & 29.366 & 27.772 & 27.925 \\
-	Rotations & 13     & 13     & 13     & 14     & 14     & 13     & 13     \\
-	Radius    & 0.344  & 0.343  & 0.343  & 0.330  & 0.334  & 0.340  & 0.342  \\
+    \hline
+    $p_\mathrm{F}$ [bar] & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 4      \\
+    $d_\mathrm{F}$ [m]   & 28.060 & 27.980 & 27.980 & 29.044 & 29.366 & 27.772 & 27.925 \\
+    $n_\mathrm{F}$ []    & 13     & 13     & 13     & 14     & 14     & 13     & 13     \\
+    \hline
 	Rear\\
-	Pressure  & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 3.5    \\
-	Distance  & 27.850 & 27.835 & 27.768 & 29.788 & 29.212 & 29.774 & 27.884 \\
-	Rotations & 13     & 13     & 13     & 14     & 14     & 14     & 13     \\
-	Radius    & 0.341  & 0.341  & 0.340  & 0.339  & 0.332  & 0.338  & 0.341
+    \hline
+	$p_\mathrm{R}$ [bar] & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 3.5    \\
+	$d_\mathrm{R}$       & 27.850 & 27.835 & 27.768 & 29.788 & 29.212 & 29.774 & 27.884 \\
+	$n_\mathrm{R}$ []    & 13     & 13     & 13     & 14     & 14     & 14     & 13     \\
 	\end{tabular}
-	\caption{}
+	\caption{Raw data for the wheel radius measurements.}
 	\label{tab:wheelRadius}
 \end{table}
-\subsection{Head tube angle}
+\subsection{HEAD TUBE ANGLE}
 The head tube angle was measured directly using an electronic level with a
-$\pm0.2^{\circ}$ accuracy. The bicycle was fixed vertically, the front wheel
-was aligned with the rear frame and the bicycle was unloaded. The steer axis
+$\pm0.2^{\circ}$ accuracy. The bicycle frame was fixed perdenpicular to the
+ground, the steering angle was set to the nominal, the tire pressures were are
+recommended levels and the bicycle was unloaded. The steer axis
 tilt $\lambda$ is the complement to the head tube angle.
 \begin{equation}
-    \lambda=\frac{\pi}{180^{\circ}}(90^{\circ}-\lambda_{ht})
+    \lambda\pm\sigma_\lambda
+    =\frac{\pi}{180^{\circ}}(90^{\circ}-\lambda_{ht})
+    \pm\left(\frac{\pi}{180^\circ}\right)\sigma_{\lambda_{ht}}
 \label{eq:headTubeAngle}
-\end{equation}
-\begin{equation}
-    \sigma_\lambda^{2}=-\sigma_{\lambda_{ht}}^{2}\left(\frac{\pi}{180^\circ}\right)^2
-    \label{eq:HeadTubeAngleVar}
-\end{equation}
 \begin{table}

### 2010-10-13 papers/BMD2010/PhysicalParametersPaper/paper.tex
about to delete the mini tables
taking a break

@@
 \end{figure}
-\begin{table}
-    \centering
-	\begin{tabular}{llllllll}
-	Bicycle   & B      & B*     & C      & G      & P      & S      & Y \& Y*\\
-    \hline
-	Front\\
-    \hline
-    $p_\mathrm{F}$ [bar] & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 4      \\
-    $d_\mathrm{F}$ [m]   & 28.060 & 27.980 & 27.980 & 29.044 & 29.366 & 27.772 & 27.925 \\
-    $n_\mathrm{F}$ []    & 13     & 13     & 13     & 14     & 14     & 13     & 13     \\
-    \hline
-	Rear\\
-    \hline
-	$p_\mathrm{R}$ [bar] & 3.5    & 3.5    & 4      & 4      & 6.9    & 4.5    & 3.5    \\
-	$d_\mathrm{R}$       & 27.850 & 27.835 & 27.768 & 29.788 & 29.212 & 29.774 & 27.884 \\
-	$n_\mathrm{R}$ []    & 13     & 13     & 13     & 14     & 14     & 14     & 13     \\
-	\end{tabular}
-	\caption{Raw data for the wheel radius measurements.}
-	\label{tab:wheelRadius}
-\end{table}
 \subsection{HEAD TUBE ANGLE}
@@
 \label{eq:headTubeAngle}
-\begin{table}
-	\begin{tabular}{lllllllll}
-    \hline
-	Bicycle   & B      & B*     & C      & G      & P      & S      & Y      & Y*\\
-	Head tube angle & 67.1 & 67.1 & 69.0 & 71.1 & 74.2 & 73.1 & 72.7 & 70.6\\
-	Steer axis tilt & 0.400 & 0.400 & 0.367 & 0.330 & 0.276 & 0.295 & 0.302 & 0.339
-	\end{tabular}
-	\caption{Head tube angle and steer axis tilt}
-	\label{tab:lambda}
-\end{table}
-
-\subsection{Trail}
+\end{equation}
+\subsection{TRAIL}
 Trail is difficult to measure directly so we instead chose to measure the fork
@@
 \end{equation}
-\begin{table}
-	\begin{tabular}{lllllllll}
-	Bicycle     & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
-    \hline
-	Fork offset & 0.071 & 0.071 & 0.045 & 0.039 & 0.032 & 0.045 & 0.057 & -0.057\\
-	Trail       & 0.069 & 0.068 & 0.083 & 0.072 & 0.062 & 0.056 & 0.047 & 0.180
-	\end{tabular}
-	\caption{Head tube angle and steer axis tilt}
-	\label{tab:trail}
-\end{table}
 \subsection{Wheelbase}
@@
 in the same upright fixed position used when measuring the head tube angle.
-\begin{table}
-	\begin{tabular}{lllllllll}
-	Bicycle    & B     & B*    & C     & G     & P     & S     & Y     & Y*\\
-    \hline
-	Wheel base & 1.121 & 1.121 & 1.101 & 1.070 & 0.989 & 1.037 & 1.089 & 0.985
-	\end{tabular}
-	\caption{Wheelbase values.}
-	\label{tab:wheelbase}
-\end{table}
-
 \section{Mass}
@@
     \begin{center}
-        \includegraphics[width=4in]{../../../images/massScale.jpg}
+        \includegraphics[width=3in]{../../../images/massScale.jpg}
     \end{center}
@@
 \end{figure}
-\begin{table}
-    \centering
-	\begin{tabular}{llllllllll}
-		Bicycle &     $\sigma_m$ & B     & B*    & C     & G     & P    & S     & Y     & Y*\\
-        \hline
-		total   &    0.1       & 18.5  & 23.5  & 20.9  & 10.5  & 9.9  & 17.6  & 10.2  & 10.2\\
-        \hline
-		frame   &    0.02      & 9.86  & 14.71 & 9.18  & 4.48  & 4.49 & 7.22  & 3.31  & 3.31\\
-		fork    &    0.02      & 3.22  & 3.22  & 4.57  & 2.52  & 2.27 & 3.04  & 2.45  & 2.45\\
-		rear wheel & 0.02      & 3.11  & 3.11  & 3.96  & 1.94  & 1.38 & 3.96  & 2.57  & 2.57\\
-		front wheel& 0.02      & 2.02  & 2.02  & 3.55  & 1.50  & 1.58 & 3.33  & 1.90  & 1.90\\
-		total sum  & 0.04      & 18.21 & 23.06 & 21.26 & 10.44 & 9.72 & 17.55 & 10.23 & 10.23
-	\end{tabular}
-	\caption{Mass of bicycles and parts}
-	\label{tab:mass}
-\end{table}
 \section{Center of Mass}
@@
     \begin{center}
-        \includegraphics[width=4in]{../../../images/pendDist.jpg}
+        \includegraphics[width=3in]{../../../images/pendDist.jpg}
     \end{center}
@@
 can be used to populate the canonical form
-\begin{equation}
-\mathbf{M\ddot{q}}+v\mathbf{C}_1\mathbf{\dot{q}}+\left[g\mathbf{K}_0+v^2\mathbf{K}_2\right]\mathbf{q}=0
-\label{eq:canonical}
-\end{equation}
 of the linear benchmark equations of motion presented in
-\cite{Meijaard2007}. The coefficient matrices for the
-example rider and bicycle follow in Eqs.~\ref{eq:M}-\ref{eq:K2}.
+\cite{Meijaard2007}.
 \subsection{Eigenvalues}
@@
     \caption{The root loci with speed as the parameter for the Crescendo.}
-    \label{fig:StratosFork}
+    \label{fig:rootloci}
 \end{figure}

### 2010-10-13 papers/BMD2010/PhysicalParametersPaper/paper.tex
cleaning up the paper, combined some subfigures

@@
 \end{tabular} \\ \vspace{2.5ex}
-
 \end{center}
-
 \section*{ABSTRACT}
@@
 \subsection{HEAD TUBE ANGLE}
+\label{sec:headtube}
 The head tube angle was measured directly using an electronic level with a
@@
 \end{equation}
-\subsection{Wheelbase}
-The wheelbase was measured directly using a tape measure while the bicycle was
-in the same upright fixed position used when measuring the head tube angle.
-\section{Mass}
+\subsection{WHEELBASE}
+We measured the wheelbase when the bicycle was in the same nominal
+configuration as described in Sec.~\ref{sec:headtube}. We used a tape measure
+to measure the distance from one wheel axle center to the other with a 0.002 m
+accuracy.
+\section{MASS}
 The total mass of each bicycle was measured using a spring scale with a
-resolution of 100g. Then each of the four bicycle parts were measured using a
-Molen 20kg scale with a resolution of 20g.
+resolution of 100 grams. The total mass was only used for comparison purposes.
+Each of the four bicycle parts were measured using a Molen 20 kilogram scale with a
+resolution of 20 grams. The accuracy was conservatively assumed to also be
+$\pm20$ grams.
 \begin{figure}[htbp]
@@
 \end{figure}
-\section{Center of Mass}
-\subsection{Wheels}
-The centers of mass of the wheels are assumed to be at their geometrical centers.
-\subsection{Rear frame}
-The rear frame was hung in three orientations as a torsional pendulum (the
-pendulum was used to measure moments of inertia). I
-assumed that the frame was laterally symmetric. The
-frame could rotate about a joint such that gravity aligned the center of mass
-with the pendulum axis. The orientation angle of the headtube $\alpha$
-Figure~\ref{fig:angles} relative
-to the earth was
-measured using the digital level Figure~\ref{fig:level}. A string was aligned with the pendulum axis
+\section{CENTER OF MASS}
+\subsection{WHEELS}
+The centers of mass of the wheels are assumed to be at their geometrical
+centers to comply with the Whipple model.
+\subsection{REAR FRAME}
+The rear frame was hung in three orientations as a torsional pendulum (both for
+the center of mass measurements and the moment of inertia measurements
+described in Sec.~\ref{sec:moi}). We assumed that the frame was laterally
+symmetric, complying with the Whipple model. The frame could rotate about a joint
+such that gravity aligned the center of mass with the pendulum axis. The
+orientation angle of the headtube, $\alpha_\mathrm{B}$, Fig.~\ref{fig:angles} relative
+to the earth was measured using a digital level ($\pm0.2^{\circ}$ accuracy),
+Figure~\ref{fig:level}. A string was aligned with the pendulum axis
 and allowed to pass by the frame. The horizontal distance $a_\mathrm{B}$ between the rear
-axle and the string (CoM line) was measured by aligning a ruler perpendicular to
-the string. The distance $a$ was negative if the string fell to the right of
+axle and the string was measured by aligning a ruler perpendicular to
+the string. The distance $a_\mathrm{B}$ was negative if the string fell to the right of
 the rear axle and positive if it fell to the left of the rear axle. These
@@
 global reference frame.
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=3in]{../../../images/pendDist.jpg}
-    \end{center}
-    \caption{Measuring the distance from the pendulum axis to the rear wheel
-    axle using level ruler.}
-    \label{fig:PendDist}
+\begin{figure}[tb]
+    \centering
+    \subfloat[]{\label{fig:angles}\includegraphics[width=3in]{../../../figures/angles.pdf}}
+    \subfloat[]{\label{fig:triangle}\includegraphics[width=3in]{../../../figures/triangle.pdf}}
+    \caption{\subref{fig:angles} Pictorial description of the angles and dimensions that related
+    the nominal bicycle reference frame $XYZ_B$ with the pendulum reference frame
+    $XYZ_P$. \subref{fig:triangle} Exaggerated intersection of the three pendulum axes and the
+    location of the center of mass.}
 \end{figure}
@@
 	\begin{center}
-		\includegraphics[width=4in]{../../../images/level.jpg}
+		\includegraphics[width=3in]{../../../images/level.jpg}
 	\end{center}
@@
 \end{figure}
-The frame rotation angle $\beta$ is defined as rotation of the frame in the
+\begin{figure}[htbp]
+    \begin{center}
+        \includegraphics[width=3in]{../../../images/pendDist.jpg}
+    \end{center}
+    \caption{Measuring the distance from the pendulum axis to the rear wheel
+    axle using level ruler.}
+    \label{fig:PendDist}
+\end{figure}
+The frame rotation angle $\beta_\mathrm{B}$ is defined as rotation of the frame in the
 nominal configuration to the hanging orientation, rotated about the $Y$ axis.
-\begin{figure}[tb]
-	\begin{center}
-		\includegraphics[width=4in]{../../../figures/angles.pdf}
-	\end{center}
-	\caption{Pictorial description of the angles and dimensions that related
-    the nominal bicycle reference frame $XYZ_B$ with the pendulum reference frame
-    $XYZ_P$.}
-	\label{fig:angles}
-\end{figure}
 \begin{equation}
@@
 \end{equation}
-The pendulum axis $X_P$ is simply a line in the nominal bicycle reference frame with a slope $m$ and a z-intercept $b$. The slope can be shown to be
+The center of mass can be found by realizing that the pendulum axis $X_P$ is
+simply a line in the nominal bicycle reference frame with a slope $m$ and a
+z-intercept $b$ where the $i$ subscript corresponds the different frame
+orientations~\ref{fig:triangle}. The slope can be shown to be
 \begin{equation}
@@
 \begin{equation}
-   y = m_ix+b_i
+   z = m_ix+b_i
    \label{eq:line}
 \end{equation}
-The CoM location can be calculated by finding the intersection of these three
+The mass center location can be calculated by finding the intersection of these three
 lines. Two approaches were used used to calculate the center of mass. Intuition
-leads me to think that the center of mass is located at the centroid of the
+lead us to think that the center of mass is located at the centroid of the
 triangle made by the three intersecting lines. The centroid can be found by
@@
 three intersection points.
-\begin{figure}[htpb]
-    \begin{center}
-        \includegraphics{../../../figures/triangle.pdf}
-    \end{center}
-    \caption{Exaggerated intersection of the three pendulum axes and the
-    location of the center of mass.}
-    \label{fig:triangle}
-\end{figure}
 \begin{equation}
@@
 \section{Moment of Inertia}
+\label{sec:moi}
 The moments of inertia of the wheels, frame and fork were measured by taking

### 2010-10-14 papers/BMD2010/PhysicalParametersPaper/paper.tex
sizing figures mainly

@@
 \end{equation}
-\section{Moment of Inertia}
+\section{MOMENT OF INERTIA}
 \label{sec:moi}
@@
 example of the beating like phenomena.
-\begin{figure}[htpb]
+\begin{figure}[tpb]
     \begin{center}
@@
 radial moment of inertia can can calculated as such:
-
-The torsional pendulum was calibrated using a known moment of inertia. 
-\begin{figure}[htbp]
+\begin{figure}[tbp]
     \begin{center}
@@
 \end{figure}
+The torsional pendulum was calibrated using a known moment of inertia.
 A torsional pendulum almost identical to the one used in
@@
 \end{equation}
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/CrescendoFwheelTorsionalFirst.jpg}
-    \end{center}
-    \caption{The front wheel of the Crescendo hung as a torsional pendulum.}
+\begin{figure}[tbp]
+    \centering
+    \subfloat[]{
         \label{fig:FwheelTor}
+        \includegraphics[width=3in]{../../../images/CrescendoFwheelTorsionalFirst.jpg}
+    }
+    \quad
+    \subfloat[]{
+        \label{fig:wheelIyy}
+        \includegraphics[width=3in]{../../../images/wheelIyy.jpg}
+    }
+    \caption{\subref{fig:FwheelTor} The front wheel of the Crescendo hung as a
+    torsional pendulum. \subref{fig:wheelIyy} A wheel hung as a compound pendulum.}
+    \label{fig:wheelPend}
 \end{figure}
@@
 gravity. Table  gives the calculated values.
-\begin{figure}[htpb]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/wheelIyy.jpg}
-    \end{center}
-    \caption{A wheel hung as a compoun pendulum.}
-    \label{fig:wheelIyy}
-\end{figure}
 \subsection{Frame}

### 2010-10-14 PhysicalParameters.py

@@ -531,6 +531,8 @@ def fit_data():
     filenames.sort()
     period = {}
     #for name in ['YellowRevForkTorsionalFirst1.p']:
+    #for name in ['BrowserFrameCompoundFirst1.p']:
+    #for name in ['CrescendoForkTorsionalFirst2.p']: # beating
     #for name in ['YellowFwheelCompoundFirst1.p']:
     #for name in ['StratosFrameCompoundFirst2.p']:
     for name in filenames:
@@ -567,20 +569,22 @@ def fit_data():
         f = wd/2./pi
         T = 1./f
         fig = plt.figure(1)
-        plot_osfit(x, y, lscurve, p1, rsq, T, fig=fig)
-        plt.savefig(direct + name[:-2] + '.png')
-        plt.close()
         # add a star in the R value is low
         if rsq <= 0.99:
-            rsq = str(rsq) + '*'
+            rsqstr = str(rsq) + '*'
+            m = max(x)
         else:
-            pass
+            rsqstr = str(rsq)
+            m = 5.
+        plot_osfit(x, y, lscurve, p1, rsq, T, m=m, fig=fig)
+        plt.savefig(direct + name[:-2] + '.png')
+        plt.close()
         # include the notes for the experiment
         try:
             note = pendDat['notes']
         except:
             note = ''
-        line = name + ',' + str(T) + ',' + str(rsq) + ',' + str(sigma) + ',' + str(note) + '\n'
+        line = name + ',' + str(T) + ',' + rsqstr + ',' + str(sigma) + ',' + str(note) + '\n'
         periodfile.write(line)
         print line
         # if the filename is already in the period dictionary...
@@ -1277,7 +1281,7 @@ def uround(value):
         s = str(value)
     return s
 
-def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
+def plot_osfit(t, ym, yf, p, rsq, T, m=None, fig=None):
     '''Plot fitted data over the measured
 
     Parameters:
@@ -1293,17 +1297,39 @@ def plot_osfit(t, ym, yf, p, rsq, T, fig=None):
         The r squared value of y (the fit)
     T : float
         The period
+    m : float
+        The maximum value to plot
 
     Returns:
     --------
     fig : the figure
 
     '''
+    # figure properties
[…]

### 2010-10-14 papers/BMD2010/PhysicalParametersPaper/paper.tex
adding more stuff

@@
 \usepackage{amsmath}
+\usepackage{pdflscape}
 
@@
 
+% title
 \begin{center}
@@
 \end{figure}
+\section{PARAMETERS}
+The 25 parameters can be estimated by many techniques. Where possible we
+measured the benchmark parameter directly.
 \section{ACCURACY}
@@
 global reference frame.
-\begin{figure}[tb]
+\begin{figure}[tbp]
     \centering
     \subfloat[]{\label{fig:angles}\includegraphics[width=3in]{../../../figures/angles.pdf}}
-    \subfloat[]{\label{fig:triangle}\includegraphics[width=3in]{../../../figures/triangle.pdf}}
+    \subfloat[]{\label{fig:triangle}\includegraphics[width=2.75in]{../../../figures/triangle.pdf}}
     \caption{\subref{fig:angles} Pictorial description of the angles and dimensions that related
@@
 \end{figure}
-\begin{figure}[tb]
+\begin{figure}[tbp]
 	\begin{center}
-		\includegraphics[width=3in]{../../../images/level.jpg}
+		\includegraphics[width=3in]{../../../images/YellowFrameTorsionalThird.jpg}
 	\end{center}
@@
     \begin{center}
-        \includegraphics[width=4in]{../../../images/fixture.jpg}
+        \includegraphics[width=3in]{../../../images/fixture.jpg}
     \end{center}
@@
         \label{fig:FwheelTor}
-        \includegraphics[width=3in]{../../../images/CrescendoFwheelTorsionalFirst.jpg}
+        \includegraphics[width=2.75in]{../../../images/CrescendoFwheelTorsionalFirst.jpg}
     }
@@
         \label{fig:wheelIyy}
-        \includegraphics[width=3in]{../../../images/wheelIyy.jpg}
+        \includegraphics[width=2.75in]{../../../images/wheelIyy.jpg}
     }
@@
 and the unknown global inertia tensor can be solved for.
-\begin{figure}[htbp]
+\begin{figure}[tbp]
+    \centering
+    \subfloat[]{
+        \label{fig:frameCompound}
+        \includegraphics[width=2.75in]{../../../images/YellowFrameCompoundFirst.jpg}
+        }
+    \subfloat[]{
+        \label{fig:forkCompound}
+        \includegraphics[width=2.75in]{../../../images/BrowserInsForkCompoundFirst.jpg}
+        }
+        \caption{\subref{fig:frameCompound} The rear frame hung as a compound
+        pendulum. \subref{fig:forkCompound} The Browser fork hung as a
+        compound pendulum.}
+        \label{fig:compound}
+\end{figure}
+\subsection{Fork}
+\begin{figure}[tbp]
+    \centering
+    \subfloat[]{
+        \label{fig:StratosFork}
+        \includegraphics[width=2.75in]{../../../images/StratosForkTorsionalThird.jpg}
+        }
+    \caption{\subref{fig:StratosFork} The Stratos fork and handlebar assembly hung as a torsional
+        pendulum.}
+    \label{fig:torsional}
+\end{figure}
+\section{LINEAR ANALYSIS}
+Once all of the bicycle parameters have been calculated the canonical matrices
+can be formed and the linear dynamics of the bicycles can be explored. We also
+added the same rigid rider to each bicycle for further comparison.
+\begin{table}[tbp]
+    \centering
+    \caption{Rider parameters}
+    \begin{tabular}{ll}
+        Parameter & Value\\
+        \hline
+        $m_\mathrm{P}$ [kg] & 72\\
+        $x_\mathrm{P}$ [m] & 0.2909\\
+        $z_\mathrm{P}$ [m] & -1.1091\\
+        $I_\mathrm{P}$ [$\mathrm{kg\ m}^2$] &
+            $\left[
+            \begin{array}{ccc}
+                 7.9985 & 0      & -1.9272\\
+                 0      & 8.0689 & 0\\
+                -1.9272 & 0      & 2.3624
+            \end{array}
+            \right]$
+    \end{tabular}
+\end{table}
+
+\subsection{EIGENVALUES}
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=4in]{../../../plots/Bike/eig_plot.pdf}
+    \caption{The eigenvalues verus speed for all eight bicycles.}
+    \label{fig:bikeEigPlot}
+\end{figure}
+\begin{figure}[tbp]
     \begin{center}
-        \includegraphics[width=4in]{../../../images/YellowFrameCompoundFirst.jpg}
-    \end{center}
-    \caption{The rear frame hung as a compound pendulum.}
-    \label{fig:FrameCompound}
-\end{figure}
-
-\subsection{Fork}
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/StratosForkTorsionalThird.jpg}
-    \end{center}
-    \caption{The Stratos fork and handlebar assembly hung as a torsional
-    pendulum.}
-    \label{fig:StratosFork}
-\end{figure}
-\section{Parameters}
-\section{Linear Analysis}
-\subsection{Canonical Matrices}
-The final results are presented in the form used by the benchmark model. These
-can be used to populate the canonical form
-of the linear benchmark equations of motion presented in
-\cite{Meijaard2007}.
-\subsection{Eigenvalues}
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../plots/Bike/CrescendoRootLoci.png}
+        \includegraphics[width=4in]{../../../plots/Bike/CrescendoRootLoci.pdf}
     \end{center}
@@
 \end{figure}
-
-\subsection{Frequency Response}
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=4in]{../../../plots/BikeRider/eig_plot.pdf}
+    \caption{The eigenvalues verus speed for all eight bicycles with the same
+    rider.}
+    \label{fig:bikeRiderEigPlot}
+\end{figure}
+\subsection{FREQUENCY RESPONSE}
 \bibliographystyle{acm}
 \bibliography{bicycle}
+\appendix
+\section{PARAMETERS}
+\subsection{BICYCLE}
+\begin{landscape}
+\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
+\end{landscape}
+\subsection{BICYCLE WITH RIDER}
+%\begin{landscape}
+%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
+%\end{landscape}
+\section{CANONICAL MATRICES}
+\subsection{BICYCLE}
+\begin{landscape}
+\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
+\end{landscape}
+\subsection{BICYCLE WITH RIDER}
+%\begin{landscape}
+%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
+%\end{landscape}
 \end{document}

### 2010-10-14 PhysicalParameters.py

@@ -373,7 +373,15 @@ def bike_bode_plots(typ='Bike', speeds=None):
 
     #plt.show()
 
-def bike_eig_plots(typ='Bike'):
+def bike_eig_plots(typ='Bike', filetype='pdf'):
+    '''
+    Parameters:
+    -----------
+    typ : string
+        Bike, BikeRider, BikeLegs
+    filetype: string
+        a matplotlib graphics export type
+    '''
 
     direct = 'data/' + typ + '/Canonical/'
     fend = 'Can.p'
@@ -396,7 +404,7 @@ def bike_eig_plots(typ='Bike'):
     vc = np.zeros(nBk)
 
     # figure properties
-    figwidth = 3.25 # in inches
+    figwidth = 4. # in inches
     goldenMean = (np.sqrt(5)-1.0)/2.0
     figsize = [figwidth, figwidth*goldenMean]
     params = {#'backend': 'ps',
@@ -448,7 +456,7 @@ def bike_eig_plots(typ='Bike'):
         plt.title('{name}\nEigenvalues vs Speed'.format(name=data['bikes'][i]))
         plt.xlabel('Speed [m/s]')
         plt.ylabel('Real and Imaginary Parts of the Eigenvalue [1/s]')
-        plt.savefig(directp + '/' + name + 'EigPlot.png')
+        plt.savefig(directp + '/' + name + 'EigPlot.' + filetype)
         # plot root loci
         plt.figure(nBk + i)
         for j in range(len(evals[0, :])):
@@ -458,7 +466,7 @@ def bike_eig_plots(typ='Bike'):
         plt.grid()
         plt.axis('equal')
         plt.title('{name}\nEigenvalues vs Speed'.format(name=data['bikes'][i]))
-        plt.savefig(directp + '/' + name + 'RootLoci.png')
+        plt.savefig(directp + '/' + name + 'RootLoci.' + filetype)
         # plot all bikes on the same plot
         plt.figure(2*nBk + 1)
         plt.plot(vel, np.abs(np.imag(wea['evals'])), color=colors[i], label='_nolegend_', linestyle='--')
@@ -475,7 +483,7 @@ def bike_eig_plots(typ='Bike'):
     plt.xlabel('Speed [m/s]')
     plt.ylabel('Real and Imaginary Parts of the Eigenvalue [1/s]')
     try:
-        plt.savefig(directp + '/eig_plot.png')
[…]

### 2010-10-14 papers/BMD2010/PhysicalParametersPaper/paper.tex
more paper work

@@
 can be formed and the linear dynamics of the bicycles can be explored. We also
-added the same rigid rider to each bicycle for further comparison.
+added the same rigid rider to each bicycle for further comparison. The rigid
+rider was assumed to be in same position and posture for each bicycle relative
+to the rear wheel contact point.
 \begin{table}[tbp]
     \centering
-    \caption{Rider parameters}
     \begin{tabular}{ll}
@@
     \end{tabular}
+    \caption{Mass, center of mass and moment of inertia for the Person relative
+    to the benchmark coordinate system from~\cite{Moore2009a}}
+    \label{tab:riderParam}
 \end{table}
-
 \subsection{EIGENVALUES}
@@
 \end{figure}
+The eigenvalues can be plotted versus forward speed.
+Figure~\ref{fig:bikeEigPlot} shows that the bikes have the typical
+characteristics of the Whipple bicycle model: four real roots at very slow
+speeds, two of which are unstable; a complex pair that is unstable and lower
+speeds and stable at higher speeds; and a root that is sligthly unstable at
+higher speeds. The one noticable difference is that the capsize and caster
+modes colasece into a complex pair between about 0.5 and 3 m/s. The frequency
+of oscillation is similar to that of the weave mode. The root locus in the real
+and imaginary plane, Fig.~\ref{fig:rootloci}, shows that the mode damps out
+quickly. Examining the eigenvectors reveals that the mode is steer leading
+roll with a 90 degree phase, both of their magitudes being similar,
+Fig~\ref{fig:evec}.
 \begin{figure}[tbp]
@@
 \begin{figure}[tbp]
+    \begin{center}
+        \includegraphics[width=3in]{../../../plots/cres.pdf}
+    \end{center}
+    \caption{The root loci with speed as the parameter for the Crescendo.}
+    \label{fig:evec}
+\end{figure}
+With the rider added, the second complex pair disappears and the bikes have the
+typical characteristics of the Whipple model. Reversing the fork on the yellow
+bike lowers the weave critical speed. Also the addition of weight to the rear
+rack of the Browser does blank.
+\begin{figure}[tbp]
     \centering
@@
 \subsection{FREQUENCY RESPONSE}
+The frequency response of the bicycles and bicycle with rider also reveal some
+interesting things. In the steer torque to roll Bode diagram the magnitude
+diffence among bicycles can vary up to 10 dB (0.148 radians per newton meter of
+torque).
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=4in]{../../../plots/Bike/Bode/Tdel2phi.pdf}
+    \caption{The eigenvalues verus speed for all eight bicycles with the same
+    rider.}
+    \label{fig:bikeBode}
+\end{figure}
+\begin{figure}[tbp]
+    \centering
+    \includegraphics[width=4in]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
+    \caption{The eigenvalues verus speed for all eight bicycles with the same
+    rider.}
+    \label{fig:bikeRiderBode}
+\end{figure}
+
 \bibliographystyle{acm}
@@
 \appendix
+\section{RAW DATA}
+%\begin{landscape}
+%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
+%\end{landscape}
 \section{PARAMETERS}

### 2010-10-14 PhysicalParameters.py

@@ -12,6 +12,37 @@ import scipy.io.matlab.mio as mio
 import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 
+def cres_evec():
+    f = open('data/cres.p', 'r')
+    cres = pickle.load(f)
+    f.close()
+    # get the 100th evec corresponding to the first eigenvalue
+    evec = cres['evecs'][100, : , 0]
+    # figure properties
+    figwidth = 3. # in inches
+    goldenMean = (np.sqrt(5)-1.0)/2.0
+    figsize = [figwidth, figwidth*goldenMean]
+    params = {
+        'axes.labelsize': 8,
+        'text.fontsize': 8,
+        'axes.titlesize': 8,
+        'legend.fontsize': 6,
+        'xtick.labelsize': 6,
+        'ytick.labelsize': 6,
+        'figure.figsize': figsize
+        }
+    plt.rcParams.update(params)
+    plt.figure(figsize=figsize)
+    plt.axes([0.15, 0.15, 0.95-0.15, 0.85-0.15])
+    plt.grid()
+    for vec in evec:
+        plt.plot([0., vec.real], [0., vec.imag])
+    plt.title('$v=1.001$ m/s, $\lambda=-3.83+0.48j$')
+    plt.xlabel('Real')
+    plt.ylabel('Imaginary')
+    plt.legend(['Roll Rate', 'Steer Rate', 'Roll Angle', 'Steer Angle'])
+    plt.savefig('plots/cres.pdf')
+
 def mat2dic():
 
     # load the main data file into a dictionary
@@ -367,7 +398,7 @@ def bike_bode_plots(typ='Bike', speeds=None):
             print i, colors[i]
             plt.setp(line, color=colors[i])
             plt.setp(v.ax2.lines[i], color=colors[i])
-        v.savefig(direct + '/' + k + '.png')
+        v.savefig(direct + '/' + k + '.pdf')
 
     return plots
 
@@ -381,8 +412,10 @@ def bike_eig_plots(typ='Bike', filetype='pdf'):
         Bike, BikeRider, BikeLegs
     filetype: string
         a matplotlib graphics export type
+
     '''
 
+    # load data from here
     direct = 'data/' + typ + '/Canonical/'
     fend = 'Can.p'
 
@@ -391,41 +424,37 @@ def bike_eig_plots(typ='Bike', filetype='pdf'):
     data = pickle.load(f)
     f.close()
 
[…]

### 2010-10-15 PhysicalParameters.py
found an error in the tor and com assigments to the wheel torsional inertias

@@ -656,6 +656,7 @@ def tor_com():
     # list of the orientation angles
     fst = ['First', 'Second', 'Third']
     # list of the bicycle parts
+    # this is the order that they are put into tor and com
     fffr = ['Frame', 'Fork', 'Fwheel', 'Rwheel']
     # list of type of pendulums
     tc = ['Torsional', 'Compound']
@@ -845,6 +846,20 @@ def calc_parameters():
     com[2, 7] = com[2, 6]
     com[3, 7] = com[3, 6]
 
+    ddU['TB1'] = tor[0]
+    ddU['TB2'] = tor[1]
+    ddU['TB3'] = tor[2]
+    ddU['TH1'] = tor[3]
+    ddU['TH2'] = tor[4]
+    ddU['TH3'] = tor[5]
+    ddU['TF1'] = tor[6]
+    ddU['TR1'] = tor[9]
+    ddU['TRod'] = avgPer['rodPer']
+    ddU['TByy'] = com[0]
+    ddU['THyy'] = com[1]
+    ddU['TFyy'] = com[2]
+    ddU['TRyy'] = com[3]
+
     tRod = avgPer['rodPer']
     # calculate the stiffness of the torsional pendulum
     iRod = tube_inertia(ddU['lRod'], ddU['mRod'], ddU['rRod'], 0.)[1]
@@ -858,14 +873,15 @@ def calc_parameters():
 
     # calculate the wheel y inertias
     par['g'] = 9.81*np.ones(ddU['forkMass'].shape, dtype=float)
-    par['IRyy'] = com_inertia(par['mR'], par['g'], ddU['rWheelPendLength'], com[3, :])
     par['IFyy'] = com_inertia(par['mF'], par['g'], ddU['fWheelPendLength'], com[2, :])
+    par['IRyy'] = com_inertia(par['mR'], par['g'], ddU['rWheelPendLength'], com[3, :])
 
     # calculate the wheel x/z inertias
-    par['IRxx'] = tor_inertia(k, tor[6, :])
-    par['IFxx'] = tor_inertia(k, tor[9, :])
+    par['IFxx'] = tor_inertia(k, tor[6, :])
+    par['IRxx'] = tor_inertia(k, tor[9, :])
 
     # calculate the y inertias for the frame and fork
+    # the coms may be switched here
     framePendLength = (frameCoM[0, :]**2 + (frameCoM[1, :] + par['rR'])**2)**(0.5)
[…]

### 2010-10-15 PhysicalParameters.py
now saves the average periods to the udata file

@@ -860,6 +860,10 @@ def calc_parameters():
     ddU['TFyy'] = com[2]
     ddU['TRyy'] = com[3]
 
+    f = open('data/udata.p', 'w')
+    pickle.dump(ddU, f)
+    f.close()
+
     tRod = avgPer['rodPer']
     # calculate the stiffness of the torsional pendulum
     iRod = tube_inertia(ddU['lRod'], ddU['mRod'], ddU['rRod'], 0.)[1]

### 2010-10-15 PhysicalParameters.py
leaving the library

@@ -206,7 +206,7 @@ def make_tables(typ='Bike'):
     for i, name in enumerate(data['bikes']):
         fname = ''.join(name.split())
         # open the new file
-        f = open('tables/' + 'ParameterTableTemplate.tex', 'r')
+        f = open('tables/ParameterTableTemplate.tex', 'r')
         fn = open(direct + fname + 'Par.tex', 'w')
         for line in f:
             #print line
@@ -248,6 +248,8 @@ def make_tables(typ='Bike'):
     template = open('tables/MasterParTableTemplate.tex', 'r')
     final = open(direct + 'MasterParTable.tex', 'w')
 
+    os.system('cp papers/BMD2010/PhysicalParametersPaper/bmd2010p.cls ' + direct)
+
     abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
     for line in template:
         if line[0] == '%':
@@ -258,7 +260,11 @@ def make_tables(typ='Bike'):
                 if varname == 'bike':
                     fline = fline + ' & \multicolumn{2}{c}{' + abbrev[i] + '}'
                 else:
-                    val, sig = uround(par[varname][i]).split('+/-')
+                    try:
+                        val, sig = uround(par[varname][i]).split('+/-')
+                    except ValueError:
+                        val = str(round(par[varname][i], 3))
+                        sig = '0.0'
                     fline = fline + ' & ' + val + ' & ' + sig
             fline = fline + r'\\' + '\n'
             final.write(fline)
@@ -276,6 +282,8 @@ def make_tables(typ='Bike'):
 
     template = open('tables/MasterCanTableTemplate.tex', 'r')
     final = open(direct + 'MasterCanTable.tex', 'w')
+
+    os.system('cp papers/BMD2010/PhysicalParametersPaper/bmd2010p.cls ' + direct)
 
     abbrev = ['B', 'B*', 'C', 'G', 'P', 'S', 'Y', 'Y*']
     for line in template:

### 2010-10-16 papers/BMD2010/PhysicalParametersPaper/paper.tex
found an error in the tor and com assigments to the wheel torsional inertias
leaving the library
rough draft ready

@@
 \end{description}
-\begin{figure}[htbp]
+\begin{figure}[tbp]
     \centering
@@
 \section{MASS}
-The total mass of each bicycle was measured using a spring scale with a
-resolution of 100 grams. The total mass was only used for comparison purposes.
-Each of the four bicycle parts were measured using a Molen 20 kilogram scale with a
-resolution of 20 grams. The accuracy was conservatively assumed to also be
-$\pm20$ grams.
-\begin{figure}[htbp]
+\begin{figure}[tbp]
     \begin{center}
@@
 \end{figure}
+The total mass of each bicycle was measured using a spring scale with a
+resolution of 100 grams. The total mass was only used for comparison purposes.
+Each of the four bicycle parts were measured using a Molen 20 kilogram scale with a
+resolution of 20 grams. The accuracy was conservatively assumed to also be
+$\pm20$ grams.
 \section{CENTER OF MASS}
@@
 \begin{figure}[tbp]
-	\begin{center}
-		\includegraphics[width=3in]{../../../images/YellowFrameTorsionalThird.jpg}
-	\end{center}
-	\caption{The digital level was mounted to a straight edge that was aligned
+    \centering
+    \subfloat[]{\label{fig:level}\includegraphics[width=2.75in]{../../../images/YellowFrameTorsionalThird.jpg}}
+    \quad
+    \subfloat[]{\label{fig:PendDist}\includegraphics[width=2.75in]{../../../images/pendDist.jpg}}
+    \caption{\subref{fig:level} The digital level was mounted to a straight edge that was aligned
     with the headtube of the bicycle frame. This was done without allowing the
@@
     submerging a low hanging area of the frame into a bucket of water to
-    decrease the oscillation.}
-	\label{fig:level}
-\end{figure}
-\begin{figure}[htbp]
-    \begin{center}
-        \includegraphics[width=3in]{../../../images/pendDist.jpg}
-    \end{center}
-    \caption{Measuring the distance from the pendulum axis to the rear wheel
-    axle using level ruler.}
-    \label{fig:PendDist}
+    decrease the oscillation. \subref{fig:PendDist} Measuring the distance from
+    the pendulum axis to the rear wheel axle using level ruler.}
 \end{figure}
@@
 compound and torsional pendulums and measuring their period of oscillation when
-perturb at small angles. The rate of oscillation was measured was measured
+perturb at small angles. The rate of oscillation was measured
 using a \href{http://www.siliconsensing.com/CRS03}{Silicon Sensing CRS03 100
@@
 Instruments USB-6008 12 bit data acquisition unit} and
-Matlab. The measurement duration were either 15 or 30 secs and each moment of
+\textsc{Matlab}. The measurement duration were either 15 or 30 secs and each moment of
 inertia measurement was performed three times. No extra care was taken to
@@
 \end{figure}
-The physical phenomena that was observed for data sets such as these were that
+The physical phenomena observed for corresponding to data sets such as these were that
 the bicycle frame or fork was perturbed torsionally. After set into motion the
@@
 prevent the swinging motion of the pendulum without damping the torsional
-motion. Roland and Massing~\cite{Roland1971} noted that they had this same problem and solved it
-by hold the torsional rod in a fixed bearing.
+motion.
 
@@
 where $\mathbf{H}$ is the Hessian~\cite{Hubbard1989b}. $\mathbf{U}$ is a $5\times5$ matrix with the
-variances of each of the five fit parameters along the diagonal. This assumes
-that the parameters are uncorrelated [note for Jason: you may have to take into
-account correlation for this..maybe checkout the R function for nls]. The
-variance of $T$ can be computed using the variance of $\zeta$ and $\omega$. It
+variances of each of the five fit parameters along the diagonal.
+The variance of $T$ can be computed using the variance of $\zeta$ and $\omega$. It
 is important to note that the uncertainties in the period are very low
-($<10e-4$), even for the fits with low $r^2$ values due to the beating issue.
-\subsection{Torsional Pendulum}
+($<10e-4$), even for the fits with low $r^2$ values.
+\subsection{TORSIONAL PENDULUM}
 A torsional pendulum was used to measure all of moments of inertia about axes
@@
 various lower clamps.
-\begin{figure}[htbp]
+\begin{figure}[tbp]
     \begin{center}
-        \includegraphics[width=3in]{../../../images/fixture.jpg}
+        \includegraphics[width=2in]{../../../images/fixture.jpg}
     \end{center}
-    \caption{The rigid pendulumfixture mounted to a concrete column.}
+    \caption{The rigid pendulum fixture mounted to a concrete column.}
     \label{fig:fixture}
 \end{figure}
-A 5mm diameter, 1m long mild steel rod was used as the torsion
+A 5 mm diameter, 1 m long mild steel rod was used as the torsion
 spring. A lightweight, low relative moment of inertia clamp was constructed
-that could clamp the rim and the tire. The wheel was hung freely such that the
+that could clamp the rim and the tire. The moments of inertia of the clamps
+were neglected. The wheel was hung freely such that the
 center of mass aligned with the torsional pendulum axis and then it was
-secured. The wheel was then perturbed and it oscillated about the pendulum axis.
-The rate gyro was mounted on the clamp in line with the pendulum axis. The
-radial moment of inertia can can calculated as such:
-\begin{figure}[tbp]
-    \begin{center}
-        \includegraphics[width=4in]{../../../images/rod.jpg}
-    \end{center}
-    \caption{The steel calibration rod.}
-    \label{fig:rod}
-\end{figure}
-The torsional pendulum was calibrated using a known moment of inertia.
-A torsional pendulum almost identical to the one used in
+secured. The wheel was then perturbed and it oscillated about the pendulum
+axis. The rate gyro was mounted on the clamp orientated along the pendulum
+axis.
+
+The torsional pendulum was calibrated using a known moment of
+inertia~\ref{fig:rod}. A torsional pendulum almost identical to the one used in
 \cite{Kooijman2006} was used to measure the averaged period $\overline{T}_i$ of
@@
 and allowed to oscillate about the pendulum axis through at least ten periods.
-The time of oscillation was recorded via a stopwatch ($\pm1$~s). This was done
-three times for each frame and the recorded times were averaged. The
-coefficient of elasticity $k$ for the torsional pendulum had previously been
-measured in \cite{Kooijman2006} and found to be $k=5.01\pm0.01$
-$\frac{\textrm{Nm}}{\textrm{rad}}$.
-
-\subsection{Wheels}
-The wheels are assumed to be symmetric laterally and about any radial axis. Thus only two
-moment's of inertia are required. The moment of inertia about the axle was
-measured by hanging the wheel as a compound pendulum. The wheel was hung on a
+This was done at least three times for each frame and the recorded periods were
+averaged.
+\begin{figure}[tbp]
+    \begin{center}
+        \includegraphics[width=4in]{../../../images/rod.jpg}
+    \end{center}
+    \caption{The steel calibration rod. The moment of inertia of the rod,
+    $I=\frac{m}{12}(3r^2+l^2)$, can be used to estimate the stiffness of the
+    pendulum, $k=\frac{4I\pi^2}{\overline{T}^2}$, with $k=5.62\pm0.02$ $\frac{\textrm{Nm}}{\textrm{rad}}$}
+    \label{fig:rod}
+\end{figure}
+\subsection{WHEELS}
+Finding the full inertia tensors of the wheels is less complex because the wheels
+are assumed symmetric about three orthogonal planes so there are no products of
+inertia. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring
+the averaged period of oscillation about an axis in the $XZ$-plane using the
+torsional pendulum setup and Eq.~\ref{eq:torPend}.The wheels are assumed to be symmetric laterally and about any radial axis. Thus only two
+moment's of inertia are required for the benchmark parameters set. The moment
+of inertia about the axle was measured by hanging the wheel as a compound
+pendulum, Fig.~\ref{fig:wheelIyy}. The wheel was hung on a
 horizontal rod and perturbed to oscillate about the axis of the rod. This rate
@@
 axle axis. The wheels tended to precess at the contact point about the vertical
-axis. This adds a very low frequency component of rate along the vertical axis,
-but should not affect the compound pendulum axis. A better fixture for the
-wheel could prevent the precession. The pendulum arm length is the distance
+axis whixh added a very low frequency component of rate along the vertical
+radial axis, but should not affect the period estimation about the compound
+pendulum axis. A fixture that prevented the precession would be preferable for
+future measurements. The pendulum arm length is the distance
 from the rod/rim contact point to the mass center of the wheel. The inner
-diameter of the rim was measured and divided by two to get $l$. The moment of
+diameter of the rim was measured and divided by two to get $l_\mathrm{F,R}$. The moment of
 inertia about the axle is calculated by:
@@
     m_\mathrm{R}l^2
+    \label{eq:CompoundInertia}
+\end{equation}
+\begin{equation}
+    need the sigma for IRyy
 \end{equation}
@@
 The radial moment of inertia was measured by hanging the wheel as a torsional
-pendulum. A 5mm diameter, 1m long mild steel rod was used as the torsion
-spring. A lightweight, low relative moment of inertia clamp was constructed
-that could clamp the rim and the tire. The wheel was hung freely such that the
-center of mass aligned with the torsional pendulum axis and then it was
-secured. The wheel was then perturbed and it oscillated about the pendulum axis.
-The rate gyro was mounted on the clamp in line with the pendulum axis. The
-radial moment of inertia can can calculated as such:
+pendulum, Fig.~\ref{fig:FwheelTor}. The wheel was hung freely such that the center of mass aligned with
+the torsional pendulum axis and then it was secured in position. The wheel was then
+perturbed and it oscillated about the pendulum axis. The radial moment of inertia can can calculated as such:
 \begin{equation}
@@
 \end{equation}
-Finding the inertia tensors of the wheels is less complex because the wheels
-are symmetric about three orthogonal planes so there are no products of
-inertia. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring
-the averaged period of oscillation about an axis in the $XZ$-plane using the
-torsional pendulum setup and Eq.~\ref{eq:torPend}. The $I_{yy}$ moment of
-inertia was calculated with a compound pendulum as described in
-\cite{Kooijman2006} and shown in Fig. using
 \begin{equation}
-	I_{yy}=\left(\frac{\overline{T}}{2\pi}\right)^2mgl-ml^2
-\label{eq:comPend}
+    need the sigma for Ixx
 \end{equation}
-where $l=0.303\pm0.002$ m is the pendulum length, $m$ is the mass of the wheel,
-$\overline{T}$ is the averaged period and $g$ is the local acceleration due to
-gravity. Table  gives the calculated values.
-\subsection{Frame}
+\subsection{FRAME}
 Three measurements were made to estimate the globally referenced moments and
@@
 frame was typically hung from the three main tubes: seat tube, down tube
-and top tube. The rear fender prevented easy connection to the seat tube on
+and top tube, Fig.~\ref{fig:level}. The rear fender prevented easy connection to the seat tube on
 some of the bikes and the clamp was attached to the fender. The fender was
-generally less rigid than the frame tube. For best accuracy with only three orientation angles, the frame
-should be hung at three angles that are $120^\circ$ apart. The three tubes on
-the frame generally provide that the orientation angles were spread evenly at
-about $120^\circ$. Taking data at more orientation angles could improve the accuracy further and
-is gernally possible with standard diamond frame bicylces.
+generally less rigid than the frame tube. For best accuracy with only three
+orientation angles, the frame should be hung at three angles that are
+$120^\circ$ apart. The three tubes on the frame generally provide that the
+orientation angles were spread evenly at about $120^\circ$. Furthermore, taking
+data at more orientation angles could improve the accuracy further and
+is generally possible with standard diamond frame bicylces.
 
-Three moments of inertia $J_{i}$ about the pendulum axes were calculated with
+Three moments of inertia $J_{i}$ about the pendulum axes were calculated using
+Eq.~\ref{eq:torPend}.
 \begin{equation}
@@
 \end{equation}
-
+\begin{equation}
+    need sigma
+\end{equation}
 The moments and products of inertia of the rear frame and handlebar/fork
@@
 \begin{equation}
-	\mathbf{J}_i=\mathbf{R}_i^T\mathbf{IR}_i
+	\mathbf{J}_i=\mathbf{R}_i\mathbf{IR}_i^T
 \label{eq:rotIn}
@@
 $\mathbf{I}$, is the inertia tensor in the global reference frame and
-$\mathbf{R}$ is the rotation matrix relating the two frames. The global inertia
+$\mathbf{R}$ is the rotation matrix relating the two frames,
+Fig.~\ref{fig:angles}. The global inertia
 tensor is defined as
@@
 	\begin{array}{rr}
-		I_{xx}  & -I_{xz}\\
-		-I_{xz} & I_{zz}
+		I_{xx}  & I_{xz}\\
+		I_{xz} & I_{zz}
 	\end{array}
@@
 assumed to be laterally symmetric and the $y$ axis of the pendulum reference is
-the same as the $y$ axis of the benchmark reference frame. The simple rotation matrix about the $Y$-axis
+the same as the $y$ axis of the benchmark reference frame. The simple rotation
+matrix about the $Y$-axis
 can similarly be reduced to a $2\times2$ matrix where $s_{\beta i}$ and
@@
 	\begin{array}{rr}
-		c_{\beta i} & s_{\beta i}\\
-		-s_{\beta i} & c_{\beta i}
+		c_{\beta i} & -s_{\beta i}\\
+		s_{\beta i} & c_{\beta i}
 	\end{array}
@@
 \end{equation}
-The first entry of $\mathbf{J}_i$ in Eq.~\ref{eq:rotIn} is the moment of inertia about the pendulum axis and is written explicitly as 
+The first entry of $\mathbf{J}_i$ in Eq.~\ref{eq:rotIn} is the moment of
+inertia about the pendulum axis and is written explicitly as
 \begin{equation}
-	J_{i}=c^{2}_{\beta i}I_{xx}+2s_{\beta i}c_{\beta i}I_{xz}+s^{2}_{\beta i}I_{zz}\textrm{.}
+	J_{i}=c^{2}_{\beta i}I_{xx}-2s_{\beta i}c_{\beta i}I_{xz}+s^{2}_{\beta i}I_{zz}\textrm{.}
 \label{eq:inRelComp}
@@
 and the unknown global inertia tensor can be solved for.
+The inertia of the frame about an axis normal to the plane of symmetry was
+estimated by hanging the frame as a compound pendulum at the wheel axis,
+Fig.~\ref{fig:frameCompound}. Equation~\ref{eq:CompoundInertia} is used but
+with the mass of the frame and the frame pendulum length.
+\begin{equation}
+    l_B=\sqrt{x_B^2+(z_B+r_R)^2}
+    \label{eq:FramePendLength}
+\end{equation}
 \begin{figure}[tbp]
@@
 \end{figure}
-\subsection{Fork}
+\subsection{FORK AND HANDLEBAR}
+The inertia of the fork and handlebar is calculated in the same way as the frame. The fork is
+hung as both a torsional pendulum, Fig.~\ref{fig:StratosFork}, and as a compound pendulum,
+Fig.~\ref{fig:forkCompound}. The fork provides fewer mounting options to
+obtain at least three equally spaced orientation angles, especially if there is
+no fender. The torsional calculations follow equations~\ref{eq:torPend} through
+\ref{eq:inRel} and the compound is calculated with
+equation~\ref{eq:CompoundInertia}. The fork pendulum length is calculated using
+\begin{equation}
+    l_H=\sqrt{(x_H-w)^2+(z_B+rF)^2}
+\end{equation}
 \begin{figure}[tbp]
@@
     \centering
+    \caption{Mass, center of mass and moment of inertia for the Person relative
+    to the benchmark coordinate system from~\cite{Moore2009a}}
     \begin{tabular}{ll}
@@
     \end{tabular}
-    \caption{Mass, center of mass and moment of inertia for the Person relative
-    to the benchmark coordinate system from~\cite{Moore2009a}}
     \label{tab:riderParam}
@@
     \centering
-    \includegraphics[width=4in]{../../../plots/Bike/eig_plot.pdf}
+    \includegraphics[width=6in]{../../../plots/Bike/eig_plot.pdf}
     \caption{The eigenvalues verus speed for all eight bicycles.}
@@
     \begin{center}
-        \includegraphics[width=4in]{../../../plots/Bike/CrescendoRootLoci.pdf}
+        \includegraphics[width=6in]{../../../plots/Bike/CrescendoRootLoci.pdf}
     \end{center}
@@
     \end{center}
-    \caption{The root loci with speed as the parameter for the Crescendo.}
+    \caption{The eigenvector components for the mode with the second complex
+    pair at low speeds.}
     \label{fig:evec}
@@
     \centering
-    \includegraphics[width=4in]{../../../plots/BikeRider/eig_plot.pdf}
+    \includegraphics[width=6in]{../../../plots/BikeRider/eig_plot.pdf}
     \caption{The eigenvalues verus speed for all eight bicycles with the same
@@
 diffence among bicycles can vary up to 10 dB (0.148 radians per newton meter of
-torque).
+torque). The difference in the frequency response for the bicycle with the
+rigid rider shows less variation among the bicycles,
+Fig.~\ref{fig:bikeRiderBode}, as the rider's mass and
+inertia play a bigger roll.
 \begin{figure}[tbp]
     \centering
-    \includegraphics[width=4in]{../../../plots/Bike/Bode/Tdel2phi.pdf}
+    \includegraphics[width=6in]{../../../plots/Bike/Bode/Tdel2phi.pdf}
     \caption{The eigenvalues verus speed for all eight bicycles with the same
@@
     \centering
-    \includegraphics[width=4in]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
+    \includegraphics[width=6in]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
     \caption{The eigenvalues verus speed for all eight bicycles with the same
@@
 \end{figure}
+\section{CONCLUSION}
+We present a detailed method of accurately estimating the physical parameters
+of a bicycle need for the benchmarked Whipple bicycle
+model~\cite{Meijaard2007}. Then we measure eight different bicycles providing
+both the raw data, parameter sets and linear model coefficient matrices with
+the uncertainties for each value. Then we present a brief comparison of the
+eight bicycles using eigenanalysis and Bode analysis.
 
@@
 \appendix
-\section{RAW DATA}
+%\section{RAW DATA}
 %\begin{landscape}
@@
 %\end{landscape}
-\section{PARAMETERS}
-\subsection{BICYCLE}
+%\section{PARAMETERS}
+%\subsection{BICYCLE}
 \begin{landscape}
 \subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
+%\end{landscape}
+%\subsection{BICYCLE WITH RIDER}
+%\begin{landscape}
+\subfile{../../../tables/BikeRider/Parameters/MasterParTable.tex}
+%\end{landscape}
+%\section{CANONICAL MATRICES}
+%\subsection{BICYCLE}
+%\begin{landscape}
+\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
+%\end{landscape}
+%\subsection{BICYCLE WITH RIDER}
+%\begin{landscape}
+\subfile{../../../tables/BikeRider/Canonical/MasterCanTable.tex}
 \end{landscape}
-\subsection{BICYCLE WITH RIDER}
-%\begin{landscape}
-%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
-%\end{landscape}
-\section{CANONICAL MATRICES}
-\subsection{BICYCLE}
-\begin{landscape}
-\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
-\end{landscape}
-\subsection{BICYCLE WITH RIDER}
-%\begin{landscape}
-%\subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
-%\end{landscape}
 \end{document}

### 2010-10-16 PhysicalParameters.py

@@ -264,7 +264,7 @@ def make_tables(typ='Bike'):
                         val, sig = uround(par[varname][i]).split('+/-')
                     except ValueError:
                         val = str(round(par[varname][i], 3))
-                        sig = '0.0'
+                        sig = 'NA'
                     fline = fline + ' & ' + val + ' & ' + sig
             fline = fline + r'\\' + '\n'
             final.write(fline)
@@ -301,8 +301,8 @@ def make_tables(typ='Bike'):
                     try:
                         val, sig = uround(can[varname][indice[0], indice[1]]).split('+/-')
                     except:
-                        val = str(can[varname][indice[0], indice[1]])
-                        sig = '0.0'
+                        val = str(round(can[varname][indice[0], indice[1]], 3))
+                        sig = 'NA'
                     fline = fline + ' & ' + val + ' & ' + sig
             fline = fline + r'\\' + '\n'
             final.write(fline)
@@ -338,7 +338,7 @@ def bike_bode_plots(typ='Bike', speeds=None):
     f.close()
 
     # figure properties
-    figwidth = 5. # in inches
+    figwidth = 6. # in inches
     goldenMean = (np.sqrt(5)-1.0)/2.0
     figsize = [figwidth, figwidth*goldenMean]
     params = {#'backend': 'ps',
@@ -1858,6 +1858,7 @@ def tor_stiffness(I, T):
 
     '''
     k = 4.*I*pi**2/T**2
+    print k
     return k
 
 def inertia_components(I, alpha):

### 2010-10-19 papers/BMD2010/PhysicalParametersPoster/poster.tex
poster draft added

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% LaTeX poster template
% Created by Nathaniel Johnston
% August 2009
% http://www.nathanieljohnston.com/index.php/2009/08/latex-poster-template/
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[final]{beamer}
\usepackage[scale=1.24]{beamerposter}
\usepackage{graphicx}			% allows us to import images

%-----------------------------------------------------------
% Define the column width and poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% The separation I chose is 0.024 and I want 4 columns
% Then set onecolwid to be (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be 2*onecolwid + sepwid = 0.464
%-----------------------------------------------------------

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{1189mm}
\setlength{\paperheight}{841mm}
\setlength{\sepwid}{0.024\paperwidth}
\setlength{\onecolwid}{0.1712\paperwidth}
\setlength{\twocolwid}{0.3664\paperwidth}
\setlength{\threecolwid}{0.5616\paperwidth}
\setlength{\topmargin}{-0.5in}
\usetheme{confposter}

%-----------------------------------------------------------
% Define colours (see beamerthemeconfposter.sty to change these colour definitions)
%-----------------------------------------------------------

\setbeamercolor{block title}{fg=ngreen,bg=white}
\setbeamercolor{block body}{fg=black,bg=white}
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70}
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10}

%-----------------------------------------------------------
% Name and authors of poster/paper/research
%-----------------------------------------------------------

\title{Accurate Measurement of Bicycle Parameters}
\author{Jason K. Moore$^{*}$, Mont Hubbard$^{*}$, A. L. Schwab$^\dag$,
            J. D. G. Kooijman$^\dag$}
\institute
{
\centering
\begin{tabular}{cc}
$^*$ Mechanical and Aerospace Engineering, University of California, Davis
\quad
& $^\dag$ Laboratory for Engineering Mechanics, Delft University of Technology\\
e-mail: jkmoor@ucdavis.edu, mhubbard@ucdavis.edu
& e-mail: a.l.schwab@tudelft.nl, jodikooijman@gmail.com
\end{tabular}
}
%-----------------------------------------------------------
% Start the poster itself
%-----------------------------------------------------------
% The \rmfamily command is used frequently throughout the poster to force a serif font to be used for the body text
% Serif font is better for small text, sans-serif font is better for headers (for readability reasons)
%-----------------------------------------------------------

\begin{document}
\frame{
%\begin{frame}[t]

% Divide the frame into columns
\begin{columns}[t] % the [t] option aligns the column's content at the top

% Empty side column for spacing
\begin{column}{\sepwid}\end{column}	% empty spacer column

% First column with content
\begin{column}{\onecolwid}
  \begin{block}{Introduction}
    \rmfamily{
    Accurate measurements of a bicycle's physical parameters are required for
    realistic dynamic simulations and analysis. For the most basic models the
    geometry, mass, mass location and mass distributions of each rigid body
    must be measured. In particular, we are concerned with the measurement of
    the minimal bicycle
    parameters required for the benchmark bicycle presented
    in~\cite{Meijaard2007}. A set of 25 parameters is used to populate the
    model.
    The experimental methods described herein are based primarily on
    the work done in \cite{Kooijman2006} and \cite{Roland1971} but have been
    refined for improved accuracy and methodology. We measured the
    characteristics of six different bicycles, two of which were set up in two
    different configurations. We provide a total of eight different parameter sets
    that can be used with, but not limited to, the benchmark bicycle model. The
    accuracy of all the measurements are presented along with a comparison of
    the linear characteristics of the eight bicycles.
    }
  \end{block}
  \vskip2ex

  % Info about the model
  \begin{alertblock}{Whipple Bicycle Model}
    \rmfamily{
    The linear unforced two degree-of-freedom, $\mathbf{q}$
    = [steer and roll], linear Whipple model takes the form:
    \small{
    \begin{equation}
        \mathbf{M\ddot{q}}
        +v\mathbf{C}_1\mathbf{\dot{q}}
        +\left[g\mathbf{K}_0
        +v^2\mathbf{K}_2\right]\mathbf{q}
        =0
        \label{eq:canonical}
    \end{equation}
    }
    Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
    matrices are combinations of 25 of the bicycle's physical parameters that
    include the geometry, mass, mass location and mass distribution of the four
    rigid bodies.
    }
  \end{alertblock}

\end{column}

% Spacer column
\begin{column}{\sepwid}\end{column} % empty spacer column

% Second Column
\begin{column}{\twocolwid}
  %\setbeamercolor{block title}{fg=red,bg=white}%frame color
  %\setbeamercolor{block body}{fg=black,bg=white}%body color

  \begin{block}{Accuracy}
    \rmfamily{
    Following the vain of \cite{Roland1971} we used error propagation theory to
    calculate accuracy of the 25 benchmark bicycle parameters. We estimate
    the standard deviation of the raw measurements. If $x$ is a parameter and is a function of
    the raw measurements, $u,v,\ldots$, then $x$ is a random variable defined as
    $x=f(u,v,\ldots)$. The sample variance of $x$ is defined as
    \begin{equation}
        s_x^2 = s_u^2\left(\frac{\partial x}{\partial u}\right)^2 +
                s_v^2\left(\frac{\partial x}{\partial v}\right)^2 +
                2s_{uv}\left(\frac{\partial x}{\partial u}\right)\left(\frac{\partial x}{\partial v}\right)
                + \ldots
        \label{eqn:variance}
    \end{equation}
    where $s_*$ is the variance and $s_{**}$ is the covariance. If $u$ and $v$ are uncorrelated then $s_{uv}=0$ but the cross correlations
    must be taken into account otherwise.
    }
  \end{block}
  \begin{columns}[t, totalwidth=\twocolwid]
      \begin{column}{\onecolwid}
          \begin{block}{Bicycles}
            \rmfamily{
            The six bicycles, chosen for both variety and convenience, are as follows:
            \emph{Batavus Browser}, a Dutch style city bicycle measured with and without
            instrumentation; \emph{Batavus Stratos
            Deluxe}, a Dutch style sporty city bicycle; \emph{Batavus Crescendo Deluxe} a
            Dutch style city bicycle with a suspended fork; \emph{Gary Fisher Mountain
            Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
            track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
            road bicycle measured in two configurations, the second with the fork rotated
            in the headtube 180 degrees for larger trail.
            \\
            \begin{center}
                \begin{tabular}{ccc}
                    \includegraphics[width=0.333\onecolwid]{../../../images/browserIns_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/crescendo_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/fisher_sub.jpg}
                    \\
                    \includegraphics[width=0.333\onecolwid]{../../../images/pista_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/stratos_sub.jpg} &
                    \includegraphics[width=0.333\onecolwid]{../../../images/yellow_sub.jpg}
                \end{tabular}
            \end{center}
            }
          \end{block}
    \end{column}

    \begin{column}{\onecolwid}

        \begin{block}{Measurements}
            \rmfamily{
            We attempted to directly measure the trail, wheelbase and wheel
            diameters with various techniques. We measured the mass centers of
            the four bodies (fork, frame, and wheels) by hanging the bodies by
            their mass centers. The moments of inertia were estimated by
            hanging the bodies as both torsional and compound pendulums and
            measuring the period of oscillation which is correlated to the
            inertia.
            }
            \begin{center}
              \includegraphics[width=0.90\onecolwid]{../../../figures/angles.pdf}
            \end{center}
        \end{block}

    \end{column}
\end{columns}

\end{column}

% Spacer column
\begin{column}{\sepwid}\end{column} % empty spacer column

% The fourth column
\begin{column}{\onecolwid}
  \begin{alertblock}{Bike Eigenanalysis}
    \rmfamily{
    Distinct variation is apparent among the bicycles' eigenvalues. Frequency,
    damping, and the stable speed range can be compared to the physical
    differences of the bicycles.
    }
    \\
    \begin{center}
      \includegraphics[width=.9\onecolwid]{../../../plots/Bike/eig_plot.pdf}
    \end{center}
    \rmfamily{
    Less variation is seen among the bicycles when the mass and inertia of a
    rigid rider is added.
    }
    \\
    \begin{center}
      \includegraphics[width=.9\onecolwid]{../../../plots/BikeRider/eig_plot.pdf}
    \end{center}
  \end{alertblock}

\end{column}

% Spacer column
\begin{column}{\sepwid}\end{column} % empty spacer column

\begin{column}{\onecolwid}
  % Frequency Response
  \begin{block}{Bike Frequency Response}
    \rmfamily{
    The steer torque to roll angle Bode plot reveals up to 10 dB variation
    among the bikes along the frequencies shown.
    \\
    \begin{center}
      \includegraphics[width=.9\onecolwid]{../../../plots/Bike/Bode/Tdel2phi.pdf}
    \end{center}
    }
  \end{block}

  % Frequency Response
  \begin{block}{Bike/Rider Frequency Response}
    \centering
    \includegraphics[width=.9\onecolwid]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
  \end{block}

  % References
  \begin{block}{References}
    \small{\rmfamily{
    \bibliographystyle{plain}
    \bibliography{bicycle}
    }}
  \end{block}
\end{column}
%\vskip2.5ex

% Last spacer column
\begin{column}{\sepwid}\end{column}			% empty spacer column

\end{columns}
}
%\end{frame}
\end{document}

### 2010-10-19 papers/BMD2010/PhysicalParametersPoster/poster.tex
bode plot titles, spaces in a m file, poster final draft

@@
   \begin{alertblock}{Whipple Bicycle Model}
-    \rmfamily{
+    \small{\rmfamily{
     The linear unforced two degree-of-freedom, $\mathbf{q}$
     = [steer and roll], linear Whipple model takes the form:
-    \small{
     \begin{equation}
@@
     \end{equation}
-    }
     Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
@@
     rigid bodies.
-    }
+    }}
   \end{alertblock}
@@
     \end{equation}
-    where $s_*$ is the variance and $s_{**}$ is the covariance. If $u$ and $v$ are uncorrelated then $s_{uv}=0$ but the cross correlations
+    where $s_u$ is the variance and $s_{uv}$ is the covariance. If $u$ and $v$ are uncorrelated then $s_{uv}=0$ but the cross correlations
     must be taken into account otherwise.
@@
 
-        \begin{block}{Measurements}
-            \rmfamily{
-            We attempted to directly measure the trail, wheelbase and wheel
-            diameters with various techniques. We measured the mass centers of
-            the four bodies (fork, frame, and wheels) by hanging the bodies by
-            their mass centers. The moments of inertia were estimated by
-            hanging the bodies as both torsional and compound pendulums and
-            measuring the period of oscillation which is correlated to the
-            inertia.
-            }
-            \begin{center}
-              \includegraphics[width=0.90\onecolwid]{../../../figures/angles.pdf}
-            \end{center}
-        \end{block}
-
-    \end{column}
-\end{columns}
-
-\end{column}
-
-% Spacer column
-\begin{column}{\sepwid}\end{column} % empty spacer column
-
-% The fourth column
-\begin{column}{\onecolwid}
-  \begin{alertblock}{Bike Eigenanalysis}
+  \begin{alertblock}{Bicycle Eigenanalysis}
     \rmfamily{
@@
     damping, and the stable speed range can be compared to the physical
-    differences of the bicycles.
+    differences of the bicycles. The bicycles also exhibit a two imaginary root
+    pairs at low speeds.
     }
@@
     \rmfamily{
-    Less variation is seen among the bicycles when the mass and inertia of a
-    rigid rider is added.
+    Less variation is seen among the bicycles when a
+    rigid rider is added, except in the stable sppeed ranges.
     }
@@
     \end{column}
+\end{columns}
+\end{column}
+
+% Spacer column
+\begin{column}{\sepwid}\end{column} % empty spacer column
+
+% The fourth column
+\begin{column}{\onecolwid}
+        \begin{block}{Measurements}
+            \rmfamily{
+            We estimated the wheel radii by measuring the distance traveled
+            by the loaded wheels, the trail by directly measuring the fork
+            offset and the wheelbase and headtube angle by direct measurement.
+            We measured the mass of the four bodies (fork, frame, and
+            wheels) directly using an precision scale. We found the location of
+            the mass center by hanging the bodies in multiple orientations
+            through their mass centers.
+            }
+            \\
+            \begin{center}
+              \includegraphics[width=0.8\onecolwid]{../../../figures/angles.pdf}
+            \end{center}
+            \rmfamily{
+            The in- and out-of-symmetric plane moments of inertia were estimated by
+            hanging the bodies as torsional and compound pendulums,
+            respectively. We then estimated the period of oscillation by
+            fitting a decaying osciallation function to a voltage signal from a
+            rate gyro after the pendulums were perturbed. The period can then
+            be correlated to the inertia.
+            }
+            \\
+            \begin{center}
+              \includegraphics[width=0.90\onecolwid]{../../../plots/PendFit/BrowserFrameCompoundFirst1.pdf}
+            \end{center}
+        \end{block}
+
+\end{column}
 
@@
   % Frequency Response
-  \begin{block}{Bike Frequency Response}
-    \rmfamily{
+  \begin{alertblock}{Bicycle Frequency Response}
+    \small{\rmfamily{
     The steer torque to roll angle Bode plot reveals up to 10 dB variation
-    among the bikes along the frequencies shown.
+    among the bikes and variation in phase.
+    }}
     \\
     \begin{center}
-      \includegraphics[width=.9\onecolwid]{../../../plots/Bike/Bode/Tdel2phi.pdf}
+      \includegraphics[width=.85\onecolwid]{../../../plots/Bike/Bode/Tdel2phi.pdf}
+      \\
+      \includegraphics[width=.85\onecolwid]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
     \end{center}
-    }
-  \end{block}
-
-  % Frequency Response
-  \begin{block}{Bike/Rider Frequency Response}
-    \centering
-    \includegraphics[width=.9\onecolwid]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
-  \end{block}
+  \end{alertblock}
 

### 2010-10-19 PhysicalParameters.py

@@ -365,13 +365,19 @@ def bike_bode_plots(typ='Bike', speeds=None):
               #'#C0C0C0']
 
     plotlist = ['Tdel2phi', 'Tdel2del', 'Tphi2phi', 'Tphi2del']
+    latexlist = ['$T_\delta/\phi$',
+                 '$T_\delta/\delta$',
+                 '$T_\phi/\phi$',
+                 '$T_\phi/\delta$']
 
     plots = {}
 
+    # make figures for each Bode plot
     for i, plot in enumerate(plotlist):
         plots[plot] = plt.figure(num=i, figsize=figsize)
 
     direct = 'data/' + typ + '/Canonical/'
+
     # for each plot
     for j, plot in enumerate(plotlist):
         for i, name in enumerate(data['shortnames']):
@@ -392,12 +398,11 @@ def bike_bode_plots(typ='Bike', speeds=None):
             if plot.split('2')[1] == 'del': CEE = C[3]
             elif plot.split('2')[1] == 'phi': CEE = C[2]
             mag, phase, plots[plot] = bode(ABCD=(A, BEE, CEE, 0.), w=freq,
-                    fig=plots[plot])
+                    fig=plots[plot], title=space_out_camel_case(typ) + ' ' + latexlist[j])
 
     direct = 'plots/' + typ + '/Bode'
     if not os.path.isdir(direct):
         os.system('mkdir ' + direct)
-
 
     for k, v in plots.items():
         print len(v.ax1.lines)
@@ -574,7 +579,7 @@ direct = 'plots/PendFit/'
 if not os.path.isdir(direct):
     os.system('mkdir ' + direct)
 
-def fit_data():
+def fit_data(filetype='.pdf'):
 
     dirs, subdirs, filenames = list(os.walk('data/pendDat/p'))[0]
     periodfile = open('data/period.txt', 'w')
@@ -627,7 +632,7 @@ def fit_data():
             rsqstr = str(rsq)
             m = 5.
         plot_osfit(x, y, lscurve, p1, rsq, T, m=m, fig=fig)
-        plt.savefig(direct + name[:-2] + '.png')
+        plt.savefig(direct + name[:-2] + filetype)
         plt.close()
         # include the notes for the experiment
         try:

### 2010-10-19 matlab/bike_inertia_jason_browser_mod.m

@@ -132,14 +132,14 @@ ele(26,1:7) = { 20 24 'right lower arm'  'cylinder' 'framex'  cla/2/pi 0      };
 figure(1)
 axis([-rr-0.1 wb+rf+0.1 0 grid(ele{22,1},3)+rh+0.1])
 hold on
-for i = 1:size(ele,1)  
+for i = 1:size(ele,1)
     switch ele{i,4}
         case {'ring','sphere'} % plot circles
             center = [grid(ele{i,1},1);grid(ele{i,1},3)];
             radius = ele{i,6};
             thetaplot = 0:0.001:2*pi;
             xplot = radius*cos(thetaplot)+center(1);
-            yplot = radius*sin(thetaplot)+center(2);            
+            yplot = radius*sin(thetaplot)+center(2);
             plot(xplot,yplot)
         case {'tube','rprism','cylinder'} % plot lines
             if i == 14 % don't plot handle bar
@@ -158,8 +158,8 @@ for i = 1:size(ele,1)
                 z1 = grid(ele{i,2},3);
                 xplot=x1:0.001:x2;
                 yplot=(z2-z1)/(x2-x1).*(xplot-x1)+z1;
-                plot(xplot,yplot)                
-            end                
+                plot(xplot,yplot)
+            end
     end
 end
 hold off

### 2010-10-19 PhysicalParameters.py
poster final draft

@@ -526,7 +526,7 @@ def bike_eig_plots(typ='Bike', filetype='pdf'):
     plt.legend(loc='lower right')
     plt.ylim((-10, 10))
     plt.xlim((0, 10))
-    plt.title('Eigenvalues vs Speed')
+    plt.title(space_out_camel_case(typ) + ' Eigenvalues vs Speed')
     plt.xlabel('Speed [m/s]')
     plt.ylabel('Real and Imaginary Parts of the Eigenvalue [1/s]')
     try:

### 2010-10-19 papers/BMD2010/PhysicalParametersPoster/poster.tex
fixed the legend colors on the bode plot

@@
 
-\setbeamercolor{block title}{fg=ngreen,bg=white}
+\setbeamercolor{block title}{fg=ucdgold,bg=white}
 \setbeamercolor{block body}{fg=black,bg=white}
-\setbeamercolor{block alerted title}{fg=white,bg=dblue!70}
-\setbeamercolor{block alerted body}{fg=black,bg=dblue!10}
+\setbeamercolor{block alerted title}{fg=white,bg=ucdblue}
+\setbeamercolor{block alerted body}{fg=black,bg=ucdblue!10}
 

### 2010-10-19 PhysicalParameters.py

@@ -404,13 +404,15 @@ def bike_bode_plots(typ='Bike', speeds=None):
     if not os.path.isdir(direct):
         os.system('mkdir ' + direct)
 
+    # set the colors for the lines to match the bike
     for k, v in plots.items():
         print len(v.ax1.lines)
-        v.ax2.legend(data['bikes'], 'lower right')
         for i, line in enumerate(v.ax1.lines):
             print i, colors[i]
             plt.setp(line, color=colors[i])
             plt.setp(v.ax2.lines[i], color=colors[i])
+        # set the legend
+        v.ax2.legend(data['bikes'], 'lower right')
         v.savefig(direct + '/' + k + '.pdf')
 
     return plots

### 2010-10-19 papers/BMD2010/PhysicalParametersPoster/poster.tex
poster final draft again and I'm sitting in 't klooster in delft drinking fine beer and eating fine cheese

@@
 
-  \begin{alertblock}{Bicycle Eigenanalysis}
+  \begin{alertblock}{Eigenanalysis}
     \rmfamily{
@@
     Less variation is seen among the bicycles when a
-    rigid rider is added, except in the stable sppeed ranges.
+    rigid rider is added, except in the stable speed ranges.
     }
@@
             respectively. We then estimated the period of oscillation by
-            fitting a decaying osciallation function to a voltage signal from a
+            fitting a decaying oscillation function to a voltage signal from a
             rate gyro after the pendulums were perturbed. The period can then
@@
   % Frequency Response
-  \begin{alertblock}{Bicycle Frequency Response}
+  \begin{alertblock}{Frequency Response}
     \small{\rmfamily{

### 2010-10-19 papers/BMD2010/PhysicalParametersPaper/paper.tex
went through the abstract

@@
 \begin{tabular}{c}
-$^*$ Mechanical and Aerospace Engineering\\
+$^*$ Department of Mechanical and Aerospace Engineering\\
 University of California, Davis\\
@@
 Accurate measurements of a bicycle's physical parameters are required for
-realistic dynamic simulations and analysis. For the most basic models the
-geometry, mass, mass location and mass distributions must be measured. More complex models
-require measurements of tire characteristics, human characteristics, friction, stiffness, damping, etc. This
-paper concerns the measurement of the minimal bicycle parameters required for
-the benchmark bicycle presented in~\cite{Meijaard2007}. This
+realistic dynamic simulations and analysis. The most basic models require the
+geometry, mass, mass location and mass distributions for the rigid bodies. More
+complex models require estimates of tire characteristics, human
+characteristics, friction, stiffness, damping, etc. In this
+paper we present the measurement of the minimal bicycle parameters required for
+the benchmark Whipple bicycle model presented in~\cite{Meijaard2007}. This
 model is composed of four rigid bodies, has ideal rolling and frictionless joints,
-and is assumed to be laterally symmetric. A set of 25
-parameters is used to describe the geometry, mass, mass location and
-mass distribution of each of the rigid bodies. The experimental methods
-described herein are based primarily on the work
-done in~\cite{Kooijman2006} but have been refined for improved accuracy and
-methodology. Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
+and is laterally symmetric. A set of 25 parameters is used to describe the
+geometry, mass, mass location and mass distribution of each of the rigid
+bodies. The experimental methods used to estimate the parameters described
+herein are based primarily on the work done in~\cite{Kooijman2006} but have
+been refined for improved accuracy and methodology.
+
+Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
 similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
-similar techniques with scooters. We measured the characteristics of six
+similar techniques with scooters. We measured the physical characteristics of six
 different bicycles, two of which were set up in two different configurations.
-This is a total of eight different parameter sets that can be used with, but not
-limited to, the benchmark bicycle model. The accuracy of all the measurements
-are presented up through the eigenvalue prediction of the linear model and are
-based on error propagation theory with correlations taken into account.
 
@@
 in the headtube 180 degrees for larger trail.
+
+These eight different parameter sets can be used with, but not
+limited to, the benchmark bicycle model. The accuracy of all the measurements
+are presented up through the eigenvalue prediction of the linear model. The
+accuracies are based on error propagation theory with correlations taken into account.
 \begin{keywords}
@@
 eight bicycles using eigenanalysis and Bode analysis.
+\section{ACKNOWLEDGEMENTS}
+This work was partially funded by the National Science Foundation.
 

### 2010-10-19 papers/BMD2010/PhysicalParametersPaper/paper.tex
editing the document

@@
 herein are based primarily on the work done in~\cite{Kooijman2006} but have
-been refined for improved accuracy and methodology.
+been refined for improved accuracy and methodology. Koojiman's work was
+preceded by \cite{Roland1971} who measured a bicycle in a
+similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
+similar techniques with scooters.
 
-Koojiman's work was preceded by \cite{Roland1971} who measured a bicycle in a
-similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
-similar techniques with scooters. We measured the physical characteristics of six
-different bicycles, two of which were set up in two different configurations.
-
-The six bicycles, chosen for both variety and convenience, are as follows:
+We measured the physical characteristics of six different bicycles, two of
+which were set up in two different configurations. The six bicycles, chosen for
+both variety and convenience, are as follows:
 \emph{Batavus Browser}, a Dutch style city bicycle measured with and without
@@
 are presented up through the eigenvalue prediction of the linear model. The
-accuracies are based on error propagation theory with correlations taken into account.
+accuracies are based on error propagation theory with correlations taken into
+account.
+
 \begin{keywords}
@@
 \section{INTRODUCTION}
-
 This work is intended to document the indirect measurement of eight real
 bicycles' physical parameters. The physical parameters measured are those that
-are needed for the benchmark bicycle model~\cite{Meijaard2007}. The work is
-based off of the techiques used to measure the instrumented bicycle in
-\cite{Kooijman2006} and \cite{Kooijman2008}. I used the same techniques as
-Kooijman to measure a second instrumented bicycle~\cite{Moore2009a}. This
-method had low accuracy and the $I_{yy}$ moments of inertia were
-not measured for the fork and frame. Furthermore, very little data exists on
-the physical parameters of different types of bicycles and this work aims to
-provide a small sample of bicycles.
+are needed for the benchmark Whipple bicycle model presented in
+\cite{Meijaard2007}. The work is based on the techiques used to measure the
+instrumented bicycles in \cite{Kooijman2006}, \cite{Kooijman2008} and
+\cite{Moore2009a}. We improve upon these methods by both increasing and reporting
+the accuracies of the measurements and by measuring the complete moments of
+inertia of the frame and fork needed for anaylsis of the nonlinear model.
+Furthermore, very little data exists on the physical parameters of different
+types of bicycles and this work aims to provide a small sample of bicycles.
 

### 2010-10-20 papers/BMD2010/PhysicalParametersPoster/poster.tex
poster final draft

@@
     \rmfamily{
-    Accurate measurements of a bicycle's physical parameters are required for
+    Accurate estimates of a bicycle's physical parameters are required for
     realistic dynamic simulations and analysis. For the most basic models the
-    geometry, mass, mass location and mass distributions of each rigid body
+    geometry, mass, mass location and mass distribution of each rigid body
     must be measured. In particular, we are concerned with the measurement of
-    the minimal bicycle
-    parameters required for the benchmark bicycle presented
-    in~\cite{Meijaard2007}. A set of 25 parameters is used to populate the
-    model.
+    the non-minimal set of 25 parameters required for the benchmark bicycle
+    presented in~\cite{Meijaard2007}.
     The experimental methods described herein are based primarily on
@@
 \begin{column}{\twocolwid}
-  %\setbeamercolor{block title}{fg=red,bg=white}%frame color
-  %\setbeamercolor{block body}{fg=black,bg=white}%body color
-
   \begin{block}{Accuracy}
     \rmfamily{
-    Following the vain of \cite{Roland1971} we used error propagation theory to
-    calculate accuracy of the 25 benchmark bicycle parameters. We estimate
+    Following in the vain of \cite{Roland1971} we used error propagation theory to
+    calculate the accuracy of the estimates of 25 benchmark bicycle parameters.
+    We first estimate
     the standard deviation of the raw measurements. If $x$ is a parameter and is a function of
@@
     The steer torque to roll angle Bode plot reveals up to 10 dB variation
-    among the bikes and variation in phase.
+    among the bikes and variation in phase below 10 rad/s.
     }}

### 2010-10-20 papers/BMD2010/PhysicalParametersPoster/poster.tex
final poster

@@
     different configurations. We provide a total of eight different parameter sets
-    that can be used with, but not limited to, the benchmark bicycle model. The
-    accuracy of all the measurements are presented along with a comparison of
+    that can be used with, but are not limited to, the benchmark bicycle model. The
+    accuracies of all the measurements are presented along with a comparison of
     the linear characteristics of the eight bicycles.
@@
     \end{equation}
-    Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
+    where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
     matrices are combinations of 25 of the bicycle's physical parameters that
@@
     \rmfamily{
-    Following in the vain of \cite{Roland1971} we used error propagation theory to
+    Following in the vein of \cite{Roland1971} we used error propagation theory to
     calculate the accuracy of the estimates of 25 benchmark bicycle parameters.
@@
             Bike}, a hardtail mountain bicycle; \emph{Bianchi Pista}, a modern steel frame
-            track racing bicycle; and \emph{Yellow Bicycle}, a stripped down aluminum frame
+            track racing bicycle; and \emph{Yellow Bicycle}, a stripped-down aluminum frame
             road bicycle measured in two configurations, the second with the fork rotated
@@
     damping, and the stable speed range can be compared to the physical
-    differences of the bicycles. The bicycles also exhibit a two imaginary root
+    differences of the bicycles. The bicycles also exhibit two complex root
     pairs at low speeds.
@@
             We measured the mass of the four bodies (fork, frame, and
-            wheels) directly using an precision scale. We found the location of
+            wheels) directly using a precision scale. We found the location of
             the mass center by hanging the bodies in multiple orientations

### 2010-10-20 PhysicalParameters.py

@@ -398,7 +398,7 @@ def bike_bode_plots(typ='Bike', speeds=None):
             if plot.split('2')[1] == 'del': CEE = C[3]
             elif plot.split('2')[1] == 'phi': CEE = C[2]
             mag, phase, plots[plot] = bode(ABCD=(A, BEE, CEE, 0.), w=freq,
-                    fig=plots[plot], title=space_out_camel_case(typ) + ' ' + latexlist[j])
+                    fig=plots[plot], title='+'.join(space_out_camel_case(typ).split()) + ' ' + latexlist[j])
 
     direct = 'plots/' + typ + '/Bode'
     if not os.path.isdir(direct):
@@ -528,7 +528,7 @@ def bike_eig_plots(typ='Bike', filetype='pdf'):
     plt.legend(loc='lower right')
     plt.ylim((-10, 10))
     plt.xlim((0, 10))
-    plt.title(space_out_camel_case(typ) + ' Eigenvalues vs Speed')
+    plt.title('+'.join(space_out_camel_case(typ).split()) + ' Eigenvalues vs Speed')
     plt.xlabel('Speed [m/s]')
     plt.ylabel('Real and Imaginary Parts of the Eigenvalue [1/s]')
     try:

### 2010-10-20 PhysicalParameters.py
eeeH'

@@ -387,7 +387,7 @@ def bike_bode_plots(typ='Bike', speeds=None):
             can = pickle.load(f)
             f.close()
             # calulate the A and B matrices
-            A, B = abMatrix(can['M'], can['C1'], can['K0'], can['K2'], 1., 9.81)
+            A, B = abMatrix(can['M'], can['C1'], can['K0'], can['K2'], 2., 9.81)
             A = unumpy.nominal_values(A)
             B = unumpy.nominal_values(B)
             # y is [phidot, deldot, phi, del]
@@ -398,7 +398,9 @@ def bike_bode_plots(typ='Bike', speeds=None):
             if plot.split('2')[1] == 'del': CEE = C[3]
             elif plot.split('2')[1] == 'phi': CEE = C[2]
             mag, phase, plots[plot] = bode(ABCD=(A, BEE, CEE, 0.), w=freq,
-                    fig=plots[plot], title='+'.join(space_out_camel_case(typ).split()) + ' ' + latexlist[j])
+                    fig=plots[plot],
+                    title='+'.join(space_out_camel_case(typ).split()) + ' ' +
+                    latexlist[j] + ' @ 2 m/s')
 
     direct = 'plots/' + typ + '/Bode'
     if not os.path.isdir(direct):

### 2010-10-21 papers/BMD2010/PhysicalParametersPaper/paper.tex
paper final draft for conference

@@
 model is composed of four rigid bodies, has ideal rolling and frictionless joints,
-and is laterally symmetric. A set of 25 parameters is used to describe the
+and is laterally symmetric. A set of 25 parameters describes the
 geometry, mass, mass location and mass distribution of each of the rigid
@@
 preceded by \cite{Roland1971} who measured a bicycle in a
-similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who have used
+similar fashion and both~\cite{Dohring1953} and~\cite{Singh1971} who used
 similar techniques with scooters.
@@
 
-These eight different parameter sets can be used with, but not
+These eight different parameter sets can be used with, but are not
 limited to, the benchmark bicycle model. The accuracy of all the measurements
@@
 This work is intended to document the indirect measurement of eight real
-bicycles' physical parameters. The physical parameters measured are those that
-are needed for the benchmark Whipple bicycle model presented in
-\cite{Meijaard2007}. The work is based on the techiques used to measure the
+bicycles' physical parameters. The physical parameters measured are those
+needed for the benchmark Whipple bicycle model presented in
+\cite{Meijaard2007}. The work is based on techniques used to measure the
 instrumented bicycles in \cite{Kooijman2006}, \cite{Kooijman2008} and
@@
 the accuracies of the measurements and by measuring the complete moments of
-inertia of the frame and fork needed for anaylsis of the nonlinear model.
+inertia of the frame and fork needed for analysis of the nonlinear model.
 Furthermore, very little data exists on the physical parameters of different
@@
 
-D\"{o}hring measured the physical parameters of a scooter~\cite{Dohring1953}.
-Singh and Goel measured the physical parameters of a scooter~\cite{Singh1971}.
-Roland and Massing measured the physical parameters of a bicycle in much the
-same was as presented, including calculations of uncertainty from the indirect
-measurement techniques~\cite{Roland1971}. Patterson used a swing to measure the
-inertia of a bicycle and rider~\cite{Patterson2004}. This work is based off of the work
-done by Kooijman~\cite{Kooijman2006}. I worked with Jodi Kooijman for a year and
-was able to use much of the same apparatus and refine the measurement
-technique.
+D\"{o}hring~\cite{Dohring1953} and Singh and Goel~\cite{Singh1971} measured the
+physical parameters of scooters.
+Roland and Massing~\cite{Roland1971} measured the physical parameters of a
+bicycle in much the
+same way as is presented, including calculations of uncertainty from the indirect
+measurement techniques. Patterson~\cite{Patterson2004} used a swing to measure the
+inertia of a bicycle and rider. The present work is based on the work
+done by Kooijman~\cite{Kooijman2006} using much of the same apparatus and
+refining the measurement technique.
 \section{BENCHMARK BICYCLE MODEL}
@@
 \end{equation}
-Where the entries to the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
-matrices are combinations of 25 of the bicycle's physical parameters that
+where the entries of the $\mathbf{M}$, $\mathbf{C}_1$, $\mathbf{K}_0$ and $\mathbf{K}_2$
+matrices are combinations of 25 bicycle physical parameters that
 include the geometry, mass, mass location and mass distribution of the four
 rigid bodies. The 25 parameters presented in \cite{Meijaard2007} are not
-necessarily a minimun set for the Whipple model, as shown in \cite{Sharp2008},
-but are useful as they represent more easily measurable quantities.
+necessarily a minimum set for the Whipple model, as shown in \cite{Sharp2008},
+but are useful as they represent more intuitively measurable quantities.
 Furthermore, many more parameters are not needed due to the assumptions of the
@@
 We choose to measure the physical parameters of six bicycles
-Figs.~\ref{fig:bicycles}. The three Batavus bicycles were donated by the
+Fig.~\ref{fig:bicycles}. The three Batavus bicycles were donated by the
 manufacturer. We asked for a bicycle that they considered stable and one that
@@
 Stratos was ``nervous''. The Fisher and the Pista were chosen to provide some
-variety, a mountatin and road bike. The yellow bike is used to demonstrate
+variety, a mountain and road bike. The yellow bike is used to demonstrate
 bicycle stability.
 \begin{description}
-    \item[Batavus Browser (B, B*)]{The Batavus Browser Fig.~\ref{fig:browser} is an averaged priced Dutch city
+    \item[Batavus Browser (B, B*)]{The Batavus Browser Fig.~\ref{fig:browser}
+        is an average priced Dutch city
         bike. It has a steel frame, a three speed internal rear hub, handle
-        bars for an upright posture and includes various accerories for utility
-        puposes. We measured the physical proprieties of the stock Browser model and
+        bars for an upright posture and includes various accessories for utility
+        purposes. We measured the physical proprieties of the stock Browser model and
         also equipped with the instrumentation used in our other
@@
         Fig.~\ref{fig:crescendo} is also a
-        Dutch city bike that is setup for touring. It has an aluminum frame, an
-        eight speed internal hub, upright handlebars, accesories for utility
+        Dutch city bike for touring. It has an aluminum frame, an
+        eight speed internal hub, upright handlebars, accessories for utility
         and a suspension fork and suspension seatpost.}
     \item[Gary Fisher Ziggurat (G)]{The Gary Fisher Ziggurat Fig.~\ref{fig:fisher} is a modern
-        lightweight front suspended mountain bike. It has a ?? frame, large low
-        pressure mountain bike tires, and is setup for racing with few extra accerories.}
+        lightweight front suspended mountain bike. It has a aluminum frame, large low
+        pressure mountain bike tires, and is for racing with few extra accessories.}
     \item[Bianchi Pista (P)]{The Bianchi Pista Fig.~\ref{fig:pista} is a modern lightweight steel track
@@
         Fig.~\ref{fig:stratos} is a sporty Dutch city bicycle. The frame is
-        made of aluminum and it has a seven speed internal hub. It has mountain
+        aluminum. It has a seven-speed internal hub and mountain
         style handle bars for a less upright seating posture, but also includes
-        accesories for utility such as a rear rack, fenders, light and
+        accessories for utility such as a rear rack, fenders, light and
         chainguard.}
-    \item[Yellow Bicycle (Y, Y*)]{The yellow bicycle Fig.~\ref{fig:yellow} used in the lab to
+    \item[Yellow Bicycle (Y, Y*)]{The yellow bicycle Fig.~\ref{fig:yellow} is used in the lab to
         demonstrate that a bicycle is stable at certain speeds. It is an
-        aluminum road frame of unknown make with the majority of components
+        aluminum road frame of unknown make with the most of components
         removed. The wheels, drop handlebar, seat, seat post and bottom bracket
-        are the only parts on the bike. The bicycle was measured with both the
-        fork in normal position and the fork reversed. The fork was reversed to
+        are the only parts on the bike. This bicycle was measured with both the
+        fork in normal position and reversed. The fork was reversed to
         ''decrease the minimal stable speed``~\cite{Kooijman2006} of the bicycle.}
 \end{description}
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \centering
@@
 \section{PARAMETERS}
-The 25 parameters can be estimated by many techniques. Where possible we
+The 25 parameters can be estimated using many techniques. Where possible we
 measured the benchmark parameter directly.
 \section{ACCURACY}
-We took great care too improve upon the accuracy and report the accuracy of the
-measurements of the bicycle parameters. Following the vain of
+We took great care to improve and report the accuracy of the
+measurements of the parameters. Following the thrust of
 ~\cite{Roland1971} we used error propagation theory to calculate accuracy of
@@
 the actual measurements taken. If $x$ is a parameter and is a function of
-the measurments, $u,v,\ldots$, then $x$ is a random variable defined as
+the measurements, $u,v,\ldots$, then $x$ is a random variable defined as
 $x=f(u,v,\ldots)$. The sample variance of $x$ is defined as
@@
 \end{equation}
-Using the definitions for variance and covariance
+Using the definitions for variance and covariance,
 Equation~\ref{eqn:sampleVariance} can be simplified to
@@
 If $u$ and $v$ are uncorrelated then $s_{uv}=0$. Most of the calculations
-hereafter have uncorrelate variables but a few do not and the covariance has to
+hereafter have uncorrelated variables but a few do not and the covariance has to
 be taken into account. Equation~\ref{eqn:variance} can be
@@
 \subsection{WHEEL RADII}
-The radius of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were
+The radii of the front $r_\mathrm{F}$ and rear $r_\mathrm{R}$ wheels were
 estimated by measuring the linear distance traversed along the ground through
@@
 the measurements were taken with a 72kg rider seated on the bicycle. A 30 meter
-tape measure (resolution: 2mm) was pulled taught and taped on a flat level smooth floor. The tire
+tape measure (resolution: 2mm) was pulled tight and taped on a flat level smooth floor. The tire
 was marked with chalk and aligned with the tape measure
@@
 The head tube angle was measured directly using an electronic level with a
-$\pm0.2^{\circ}$ accuracy. The bicycle frame was fixed perdenpicular to the
-ground, the steering angle was set to the nominal, the tire pressures were are
+$\pm0.2^{\circ}$ accuracy. The bicycle frame was fixed perpendicular to the
+ground, the steering angle was set to the nominal, tire pressures were at
 recommended levels and the bicycle was unloaded. The steer axis
@@
 the center of the head tube and the height of the center of the axle axis. The
-fork blades where aligned such that the axle axis was parallel to the table
+fork blades were aligned such that the axle axis was parallel to the table
 surface.
@@
 \subsection{WHEELBASE}
-We measured the wheelbase when the bicycle was in the same nominal
-configuration as described in Sec.~\ref{sec:headtube}. We used a tape measure
+We measured the wheelbase with the bicycle in nominal
+configuration described in Sec.~\ref{sec:headtube}. We used a tape measure
 to measure the distance from one wheel axle center to the other with a 0.002 m
@@
 \section{MASS}
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \begin{center}
@@
 global reference frame.
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \centering
@@
 \end{figure}
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \centering
@@
     \subfloat[]{\label{fig:PendDist}\includegraphics[width=2.75in]{../../../images/pendDist.jpg}}
-    \caption{\subref{fig:level} The digital level was mounted to a straight edge that was aligned
+    \caption{\subref{fig:level} The digital level was mounted to a straight edge aligned
     with the headtube of the bicycle frame. This was done without allowing the
@@
 z-intercept $b$ where the $i$ subscript corresponds the different frame
-orientations~\ref{fig:triangle}. The slope can be shown to be
+orientations Fig.~\ref{fig:triangle}. The slope can be shown to be
 \begin{equation}
@@
 \end{equation}
-Only two lines are required to calculate the center of mass of the laterally
+Theoretically, the center of mass lies on each line but due to experimental
+error, if there are more than two lines, the lines do not cross all at the same point. Only two lines
+are required to calculate the center of mass of the laterally
 symmetric frame, but more orientations increase the center of mass measurement
@@
 \end{equation}
-Alternatively the three lines can be treated as an over determined linear
+Alternatively, the three lines can be treated as an over determined linear
 system and the least squares method is used to find a unique solution. This
-solution is not at the centroid of the triangle made by the intersecting lines.
-I am not sure which solution is theoretically the correct one for the location
-of the center of mass.
+solution is not the same as the triangle centroid method.
 \begin{equation}
@@
 \end{equation}
+The solution with the higher accuracy is the preferred one.
 \subsection{Fork}
@@
 advantage of the assumed symmetry of the parts and by hanging the parts as both
-compound and torsional pendulums and measuring their period of oscillation when
-perturb at small angles. The rate of oscillation was measured
+compound and torsional pendulums and measuring their periods of oscillation when
+perturbed at small angles. The rate of oscillation was measured
 using a \href{http://www.siliconsensing.com/CRS03}{Silicon Sensing CRS03 100
@@
 Instruments USB-6008 12 bit data acquisition unit} and
-\textsc{Matlab}. The measurement duration were either 15 or 30 secs and each moment of
+\textsc{Matlab}. The measurement durations were either 15 or 30 secs and each moment of
 inertia measurement was performed three times. No extra care was taken to
@@
 drains slowly), or account for drift. The raw voltage signal was used to
-determine only the period of oscillation which is need for the moment of inertia
+determine only the period of oscillation which is needed for the moment of inertia
 calculations.
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \begin{center}
@@
 \end{figure}
-The following function Eqn~\ref{eqn:decayOs} was fit to the data using a nonlinear least squares fit
-routine for each experiment.
+The function Eqn~\ref{eqn:decayOs} was fit to the data using a nonlinear least squares fit
+routine for each experiment to determine the quantities $A$, $B$, $C$, $\zeta$,
+and $\omega$.
 \begin{equation}
@@
 \end{figure}
-The physical phenomena observed for corresponding to data sets such as these were that
+
+The physical phenomenon observed corresponding to data sets such as these was that
 the bicycle frame or fork was perturbed torsionally. After set into motion the
@@
 \end{equation}
-But the fit predicts that $\omega_1$ and $\omega_2$ have very similar
+But the fit predicts that $\omega_1$ and $\omega_2$ are very similar
 frequencies. There was no easy way to choose which of the two $\omega$'s was
@@
 the torsional pendulum as a laterally flexible beam to determine this, but we
-thought the improvement in the accuracy of the period calculation would not
+thought accuracy of the period calculation would not
 improve enough for the effort required. Future experiments should simply
@@
     \label{eqn:periodDamped}
-\end{equation}
-and for light damping
-\begin{equation}
-    T \approx \frac{2\pi}{\omega_n}
-    \label{eqn:period}
 \end{equation}
@@
 is important to note that the uncertainties in the period are very low
-($<10e-4$), even for the fits with low $r^2$ values.
+($<1e-4$), even for the fits with low $r^2$ values.
 \subsection{TORSIONAL PENDULUM}
-A torsional pendulum was used to measure all of moments of inertia about axes
-in the lateral symmetric plane of each of the wheels, fork and frame. The
+A torsional pendulum was used to measure all moments of inertia about axes
+in the laterally symmetric plane of each of the wheels, fork and frame. The
 pendulum is made up of a rigid mount, an upper clamp, a torsion rod, and
 various lower clamps.
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \begin{center}
@@
 were neglected. The wheel was hung freely such that the
-center of mass aligned with the torsional pendulum axis and then it was
-secured. The wheel was then perturbed and it oscillated about the pendulum
-axis. The rate gyro was mounted on the clamp orientated along the pendulum
+center of mass aligned with the torsional pendulum axis and then
+secured. The wheel was then perturbed and oscillated about the pendulum
+axis. The rate gyro was mounted on the clamp oriented along the pendulum
 axis.
@@
 The torsional pendulum was calibrated using a known moment of
-inertia~\ref{fig:rod}. A torsional pendulum almost identical to the one used in
-\cite{Kooijman2006} was used to measure the averaged period $\overline{T}_i$ of
+inertia Fig.~\ref{fig:rod}. A torsional pendulum almost identical to the one used in
+\cite{Kooijman2006} was used to measure the average period $\overline{T}_i$ of
 oscillation of the rear frame at three different
@@
 averaged.
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \begin{center}
@@
 Finding the full inertia tensors of the wheels is less complex because the wheels
-are assumed symmetric about three orthogonal planes so there are no products of
-inertia. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring
+are assumed symmetric about three orthogonal planes so products of
+inertia are zero. The $I_{xx}=I_{zz}$ moments of inertia were calculated by measuring
 the averaged period of oscillation about an axis in the $XZ$-plane using the
-torsional pendulum setup and Eq.~\ref{eq:torPend}.The wheels are assumed to be symmetric laterally and about any radial axis. Thus only two
-moment's of inertia are required for the benchmark parameters set. The moment
+torsional pendulum setup and Eq.~\ref{eq:torPend}.The wheels are assumed to be
+laterally symmetric and about any radial axis. Thus only two
+moments of inertia are required for the set of benchmark parameters. The moment
 of inertia about the axle was measured by hanging the wheel as a compound
@@
 horizontal rod and perturbed to oscillate about the axis of the rod. This rate
-gyro was attached to the spokes near the hub and orientated mostly along the
+gyro was attached to the spokes near the hub and oriented mostly along the
 axle axis. The wheels tended to precess at the contact point about the vertical
-axis whixh added a very low frequency component of rate along the vertical
-radial axis, but should not affect the period estimation about the compound
-pendulum axis. A fixture that prevented the precession would be preferable for
+axis which added a very low frequency component of rate along the vertical
+radial axis, but this should not affect the period estimation about the compound
+pendulum axis. A fixture that prevented precession would be preferable for
 future measurements. The pendulum arm length is the distance
@@
 diameter of the rim was measured and divided by two to get $l_\mathrm{F,R}$. The moment of
-inertia about the axle is calculated by:
+inertia about the axle is calculated from:
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-    need the sigma for IRyy
-\end{equation}
-\begin{figure}[tbp]
+%\begin{equation}
+    %need the sigma for IRyy
+%\end{equation}
+\begin{figure}[tb]
     \centering
@@
 pendulum, Fig.~\ref{fig:FwheelTor}. The wheel was hung freely such that the center of mass aligned with
-the torsional pendulum axis and then it was secured in position. The wheel was then
-perturbed and it oscillated about the pendulum axis. The radial moment of inertia can can calculated as such:
+the torsional pendulum axis and then secured. The wheel was then
+perturbed and oscillated about the vertical pendulum axis. The radial moment of inertia can can calculated as such:
 \begin{equation}
@@
 \end{equation}
-\begin{equation}
-    need the sigma for Ixx
-\end{equation}
+%\begin{equation}
+    %need the sigma for Ixx
+%\end{equation}
 \subsection{FRAME}
@@
 orientation angles were spread evenly at about $120^\circ$. Furthermore, taking
-data at more orientation angles could improve the accuracy further and
-is generally possible with standard diamond frame bicylces.
+data at more orientation angles could improve the accuracy and
+is generally possible with standard diamond frame bicycles.
 
@@
 \end{equation}
-\begin{equation}
-    need sigma
-\end{equation}
+%\begin{equation}
+    %need sigma
+%\end{equation}
 The moments and products of inertia of the rear frame and handlebar/fork
@@
 \end{equation}
-Calculating all three $J_{i}$ allows one to form
+Similarly, calculating all three $J_{i}$ allows one to form
 \begin{equation}
@@
 \end{equation}
-and the unknown global inertia tensor can be solved for.
+and the moments of inertia can be solved for.
 The inertia of the frame about an axis normal to the plane of symmetry was
@@
 \end{equation}
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \centering
@@
         \label{fig:frameCompound}
-        \includegraphics[width=2.75in]{../../../images/YellowFrameCompoundFirst.jpg}
+        \includegraphics[width=2in]{../../../images/YellowFrameCompoundFirst.jpg}
         }
@@
         \label{fig:forkCompound}
-        \includegraphics[width=2.75in]{../../../images/BrowserInsForkCompoundFirst.jpg}
+        \includegraphics[width=2in]{../../../images/BrowserInsForkCompoundFirst.jpg}
         }
-        \caption{\subref{fig:frameCompound} The rear frame hung as a compound
-        pendulum. \subref{fig:forkCompound} The Browser fork hung as a
+        \caption{\subref{fig:frameCompound} Rear frame hung as a compound
+        pendulum. \subref{fig:forkCompound} Browser fork hung as a
         compound pendulum.}
@@
 no fender. The torsional calculations follow equations~\ref{eq:torPend} through
-\ref{eq:inRel} and the compound is calculated with
+\ref{eq:inRel} and the compound pendulum calculations is calculated with
 equation~\ref{eq:CompoundInertia}. The fork pendulum length is calculated using
@@
 \end{equation}
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \centering
-    \subfloat[]{
+    \includegraphics[width=2.75in]{../../../images/StratosForkTorsionalThird.jpg}
+    \caption{The Stratos fork and handlebar assembly hung as a torsional
+        pendulum.}
     \label{fig:StratosFork}
-        \includegraphics[width=2.75in]{../../../images/StratosForkTorsionalThird.jpg}
-        }
-    \caption{\subref{fig:StratosFork} The Stratos fork and handlebar assembly hung as a torsional
-        pendulum.}
-    \label{fig:torsional}
 \end{figure}
 \section{LINEAR ANALYSIS}
-Once all of the bicycle parameters have been calculated the canonical matrices
-can be formed and the linear dynamics of the bicycles can be explored. We also
+Once all bicycle parameters have been calculated the canonical matrices
+can be formed and the linear dynamics of the bicycles can be explored. The
+values of the canonical matrices can be found in the second table in Appendix~\ref{sec:partables}. We also
 added the same rigid rider to each bicycle for further comparison. The rigid
-rider was assumed to be in same position and posture for each bicycle relative
+rider was assumed to be in the same position and posture for each bicycle relative
 to the rear wheel contact point.
-\begin{table}[tbp]
+\begin{table}[tb]
     \centering
-    \caption{Mass, center of mass and moment of inertia for the Person relative
+    \caption{Mass, center of mass and moment of inertia for the rider relative
     to the benchmark coordinate system from~\cite{Moore2009a}}
@@
     \includegraphics[width=6in]{../../../plots/Bike/eig_plot.pdf}
-    \caption{The eigenvalues verus speed for all eight bicycles.}
+    \caption{Eigenvalues versus speed for all eight bicycles without the rider.}
     \label{fig:bikeEigPlot}
 \end{figure}
-The eigenvalues can be plotted versus forward speed.
+The eigenvalues of the bicycles with (Fig.~\ref{fig:bikeEigPlot}) and without
+(Fig.~\ref{fig:bikeRiderEigPlot} the rider can be plotted versus forward speed.
 Figure~\ref{fig:bikeEigPlot} shows that the bikes have the typical
-characteristics of the Whipple bicycle model: four real roots at very slow
-speeds, two of which are unstable; a complex pair that is unstable and lower
-speeds and stable at higher speeds; and a root that is sligthly unstable at
-higher speeds. The one noticable difference is that the capsize and caster
-modes colasece into a complex pair between about 0.5 and 3 m/s. The frequency
-of oscillation is similar to that of the weave mode. The root locus in the real
+characteristics of the benchmark bicycle: four real roots at very slow
+speeds, two of which are unstable; a complex pair that is unstable at lower
+speeds and stable at intermediate speeds; and a root that is mildly unstable at
+higher speeds. The one noticeable difference is that the capsize and caster
+modes are contained in a complex pair between about 0.5 and 3 m/s. The frequency
+of oscillation is of comparable magnitude to that of the weave mode. But, the root locus in the real
 and imaginary plane, Fig.~\ref{fig:rootloci}, shows that the mode damps out
 quickly. Examining the eigenvectors reveals that the mode is steer leading
-roll with a 90 degree phase, both of their magitudes being similar,
+roll with a 90 degree phase, both of their magnitudes being similar,
 Fig~\ref{fig:evec}.
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \begin{center}
@@
 \end{figure}
-\begin{figure}[tbp]
+\begin{figure}[tb]
     \begin{center}
@@
     \end{center}
-    \caption{The eigenvector components for the mode with the second complex
-    pair at low speeds.}
+    \caption{Eigenvector components for the second complex mode
+    pair at low speed.}
     \label{fig:evec}
@@
 With the rider added, the second complex pair disappears and the bikes have the
-typical characteristics of the Whipple model. Reversing the fork on the yellow
-bike lowers the weave critical speed. Also the addition of weight to the rear
-rack of the Browser does blank.
+typical characteristics of the benchmark bicycle model. Reversing the fork on the yellow
+bike lowers the weave critical speed and increases the stable speed range.
+Also, the addition of weight to the rear rack of the Browser does little to the
+eigenvalues.
 \begin{figure}[tbp]
@@
     \includegraphics[width=6in]{../../../plots/BikeRider/eig_plot.pdf}
-    \caption{The eigenvalues verus speed for all eight bicycles with the same
-    rider.}
+    \caption{Eigenvalues versus speed for all eight bicycles with the same
+    rigid rider.}
     \label{fig:bikeRiderEigPlot}
@@
 \subsection{FREQUENCY RESPONSE}
-The frequency response of the bicycles and bicycle with rider also reveal some
-interesting things. In the steer torque to roll Bode diagram the magnitude
-diffence among bicycles can vary up to 10 dB (0.148 radians per newton meter of
-torque). The difference in the frequency response for the bicycle with the
+The frequency response of the bicycles (Fig.~\ref{fig:bikeBode}) and bicycle
+with rider (Fig.~\ref{fig:bikeRiderBode}) also reveal some
+interesting things. In the steer-torque-to-roll Bode diagram the magnitude
+difference among bicycles can vary up to 10 dB (or about 8.5 degrees per Newton-meter of
+torque) for the particular speed shown. The difference in the frequency response for the bicycle with the
 rigid rider shows less variation among the bicycles,
-Fig.~\ref{fig:bikeRiderBode}, as the rider's mass and
-inertia play a bigger roll.
+Fig.~\ref{fig:bikeRiderBode}, as the rider's mass and inertia play a larger roll.
 \begin{figure}[tbp]
@@
     \includegraphics[width=6in]{../../../plots/Bike/Bode/Tdel2phi.pdf}
-    \caption{The eigenvalues verus speed for all eight bicycles with the same
-    rider.}
+    \caption{The frequency response for steer-torque-to-roll for all eight
+    bicycles without the rider at 2 m/s.}
     \label{fig:bikeBode}
@@
     \includegraphics[width=6in]{../../../plots/BikeRider/Bode/Tdel2phi.pdf}
-    \caption{The eigenvalues verus speed for all eight bicycles with the same
-    rider.}
+    \caption{The frequency response for steer-torque-to-roll for all eight
+    bicycles with the same rigid rider at 2 m/s.}
     \label{fig:bikeRiderBode}
@@
 \section{CONCLUSION}
-We present a detailed method of accurately estimating the physical parameters
-of a bicycle need for the benchmarked Whipple bicycle
-model~\cite{Meijaard2007}. Then we measure eight different bicycles providing
-both the raw data, parameter sets and linear model coefficient matrices with
-the uncertainties for each value. Then we present a brief comparison of the
-eight bicycles using eigenanalysis and Bode analysis.
+We have presented a detailed method to accurately estimate the physical parameters
+of a bicycle needed for the benchmarked Whipple bicycle
+model~\cite{Meijaard2007}. We measured eight different bicycles providing
+both the parameter sets and linear model coefficient matrices
+for the bicycles alone and the bicycles with the same
+rigid rider. The uncertainties in the parameters and matrix coefficients are
+included for the bicycle alone. Finally, we have presented a brief comparison of the
+eight bicycles using eigenanalysis and Bode frequency response.
 \section{ACKNOWLEDGEMENTS}
-This work was partially funded by the National Science Foundation.
+This material is based upon work partially supported by the National Science
+Foundation under Grant No. 0928339.
 
@@
 \appendix
+\section{PARAMETER TABLES}
+\label{sec:partables}
+The tabulated values for the both the physical parameters and the canonical matrix
+coefficients are shown in the following four tables. The uncertainties in the
+estimations of both the parameters and coefficients are also shown for the
+bicycle without a rider.
 %\section{RAW DATA}
@@
 \begin{landscape}
+
+\begin{table}[tb]
+\caption{The parameters for the eight bicycles with uncertainties in the
+estimations.}
 \subfile{../../../tables/Bike/Parameters/MasterParTable.tex}
-%\end{landscape}
-%\subsection{BICYCLE WITH RIDER}
-%\begin{landscape}
+\label{tab:bicyclePar}
+\end{table}
+
+\begin{table}[tb]
+\caption{The canonical matrix coefficients for the eight bicycles with the
+uncertainty in the estimations.}
+\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
+\label{tab:bicycleCan}
+\end{table}
+
+\begin{table}[tb]
+\caption{The parameters for the eight bicycles with the same rigid rider.}
 \subfile{../../../tables/BikeRider/Parameters/MasterParTable.tex}
-%\end{landscape}
-%\section{CANONICAL MATRICES}
-%\subsection{BICYCLE}
-%\begin{landscape}
-\subfile{../../../tables/Bike/Canonical/MasterCanTable.tex}
-%\end{landscape}
-%\subsection{BICYCLE WITH RIDER}
-%\begin{landscape}
+\label{tab:bicycleRiderPar}
+\end{table}
+
+\begin{table}[tb]
+\caption{The canonical matrix coefficients for the eight bicycles with the
+rigid rider.}
 \subfile{../../../tables/BikeRider/Canonical/MasterCanTable.tex}
+\label{tab:bicycleRiderCan}
+\end{table}
+
 \end{landscape}

### 2011-04-28 PhysicalParameters.py
added the templates

@@ -140,7 +140,10 @@ def mat2dic():
                  #else it is the data which is formatted correctly
                 else:
                     pendDat[k] = v
-        f = open('data/pendDat/p/' + name[:-4] + '.p', 'w')
+        pdir = os.path.join('data', 'pendDat', 'p')
+        if not os.path.exists(pdir):
+            os.makedirs(pdir)
+        f = open(os.path.join(pdir, name[:-4] + '.p'), 'w')
         pickle.dump(pendDat, f, protocol=2)
         f.close()
 

### 2011-05-11 other/fit_data_example.py
moved things around, putting the old code in a sub folder

@@ -1,43 +1,0 @@
-from pylab import *
-from scipy import *
-import scipy.optimize as op
-
-# Generate data points with noise
-num_points = 150
-Tx = linspace(5., 8., num_points)
-Ty = Tx
-
-tX = 11.86*cos(2*pi/0.81*Tx-1.32) + 0.64*Tx+4*((0.5-rand(num_points))*exp(2*rand(num_points)**2))
-tY = -32.14*cos(2*pi/0.8*Ty-1.94) + 0.15*Ty+7*((0.5-rand(num_points))*exp(2*rand(num_points)**2))
-# Fit the first set
-fitfunc = lambda p, x: p[0]*cos(2*pi/p[1]*x+p[2]) + p[3]*x # Target function
-errfunc = lambda p, x, y: fitfunc(p, x) - y # Distance to the target function
-p0 = [-15., 0.8, 0., -1.] # Initial guess for the parameters
-p1, success = op.leastsq(errfunc, p0[:], args=(Tx, tX))
-
-time = linspace(Tx.min(), Tx.max(), 100)
-plot(Tx, tX, "ro", time, fitfunc(p1, time), "r-") # Plot of the data and the fit
-# Fit the second set
-p0 = [-15., 0.8, 0., -1.]
-p2,success = op.leastsq(errfunc, p0[:], args=(Ty, tY))
-
-time = linspace(Ty.min(), Ty.max(), 100)
-plot(Ty, tY, "b^", time, fitfunc(p2, time), "b-")
-
-# Legend the plot
-title("Oscillations in the compressed trap")
-xlabel("time [ms]")
-ylabel("displacement [um]")
-legend(('x position', 'x fit', 'y position', 'y fit'))
-
-ax = axes()
-
-text(0.8, 0.07,
-             'x freq :  %.3f kHz \n y freq :  %.3f kHz' % (1/p1[1],1/p2[1]),
-                  fontsize=16,
-                       horizontalalignment='center',
-                            verticalalignment='center',
-                                 transform=ax.transAxes)
-
-show()
-

### 2011-05-11 PhysicalParameters/other/fit_data_example.py
some init files

@@ -1,0 +1,43 @@
+from pylab import *
+from scipy import *
+import scipy.optimize as op
+
+# Generate data points with noise
+num_points = 150
+Tx = linspace(5., 8., num_points)
+Ty = Tx
+
+tX = 11.86*cos(2*pi/0.81*Tx-1.32) + 0.64*Tx+4*((0.5-rand(num_points))*exp(2*rand(num_points)**2))
+tY = -32.14*cos(2*pi/0.8*Ty-1.94) + 0.15*Ty+7*((0.5-rand(num_points))*exp(2*rand(num_points)**2))
+# Fit the first set
+fitfunc = lambda p, x: p[0]*cos(2*pi/p[1]*x+p[2]) + p[3]*x # Target function
+errfunc = lambda p, x, y: fitfunc(p, x) - y # Distance to the target function
+p0 = [-15., 0.8, 0., -1.] # Initial guess for the parameters
+p1, success = op.leastsq(errfunc, p0[:], args=(Tx, tX))
+
+time = linspace(Tx.min(), Tx.max(), 100)
+plot(Tx, tX, "ro", time, fitfunc(p1, time), "r-") # Plot of the data and the fit
+# Fit the second set
+p0 = [-15., 0.8, 0., -1.]
+p2,success = op.leastsq(errfunc, p0[:], args=(Ty, tY))
+
+time = linspace(Ty.min(), Ty.max(), 100)
+plot(Ty, tY, "b^", time, fitfunc(p2, time), "b-")
+
+# Legend the plot
+title("Oscillations in the compressed trap")
+xlabel("time [ms]")
+ylabel("displacement [um]")
+legend(('x position', 'x fit', 'y position', 'y fit'))
+
+ax = axes()
+
+text(0.8, 0.07,
+             'x freq :  %.3f kHz \n y freq :  %.3f kHz' % (1/p1[1],1/p2[1]),
+                  fontsize=16,
+                       horizontalalignment='center',
+                            verticalalignment='center',
+                                 transform=ax.transAxes)
+
+show()
+

### 2011-05-11 PhysicalParameters/__init__.py
updating the init files

@@ -1,0 +1,1 @@
+from PhysicalParameters import *
